---
title: "Microplate_Sulfide_Code"
author: "Alia Al-Haj"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    df_print: paged
geometry: left=2cm,right=2cm,top=1cm,bottom=2cm
output_dir: "2025"
---

## Information

```{r}
#######################################################################################
####### GENX2 
####### Brackish
####### Data Analysis Code: Porewater sulfide 
####### 
##############################################################################################


############################# Information ##################################
#Author: Alia N. Al-Haj/Melanie Giessner
#Edited: 202510


```

#Load Libraries

```{r,include=FALSE}
#packages:
library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
library(purrr)
library(tidyverse)
library(here)
library(data.table)
library(matrixStats)
library(gridExtra)
library(grid)
library(plater)
library(raster)
library(knitr)  
library(readxl)

```

## Run info

```{r run info}
#identify which section you are in
cat("Run Information")

###things that need to be changed
Date_Run = "20250827"
Plate="Plate3"
Run_by = "Melanie Giessner"#Instrument user
Script_run_by ="Melanie Giessner"#Code user
Project = "GENX2"
Experiment="Experiment 4"
Run_notes=""#any notes from run

#list Stds that should be excluded
stds_to_remove<-c(NA)
#list of points that should be excluded
Points_to_remove<-c(NA)

```

##file names

```{r files names,include=FALSE}

###### File Names - - PLEASE CHECK, should automatically update based on Date and Plate Number
#file path and name for raw summary data file 
file_path_raw = paste0("2025/",Experiment,"/Tidy Data/",Date_Run,"_GENX2_H2S_",Plate,".csv")

#file path name for processed data
summarizeddata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",Plate,"_Summary_Data.csv")
fulldata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",Plate,"_Data.csv")
shortdata_path=paste0("2025/",Experiment,"/QAQC'd data/GENX2_H2S_",Date_Run,"_",Plate,"_short_Data.csv")
need_dilute=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",Plate,"_SamplesNeedDilution_Data.csv")
high_CV=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",Plate,"_SamplesHighCV_Data.csv")

###QAQC log path
log_path="S:/GlobalChangeEco/porewater_r scripts & Data/BGC_GCE_PorewaterAnalysis/Sulfide/Sulfide_STD_QAQC.csv"
```

```{r Set Up Code - constants and QAQC cutoffs, include=FALSE}

#Flag cutoffs
  r2_cutoff = 0.985            #this is the level below which we want to rerun or consider a curve 
  cv_flag_stds = 10           #this is the maximum cv allowed for standards
  p_value_chkstds = 0.05      #the p-value for the t-test between check standards must be greater than this
  p_value_MC = 0.05           #the p-value for the t-test between matrix checks and top std must be greater than this
  cv_flag_sample = 10         #this is the maximum cv allowed for samples
  dups_perc_diff = 15.5       #this is the maximum percent difference allowed between duplicates
  high_recovery_cutoff = 120  #this is the maximum percent recovery of SO4 allowed in spiked samples
  low_recovery_cutoff = 80    #this is the minimum percent recovery of SO4 allowed in spiked samples

#Standard concentrations - Update if running different standard curve:
# standard units are in uM
  standards <- tibble(
    IDs = c("Std 0", "Std 1", "Std 2", "Std 3", "Std 4", "Std 5"),
    Conc = c(0, 5.0, 12.5, 25.0, 50.0, 100.0))  #uM

#Spike concentration calc
  #spike for these samples was 50 uL of the 100uM standard
  Con1 <- 1000000               #conversion factor value for spike volumes (uL -> L)
  spk_std <- 100                # uM S2- standard used
  spkvol <- 50                  # uL volume of spike added
  spk_Conc <- spk_std*(spkvol/Con1)  # umoles of S2- added to each spiked sample
  sample_vol <- 250             # the sample volume without the spike is 250 uL
  cline_vol <- 20               # the total volume without the spike is 250 uL

#Top standard Concentration- Update if running different standard curve: 
   Top_STD = 100



```

##Notes


#Files Required: #csv of samples and standard absorbance (1st box), IDs (second box), and dilutions (third box): see template

## Read in data

```{r Read in tidy data,echo=FALSE, include=FALSE}
cat("Import Sample Data")
# Read in raw data csv
dat <- read_plate(
  file = file_path_raw,             # full path to the .csv file
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
)
str(dat)
head(dat)

#Change the headers
colnames(dat) <- c("Wells", "Abs", "IDs", "Dilution")
```

##Standards
```{r first look at Stds,echo=FALSE,fig.keep='none'}
#subset by H2s Stds
H2S_stds_all <- dat %>% subset(IDs %like% "Std"|IDs %like% "MC") 
head(H2S_stds_all)

H2S_stds_all <- left_join(H2S_stds_all, standards, by = "IDs")

##Plot stds and calculate the slope, intercept, and R2
H2S_stds <- na.omit(H2S_stds_all)
H2S <- ggplot(H2S_stds, aes(Conc, Abs)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+
  ggtitle("Sulfide Standard Curve")

H2S
```


```{r Remove high cv stds,echo=FALSE,fig.keep='none'}
#Calculate cv for STDs
H2S_stds1_all <- H2S_stds_all %>%
  group_by(IDs) %>%
  summarise(H2S_mean_Abs = mean(Abs), H2S_sd = sd(Abs), H2S_cv = cv(Abs), 
            Dilution = first(Dilution), 
            Conc = first(Conc))

#Flag high cvs
H2S_stds1_all$H2S_cv_flag <- ifelse(H2S_stds1_all$H2S_cv > cv_flag_stds, 'High CV', 'Within range')

# filter High CV Samples
H2S_stds1_all_HCV <- H2S_stds1_all %>%  
  filter(str_detect(H2S_cv_flag, "High CV"))

#make a dataframe for cvs
cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
colnames(cv_trial)<-c("cv1","cv2","cv3")
#make dataframe for fixed data
H2S_stds_HCV_fixed<-data.frame(matrix(ncol = 7, nrow =0 ))
colnames(H2S_stds_HCV_fixed)<-c("IDs", "H2S_mean_Abs", "H2S_sd", "H2S_cv", "Dilution",  "Conc", "H2S_cv_flag")

#this will break if there are no high cv points
if(!nrow(H2S_stds1_all_HCV)==0){
#make a loop to filter each high cv sample
for(x in 1:nrow(H2S_stds1_all_HCV)){ 
sing_ID<- H2S_stds_all %>% subset(IDs == H2S_stds1_all_HCV$IDs[x])

      #make a dataframe for cvs
      cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
      
      #make a loop to find the cv for each deleted point
      for(i in 1:nrow(sing_ID)){
        sing_ID1<-sing_ID[-i,] 
        cv_trial[i]<-cv(sing_ID1$Abs)
      }
      
      # delete the point that gives the lowest cv
      sing_ID<-sing_ID[-which.min(cv_trial),]
        H2S_stds_HCV_fixed<-rbind(H2S_stds_HCV_fixed,sing_ID)
      }
H2S_stds1_all_fixed <- H2S_stds_HCV_fixed %>%
  group_by(IDs) %>%
  summarise(H2S_mean_Abs = mean(Abs), H2S_sd = sd(Abs), H2S_cv = cv(Abs), 
            Dilution = first(Dilution), 
            Conc = first(Conc))
#Flag high cvs
H2S_stds1_all_fixed$H2S_cv_flag <- ifelse(H2S_stds1_all_fixed$H2S_cv > cv_flag_stds, 'High CV', 'Within range')
#bind data frames back together
H2S_stds1_all <- subset(H2S_stds1_all, (!IDs %in% H2S_stds1_all_fixed$IDs ))
H2S_stds1_all <- rbind(H2S_stds1_all_fixed, H2S_stds1_all)

H2S_stds_all <- subset(H2S_stds_all, (!IDs %in% H2S_stds_HCV_fixed$IDs ))
H2S_stds_all <- rbind(H2S_stds_HCV_fixed, H2S_stds_all)
}
# filter High CV Samples
stds_HCV <- H2S_stds1_all %>%  
  filter(str_detect(H2S_cv_flag, "High CV")) 

knitr::kable(stds_HCV, caption = "High CV Samples")
```


```{r Plot Standard Curve, echo=FALSE,fig.keep='none'}

#Remove check standards which have NAs in the Conc colum
H2S_stds <- na.omit(H2S_stds_all)
H2S_stds1 <- na.omit(H2S_stds1_all)
head(H2S_stds1)

H2S_stds<-H2S_stds %>% subset(!IDs%in%stds_to_remove)

##Plot stds and calculate the slope, intercept, and R2 
H2S <- ggplot(H2S_stds, aes(Conc, Abs)) + geom_point() + geom_smooth(method = "lm", se = FALSE)+
  ggtitle("Sulfide Standard Curve")

H2S

H2S_lm <- lm(H2S_stds$Abs ~ H2S_stds$Conc)
summary(H2S_lm)
cf <- coef(H2S_lm)


#create data frame 
Date <- Date_Run
R2 <-summary(H2S_lm)$adj.r.squared
Slope <- cf[2]
Intercept <- cf[1]
qplate <- data.frame(Date, Project , R2, Slope, Intercept, Top_STD, row.names = NULL)
head(qplate)


#Was the standard curve ok for this plate? 
#Write out to the user whether or not the r2 is above the cutoff
if (qplate$R2 < (r2_cutoff)) {
print("Std Curve r2 is below cutoff! - REASSESS")
} else {
print("Std Curve r2 GOOD")
}
  
#write out a flag to the sample dataframe if the r2 is above the cutoff of 0.98
##This creates the QAQC flag column
dat <- dat %>%
  mutate(
    QAQC_flag = if (qplate$R2 <= r2_cutoff) {
      "Std curve r2 low"
    } else {
      ""
    }
  )

```

# Checking STD Data against QAQC file

```{r Assess Standard Curve, message=FALSE, warning = FALSE, echo=FALSE,fig.keep='none'}
cat("Check std curve against log")

#read in datafile with all the slopes
qlogO <- read.csv(log_path)
qlogO <- qlogO[,c(2:7)]
colnames(qlogO) <- c("Date", "Project", "R2", "Slope", "Intercept", "Top_STD")
qlogO$Date<- as.character(qlogO$Date)
#head(qlogO)


#add data to file and make sure it is not already in the log
qplate<- anti_join(qplate,qlogO ,by = c("Date", "Project", "R2", "Slope", "Intercept"))
qlog<-rbind(qplate,qlogO)
head(qlog)


##plot the slopes to make sure there are no crazy outliers 
slope1 <- ggplot(data=qlog, aes(x=Date, y=Slope)) +
           geom_hline(yintercept= (mean(qlog$Slope)+ (2*sd(qlog$Slope))), linetype="dashed", color = "red", linewidth=2)+
            geom_hline(yintercept= (mean(qlog$Slope)- (2*sd(qlog$Slope))), linetype="dashed", color = "red", linewidth=2)+
            geom_point(aes(size=3)) + 
            theme_classic() + ylim(0, 0.02) + 
            theme(legend.position="none") + 
            ggtitle("Sulfide Slopes")+
            guides(x = guide_axis(angle = 70)) +
            geom_point(aes( x = qplate$Date, y = qplate$Slope,size=3), colour = "orange")
slope1

##Check if the current slope is within 2 std deviations of previous slopes
if (qplate$Slope[1] > (mean(qlog$Slope)+ (2*sd(qlog$Slope)))| qplate$Slope[1] < (mean(qlog$Slope)- (2*sd(qlog$Slope)))) {
print("Std curve slope is 2 sd different from previous slopes! - REASSESS")
} else {
print("Std curve slope is with 2 sd of previous slopes")
}

#Rerun if outside of red lines

##plot the intercepts to make sure there are no crazy outliers 
int1 <- ggplot(data=qlog, aes(x=Date, y=Intercept)) +
  geom_hline(yintercept= (mean(qlog$Intercept)+ (2*sd(qlog$Intercept))), linetype="dashed", color = "red", linewidth=2)+
  geom_hline(yintercept= (mean(qlog$Intercept)- (2*sd(qlog$Intercept))), linetype="dashed", color = "red", linewidth=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.05,0.125) + 
  theme(legend.position="none")+ 
  ggtitle("Sulfide Intercepts")+
  guides(x = guide_axis(angle = 70))+
            geom_point(aes( x = qplate$Date, y = qplate$Intercept,size=3), colour = "orange")

int1

##Check if the current intercept is within 2 std deviations of previous intercepts
if (qplate$Intercept[1] > (mean(qlog$Intercept)+ (2*sd(qlog$Intercept)))| qplate$Intercept[1] < (mean(qlog$Intercept)- (2*sd(qlog$Intercept)))) {
print("Std curve intercept is 2 sd different from previous intercepts! - REASSESS")
} else {
print("Std curve intercept is with 2 sd of previous intercepts")
}

#plot the R2s to make sure there are no crazy outliers 
Rsq1 <- ggplot(data=qlog, aes(x=Date, y=R2)) +
  geom_hline(yintercept= (0.98), linetype="dashed", color = "red", linewidth=2)+
  geom_point(aes(size=3)) + 
  theme_classic() + ylim(0.96, 1.01) + 
  theme(legend.position="none")+ 
  ggtitle("Sulfide R2s")+
  guides(x = guide_axis(angle = 70))+
            geom_point(aes( x = qplate$Date, y = qplate$R2,size=3), colour = "orange")

Rsq1


```

```{r Plot standard graphs, echo = FALSE, warning = FALSE, message = FALSE, fig.width=8, fig.height=12}

knit_print((paste("Std Curve R squared:", round(qplate$R2, 4))))

##Plot std graphs together
ggarrange(H2S,slope1,int1,Rsq1, nrow=2, ncol=2)

```

## Check standards QAQC

```{r Assess Check Standards, echo=FALSE}
cat("Compare CheckStds to Stds")
#Make sure check standards are not different from standard concentration
#subset Check Standards from the rest of the dataset
H2S_std_Chk <- H2S_stds_all %>%
  filter(grepl("ChkStd", IDs))

Chkstds_HCV <- H2S_stds1_all%>%
  filter(grepl("ChkStd", IDs)) %>%  
  filter(str_detect(H2S_cv_flag, "High CV"))
head(Chkstds_HCV)

#Are the check standards significantly different from the standards?
#subset datasets for comparison
std0 <- subset(H2S_stds, IDs == "Std 0")
Chkstd0 <- subset(H2S_std_Chk, IDs == "ChkStd 0")
std3 <- subset(H2S_stds, IDs == "Std 3")
Chkstd3 <- subset(H2S_std_Chk, IDs == "ChkStd 3")
std4 <- subset(H2S_stds, IDs == "Std 4")
Chkstd4 <- subset(H2S_std_Chk, IDs == "ChkStd 4")

#Check std 0 with T-test
if(!"Std 0" %in% stds_to_remove){
t.test.std0 <- t.test(std0$Abs,Chkstd0$Abs,var.equal = T)
t.test.std0

if(t.test.std0$p.value > p_value_chkstds){
print("Chk Std 0 GOOD")
} else {
print("Chk Std 0 is significantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
if (t.test.std0$p.value < p_value_chkstds) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; Chk Std 0 out of range"),
    "Chk Std 0 out of range"
  )
}
##Add p-values to dataframes
std0$p_value <- t.test.std0$p.value
}else{
std0$p_value <- c(NA) 
print("Std excluded")
}

#Check std 3 with T-test
if(!"Std 3" %in% stds_to_remove){
t.test.std3 <- t.test(std3$Abs,Chkstd3$Abs,var.equal = T)
t.test.std3

if(t.test.std3$p.value > p_value_chkstds){
print("Chk Std 3 GOOD")
} else {
print("Chk Std 3 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std is Bad
if (t.test.std3$p.value < p_value_chkstds) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; Chk Std 3 out of range"),
    "Chk Std 3 out of range"
  )
}
##Add p-values to dataframes
std3$p_value <- t.test.std3$p.value
}else{
  print("Std excluded")
std3$p_value <- c(NA) 
}


#Check std 4 with T-test
if(!"Std 4" %in% stds_to_remove){
t.test.std4 <- t.test(std4$Abs,Chkstd4$Abs,var.equal = T)
t.test.std4

if(t.test.std4$p.value > p_value_chkstds){
print("Chk Std 4 GOOD")
} else {
print("Chk Std 4 is signficantly different from Std - REASSESS")
}
#write out a flag to the sample dataframe if the Chk std 4is Bad
if (t.test.std4$p.value < p_value_chkstds) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; Chk Std 4 out of range"),
    "Chk Std 4 out of range"
    )
    }
##Add p-values to dataframes
std4$p_value <- t.test.std4$p.value
}else{
  std4$p_value <- c(NA) 
  print("Std excluded")
}

##Check Stds Plot
##plot MC Abs vs Std 5 Abs and color based on whether passed T test

##Combine std checks
QAstdchk <- rbind(std0, std3, std4)

#Calculate p-values
QAstdchk$p_flag <-  ifelse(QAstdchk$p_value > p_value_chkstds, 'OK', 'Rerun')

#plot output as a bar graph to easily check
ChkStdbar <- ggplot(data = QAstdchk, aes(x = IDs, y = p_value/3, fill=p_flag)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values = c("OK" = "darkgreen", "Rerun" = "darkred"))+
        theme_classic() + labs(x= "Sample ID", y="T-test p-value") + 
        theme(legend.position="none") +  geom_hline(yintercept=0.05, linetype="dashed", 
                color = "black", linewidth=1)+ 
        ggtitle("Check Standards")

ChkStdbar
```

## Matrix Check QAQC

```{r Matrix Checks,echo=FALSE}
cat("Compare Matrix Checks to Stds")
H2S_MC <- H2S_stds_all %>%  
  filter(str_detect(IDs, "MC")) 

H2S_MC_HCV <- H2S_stds1_all%>%
  filter(grepl("MC", IDs)) %>%  
  filter(str_detect(H2S_cv_flag, "High CV"))
head(H2S_MC_HCV)

##Select Std 5 for comparison to MCs
std5 <- subset(H2S_stds, IDs == "Std 5")
MC10 <- subset(H2S_MC, IDs == "MC: 10ppt S5")
MC20 <- subset(H2S_MC, IDs == "MC: 20ppt S5") 

#10ppt matrix check
if("MC: 10ppt S5" %in% MC10$IDs){
t.test.MC10 <- t.test(std5$Abs,MC10$Abs,var.equal = T)
t.test.MC10

if(t.test.MC10$p.value > p_value_MC){
print("Matrix Check 10 GOOD")
} else {
print("Matrix Check 10 is signficantly different from Std 5 - REASSESS")
}
#write out a flag to the sample dataframe if the MC is Bad
if (t.test.MC10$p.value < p_value_MC) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; MC 10 out of range"),
    "MC 10 out of range"
  )
}
##Add p-value to dataframe
MC10$p_value <- t.test.MC10$p.value
}else{
  print("Did not run 10ppt matrix")
}

#20ppt matrix check
if("MC: 20ppt S5" %in% MC20$IDs){
t.test.MC20 <- t.test(std5$Abs,MC20$Abs,var.equal = T)
t.test.MC20
 
if(t.test.MC20$p.value > p_value_MC){
print("Matrix Check 20 GOOD")
} else { 
print("Matrix Check 20 is signficantly different from Std 5 - REASSESS")
}
#write out a flag to the sample dataframe if the MC is Bad
if (t.test.MC20$p.value < p_value_MC) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; MC 20 out of range"),
    "MC 20 out of range"
  )
}
##Add p-value to dataframe
MC20$p_value <- t.test.MC20$p.value
}else{
  print("Did not run 20ppt matrix")
 }


##Matrix Checks Plot
##plot MC Abs vs Std 5 Abs and color based on whether passed T test

##Combine matrix checks
QAMC <- rbind(MC10, MC20)

#Calculate p-values
QAMC$p_flag <-  ifelse(QAMC$p_value > p_value_MC, 'OK', 'Rerun')

# head(QAdups)

#plot dups output as a bar graph to easily check - want any over 10% to be red need to work on this 
MCbar <- ggplot(data = QAMC, aes(x = IDs, y = p_value/3, fill=p_flag)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values = c("OK" = "darkgreen", "No, rerun" = "darkred"))+
        theme_classic() + labs(x= "Sample ID", y="T-test p-value") + 
        theme(legend.position="none") +  geom_hline(yintercept=0.05, linetype="dashed", 
                color = "black", linewidth=1)+ 
        ggtitle("Matrix Effects")

MCbar

```

## Calculate Sulfide Concentrations

```{r,echo=FALSE}
cat("Calculate Sulfide Concentrations")

cat("Subset the data")
#filter out stds
dat <- dat %>%  
      filter(!str_detect(IDs, "Std")) %>%  
      filter(!str_detect(IDs, "MC"))

#head(dat)

#Calculate concentrations of Sulfide
dat$Conc <- ((dat$Abs-cf[1])/cf[2])*(dat$Dilution)
#head(dat)

#Use ifelse to make any negative values equal to zero
dat$H2S_Conc_Final <- ifelse(dat$Conc <0, 0, dat$Conc)

#Use ifelse to mark samples that were below detection limit and changed to zero
dat$H2S_info <- ifelse(dat$Conc <=0, 'bdl', ifelse(dat$Abs > mean(std5$Abs), "adl", "Within_Range")) 

head(dat)
```

## Calculate Averages across wells and std. dev.

```{r,echo=FALSE}

cat("Calculate Averages across wells and std. dev.")
#head(dat)

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1 <- dat %>%
  group_by(IDs) %>%
  summarise(H2S_mean = mean(H2S_Conc_Final), H2S_sd = sd(H2S_Conc_Final), H2S_cv = cv(H2S_Conc_Final),
            QAQC_flag = first(QAQC_flag),
            Dilution = first(Dilution))

#head(dat1)

#Flag high cvs 
dat1$H2S_cv_flag <- ifelse(dat1$H2S_cv > cv_flag_sample, 'High CV', 'within range')

## to flag high dilutions 
dat1$H2S_flag <- ifelse(dat1$H2S_mean <=0, 'bdl', ifelse(dat1$H2S_mean > ((mean(std5$Abs)-cf[1])/cf[2])*(dat1$Dilution), "adl", "Within_Range"))


head(dat1)


#plot data and sd's just to check and see what they look like - just a quick first look 
H2S <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color=H2S_flag))+ 
  geom_point(size=4)+ 
       scale_color_manual(values = c("abl" = "red", "bdl" = "orange", "Within_Range" = "black")) +  theme_classic() + 
  labs(y="Sulfide (uM)", x="Sample ID",title=paste(Project, Experiment," Sulfide Data")) + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
H2S

H2S_original <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color = H2S_cv_flag))+ 
  geom_point(size=4) +  theme_classic() + 
  scale_color_manual(values = c("Within range" = "darkgreen", "High CV" = "darkred"))+
  labs(y="Sulfide (uM)", x="Sample ID") + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70)) + 
  ggtitle("Sample triplicate means and sd dev before bad reps removed") 
H2S_original

```

## Remove bad reps

```{r remove bad reps,echo=FALSE}
cat("Remove bad replicates")

#auto remove bad reps
# filter High CV Samples
dat1_HCV <- dat1 %>%  
  filter(str_detect(H2S_cv_flag, "High CV")) 

dat_HCV <- subset(dat, (IDs %in% dat1_HCV$IDs ))


#make a dataframe for cvs
cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
colnames(cv_trial)<-c("cv1","cv2","cv3")
#make dataframe for fixed data
dat_HCV_fixed<-data.frame(matrix(ncol = 8, nrow =0 ))
colnames(dat_HCV_fixed)<-c("Wells", "Abs", "IDs", "Dilution", "QAQC_flag", "Conc", "H2S_Conc_Final", "H2S_info")

#make a loop to filter each high cv sample
for(x in 1:nrow(dat1_HCV)){ 
sing_ID<- dat_HCV %>% subset(IDs == dat1_HCV$IDs[x])
#make a dataframe for cvs
      cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
      
      #make a loop to find the cv for each deleted point
      for(i in 1:nrow(sing_ID)){
        sing_ID1<-sing_ID[-i,] 
        cv_trial[i]<-cv(sing_ID1$Abs)
      }
      
      # delete the point that gives the lowest cv
      sing_ID<-sing_ID[-which.min(cv_trial),]
        dat_HCV_fixed<-rbind(dat_HCV_fixed,sing_ID)
      }

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1_HCV_fixed <- dat_HCV_fixed %>%
  group_by(IDs) %>%
  summarise(H2S_mean = mean(H2S_Conc_Final), H2S_sd = sd(H2S_Conc_Final), H2S_cv = cv(H2S_Conc_Final),
            QAQC_flag = first(QAQC_flag),
            Dilution = first(Dilution))

#Flag high cvs 
dat1_HCV_fixed$H2S_cv_flag <- ifelse(dat1_HCV_fixed$H2S_cv > cv_flag_sample, 'High CV', 'within range')

## to flag high dilutions 
dat1_HCV_fixed$H2S_flag <- ifelse(dat1_HCV_fixed$H2S_mean <=0, 'bdl', ifelse(dat1_HCV_fixed$H2S_mean > ((mean(std5$Abs)-cf[1])/cf[2])*(dat1_HCV_fixed$Dilution), "adl", "Within_Range"))

dat1 <- subset(dat1, (!IDs %in% dat1_HCV_fixed$IDs ))
dat1 <- rbind(dat1_HCV_fixed, dat1)
head(dat1)


#plot data and sd's just to check and see what they look like - just a quick first look 
H2S <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color=H2S_flag))+ 
  geom_point(size=4)+ 
       scale_color_manual(values = c("abl" = "red", "bdl" = "orange", "Within_Range" = "black")) +  theme_classic() + 
  labs(y="Sulfide (uM)", x="Sample ID",title=paste(Project, Experiment," Sulfide Data")) + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
H2S

H2S_fixed <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color = H2S_cv_flag))+ 
  geom_point(size=4) +  theme_classic() + 
  scale_color_manual(values = c("Within range" = "darkgreen", "High CV" = "darkred"))+
  labs(y="Sulfide (uM)", x="Sample ID") + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70)) + 
  ggtitle("Sample triplicate means and sd dev after bad reps removed") 
H2S_fixed

#If samples with cv >10 rerun those samples 
H2S_HighCVRerun <- subset(dat1, H2S_cv_flag == "High CV")
H2S_DiluteRerun <- subset(dat1, H2S_flag == "adl")
H2S_bdl <- subset(dat1,H2S_flag == "bdl")

knitr::kable(H2S_HighCVRerun, caption = "High CV Samples")
knitr::kable(H2S_DiluteRerun, caption = "Samples Above the Detection Limit")
knitr::kable(H2S_bdl, caption = "Samples Below the Detection Limit")


```

## Check the dups for QAQC

```{r,echo=FALSE}
cat("Assess Duplicates")

#Show me the data that we have from the calculations 
#head(dat1)

#pull out any rows that have "Dup" in the ID column
dups <- dat1 %>%  
  filter(str_detect(IDs, "Dup")) 
#head(dups)

#remove these from dat1
dat2 <- dat1 %>%  
  filter(!str_detect(IDs, "Dup")) %>%  
  filter(!str_detect(IDs, "Spike")) 
#head(dat2)

#remove the dup from these IDs so we will have duplicates 
dups$IDs<-gsub(" Dup",'',dups$IDs)
head(dups)
colnames(dups) <- c('IDs', 'mean_dup')


#put it back together with the old data set and look for duplicates 
QAdups <- merge(dat2, dups)
#head(QAdups)

QAdups$dups_chk <- ((abs(QAdups$H2S_mean-QAdups$mean_dup))/((QAdups$H2S_mean+QAdups$mean_dup)/2))*100
QAdups$dups_flag <-  ifelse(QAdups$dups_chk <dups_perc_diff, 'OK', 'Rerun')

head(QAdups)

#plot dups output as a bar graph to easily check - want any over 10% to be red need to work on this 
dupsbar <- ggplot(data = QAdups, aes(x = IDs, y = dups_chk, fill=dups_flag)) +
       geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK" ='darkgreen' , "Rerun" = 'darkred')) + 
        theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
        theme(legend.position="none") +  geom_hline(yintercept=10, linetype="dashed", 
                color = "black", linewidth=1)

dupsbar

#check for any one's that would warrant reruns 

Baddups <- subset(QAdups, dups_flag == "Rerun")

#write out a flag to the sample dataframe if any dups have percent differences out of range 
if (nrow(Baddups) > 0) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; Dup perc diff out of range"),
    "Dup perc diff out of range"
  )
}

if (nrow(Baddups) > 0) {
    dat1$QAQC_flag <- ifelse(
    dat1$QAQC_flag != "",
    paste0(dat1$QAQC_flag, "; Dup perc diff out of range"),
    "Dup perc diff out of range"
  )
}
```

## Check the spks for QAQC

```{r,echo=FALSE}

cat("Assess Spikes")

#Show me the data that we have from the calculations 
#head(dat1)

#pull out any rows that have "d" in the SampleID column
spks <- dat1 %>%  
  filter(str_detect(IDs, "Spike")) 
#head(spks)

#remove these from dat1
dat2 <- dat1 %>%  
  filter(!str_detect(IDs, "Dup")) %>%  
  filter(!str_detect(IDs, "Spike")) 
#head(dat2)

#remove the Spike from these IDs so we will have duplicates 
spks$IDs<-gsub(" Spike","",spks$IDs)
#head(spks)
colnames(spks) <- c('IDs', 'mean_spk')


#put it back together with the old data set and look for duplicates 
QAspks <- merge(dat2, spks)
#head(QAspks)

#now we need to calculate the spike concentration and calculate the spike recovery 
#spike for these samples was 10 uL of the 100uM standard
QAspks$unspkd <- (QAspks$H2S_mean/QAspks$Dilution)*((sample_vol )/Con1) #gives us the total S2- in the sample in umoles
QAspks$spkd <-    (QAspks$mean_spk/QAspks$Dilution)*((sample_vol + spkvol)/Con1)        ##total S2- in spiked sample in umoles 
QAspks$expctd_spkd <-  (QAspks$unspkd + spk_Conc)
QAspks$spk_recovery <-    (QAspks$spkd/QAspks$expctd_spkd)*100
QAspks$spks_flag <-  ifelse(QAspks$spk_recovery > low_recovery_cutoff & QAspks$spk_recovery < high_recovery_cutoff, 'OK', 'NO, rerun')  
 

head(QAspks)

#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 
spksbar <- ggplot(data = QAspks, aes(x = IDs, y = spk_recovery, fill=spks_flag)) +
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK"="darkgreen","NO, rerun"="darkred")) + 
        theme_classic() + labs(x= "Sample ID", y="Spike Recovery (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=low_recovery_cutoff, linetype="dashed", color = "black", linewidth=1) + 
        geom_hline(yintercept=high_recovery_cutoff, linetype="dashed", color = "black", linewidth=1)

spksbar

#check for any no's that would warrant reruns! 

Badspks <- subset(QAspks, spks_flag == "NO, rerun")

#write out a flag to the sample dataframe if any spks have recovery out of range 
if (nrow(Badspks) > 0) {
    dat$QAQC_flag <- ifelse(
    dat$QAQC_flag != "",
    paste0(dat$QAQC_flag, "; Spk recovery out of range"),
    "Spk recovery out of range"
  )
}

#write out a flag to the sample dataframe if any spks have recovery out of range 
if (nrow(Badspks) > 0) {
    dat1$QAQC_flag <- ifelse(
    dat1$QAQC_flag != "",
    paste0(dat1$QAQC_flag, "; Spk recovery out of range"),
    "Spk recovery out of range"
  )
}
```

## Export full data then just final data

```{r,echo=FALSE}
dat_IDs <- dat1

#Read out the summarized data
head(dat_IDs)
  write.csv(dat_IDs, file= summarizeddata_path)

#Read out all the data in dat 
head(dat)
write.csv(dat, file= fulldata_path)

#Now take out the absorbance and stuff 
head(dat)
dat3 <- dat[ ,-(1:2)]
dat3 <- dat3[ ,-(2:5)]
head(dat3)
write.csv(dat3, file= shortdata_path)

write.csv(H2S_DiluteRerun, need_dilute)
write.csv(H2S_HighCVRerun, high_CV)

#save to QAQC Data to QAQC File
write.csv(qlog, file= log_path)


```

### END
