---
title: "SMARTX: Porewater Ammonium"
author: "May 2024 Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information 
```{r run information, include=TRUE}
cat("Run Information: Stephanie J. Wilson ") #lets you know what section you're in 
#set the run date & user name 
  run_date <- "09/19/2024"
  sample_year <- "2024"
  sample_month <- "May"
  user <- "Max Wegner"

#identify the files you want to read in 
  #read in as a list to accommodate multiple runs in a month
  NH3_files <- c("Raw Data/May_2024_SMARTX_NH4_Run1.csv",
                 "Raw Data/May_2024_SMARTX_NH4_Run2.csv",
                 "Raw Data/May_2024_SMARTX_NH4_Run3.csv",
                 "Raw Data/May_2024_SMARTX_NH4_Run4.csv")

# Define the file path for QAQC log file - NO Need to change just check year 
  file_path <- "Raw Data/SEAL_SMARTX_QAQC_Log_2024.csv"
  final_path <- "Processed Data/GCReW_SMARTX_Porewater_NH4_202405.csv"

#record any notes about the run or anything other info here: 
  run_notes <- "Check standards were double the expected concentration for these runs. "
  
#Set up file path for metadata 
  #downloaded metadata csv - downloaded from Google drive as csv for this year
  #Raw_Metadata = "Raw Data/Sample_IDs_LTREB_SMARTX_2024.csv"
  #Raw_Metadata = "..GCReW/GCReW_Project_Treatment_Metadata.csv"
  library(here)
  Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")
  
  #Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

  cat(run_notes)
```

##Setup
```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#load/install packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  ggplot2,
  ggpubr,
  tidyr,
  tidyverse,
  lubridate,
  tinytex,
  stringr,
  writexl,
  readr,
  readxl,
  purrr,
  tinytex,
  broom)

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
P_mw <- 30.973762  # molecular weight of P 
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 

#Detection limit and top standards for flagging 
NOx_dl <- 0.025  #mgL
NH3_dl <- 0.02  #mgL
PO4_dl <- 0.003 #mgL

NOx_top <- 1  #mgL
NH3_top <- 2  #mgL
PO4_top <- 0.3 #mgL

#Check Standard concnetrations 
NOx_CCV <- 0.5 
NH3_CCV <- 2.0 
PO4_CCV <- 0.15 

#Spike Concentrations 
NOx_Spk <- 0.5 #ppm or mg/L
NH3_Spk <- 1
PO4_Spk <- 0.152 

#expected ranges for sample concentrations used for flags 
r2_cutoff = 0.990
chk_flag = 0.25
chk_conc_flag = 15 #cutoff level for percent difference of check stds vs. expected concentration 
chks_flag = 60
rep_flag = 25  #this is a 25% error between samples
#blank_flag is calculated based on samples later in this code

#pe check Concentrations
NH3_pe <- 1.034
NOx_pe <- 1.51
PO4_pe <- 0.824

```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata)

#clean up columns in metadata file:
  #Pull out only SMARTX Rows 
  #Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("SMARTX", Project, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Project, "_", Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
    rename(
    Sample_ID = Sample.ID,
  ) %>%
  select(-N_Treatment)

#create a column for C3 or C4 plants 
#metadata <- metadata %>%
#  mutate(
#    Nickname = case_when(
#      startsWith(as.character(Plot), "3") ~ "C3",
#      startsWith(as.character(Plot), "6") ~ "C3",
#      startsWith(as.character(Plot), "4") ~ "C4",
#      TRUE ~ NA_character_  # if other values exist
 #   )
#  )

#create a column for dominant plant communities  
#metadata <- metadata %>%
#  mutate(
 #   Community = case_when(
 #     startsWith(as.character(Plot), "3") ~ "SCAM",
  #    startsWith(as.character(Plot), "6") ~ "SCAM",
#      startsWith(as.character(Plot), "4") ~ "SPPA, DISP",
  #    TRUE ~ NA_character_  # if other values exist
 #   )
 # )

#Bring in the collection date and any field notes if relevant 
#collection_metadata <- read.csv(Collection_Dates)
  #then we will merge the collection date for SMARTX based on the project and the chambers 

```

\newpage

## Import Data & Clean
```{r Import Data, include=FALSE}
cat("Import Data")

#set file path for data - in run info chunk 
path <- ("file path")

#Read in Raw Data 
data_NH3 <- map(NH3_files, read.csv)

# Combining Files
df_combo <- bind_rows(data_NH3)

# Select needed columns & rename 
df <- df_combo %>%
  select(Sample_Name = 4, 
         Run_Number = 5,
         Conc = 6, 
         Absorbance = 7, 
         Dilution = 9, 
         Unit = 12,
         Test = 13, 
         Run_Time = 15)

#Checking column headers
head(df)

df_all <- df %>%
  mutate(Run_Number = as.numeric(Run_Number))

## Extra code to put SMARTX at the beginning of samples that don't have it 
df_all <- df_all %>%
  mutate(
    Sample_Name = case_when(
      str_detect(Sample_Name, "^\\d{3}-\\d{2,3}$") &  # matches 123-45, 456-120, etc.
      !str_starts(Sample_Name, "SMARTX_") ~ paste0("SMARTX_", Sample_Name),
      
      TRUE ~ Sample_Name  # leave all others unchanged
    )
  )

```



## Assessing standard Curves
  #Pull out standards data  
```{r Pull out Standard Curves, echo=FALSE}
cat("Assess Standard Curves")

#Pull out standards 
stds <- df_all %>%
  filter(str_detect(Sample_Name, "Standard"))

#Making standards dataframe based on the test
stds_NH3 <- stds[stds$Test == "Ammonia 2", ]

#Inputting Standard Values Based on Protocol (LINK PROTOCOL)
stds_NH3 <- stds_NH3 %>%
  mutate(`Conc` = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard .0389" ~ 0.0389,
    Sample_Name == "Standard .1000" ~ 0.1,
    Sample_Name == "Standard .2000" ~ 0.2,
    Sample_Name == "Standard .5000" ~ 0.5,
    Sample_Name == "Standard 1.0000" ~ 1.0,
    Sample_Name == "Standard 1.5000" ~ 1.5,
    Sample_Name == "Standard 2.0000" ~ 2,
    TRUE ~ `Conc`),  # leave unchanged if no match
    Run_Date = format(lubridate::mdy_hm(Run_Time), "%m-%d-%Y"))

# identifying x and y
x2 <- stds_NH3$Absorbance
y2 <- stds_NH3$Conc

#generating line of best fit aka standard curves
#NH3
lin_reg_NH3 <- lm(y2 ~ x2)
NH3_coefs <- coef(lin_reg_NH3)                        # intercept and slope
NH3_r2 <- summary(lin_reg_NH3)$r.squared              # R squared
# Store in a clean dataframe
std_curve_NH3_1 <- data.frame(
  Test = "NH3",
  Intercept = NH3_coefs["(Intercept)"],
  Slope = NH3_coefs[2],
  R_squared = NH3_r2
)

#Combining standards & curve data frames
stds_combo <- bind_rows(stds_NH3)
std_curve_combo <- bind_rows( std_curve_NH3_1)


```
  #Plot standards data 
```{r Assess Standard Curves, echo=FALSE}
cat("Assess Standard Curves")

#Plot standard Curve or Curves 
  #NH3 
std_curve_NH3 <- ggplot(stds_NH3, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgoldenrod") +
  theme_bw() + 
  facet_wrap(~Run_Date) +
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "NH3 Standard Curve")+ 
  annotate("text", x = 1.7, y = max(stds_NH3$Absorbance), label = paste("r^2 =", round(NH3_r2, 4)),
          color = "black", size = 4)
print(std_curve_NH3)

############## Report on Cutoffs 

#Report out a flag if the run has an R2 lower than appropriate 
#Write out to the user whether or not the r2 is above the cutoff of 0.98
  ifelse(std_curve_NH3_1$R_squared <= r2_cutoff, 
         "NH3 Curve r2 is below cutoff! - REASSESS", 
         "NH3 Curve r2 GOOD - PROCEED")

#check for the a slope QAQC file, if there is not one, make one 
if (file.exists(file_path)) {
  # If it exists, read it back into R
  stds_log <- read.csv(file_path)
  print("QAQC log file exists and has been read into the code.")
  } else {
  # If it does not exist, create the CSV file
  stds_log <- as.data.frame(matrix(ncol = 6, nrow = 0))
  colnames(stds_log) <- c("Test", "Intercept", "Slope", "R_squared", "run_date", "user")
  
  #maybe add here the last slopes from the previous year
  
  # Write add_log to CSV
  write.csv(stds_log, file = file_path, row.names = FALSE)
  print("QAQC log file does not exist. It has been created.")
  }
  
  #getting rid of first "X" column
  #stds_log <- stds_log %>%
  #  select(-1)
  
#create a dataframe that can be appended to the QAQC log 
add_log <- as.data.frame(bind_rows(std_curve_NH3_1))
add_log$run_date <- run_date
add_log$user <- user

#compare slopes to previous runs (from log) in order to assess drift 
stds_log$run_date <- as.character(stds_log$run_date)
add_log$run_date  <- as.character(add_log$run_date)     #getting the run_date column in the same type to merge

#Filter to only rows in Slopes that are NOT already in log (by run_date + analyte) but need to add if none
if (nrow(stds_log) == 0) {
  # If log is empty, everything in add_log is new
  new_rows <- add_log
} else {
  # Otherwise, compare to existing log
  new_rows <- anti_join(add_log, stds_log, by = c("run_date", "Test"))
}


# Append the new, non-duplicate rows to log
log <- bind_rows(stds_log, new_rows)


#Plot the slopes through time 
Slopes_chk <- ggplot(log, aes(run_date, Slope, col=Test)) + 
  geom_point(size=4) + 
  geom_line(aes(group = Test)) + 
  theme_bw() + 
  labs(title="Slope Drift Assessment", x="Run Date", y="Slope") +
  scale_color_manual(values=c( "darkgoldenrod"))+
  ylim(1.5,2)
Slopes_chk

#averaging the slopes from the log and printing
avg_slopes <- log %>%
  group_by(Test) %>%
  summarise(avg_slope = mean(Slope, na.rm = TRUE))
knitr::kable(avg_slopes, caption = "Average Slope by Analyte", digits = 3)


#write out the log file with the added lines for this run date 
write.csv(log, file_path)

#write out a flag to the sample dataframe if the r2 is above the cutoff of 0.98
df_all <- df_all %>%
  mutate(
    NH3_flag = if (NH3_r2 <= r2_cutoff) {
      "NH3 r2 low"
    } else {
      ""
    }
  )



```
 
 \newpage


## Dilution Corrections - ensure the latest dilution is kept
```{r Dilution Corrections, echo=FALSE}

cat("Dilution Corrections")

#Calculate the concentration when accounting for the dilution factor if applicable

#checking to see if duplicate samples (mislabeled/wrongly inputted/reran)
df_duplicates <- df_all %>% 
  filter(!if_all(everything(), is.na)) %>%                    #removing NA rows
  filter(!(Sample_Name %in% c("Duplicate",                    #removing duplicates, autospikes, ccv/bs 
                              "CCV", "CCB", "1mg/L ammonia", "DI",
                              "Blank", "Auto Spike", 
                              "0.15 mg/L")) &
  !startsWith(Sample_Name, "Standard ") &               #removing standards
  !startsWith(Sample_Name, "Nitrite ") &                #removing standards
  !startsWith(Sample_Name, "Nitrate ") &                #removing standards
  !startsWith(Sample_Name, "Abs_Chk_") &                #removing abs_chks
  !startsWith(Sample_Name, "peChk_") &                 #removing peChks
  !grepl("^\\d{1,2}/\\d{1,2}/\\d{4}\\s+\\d{1,2}:\\d{2}$", Sample_Name) & #removing naything that is a date
  !grepl("^\\d+$", Sample_Name)) %>%                    #removing any rows with just numbers in the sample name column
  group_by(Test, Sample_Name)%>%                              #grouping by test & sample name to avoid fake dups
  filter(n() > 1) %>%                                         #count how many times each Sample_Name appears
  ungroup() 

#pulling them out into a string and printing them out
dil_dups_string <- df_duplicates %>%
  distinct(Sample_Name) %>%
  pull(Sample_Name) %>%
  paste(collapse = ", ")

#Report duplicated samples
if (dil_dups_string != "") {
  message("Duplicated samples: ", dil_dups_string)
  
  bad_dups <- df_duplicates %>%
    group_by(Sample_Name) %>%
    summarise(
      n_dups = n(),
      has_dilution = any(Dilution > 0, na.rm = TRUE)
    ) %>%
    filter(n_dups > 1 & !has_dilution) %>%
    pull(Sample_Name)
  
  if (length(bad_dups) > 0) {
    message("Duplicated samples with NO dilution present (possible input error): ",
            paste(bad_dups, collapse = ", "))
  } else {
    message("All duplicated samples have valid dilutions. No naming issues detected.")
  }
  
} else {
  message("No duplicated samples.")
}

df_all_cor <- df_all %>%
  arrange(row_number()) %>%
  mutate(
    Dilution = case_when(
      Dilution == "0" ~ 1.0,
      TRUE ~ as.numeric(Dilution)
    ),
    Pair_ID = ifelse(
      grepl("Duplicate", Sample_Name, ignore.case = TRUE),
      paste0(lag(Sample_Name), "_", lag(Run_Number)),  # pair only if same run
      paste0(Sample_Name, "_", Run_Number)
    ),
    Original_Name = Sample_Name
  ) %>%
  group_by(Pair_ID, Test, Run_Number) %>%  # group by Pair_ID, Test, and Run_Number
  filter(
    Dilution == ifelse(any(Dilution > 1, na.rm = TRUE), max(Dilution, na.rm = TRUE), 1)
  ) %>%
  ungroup()




```


## Performance Check
```{r pe Check, echo = FALSE}

NH3_peChk <- df_all_cor %>%
  filter(Test == "Ammonia 2", 
         str_detect(Sample_Name, "pechk")) %>%
  summarise(
    Mean_Chk_Conc = mean(Conc, na.rm = TRUE),
    SD_Chk_Conc = sd(Conc, na.rm = TRUE),
    CV = (SD_Chk_Conc / Mean_Chk_Conc) * 100,
    pct_diff_Chk = abs(Mean_Chk_Conc - NH3_pe) / NH3_pe,
    NH3_pe_flag = ifelse(abs(pct_diff_Chk) < 25, "YES, Pass", "NO, rerun"))

#flag write out 

ifelse(NH3_peChk$pct_diff_Chk >= chk_flag, 
       "NH3 pe Check has a % Difference >25% - REASSESS",  
       "NH3 pe Check has a % Difference <25% - PROCEED")
cat("Run mean =", NH3_peChk$Mean_Chk_Conc, "\n")
cat("Expected  =", NH3_pe, "\n\n")


```

\newpage

## Analyze the Check Standards
```{r Check Standards, echo=FALSE}

cat("Analyze Check Standards")

#Pull out check standards from raw file 
NH3_chks <- df_all %>%
  filter(str_detect(Sample_Name, "1mg/L ammonia"))%>% 
  mutate(rep = row_number())


#RSV of standards 
chks_NH3_rsv <- ((sd(NH3_chks$Conc))/mean(NH3_chks$Conc))

#write out to user about whether or not to continue 
ifelse(chks_NH3_rsv >= chk_flag, 
       "NH3 CHECK STANDARD RSD TOO HIGH - REASSESS",
       "NH3 Check Standard RSD within Range - PROCEED")

#calculate percent difference between check standards & expected concentration 
#flag if the percent difference is over X% (defined in setup)
#calculate percent difference of check standards 

NH3_chks$NH3_diff <- ((NH3_chks$Conc - NH3_CCV)/((NH3_chks$Conc + NH3_CCV)/2)) * 100
NH3_chks$NH3_diff_flag <-  ifelse(NH3_chks$NH3_diff <= chk_conc_flag, 'YES', 'NO, rerun')


#plot the check standards
NH3_chks_plot <-  ggplot(data = NH3_chks, aes(x = rep, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
       theme_classic() + 
       labs(x= " ", y="NH3  (mg/L)", title="Check Stds: NH3") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=NH3_CCV,
              linetype="dashed",  color = "black", linewidth=1) + 
       guides(fill=guide_legend(title="% Difference <10%"))

print(NH3_chks_plot)

#calculate the percent of check standards that are within the range based on the flag 
NH3_chks_percent <- (sum(NH3_chks$NH3_diff_flag == "YES")/nrow(NH3_chks))*100

#report out if flags indicate need for rerun
ifelse(NH3_chks_percent >= chks_flag,
       ">60% of NH3 Check Standards are within range of expected concentration - PROCEED",
       "<60% of NH3 Check Standards are within range of expected concentration - REASSESS")

#write out a flag to the sample dataframe if less than 60% of the checks are within the expected CV
if (NH3_chks_percent <= chks_flag) {  
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 checks out of range"),
    "NH3 checks out of range"
  )
}

```

\newpage

## Analyze Blanks 
```{r Analyze Blanks, echo=FALSE}

cat("Assess Blanks")

#Pull out check standards from raw file 
NH3_blks <- df_all %>%
  filter(Test == "Ammonia 2")%>% 
  filter(str_detect(Sample_Name, "Blank"))%>% 
  mutate(rep = row_number())


#Pull out samples from df_all to calc quantile
samples <- df_all %>%  
  filter(str_detect(Sample_Name, c("SMARTX"))) 

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

NH3_only <- samples %>%
  filter(Test == "Ammonia 2")
blk_flag_NH3 <- quantile(NH3_only$Conc, prob=c(.25))   #this gives you the lower 25% quartile of the data 
NH3_blks$NH3_diff_flag <-  ifelse(NH3_blks$Conc <= blk_flag_NH3, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag 
NH3_blks_percent <- (sum(NH3_blks$NH3_diff_flag == "YES")/nrow(NH3_blks))*100

#report out if flags indicate need for rerun
ifelse(NH3_blks_percent >= 60,
       ">60% of NH3 Blanks are below the lower 25% quartile of samples - PROCEED",
       "<60% of NH3 blaks are lower 25% quartile of samples - REASSESS")

#plotting the blanks compared to the lower 25% of conc (what the flag is)
NH3_blks_plot <-  ggplot(data = NH3_blks, aes(x = rep, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity') + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="NH3  (mg/L)", title="NH3 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=as.numeric(blk_flag_NH3), linetype="dashed",
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

print(NH3_blks_plot)


#find average of run blanks for flagging samples later 
# Compute averages
blk_avg_NH3 <- mean(NH3_blks$Conc, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(
  Test = c("NH3"),
  Blank_Mean_Conc = c(blk_avg_NH3)
)

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Concentration of Blanks", digits = 4)


#write out a flag to the sample dataframe if more than 60% of the blanks are above the lower 25% quantile of samples
if (NH3_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 blanks out of range"),
    "NH3 blanks out of range"
  )
}


```

\newpage

## Analyze Duplicates
```{r Analyze Duplicates, echo=FALSE}

cat("Analyze Duplicates")

#Getting dataframe of just duplicated samples
df_NH3_dup <- df_all_cor %>%
  filter(Test == "Ammonia 2") %>%
  filter(
     !grepl("CCV|CCB|Blank|1mg/L ammonia|10ppt|Standard|Auto Spike", Sample_Name, ignore.case = TRUE), 
     #^removes autospice, ccv, ccb, standards
     !is.na(Sample_Name),                                         # Remove NA
     !grepl("^\\d+$", Sample_Name),                               # removes all numbers
     !grepl("^\\d{1,2}/\\d{1,2}/\\d{2,4}", Sample_Name)) %>%      # removes all date/time
  mutate(`Duplicate?` = if_else(                                  # adds duplicate column
     grepl("Duplicate", Sample_Name, ignore.case = TRUE),         # keeps sample ID if there
     lag(Sample_Name),                                            # get previous row's SampleID
     Sample_Name)) %>%                                            # keeps sample ID if not duplicate
  group_by(`Duplicate?`) %>%                                      # groups by duplicate 
  filter(n() > 1) %>%                                             # filters anything that appears more than once 
  ungroup() %>%                      
  transmute(   
     Pair_ID = `Duplicate?`,                                      # changing heading to pair_id
     Sample_Name,                                                 #keeping sample name and conc
     Conc,                                                        #if sample name duplicate then put duplicate 
     Type = if_else(str_detect(Sample_Name, "Duplicate"), "Duplicate", "Original"))  #, if not put original to ID 

# Duplicate Stats
NH3_dup_chk <- df_NH3_dup %>%
  group_by(Pair_ID) %>%  # Use Pair_ID as it groups Originals and Duplicates
  filter(n_distinct(Type) == 2) %>%  # Keep only groups that have both Original and Duplicate
  summarise(
    OG_Conc = Conc[Type == "Original"],
    Dup_Conc = Conc[Type == "Duplicate"],
    Mean_Conc = mean(c(OG_Conc, Dup_Conc)),
    SD_Conc = sd(c(OG_Conc, Dup_Conc)),
    CV = (SD_Conc / Mean_Conc) * 100,
    pct_diff = abs(Dup_Conc - OG_Conc) / Mean_Conc * 100,
    NH3_diff_flag = ifelse(abs(CV) < 10, "YES", "NO, rerun"),
    .groups = "drop") 


#calculate the percent of check standards that are within the range based on the flag 
NH3_dup_percent <- (sum(NH3_dup_chk$NH3_diff_flag == "YES")/nrow(NH3_dup_chk))*100

#dup flag report out
ifelse(NH3_dup_percent >= chks_flag, 
       ">60% of NH3 Duplicates have a CV <10% - PROCEED",
       "<60% of NH3 Duplicates have a CV <10% - REASSESS")

#Create row numbers for plotting
NH3_dup_chk <- NH3_dup_chk %>%
  mutate(row_num = row_number())

#plot dups output as a bar graph to easily check - want any over 10% to be red need to work on this 
NH3_dups_plot <- ggplot(data = NH3_dup_chk, aes(x =Pair_ID, y =CV, fill=NH3_diff_flag)) +
        geom_bar(stat = 'identity') + 
        theme_classic() + 
        labs(x= "Sample ID", y="CV between dups", title = "NH3 Duplicates") + 
        scale_fill_manual(values = c("YES" = "darkslateblue", "NO, rerun" = "darkred")) +
        theme(legend.position="none") +  
        geom_hline(yintercept=10, linetype="dashed", 
              color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="<10%:"))+
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))

print(NH3_dups_plot)

#write out a flag to the sample dataframe if more than 60% of the dups have CVs out of range 
if (NH3_dup_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 dups out of range"),
    "NH3 dups out of range"
  )
}


```

\newpage

## Spikes
```{r Analyze Spikes, echo=FALSE}

#getting spikes and samples pulled 
df_NH3_spk <- df_all_cor %>%
  filter(Test == "Ammonia 2") %>%
  filter(
     !grepl("CCV|CCB|Blank|1mg/L ammonia|Standard|Duplicate", Sample_Name, ignore.case = TRUE),
     #^removes dup, ccv, ccb, standards
     !is.na(Sample_Name),                                         # Remove NA
     !grepl("^\\d+$", Sample_Name),                               # removes all numbers
     !grepl("^\\d{1,2}/\\d{1,2}/\\d{2,4}", Sample_Name)) %>%      # removes all date/time
  mutate(`Spike?` = if_else(                                      # adds duplicate column
     grepl("Auto Spike", Sample_Name, ignore.case = TRUE),        # keeps sample ID if there
     lag(Sample_Name),                                            # get previous row's SampleID
     Sample_Name)) %>%                                            # keeps sample ID if not duplicate
  group_by(`Spike?`) %>%                                          # groups by duplicate
  filter(n() > 1) %>%                                             # filters anything that appears more than once
  ungroup()  %>%
  transmute(
    Pair_ID = `Spike?`,
    Sample_Name,
    Conc,
    Type = if_else(str_detect(Sample_Name, "Auto Spike"), "Auto_Spike", "Original"))
 

#checking and flagging against the expected spike conc
df_spk_NH3_chk <- df_NH3_spk %>%
  select(Pair_ID, Conc, Type) %>%
  pivot_wider(
    names_from = Type,
    values_from = Conc,
    values_fn = mean,  # in case there are duplicates
    names_prefix = "" ) %>%
  rename(
    Spike_Conc = Auto_Spike,
    OG_Conc = Original) %>%
  filter(!is.na(Spike_Conc), !is.na(OG_Conc)) %>%
  rowwise() %>%# keep only rows with both
   mutate(
    Exp_Conc = OG_Conc + NH3_Spk,
    Mean_Conc = mean(c(Spike_Conc, Exp_Conc), na.rm = TRUE),
    Pct_Diff = abs(Spike_Conc - Exp_Conc) / Mean_Conc * 100,
    SD_Conc = sd(c(Spike_Conc, Exp_Conc)),
    CV = (SD_Conc / Mean_Conc) * 100,
    Spike_diff_flag = ifelse(CV < 50, "YES", "NO, check"),
    Test = "Ammonia 2") %>%
    ungroup()

# Combining Auto Spike data frames
df_spk_combo <- bind_rows(df_spk_NH3_chk)

#calculate the percent of check standards that are within the range based on the flag
NH3_spk_percent <- (sum(df_spk_NH3_chk$Spike_diff_flag == "YES")/nrow(df_spk_NH3_chk))*100
 
#spk flag report out
ifelse(NH3_spk_percent >= chks_flag, 
       ">60% of Spikes have a CV <50% - PROCEED",
       "<60% of Carbon Spikes have a CV <50% - REASSESS")

#plot spikes output as a bar graph to easily check - want any over 10% to be red need to work on this 
NH3_spk_plot <- ggplot(data = df_spk_NH3_chk, aes(x =Pair_ID, y =CV, fill=Spike_diff_flag)) +
        geom_bar(stat = 'identity') + 
        theme_classic() + 
        labs(x= "Sample ID", y="NH3 N(mg/L)", title = "NH3 Auto Spike CV") + 
        scale_fill_manual(values = c("YES" = "darkcyan", "NO, check" = "darkred")) +
        theme(legend.position="none")  + 
        geom_hline(yintercept=50, linetype="dashed", 
            color = "black", size=1)  +
        guides(fill=guide_legend(title="CV Between Spks <50%"))+
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
 
print(NH3_spk_plot)


#write out a flag to the sample dataframe if more than 60% of the dups have CVs out of range
if (NH3_spk_percent <= chks_flag) {  
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 spikes out of range"),
    "NH3 spikes out of range"
  )
}

```

\newpage

## Matrix Effects  
```{r Analyze Matrix Effects, echo=FALSE}

df_matrix <- df_spk_combo%>%
  filter(str_detect(Pair_ID, "10ppt")) 
  
df_matrix_NH3 <- df_matrix %>%
  filter(str_detect(Test, "Ammonia")) 
  ifelse(
  all(df_matrix_NH3$Spike_diff_flag == "YES"),
  "NO NH3 Matrix Effect, PROCEED",
  ">20% CV in ASW NH3 matrix effect checks - REASSESS"
)

#write out a flag to the sample dataframe if more than 60% of the dups have CVs out of range
if (any(df_matrix$Spike_diff_flag[df_matrix$Test == "Ammonia 2"] != "YES")) {
  
    df_all_cor$NOx_flag <- ifelse(
    df_all_cor$NOx_flag != "",
    paste0(df_all_cor$NOx_flag, "; NH3 matrix check out of range"),
    "NH3 matrix check out of range"
  )

}


```

## Unit Converted Data Column Added (mg/L to uM )
```{r Add unit conversion, include=FALSE}

#Pull out samples 
samples <- df_all_cor %>%  
  filter(str_detect(Sample_Name, c("SMARTX")))     

#Convert values based on the Test ID 
samples <- samples %>%
  mutate(
    Conc_uM = case_when(
      Test == "Vanadium NOx" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Ammonia 2" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "o-PHOS 0.3" ~ (((as.numeric(samples$Conc))/Con1)/P_mw)*Con2,
      TRUE ~ as.numeric(NA)  # fallback in case of unexpected value
    ),
    # Replace negatives with 0 in 'value' and 'value_converted'
    Conc = pmax(Conc, 0),
    Conc_uM = pmax(Conc_uM, 0)
  )

head(samples)

```

## Sample Flagging - Within range of standard curve
```{r Sample Flagging, echo=FALSE}

cat("Sample Flagging")

#Flagging data if the concentration is outside the standards range 
samples_flagged <- samples %>%
  mutate(
    Conc_flag = case_when(
      Test == "Vanadium NOx" & Conc < NOx_dl ~ "bdl",
      Test == "Vanadium NOx" & Conc > NOx_top ~ "adl",
      Test == "Vanadium NOx" ~ "Within_Range",
      
      Test == "Ammonia 2" & Conc > NH3_top & Dilution == 1 ~ "adl",       # will flag if it wasn't diluted 
      Test == "Ammonia 2" & Conc < NH3_dl ~ "bdl",
        # Test == "Ammonia 2" & Conc > NH3_top ~ "adl",     #this would flag if above 2, but we want that only if not diluted
      Test == "Ammonia 2" & Dilution != 1 ~ "Within_Range_Diluted",# your new condition
      Test == "Ammonia 2" ~ "Within_Range",
      
      Test == "o-PHOS 0.3" & Conc < PO4_dl ~ "bdl",
      Test == "o-PHOS 0.3" & Conc > PO4_top ~ "adl",
      Test == "o-PHOS 0.3" ~ "Within_Range",
      
      TRUE ~ NA_character_
    )
  )

```

## Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata 
all_present <- all(metadata$Sample_Name %in% samples_flagged$Sample_Name)

if (all_present) {
  message("All sample IDs are present in data.")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, samples_flagged$Sample_Name)
  print(missing_ids)
}


#merge metadata with sample run data 
merged_data <- samples_flagged %>%
  left_join(metadata, by = c("Sample_Name" = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included? 

df_all_clean <- merged_data 

```

\newpage

## Visualize Data
```{r Visualize Data, echo=FALSE}
cat("Visualize Data")

### Ammonia 
NH3_forplot <- df_all_clean %>%
  filter(Test == "Ammonia 2")

#C3 plots
NH3_forplot_C3 <- NH3_forplot %>%
  filter(Nickname == "C3")

#group the data for plotting
NH3_forplot_C3 <- NH3_forplot_C3 %>%
  group_by(Depth_cm) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NH3_plot_C3 <- ggplot(data = NH3_forplot_C3, aes(x = as.factor(Temp_Treatment), y = Conc, 
                                                     fill=as.factor(Temp_Treatment))) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") + 
       facet_grid(~ Depth_cm, scales="free_x") +
        #facet_grid(Depth_cm ~ ., scales = "free_x") + 
        scale_fill_manual(values = c("0" = "blue", 
                                     "1.7" = "yellow", 
                                     "3.4" = "orange",
                                     "5.1" = "red")) +
        theme_classic() + 
        labs(x= " ", y="NH3 (mg/L)", title="C3: Porewater NH3", fill="+Temp. (C)") + 
          theme(legend.position = "right") +
          scale_x_discrete(drop = TRUE)+
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5), 
              panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8))
print(viz_NH3_plot_C3) 

#C4 plots
NH3_forplot_C4 <- NH3_forplot %>%
  filter(Nickname == "C4")

#group the data for plotting
NH3_forplot_C4 <- NH3_forplot_C4 %>%
  group_by(Depth_cm) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

viz_NH3_plot_C4 <- ggplot(data = NH3_forplot_C4, aes(x = as.factor(Temp_Treatment), y = Conc, fill=as.factor(Temp_Treatment))) +
       geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
                 color="black") + 
        #facet_grid(~ Depth_cm, scales="free_x") +
        facet_grid(Depth_cm ~ ., scales = "free_x") + 
        scale_fill_manual(values = c("0" = "blue", 
                                     "1.7" = "yellow", 
                                     "3.4" = "orange",
                                     "5.1" = "red")) +
        theme_classic() + 
        labs(x= " ", y="NH3 (mg/L)", title="C4: Porewater NH3", fill="+Temp. (C)") + 
          theme(legend.position = "right") +
          scale_x_discrete(drop = TRUE)+
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5), 
              panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8))
print(viz_NH3_plot_C4) 

```

## Export Processed Data
```{r Export Processed Data, include=FALSE}

cat("Export Processed Data")

#pivot the data set wider to make it wide format 
final_data1 <- df_all_clean %>%
  pivot_wider(
    id_cols = Sample_Name,
    names_from = Test,
    values_from = Conc,
    names_glue = "{Test}_Conc"
  )

final_data2 <- df_all_clean %>%
  pivot_wider(
    id_cols = Sample_Name,
    names_from = Test,
    values_from = Conc_uM,
    names_glue = "{Test}_Conc_uM"
  )

final_data3 <- df_all_clean %>%
  pivot_wider(
    id_cols = Sample_Name,
    names_from = Test,
    values_from = Conc_flag,
    names_glue = "{Test}_Conc_flag"
  )

#take out only the columns we want to merge 
df_all_clean_cols <- df_all_clean %>%
  select(Sample_Name, Sample_ID, Project, Plot, Depth_cm, CO2_Treatment, Temp_Treatment, 
         Nickname, Community,  Year, Month,
         #Day, Time..24hr., Time.Zone_EDT.EST,
        NH3_flag)

df_all_clean_cols_one_row <- df_all_clean_cols %>%
  group_by(Sample_Name) %>%
  slice(1) %>%  # or use summarise() if you want to aggregate
  ungroup()

#merge these together
data_list <- list(df_all_clean_cols_one_row, final_data1, final_data2, final_data3)

final_data4 <- reduce(data_list, full_join, by = c("Sample_Name"))


#Add project information 
final_data_labeled <- final_data4 %>% 
  mutate(
    Site = "GCReW",   # new column with same value on every row
    Run_notes = run_notes, 
    Analysis_rundate = print(run_date)# new column with notes about the run
  ) 

#Prepare data to be exported 
final_data <- final_data_labeled %>%
    rename(
    Site = Site,
    Sample_Name = Sample_Name, 
    #Time = Time..24hr., 
    #Time_Zone = Time.Zone_EDT.EST,
    #Replicate = Lysimeter,
    
    NH3_Conc_mgL = `Ammonia 2_Conc`,
    NH3_Conc_uM = `Ammonia 2_Conc_uM`,
    NH3_Conc_Flag =`Ammonia 2_Conc_flag`,
    NH3_QAQC_Flag = NH3_flag
    

    # add more rename pairs as needed
  ) %>%
  select(Site, Project, Year, Month, #Sample_Name, Sample_ID,  
         Community, Nickname, Plot, Depth_cm, CO2_Treatment, Temp_Treatment, 
         #Sample_ID, Day, Time, Time_Zone,
         NH3_Conc_mgL, NH3_Conc_uM, NH3_Conc_Flag, NH3_QAQC_Flag,
         Analysis_rundate,  Run_notes)


#Write out data frame 
  write.csv(final_data, final_path)
  

```


#end