---
title: "Chapada Checks"
author: "TE Technicians"
date: "`r Sys.Date()`"
output: html_document
---
<style>
.vscroll-plot {
    width: 1000px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>
```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
library(ggpubr)
library(gghighlight)
library(kableExtra)
library(ggrepel)

library(knitr)
library(utils)
library(pacman)

```

``` {r set knitr options,include = F}
options(width = 150)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height, options$max.width)
    )
    hook_output(x, options)
  })
})
```

``` {r load data, echo = F}
#set wd
wd =  "C:/Users/LaGorgaL.S/Documents/Chapata_Stem/daily_checks/"
#read in files from dropbox
dbfold <- "C:/Campbellsci/LoggerNet/CHAPADA_Loggerfiles/"
files <- list.files(dbfold, pattern = "*.dat", full.names = TRUE)
names(files) <- gsub(".dat", "", basename(files))

#read into list to stay organized, format head function
chapada_data <- lapply(files, function(file) {
 data <- fread(file,skip=1, header = TRUE)[-c(1,2)]
  return(data)
})
names(chapada_data) <- names(files)

#convert timestamps
chapada_data <- lapply(chapada_data, function(df) {
  df$TIMESTAMP <- as.POSIXct(df$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S")
  df<- df %>% filter(TIMESTAMP>=format(Sys.time()-2419200,"%Y-%m-%d %H:%M:%S"))
  return(df)
})
```
Check the Battery Voltage! 
12-14V is ideal. If you notice that the battery voltage is drifting, it may be time to change the battery. 

```{r batteries babyyyy, echo=F}
#aggregate and make long battery table for plotting
battery_tables <- grep("CheckTable|ChamberBME280|Redox15|Teros12Table", names(chapada_data), value = TRUE)
batt_plot_data <- bind_rows(lapply(battery_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, BattV = matches("BattV|BattV_Min|BattV_Avg")) %>%
    mutate(origin = table)
}))

ggplot(batt_plot_data, aes(x = TIMESTAMP, y = as.numeric(BattV), color = origin))+
  geom_line(na.rm = TRUE) +
  labs(title = "Battery Voltage Across Tables", y = "Voltage (V)", x = "Time") +
  theme_minimal()
```
Check that the Ringdown Times on Both Licor Are Not Changing!
The plot below shows this months's record of the ringdown times for both licor. Make sure that you are not seeing any long-term drift of the ringdown times.

```{r flux deluxe!, echo=F}
flux_tables <- grep("FLUX_7810", names(chapada_data), value = TRUE)

ringdown_data <- bind_rows(lapply(flux_tables, function(table){
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, ringdown = matches("ringdown")) %>%
    mutate(origin = table)
  }))

ggplot(ringdown_data, aes(x = TIMESTAMP, y = as.numeric(ringdown), color = origin)) +
  geom_line(na.rm = TRUE) +
  labs(title = "Ringdown Time by Instrument", y = "Ringdown", x = "Time") +
  theme_minimal()

flux_comb_tables <- grep("FLUX_COMB", names(chapada_data), value = TRUE)

flux_data <- bind_rows(lapply(flux_comb_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, Fluxing_Chamber, Flux_Status, Manifold_Timer, H2O, CO2, CH4) %>%
    mutate(origin = table)
}))

ggplot(flux_data, aes(x = TIMESTAMP, y = as.numeric(CO2), color = as.factor(Fluxing_Chamber))) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "CO2 Concentration by Chamber", y = "CO2 ppm", x = "Time") +
  theme_minimal()
```
Record of Flux Slopes and R2 Values During Flux Flag 2
```{r flux flag 2, echo=F}
#flux slopes in table form
df <- chapada_data[["Chapada_High_FLUX_COMB"]]
df <- df %>% filter(Flux_Status == 2, Manifold_Timer >= 180 & Manifold_Timer <= 540)

lm_results <- df %>%
  group_by(Fluxing_Chamber) %>%
  summarise(
    CO2_slope = coef(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))[2],
    CO2_r2 = summary(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))$r.squared,
    CH4_slope = coef(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))[2],
    CH4_r2 = summary(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))$r.squared
  )

kable(lm_results, caption = "Flux Slopes and R^2 by Chamber") %>%
  kable_styling()
```
Check out BME performance in each chamber:
```{r high fives for hives!, echo=F}
temp_tables <- grep("ChamberBME280", names(chapada_data), value = TRUE)

temp_data <- bind_rows(lapply(temp_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, matches("Temp")) %>%
    pivot_longer(cols = -TIMESTAMP, names_to = "Sensor", values_to = "Temp") %>%
     filter(TIMESTAMP>as.POSIXct("2025-09-22 23:00:00"))%>%
    mutate(origin = table)
}))

ggplot(temp_data, aes(x = TIMESTAMP, y = as.numeric(Temp), color = Sensor)) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "BME Temperature Trends", y = "Temperature (°C)", x = "Time") +
  theme_minimal()

rh_tables <- grep("ChamberBME280", names(chapada_data), value = TRUE)

rh_data <- bind_rows(lapply(rh_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, matches("RH")) %>%
    pivot_longer(cols = -TIMESTAMP, names_to = "Sensor", values_to = "RH") %>%
    mutate(origin = table)
}))

ggplot(rh_data, aes(x = TIMESTAMP, y = as.numeric(RH), color = Sensor)) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "BME Relative Humidity Trends", y = "RH (%)", x = "Time") +
  ylim(y=0, 100)+
  theme_minimal()
```