---
title: "CO2xNxPhrag: Porewater Ammonium"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information 
```{r run information, include=TRUE}
cat("Run Information: ") #lets you know what section you're in 
#set the run date & user name 
  run_date <- "May"
  sample_year <- "2025"
  sample_month <- "Sept"
  user <- "Max Wegner"

#identify the files you want to read in 
  #read in as a list to accommodate multiple runs in a month
  NH3_files <- c("Raw Data/LTREB_Phrag_May_2025.csv")

# Define the file path for QAQC log file - NO Need to change just check year 
  file_path <- "Raw Data/SEAL_Phrag_QAQC_Log_2025.csv"
  final_path <- "Processed Data/Phrag_Porewater_NH4_202505.csv"

#record any notes about the run or anything other info here (Leave empty otherwise): 
  run_notes <- ""
  

  cat(run_notes)
```

##Setup
```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#a link to the Gitbook or whatever protocol you are using for this analysis 


#load/install packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  ggplot2,
  ggpubr,
  tidyr,
  tidyverse,
  lubridate,
  tinytex,
  stringr,
  writexl,
  readr,
  readxl,
  purrr,
  tinytex,
  broom,
  data.table,
  raster,
  here)

Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 

#Detection limit and top standards for flagging 
NH3_dl <- 0.02  #mgL

NH3_top <- 2  #mgL

#Check Standard concnetrations 
NH3_CCV <- 1.0 

#Spike Concentrations 
NH3_Spk <- 1

#expected ranges for sample concentrations used for flags 
r2_cutoff = 0.990
chk_flag = 0.25
chks_flag = 60
rep_flag = 25 
#^this is a 25% error between samples
dup_cutoff = 10 #precent difference limit between duplicates 
spike_cutoff = 50 #percent difference limit between spike and expected value
NH3_chk_cutoff <- 10 #cutoff levels for percent difference of check stds vs. expected concentration

#blank_flag is calculated based on samples later in this code

#pe check Concentrations
NH3_pe <- 1.034


```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata

raw_metadata <- read.csv(Raw_Metadata)
#clean up columns in metadata file:
#Pull out only SMARTX Rows
#Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("Phrag", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  ) %>%
  select(-Temp_Treatment) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2_N",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2",
      ifelse(CO2_Treatment == "Ambient" &
               N_Treatment == "Amended"  , "N" , "Ambient")
    )
  ))

#Bring in the collection date and any field notes if relevant
#collection_metadata <- read.csv(Collection_Dates)
#then we will merge the collection date for SMARTX based on the project and the chambers

```

\newpage

##Import Data & Clean
```{r Import Data, include=FALSE}
cat("Import Data")

#set file path for data - in run info chunk 
path <- ("file path")

#Read in Raw Data 
data_NH3 <- bind_rows(map(NH3_files, read.csv))


# Combining Files
df_combo <- data_NH3

# Rename columns
df1 <- df_combo %>%
  dplyr::select(Sample_Name = 4,
         Run_Number = 5,
         Conc = 6,
         Absorbance = 7,
         Dilution = 9,
         Unit = 12,
         Test = 13,
         Run_Time = 15) %>%
  mutate(Run_Number = as.numeric(Run_Number))

#Checking column headers
head(df)

df_all <- df1 %>% mutate(Run_Number = as.numeric(Run_Number),
         Run_Date = format(as.POSIXct(Run_Time,format='%m/%d/%Y %H:%M'),format='%m/%d/%Y')) %>% mutate(across(c(Test, Run_Date), as.factor))

#Initialize Flags
DF_Flags <- df_all %>% dplyr::select(Run_Date) %>% distinct() %>% na.omit()

```

\newpage

## Assessing standard Curves
```{r Pull out Standard Curves, echo=FALSE}

#Inputting Standard Values Based on Protocol (LINK PROTOCOL)
split_data <- df_all %>% dplyr::select(Sample_Name, Conc, Absorbance, Test, Run_Date) %>% filter(Sample_Name %ilike% "Standard") %>%
  mutate(Conc = case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard .0389" ~ 0.0389,
    Sample_Name == "Standard .1000" ~ 0.1,
    Sample_Name == "Standard .2000" ~ 0.2,
    Sample_Name == "Standard .5000" ~ 0.5,
    Sample_Name == "Standard 1.0000" ~ 1.0,
    Sample_Name == "Standard 1.5000" ~ 1.5,
    Sample_Name == "Standard 2.0000" ~ 2, 
   TRUE ~ NA)) %>%
  group_by(Run_Date, Test) %>% 
  na.omit() %>%
  group_split() 

#run regression and get coefficients by Run
 
Slope_raw_list <- map(
  split_data,
  ~ .x %>% mutate(
    R2 = 
      summary(lm(Conc ~ Absorbance, data = .))$adj.r.squared
     ,
    Slope = 
      coef(lm( Conc ~ Absorbance, data = . ))[2]
      ,
    Intercept = 
      coef(lm(Conc ~ Absorbance, data = . ))[1],
      )
      )



 Slope_raw <- bind_rows(Slope_raw_list)
  
 #check R2 and flag
 Slope_Std_Crvs <- Slope_raw %>% 
   dplyr::select(Run_Date, R2, Slope, Intercept, Test) %>% 
   distinct() %>% 
   mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Pass", "R2_Fail"))

 #join standard curve CVs and regression values
 
 R2_Check <- Slope_Std_Crvs %>% 
   dplyr::select(Run_Date,
                 R2,
                 Slope,
                 Intercept,
                 Test,
                 R2_Flag) %>%
   distinct()
 
 For_crv_plot <- Slope_raw %>% mutate(Conc =  as.numeric(Conc))
 
 NH3_crv_df <- For_crv_plot %>% filter(Test == "Ammonia 2")
 
 NH3_annotate <- NH3_crv_df %>% dplyr::select(R2, Run_Date) %>% distinct() %>% mutate(Conc = c(), Absorbance = max(NH3_crv_df$Absorbance))
 
 

 NH3_Curves_Plot <- ggplot(NH3_crv_df, aes(Conc, Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgoldenrod") +
  theme_bw() + 
  facet_wrap(~Run_Date) +
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "NH3 Standard Curve") +
  geom_text(data = NH3_annotate, aes(x = 0.3, y = Absorbance, label = paste("r^2 =", round(R2, 4))),
          color = "black", size = 4)
 
NH3_Curves_Plot

R2_Flags <- R2_Check %>% dplyr::select(Test, Run_Date, R2_Flag, R2, Slope) %>% arrange(Test, Run_Date) #make df for display 

R2_Flags_join <- R2_Flags %>% dplyr::select(Test, Run_Date, R2_Flag)

knitr::kable(R2_Flags) #display Regression flags and R2


```

## Update Standard Log
```{r Update Standard Curve Log, echo = FALSE}

#log <- tibble(Test = c(), Intercept= c(), Slope= c(), Quadratic= c(), Run_Date= c(), User= c()) #create standard log in proper format

old_log <- read.csv(file_path) %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% mutate(Run = "Past Runs")#read in log and format date

add_log <- R2_Check %>% dplyr::select(Test, Run_Date, R2, Slope, Intercept, Quadratic) %>%
  mutate(Test = case_when(
    Test == "Ammonia 2" ~ "NH3"
  )) %>% mutate(Run_Date = as.Date(as.character(Run_Date), format = "%m/%d/%Y")) %>% mutate(User = user) %>% mutate(Run = "Current Runs")#Format new run data for log


new_log <- rbind(old_log, add_log) %>% mutate(across(where(is.numeric), \(x) round(x, digits = 4))) %>% distinct(dplyr::select(.,-Run, -User), .keep_all = TRUE) #bind rows, round data, and remove possible duplicate data from rerunning code
Reg_Coef <- new_log %>% group_by(Test) %>% reframe(Mean_Slope = mean(Slope, na.rm = TRUE), SD_Slope = sd(Slope, na.rm = TRUE), Mean_Int = mean(Intercept, na.rm = TRUE), SD_Int = sd(Intercept, na.rm = TRUE)) %>% distinct() %>% 
  mutate(
  SD_Slope_Max = Mean_Slope + 2*SD_Slope, SD_Int_Max = Mean_Int + 2*SD_Int, SD_Quad_Max = Mean_Quad + 2*SD_Quad,
  SD_Slope_Min = Mean_Slope - 2*SD_Slope, SD_Int_Min = Mean_Int - 2*SD_Int, SD_Quad_Min = Mean_Quad - 2*SD_Quad
                                                                                                       )

new_log_plot <- left_join(new_log, Reg_Coef, by = join_by(Test)) %>% 
  mutate(
    Slope_Lim = ifelse(Slope >= SD_Slope_Max | Slope <= SD_Slope_Min, "Bad", "Good"),
    Int_Lim = ifelse(Intercept >= SD_Int_Max | Intercept <= SD_Int_Min, "Bad", "Good") ) %>% mutate(Slope_Range = paste0(Test, Slope_Lim))

Coef_Lim <- tibble(Test      = c("NH3"),
                   Min_Slope = c(1.5),
                   Max_Slope = c(2.5),
                   Min_Int   = c(-0.06),
                   Max_Int   = c(0),
                   Min_Quad  = c(NA),
                   Max_Quad  = c(NA),
                   Run_Date = median(new_log$Run_Date),
                   Run = "Past Runs")


#Plot Slope log
Slopes_chk <- ggplot(new_log_plot, aes(x = Run_Date, y = Slope, fill = Test, col = Slope_Lim, shape = Run)) +
  scale_shape_manual(values = c("Past Runs" = 21, "Current Runs" = 24)) + 
  geom_line(aes(group = Test), color = "black") + 
  theme_bw() + 
  labs(title="Slope Drift Assessment", x="Run Date", y="Slope") +
  scale_fill_manual(values=c("NH3" = "darkgoldenrod"))+
  scale_color_manual(values=c("Bad" = "red", "Good" = "transparent"))+
  geom_point(size=4) + guides(fill = guide_legend(override.aes = list(shape = 21)))



#Plot Intercept log
ints_chk <- ggplot(new_log_plot, aes(x = Run_Date, y = Intercept, fill = Test, col = Int_Lim, shape = Run)) +
  scale_shape_manual(values = c("Past Runs" = 21, "Current Runs" = 24)) + 
  geom_line(aes(group = Test), color = "black") + 
  theme_bw() + 
  labs(title="Intercept Drift Assessment", x="Run Date", y="Intercept") +
  scale_fill_manual(values=c( "NH3" = "darkgoldenrod"))+
  scale_color_manual(values=c("Bad" = "red", "Good" = "transparent"))+
  geom_point(size=4) + guides(fill = guide_legend(override.aes = list(shape = 21)))


Slopes_chk

quad_chk

ints_chk

write.csv(new_log, "Raw Data/SEAL_TEMPEST_QAQC_Log_2024.csv", row.names = FALSE) #write out log



```


## Dilution Corrections - ensure the latest dilution is kept
```{r Dilution Corrections, echo=FALSE}
df_all_cor <- df_all %>% na.omit() %>% group_by(Run_Number, Run_Date, Sample_Name, Test) %>%
  mutate(Keep = ifelse(Dilution == max(Dilution), "Keep", "Rmv")) %>% 
  filter(Keep == "Keep") %>% ungroup() %>% #keep only latest dilution
  mutate(Pair_ID = ifelse(
  grepl("Duplicate", Sample_Name, ignore.case = TRUE),
  paste0(lag(Sample_Name), "_", lag(Run_Number)),
  # pair only if same run
  paste0(Sample_Name, "_", Run_Number) #create pair ID
))

#Display duplicated samples
df_dup <- df_all %>% na.omit() %>% filter( #get rid of anything other than samples
  !Sample_Name %like% "Standard",
  !Sample_Name %like% "1mg",
  !Sample_Name %ilike% "pe",
  !Sample_Name %ilike% "ppt",
  !Sample_Name %like% "Blank",
  !Sample_Name %like% "Duplicate",
  !Sample_Name %like% "Spike",
  !Sample_Name %like% "0.15 mg",
  !Sample_Name %like% "Nitrate",
  !Sample_Name %like% "Nitrite",
  !Sample_Name %like% "CCV",
  !Sample_Name %like% "CCB"
) %>% group_by(Run_Number, Run_Date, Sample_Name, Test) %>% mutate(Duplicated = ifelse(length(Sample_Name) == 1, "No Duplicates", "Yes Duplicates")) %>% filter(Duplicated == "Yes Duplicates") %>% ungroup() #check if samples were duplicated and get those samples

df_dup_disp <- df_dup %>% dplyr::select(Sample_Name) %>% distinct() #get list of duplicated samples

df_dup_dil_samp <- df_dup %>% group_by(Run_Number, Run_Date, Sample_Name, Test) %>% mutate(Diluted = ifelse(length(unique(Dilution)) == 1, "No Dilution", "Yes Diluted")) %>% ungroup() # check if any samples were reran without dilution

df_dup_dil_disp <- df_dup_dil_samp %>% filter(Diluted == "Yes Diluted") %>% dplyr::select(Sample_Name) %>% distinct() #get list of diluted samples

df_dup_ndil_disp <- df_dup_dil_samp %>% filter(Diluted == "No Dilution") %>% dplyr::select(Sample_Name) %>% distinct() #get list of reran samples without dilutions

#Which Samples were ran multiple times and were diluted 
ifelse(
  length(df_dup_disp$Sample_Name) == 0,
  paste("No Reruns"),
  paste("Reran Samples:", df_dup_disp)
)

ifelse(
  length(df_dup_dil_disp$Sample_Name) == 0,
  paste("No Dilutions"),
  paste("Diluted Samples:", df_dup_dil_disp)
)

ifelse(
  length(df_dup_ndil_disp$Sample_Name) == 0,
  paste("No Naming Issues Detected"),
  paste("Possible Naming Errors:", df_dup_ndil_disp)
)


```


## Performance Check
```{r pe Check, echo = FALSE}


NH3_peChk <- df_all_cor %>%
  filter(Test == "Ammonia 2", #get only pe check for NH3
         Pair_ID %ilike% "pe",
         !Pair_ID %ilike% "TMP")  %>% group_by(Run_Date) %>%
  mutate(
    Mean_Chk_Conc = mean(Conc, na.rm = TRUE),
    SD_Chk_Conc = sd(Conc, na.rm = TRUE),
    CV = cv(Conc),
    pct_diff_Chk = 100 *abs(Mean_Chk_Conc - NH3_pe) / NH3_pe,
    NH3_pe_flag = ifelse(abs(pct_diff_Chk) < 25, "Performance Check Within 25% - PROCEED", "Performance Check NOT Within 25% - REASSESS")) %>% #flag data if percent difference is out of range
  dplyr::select(Pair_ID, Mean_Chk_Conc, Run_Date, pct_diff_Chk, Test, NH3_pe_flag) %>% distinct()

  
#Create Data Frames for Display


PE_NH3_flag_disp <- NH3_peChk %>% dplyr::select(Test, Run_Date, NH3_pe_flag, Mean_Chk_Conc) %>% rename(PE_Flag = NH3_pe_flag, PE_Conc = Mean_Chk_Conc) %>% mutate(PE_Target_Conc = NH3_pe)

PE_Flag_disp <- PE_NH3_flag_disp %>% arrange(Test, Run_Date)

PE_Flag_join <- PE_Flag_disp %>% dplyr::select(Test, Run_Date, PE_Flag)

knitr::kable(PE_Flag_disp)



```

 \newpage


\newpage

## Analyze the Check Standards
```{r Check Standards, echo=FALSE}

#Pull out check standards from raw file 

All_chks <- df_all %>% filter(
    str_detect(Sample_Name, "1mg/L ammonia")
) %>% group_by(Test) %>% mutate(
  rep = row_number(),
  RSV = sd(Conc) / mean(Conc),
  RSV_Flag = ifelse(
    RSV >= chk_flag,
    "RSV TOO HIGH - REASSESS",
    "RSV WITHIN RANGE - PROCEED" #Calculate RSV and create flags
  )
)  

#Create Data frame for displaying RSV flags
check_flags_disp <- All_chks %>% mutate(RSV_Cutoff = chk_flag) %>% dplyr::select(Test, Run_Date, RSV_Flag, RSV, RSV_Cutoff) %>% distinct() %>% arrange(Test, Run_Date)

knitr::kable(check_flags_disp)


Chks_diff <- All_chks %>% mutate(Diff = 
  ifelse(
    Test == "Ammonia 2",
    100 * (Conc - NH3_CCV) / NH3_CCV, -999) #Calculate Check Percent Difference
  ) %>% mutate(Diff_Flag = ifelse(Test == "Ammonia 2", ifelse(Diff <= NH3_chk_cutoff, "YES", "NO, rerun"), "TEST NOT FOUND")) #Flag data

NH3_Chks_diff <- Chks_diff %>% filter(Test == "Ammonia 2")

                                        
#plot the check standards
NH3_chks_plot <-  ggplot(data = NH3_Chks_diff, aes(x = rep, y = Conc, fill=Diff_Flag)) +
       geom_bar(stat = 'identity') + 
        scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred")) +
        labs(x= " ", y="Concentration (mg/L)", title="NH3 Check Standards") + 
        facet_wrap(Run_Date~., scales = "free") +
        scale_x_discrete(labels = NULL, breaks = NULL) +
        geom_hline(yintercept=NH3_CCV,
              linetype="dashed", color = "black", linewidth=1) + 
        guides(fill=guide_legend(title="% Difference <10%")) +
        theme_classic() +
        theme(legend.position="bottom")

NH3_chks_plot


#Create Plot for displaying check flags
Chks_flag_disp <- Chks_diff %>% group_by(Test, Run_Date) %>% mutate(Percent = 100*length(which(Diff_Flag == "YES"))/length(Diff_Flag)) %>% mutate(CHK_Flag = ifelse(Percent >= chks_flag, ">60% of Checks Pass - PROCEED", "<60% of Checks Pass - REASSESs")) %>% ungroup() %>% dplyr::select(Test, Run_Date, CHK_Flag) %>% distinct() %>% arrange(Test, Run_Date)

knitr::kable(Chks_flag_disp)

Chks_Flag_join <- Chks_flag_disp


```

\newpage

## Analyze Blanks 
```{r Analyze Blanks, echo=FALSE}


#Pull out check standards from raw file 
blks_df <- df_all %>% filter(str_detect(Sample_Name, "CCB") | str_detect(Sample_Name, "Blank")) %>% group_by(Test, Run_Date) %>% mutate(rep = row_number())


#Pull out samples from df_all to calc quantile
samples <- df_all %>% 
  filter(str_detect(Sample_Name, c("P"))) %>% 
  group_by(Test, Run_Date) %>% 
  mutate(Quant = quantile(Conc, prob=c(.25)) ) %>% 
  dplyr::select(Test, Run_Date, Quant) %>% 
  distinct() %>% 
  mutate(Quant = ifelse(Test == "Vanadium NOx", ifelse(Quant <= NOx_dl, 0.0125, Quant), Quant))  %>% #If NOx Quantile is below detection limit make blank cutoff half of detection limit
  ungroup()

blks_all <- left_join(blks_df, samples, by = join_by(Test, Run_Date))



blks_flagged <- blks_all %>% mutate(BLK_Flag = ifelse(Conc <= Quant, "YES", "NO, rerun")) %>% group_by(Test, Run_Date) %>% mutate(BLKs_Percent = 100*length(which(BLK_Flag == "YES"))/(length(BLK_Flag)) ) %>% mutate(BLK_Pct_Flag = ifelse(BLKs_Percent >= 60, ">60% of Blanks Pass - PROCEED", "<60% of Blanks Pass - REASSESS")) #Flag blanks 

#Create Data frame for displaying blank flags
blks_disp <- blks_flagged%>% group_by(Test, Run_Date) %>% mutate(Mean_Blk_Conc = mean(Conc)) %>% dplyr::select(Test, Run_Date, BLK_Pct_Flag, Mean_Blk_Conc, Quant) %>%  rename(Quantile_25 = Quant) %>% distinct() %>% arrange(Test, Run_Date)

knitr::kable(blks_disp)

Blks_Flag_join <- blks_disp %>% dplyr::select(Test, Run_Date, BLK_Pct_Flag)


NH3_blks_flagged <- blks_flagged %>% filter(Test == "Ammonia 2")

#plotting the blanks compared to the lower 25% of conc (what the flag is)

NH3_blks_plot <-  ggplot(data = NH3_blks_flagged, aes(x = rep, y = Conc, fill=BLK_Flag)) +
       geom_bar(stat = 'identity') + 
       facet_wrap(Test ~ Run_Date, scales = "free") +
       scale_x_discrete(labels = NULL, breaks = NULL) +
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
       theme_classic() + labs(x= " ", y="Concentration (mg/L)", title="NH3 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept= Quant), linetype="dashed", 
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

NH3_blks_plot


```

\newpage

## Analyze Duplicates
```{r Analyze Duplicates, echo=FALSE}

Duplicates <- df_all_cor %>% filter(Sample_Name == "Duplicate") %>%
  mutate(Sample_Num = Run_Number - 1) %>% #Get Run Number of Original Duplicated Sample
  mutate(Date_Num = paste0(Sample_Num, "_", Run_Date, "_", Test)) %>% #Create ID based on Run Date Test, and Run Number
  dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% #Select what we need
  rename(
  Dup_Conc = Conc,
  Dup_Absorbance = Absorbance,
  Dup_Dilution = Dilution
)

Dup_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test)) #Create ID for original samples

Dup_Joined <- left_join(Duplicates, Dup_ready, by = join_by(Date_Num)) #Join originals and duplicates

Dup_plot_df <- Dup_Joined %>% filter(!Sample_Name %ilike% "chk") %>% group_by(Date_Num) %>% mutate(Duplicate_Diff = 100*abs(Dup_Conc - Conc)/mean(c(Dup_Conc, Conc))) %>% #Calculate cv of duplicates
  mutate(Dup_Flag = ifelse(Duplicate_Diff >= dup_cutoff, "NO, rerun", "YES")) %>% #Flags Dups
  ungroup() %>%
  group_by(Sample_Name) %>% ungroup() %>% group_by(Sample_Name) %>% 
  mutate(
  Num_Runs = length(Sample_Name),
  Max_Dil = max(Dilution),
  Rerun = ifelse(
    Num_Runs != 1 ,
    ifelse(Dilution == Max_Dil, "KEEP", "REMOVE"),
    "KEEP"
  ) ) %>%
  filter(Rerun == "KEEP") %>% mutate(Dilution = ifelse(Dilution == 0, 1, Dilution )) %>% mutate(Within_Range = case_when(
      
      Test == "Ammonia 2" & Conc/Dilution < NH3_dl ~ "No, bdl",
      Test == "Ammonia 2" & Conc/Dilution > NH3_top    ~ "No, adl",
      Test == "Ammonia 2"                ~ "Within_Range",
      
      
      TRUE ~ NA_character_  # fallback for unexpected values
    )) %>%
  mutate(Short_Names = str_extract(Sample_Name, "(?<=_).*(?=_)")) %>% mutate(Plot_Flag = ifelse(Dup_Flag == "YES", "YES", ifelse(Within_Range == "No, bdl", "BDL", "NO, rerun")))

NH3_Dup_plot_df <- Dup_plot_df %>% filter(Test == "Ammonia 2")

#Plot Duplicates
NH3_Dup_Plot <-  ggplot(NH3_Dup_plot_df, aes(x = Short_Names, y = Duplicate_Diff, fill =  Plot_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "BDL" = "darkgreen", "NO, rerun" = "darkred"), name = "Dup Pass ?") +
  geom_hline(yintercept = dup_cutoff, linetype = "dashed" ) +
  facet_wrap(Test~Run_Date, scales = "free")  + 
  theme_classic() + 
    theme(legend.position = "bottom", strip.text = element_text( size = 8),
          axis.text.x = element_text(size = 7)
          ) +
  xlab(NULL) +
   labs(title = "NH3 Duplicate Check") + scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))

NH3_Dup_Plot


#Create Data Frame for Flagging and Displaying Duplicates QAQC
Flags_Duplicates <- Dup_plot_df %>% group_by(Run_Date, Test) %>% mutate(Dup_Flags = ifelse(length(which(Dup_Flag == "YES"))/(length(Dup_Flag)) >= 0.6,  ">60% of Dups Pass - PROCEED", "<60% of Dups Pass - REASSESS")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Dup_Flags) %>% distinct() %>% arrange(Test, Run_Date)

knitr::kable(Flags_Duplicates)

Dup_Flags_join <- Flags_Duplicates




```

\newpage

## Spikes
```{r Analyze Spikes, echo=FALSE}

Spikes <- df_all_cor %>% filter(Sample_Name == "Auto Spike") %>% mutate(Sample_Num = Run_Number - 1) %>% mutate(Date_Num = paste0(Sample_Num, "_", Run_Date, "_", Test)) %>% dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% rename(
  Spike_Conc = Conc,
  Spike_Absorbance = Absorbance,
  Spike_Dilution = Dilution
)

Spike_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test))

Spike_Joined <- left_join(Spikes, Spike_ready, by = join_by(Date_Num))

Spike_Val <- tibble(
    Test = c("Ammonia 2"),
    Spike = c(NH3_Spk)
  )

Spike_Joined2 <- left_join(Spike_Joined, Spike_Val, by = join_by(Test))

Spike_plot_df <- Spike_Joined2 %>% filter(!Sample_Name %ilike% "ppt") %>% group_by(Date_Num) %>% mutate(Spike_Exp_Recovery = Conc + Spike) %>% 
  mutate(Spike_Diff = 100*abs(Spike_Exp_Recovery - Spike_Conc)/Spike_Exp_Recovery) %>% 
  mutate(
    Spike_Flag = ifelse(
        Spike_Diff >= spike_cutoff,    #SEAL Uses 50% CV cutoff to check Spikes
      "No, rerun",
      "YES"
      )) %>% ungroup() %>%
  mutate(Short_Names = str_extract(Sample_Name, "(?<=_).*(?=_)"))

NH3_Spike_plot_df <- Spike_plot_df %>% filter(Test == "Ammonia 2")

NH3_Spike_Plot <-  ggplot(NH3_Spike_plot_df, aes(x = Short_Names, y = Spike_Diff, fill =  Spike_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "No, rerun" = "darkred"), name = "Spike CV <50% ?") +
  geom_hline(yintercept = c(spike_cutoff), linetype = "dashed" ) +
  facet_wrap(.~Run_Date, scales = "free_x") + 
  theme_classic()  + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8),
          axis.text.x = element_text(size = 7)) +
  xlab(NULL) + scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2)) +
   labs(title = "NH3 Spike Check")


NH3_Spike_Plot


Flags_Spike <- Spike_plot_df %>% group_by(Run_Date, Test) %>% mutate(Spike_Flags = ifelse(length(which(Spike_Flag == "YES"))/(length(Spike_Flag)) >= 0.6,  ">60% of Spikes have a Diff <50% - PROCEED", ">60% of Spikes have a Diff <50% - REASSESS")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Spike_Flags) %>% distinct() %>% arrange(Test, Run_Date)

knitr::kable(Flags_Spike)

Spk_Flags_join <- Flags_Spike



```

\newpage

## Matrix Effects
```{r Analyze Matrix Effects, echo=FALSE}

Matrix <- df_all_cor %>% filter(Sample_Name == "Auto Spike") %>% mutate(Sample_Num = Run_Number - 1) %>% mutate(Date_Num = paste0(Sample_Num,"_", Run_Date, "_", Test)) %>% dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% rename(Spike_Conc = Conc, Spike_Absorbance = Absorbance, Spike_Dilution = Dilution)

Spike_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test))

Spike_Joined <- left_join(Matrix, Spike_ready, by = join_by(Date_Num))

Spike_Val <- tibble(
    Test = c("Ammonia 2"),
    Spike = c(NH3_Spk)
  )

Spike_Joined2 <- left_join(Spike_Joined, Spike_Val, by = join_by(Test))

Spike_plot_df <- Spike_Joined2 %>% filter(Sample_Name %ilike% "ppt") %>% group_by(Date_Num) %>% mutate(Spike_Exp_Recovery = Conc + Spike) %>% 
  mutate(Spike_Diff = 100*abs(Spike_Exp_Recovery - Spike_Conc)/Spike_Exp_Recovery) %>% 
  mutate(
    Spike_Flag = ifelse(
        Spike_Diff >= 50,    #SEAL Uses 50% CV cutoff to check Spikes
      "No, rerun",
      "YES"
      )) %>% ungroup() %>% mutate(Sample_Name = ifelse(Sample_Name %ilike% "10ppt", "10ppt", ifelse(Sample_Name %ilike% "20ppt", "20ppt", "PPT NOT FOUND")))

NH3_Spike_plot_df <- Spike_plot_df %>% filter(Test == "Ammonia 2")


NH3_Spike_Plot <-  ggplot(NH3_Spike_plot_df, aes(x = Sample_Name, y = Spike_Diff, fill =  Spike_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "No, rerun" = "darkred"), name = "Matrix Spike Diff <50% ?") +
       scale_x_discrete(name =NULL) +
  geom_hline(yintercept = c(50), linetype = "dashed" ) +
  facet_wrap(.~Run_Date, scales = "free_x") + 
  theme_classic()  + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8),
          axis.text.x = element_text(size = 8)) +
   labs(title = "NH3 Matrix Check") +
  ylab("Matrix Spike Diff Percent")


NH3_Spike_Plot


Flags_Matrix <- Spike_plot_df %>% group_by(Run_Date, Test) %>% mutate(Matrix_Flags = ifelse(length(which(Spike_Flag == "YES"))/(length(Spike_Flag)) >= 0.6,  "Matrix Has Diff <50% - PROCEED", "Matrix Has Diff >50% - REASSESS")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Matrix_Flags) %>% distinct() %>% arrange(Test, Run_Date)%>% arrange(Test, Run_Date)

knitr::kable(Flags_Matrix)

Matrix_Flags_join <- Flags_Matrix


```

```{r Add unit conversion, include=FALSE}

#Pull out samples 
samples <- df_all_cor %>%  
  filter(str_detect(Sample_Name, c("P")))


#Convert values based on the Test ID 
samples <- samples %>%
  mutate(
    Conc_uM = case_when(
      Test == "Ammonia 2" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      TRUE ~ as.numeric(NA)  # fallback in case of unexpected value
    ),
    # Replace negatives with 0 in 'value' and 'value_converted'
    Conc = pmax(Conc, 0),
    Conc_uM = pmax(Conc_uM, 0)
  )

head(samples)

```
\newpage


##Sample Flagging - Within range of standard curve
```{r Sample Flagging, echo=FALSE, fig.align='center'}

#Flagging data if the concentration is outside the standards range 
samples_flagged1 <- samples %>% mutate(Dilution = ifelse(Dilution == 0, 1, Dilution )) %>%
  mutate(
    Conc_flag = case_when(

      Test == "Ammonia 2" & Conc/Dilution < NH3_dl ~ "bdl",
      Test == "Ammonia 2" & Conc/Dilution > NH3_top    ~ "adl",
      Test == "Ammonia 2"                ~ "Within_Range",

      TRUE ~ NA_character_  # fallback for unexpected values
    )
  )

NH3_dl_df <- samples_flagged1 %>% filter(Test == "Ammonia 2")

  
NH3_dl_plot <- ggplot(NH3_dl_df, aes(x = Sample_Name, y = Conc, fill = Conc_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("Within_Range" = "darkgreen", "adl" = "darkred", "bdl" = "navy"), name = "Detection Limit") +
       scale_x_discrete(name =NULL) +
  facet_wrap(.~Run_Date, scales = "free_x") + 
  theme_classic()  + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8),
          axis.text.x = element_blank()) +
   labs(title = "NH3 Detection Limits") +
  ylab("Concentration (mg/L)")

NH3_dl_plot

  

```

##Add QAQC Flags
```{r add QAQC Flags, echo = FALSE}
Flags_2 <- left_join(R2_Flags_join, PE_Flag_join, by = join_by(Test, Run_Date))

Flags_3 <- left_join(Flags_2, Chks_Flag_join, by = join_by(Test, Run_Date) )

Flags_4 <- left_join(Flags_3, Blks_Flag_join, by = join_by(Test, Run_Date) )

Flags_5 <- left_join(Flags_4, Spk_Flags_join, by = join_by(Test, Run_Date) )

Flags_6 <- left_join(Flags_5, Dup_Flags_join, by = join_by(Test, Run_Date) )
  
Flags_joined <- left_join(Flags_6, Matrix_Flags_join, by = join_by(Test, Run_Date) )

Flags_Bad <- Flags_joined %>% mutate(across(!1:2, \(x) ifelse(x %ilike% "PROCEED" | x %ilike% "PASS" | x %ilike% "NA" | is.na(x), "", x))) %>% distinct()

Flags_Fin <- Flags_Bad %>% mutate(QAQC_Flag = paste0(R2_Flag, PE_Flag, CHK_Flag, BLK_Pct_Flag, Dup_Flags, Spike_Flags, Matrix_Flags)) %>% dplyr::select(Test, Run_Date, QAQC_Flag)
```

## Create Log for QAQC
```{r}
#Old_Log_QAQC <- tibble(Test = c(), Mean_PE_Conc= c(), Spike_CV= c(), Top_Std_Absorbance= c(), Run_Date= c()) #create standard log in proper format
# Old_Log_QAQC <- read.csv(file_path_QAQC) %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% mutate(Run = "Past Runs")
# 
# Top_Standards <- Slope_raw %>% group_by(Test, Run_Date) %>% filter(Sample_Name == "Standard 2.0000") %>% dplyr::select(Test, Run_Date, Absorbance) %>% distinct() %>% ungroup() %>% rename(Top_Std_Absorbance = Absorbance)
# 
# PE_Check <- NH3_peChk %>% group_by(Test, Run_Date) %>% dplyr::select(Test, Run_Date, Mean_Chk_Conc) %>% rename(Mean_PE_Conc = Mean_Chk_Conc) %>% ungroup()
# 
# Spikes_Log <- Spike_plot_df %>% group_by(Test, Run_Date) %>% dplyr::select(Test, Run_Date, Spike_CV) %>% ungroup()
# 
# Joined_QAQC <- left_join(Top_Standards, PE_Check, by = join_by(Test, Run_Date)) %>%
#   left_join(Spikes_Log, by = join_by(Test, Run_Date)) %>% mutate(Run = "Current Runs") %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% rbind(Old_Log_QAQC) %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% distinct()
# 
# Spike_Log_Plot <- ggplot(Joined_QAQC, aes(x = Run_Date, y = Spike_CV)) +
#   geom_point(size = 3, aes(col = Run)) +
#  # geom_path(linewidth = 1) +
#   facet_grid(~Test, scales = "free") +
#   labs(title = "Spike") +
#   theme_classic()
# 
# Spike_Log_Plot
# 
# TopStd_Log_Plot <- ggplot(Joined_QAQC, aes(x = Run_Date, y = Top_Std_Absorbance)) +
#   geom_point(size = 3, aes(col = Run)) +
#  # geom_path(linewidth = 1) +
#   facet_grid(~Test, scales = "free") +
#   labs(title = "Top Standard Absorbance") +
#   theme_classic()
# 
# TopStd_Log_Plot
# 
# PE_Log_Plot <- ggplot(Joined_QAQC, aes(x = Run_Date, y = Mean_PE_Conc)) +
#   geom_point(size = 3, aes(col = Run)) +
#   #geom_path(linewidth = 1) +
#   facet_grid(~Test, scales = "free") +
#   labs(title = "Performance Check") +
#   theme_classic()
# 
# PE_Log_Plot


```


## Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata 
all_present <- all(metadata$Sample_Name %in% samples_flagged1$Sample_Name)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Flagged_Samples$Sample_Name)
  print(missing_ids)
}


#merge metadata with sample run data 
merged_data <- samples_flagged1 %>%
  left_join(metadata, by = c("Sample_Name" = "Sample_Name"))  %>% left_join(Flags_Fin, by = join_by(Test, Run_Date))

#might want to include a flag or something to see that all the samples are included? 

df_all_clean <- merged_data 

```

\newpage
## Organize Data
```{r Organize Data, include=FALSE}

cat("Export Processed Data")

df_all_clean_cols <- merged_data %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Sample_Name,
    Plot_Name,
    SubPlot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    Conc,
    Conc_uM,
    Conc_flag,
    NH3_flag
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = run_notes,# new column with notes about the run
    Analysis_rundate = run_date) 

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
   NH3_Conc = Conc,
   NH3_Conc_uM = Conc_uM,
    Plot = Plot_Name,
   Sample_ID = Sample_Name
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Sample_ID,
    Plot,
    SubPlot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    NH3_Conc,
    NH3_Conc_uM,
    Conc_flag,
    NH3_flag,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)


```

## Visualize Data
```{r Visualize Data, echo=FALSE}
cat("Visualize Data")

### Ammonia
NH3_forplot <- All_Clean_Data

color1 <- c("P1", "P2", "P3", "P4")
color2 <- c("P5", "P6", "P7", "P8")
color3 <- c("P10", "P9", "P11", "P12")


NH3_forplot <- NH3_forplot %>%
  mutate(
    Treatment1 = ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Amended",
      "eCO2 & eN",
      ifelse(
        CO2_Treatment == "Elevated" &
          N_Treatment == "Ambient",
        "eCO2 & aN",
        ifelse(
          CO2_Treatment == "Ambient" &
            N_Treatment == "Amended"  ,
          "aCO2 & eN" ,
          "aCO2 & aN"
        )
      )
    )
  ) %>%
  mutate(Treatment = ifelse(
    SubPlot == "A",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Amended",
      "eCO2_eN_Phrag",
      ifelse(
        CO2_Treatment == "Elevated" &
          N_Treatment == "Ambient",
        "eCO2_aN_Phrag",
        ifelse(
          CO2_Treatment == "Ambient" &
            N_Treatment == "Amended"  ,
          "aCO2_eN_Phrag" ,
          "aCO2_aN_Phrag"
        )
      )
    ),
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Amended",
      "eCO2_eN_Native",
      ifelse(
        CO2_Treatment == "Elevated" &
          N_Treatment == "Ambient",
        "eCO2_aN_Native",
        ifelse(
          CO2_Treatment == "Ambient" &
            N_Treatment == "Amended"  ,
          "aCO2_eN_Native" ,
          "aCO2_aN_Native"
        )
      )
    )
  ))  %>%
  mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2"))  %>%
  mutate(N_Treatment = ifelse(N_Treatment == "Ambient", "aN", "eN")) %>%
  mutate(
    Plot_Color = case_when(
      Plot %in% color1 ~ "Color1",
      Plot %in% color2 ~ "Color2",
      Plot %in% color3 ~ "Color3"
    )
  )

NH3_forplot <- NH3_forplot[order(NH3_forplot$Depth_cm), ]

viz_NH3_plot_v2 <- ggplot(
  data = NH3_forplot,
  aes(
    x = factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")),
    y = NH3_Conc_uM,
    fill = as.factor(Treatment),
    color = as.factor(Treatment),
    pattern = as.factor(Treatment)
  )
) +
  geom_bar_pattern(
    stat = "identity",
    position = position_dodge2(padding = 0.25, preserve = "single"),
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.025,
    pattern_key_scale_factor = 0.6
  ) +
  facet_grid(Depth_cm ~ N_Treatment, scales = "fixed") +
  scale_fill_manual(
    values = c(
      "eCO2_eN_Phrag" = "#DB5461",
      "eCO2_aN_Phrag" = "#454ADE",
      "aCO2_eN_Phrag" = "white",
      "aCO2_aN_Phrag" = "white",
      "eCO2_eN_Native" = "#DB5461",
      "eCO2_aN_Native" = "#454ADE",
      "aCO2_eN_Native" = "white",
      "aCO2_aN_Native" = "white"
    ),
    name = "Treatment:",
    breaks = c(
      "eCO2_eN_Phrag",
      "eCO2_aN_Phrag",
      "aCO2_aN_Phrag",
      "aCO2_eN_Phrag",
      "eCO2_eN_Native",
      "eCO2_aN_Native",
      "aCO2_aN_Native",
      "aCO2_eN_Native"
    ),
    labels = c(
      "eCO2_eN_Phrag",
      "eCO2_aN_Phrag",
      "aCO2_aN_Phrag",
      "aCO2_eN_Phrag",
      "eCO2_eN_Native",
      "eCO2_aN_Native",
      "aCO2_aN_Native",
      "aCO2_eN_Native"
    )
    
  ) +
  scale_color_manual(
    values = c(
      "eCO2_eN_Phrag" = "#DB5461",
      "eCO2_aN_Phrag" = "#454ADE",
      "aCO2_eN_Phrag" = "#DB5461",
      "aCO2_aN_Phrag" = "#454ADE",
      "eCO2_eN_Native" = "#DB5461",
      "eCO2_aN_Native" = "#454ADE",
      "aCO2_eN_Native" = "#DB5461",
      "aCO2_aN_Native" = "#454ADE"
    ),
    name = "Treatment:",
    breaks = c(
      "eCO2_eN_Phrag",
      "eCO2_aN_Phrag",
      "aCO2_aN_Phrag",
      "aCO2_eN_Phrag",
      "eCO2_eN_Native",
      "eCO2_aN_Native",
      "aCO2_aN_Native",
      "aCO2_eN_Native"
    ),
    labels = c(
      "eCO2_eN_Phrag",
      "eCO2_aN_Phrag",
      "aCO2_aN_Phrag",
      "aCO2_eN_Phrag",
      "eCO2_eN_Native",
      "eCO2_aN_Native",
      "aCO2_aN_Native",
      "aCO2_eN_Native"
    )
    
  ) +
  scale_pattern_manual(
    values = c(
      "eCO2_eN_Phrag" = "stripe",
      "eCO2_aN_Phrag" = "stripe",
      "aCO2_eN_Phrag" = "stripe",
      "aCO2_aN_Phrag" = "stripe",
      "eCO2_eN_Native" = "none",
      "eCO2_aN_Native" = "none",
      "aCO2_eN_Native" = "none",
      "aCO2_aN_Native" = "none"
    ),
    name = "Treatment:",
    breaks = c(
      "eCO2_eN_Phrag",
      "eCO2_aN_Phrag",
      "aCO2_aN_Phrag",
      "aCO2_eN_Phrag",
      "eCO2_eN_Native",
      "eCO2_aN_Native",
      "aCO2_aN_Native",
      "aCO2_eN_Native"
    ),
    labels = c(
      "eCO2_eN_Phrag",
      "eCO2_aN_Phrag",
      "aCO2_aN_Phrag",
      "aCO2_eN_Phrag",
      "eCO2_eN_Native",
      "eCO2_aN_Native",
      "aCO2_aN_Native",
      "aCO2_eN_Native"
    )
  ) +
  theme_classic() +
  labs(
    x = " ",
    y = "NH3 (uM)",
    title = "CO2xNxPhrag: Porewater Methane",
    fill = "CO2 Treatment:",
    color = "N Treatment:",
    pattern = "Subplot:"
  ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  ) +
  coord_flip()
print(viz_NH3_plot_v2)


```
```{r, echo = FALSE}
#make plot

NH3_forplot <- NH3_forplot %>% 
  mutate(Sub_Plot = ifelse(SubPlot == "A", "Phrag", "Native"))

viz_NH3_plot_2 <- ggplot(data = NH3_forplot, aes(x = NH3_Conc_uM, y = as.numeric(Depth_cm), color = Plot_Color, shape = Sub_Plot)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(
    values = c(
      "Color1" = "#FFD500",
      "Color2" = "#00D0FF",
      "Color3" = "#FF0075"
   
    ),
    labels = c("1", "2", "3")
  ) +
  facet_grid(Treatment1 ~ ., scales = "fixed") +
  labs(
    x = "NH3 (uM)",
    y = "Depth (cm)",
    title = "CO2xNxPhrag: Porewater NH3 by Replicate",
    color = "Replicate"
  ) +
  scale_y_reverse(breaks = c(20, 40), limits = c(45, 15)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_NH3_plot_2
```

##Export Data

```{r, echo = FALSE}

#Write out data frame 
  write.csv(All_Clean_Data, final_path)
  

```


#end