---
title: "CO2xN: Porewater Sulfide"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
    number_sections: false
geometry: left=2cm,right=2cm,top=1cm,bottom=2cm
---
## Code Set up 
```{r setup, include=FALSE}

library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
library(purrr)
library(tidyverse)
library(here)
library(data.table)
library(matrixStats)
library(gridExtra)
library(grid)
library(plater)
library(raster)
library(knitr)  
library(readxl)


```

## run information
```{r Information to be changed}

##Before or After Reruns (YES or NO)
After_Rerun = "YES"
#If NO will not write final data file
  #And Will Write a Rerun File

###things that need to be changed

sample_month = "May"
sample_year = "2025"
Date_Run = "20250527"
Run_by = "Melanie Giessner"#Instrument user
Script_run_by ="Brenden Blakley"#Code user

Run_notes=""#any notes from run

#Import Plate Data

plates<- c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate_rr")

Raw_plates <- c("Raw Data/20250527_LTREB_H2S_Plate1.csv","Raw Data/20250527_LTREB_H2S_Plate2.csv","Raw Data/20250527_LTREB_H2S_Plate3.csv","Raw Data/20250527_LTREB_H2S_Plate4.csv","Raw Data/20250527_LTREB_H2S_Plate5.csv","Raw Data/20250527_LTREB_H2S_Plate6.csv","Raw Data/20250527_LTREB_H2S_Plate7.csv","Raw Data/20250602_H2S_LTREBrr.csv")

#Import Run Date for Each Plate
Run_dates <- tibble(
    Plate = c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate_rr"),
    Run_Date = c("05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","06/02/2025")
  )

  Run_dates <- Run_dates %>% mutate(Run_Date = as.Date(Run_Date, format = "%m/%d/%Y"))


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Processed Data/LTREB_2M_H2S_QAQC_2025.csv"

# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_CO2xN_Porewater_H2S_202505.csv"

#Stds that should be excluded
#stds_to_remove<-data.frame(Plate=c("Plate4","STD2"),IDs=c("Std 4","Std 3"))


```


```{r, echo=FALSE}


#file path name for processed data
summarizeddata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_Summary_Data.csv")

fulldata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_Data.csv")

shortdata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_short_Data.csv")

Need_dilute=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_Need_dilute_Data.csv")

high_CV=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_high_CV_Data.csv")


```

```{r Set Up Code - constants and QAQC cutoffs, include=FALSE}

#Flag cutoffs
  r2_cutoff = 0.985            #this is the level below which we want to rerun or consider a curve 
  cv_flag_stds = 10           #this is the maximum cv allowed for standards
  p_value_chkstds = 0.05      #the p-value for the t-test between check standards must be greater than this
  p_value_MC = 0.05           #the p-value for the t-test between matrix checks and top std must be greater than this
  cv_flag_sample = 10         #this is the maximum cv allowed for samples
  dups_perc_diff = 15.5       #this is the maximum percent difference allowed between duplicates
  high_recovery_cutoff = 120  #this is the maximum percent recovery of SO4 allowed in spiked samples
  low_recovery_cutoff = 80    #this is the minimum percent recovery of SO4 allowed in spiked samples

#Standard concentrations - Update if running different standard curve:
# standard units are in uM
  standards <- tibble(
    IDs = c("Std 0", "Std 1", "Std 2", "Std 3", "Std 4", "Std 5"),
    Conc = c(0, 5.0, 12.5, 25.0, 50.0, 100.0)
  )  #uM
  
    chk_standards <- tibble(
    IDs = c("ChkStd 0", "ChkStd 3", "ChkStd 4"),
    Conc = c(0, 25.0, 50.0)
  )  #uM
    
    mtx_standards <- tibble(
    IDs = c("MC: 10ppt S5"),
    Conc = c(100)
  )  #uM

#Spike concentration calc
  #spike for these samples was 50 uL of the 100uM standard
  Con1 <- 1000000               #conversion factor value for spike volumes (uL -> L)
  spk_std <- 100                # uM S2- standard used
  spkvol <- 50                  # uL volume of spike added
  spk_Conc <- spk_std*(spkvol/Con1)  # umoles of S2- added to each spiked sample
  sample_vol <- 250             # the sample volume without the spike is 250 uL

#Top standard Concentration- Update if running different standard curve: 
   Top_STD = 100
   
   
#Metadata file path
Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata

raw_metadata <- read.csv(Raw_Metadata)
#clean up columns in metadata file:
#Pull out only SMARTX Rows
#Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("2-Meter", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  ) %>%
  dplyr::select(-Temp_Treatment) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2_N",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2",
      ifelse(CO2_Treatment == "Ambient" &
               N_Treatment == "Amended"  , "N" , "Ambient")
    )
  )) 
#Bring in the collection date and any field notes if relevant
#collection_metadata <- read.csv(Collection_Dates)
#then we will merge the collection date for SMARTX based on the project and the chambers

```



## read in data
```{r read in plates,echo=FALSE,include=FALSE}

#read in the csv files
dat<-read_plates(
  files = Raw_plates,#list of all file paths
  plate_names = plates ,  #list of plate names          
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
) %>%
  rename("Abs"=values,
         "IDs"=values.2,
         "Dilution"=values.3)
head(dat)
  
```

#Remove High CV Stds
```{r}

#filter for STDs
H2S_stds_raw <- dat %>% subset(IDs %like% "Std") %>% subset(!IDs %like% "Chk")

#Add in STD values
H2S_stds <- left_join(H2S_stds_raw, standards, by = "IDs")

#Calculate CVs
H2S_STD_cv <- H2S_stds %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_STD_cv_cor <- H2S_STD_cv %>% mutate(Mean_diff = abs(Abs - Mean_Abs)) %>% group_by(Plate, IDs) %>% mutate(Abs_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Abs), Abs))

H2S_STD_cv_new <- H2S_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>% mutate(Abs_SD2 = sd(Abs_rm, na.rm = TRUE)) %>% mutate(CV2 = (Abs_SD2 /
                                                                                                                                                                                Mean_Abs2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_STD_CV_fin <- H2S_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate, 
                                                Wells,                                                                                                                Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Conc,
                                                Mean_Abs,
                                                CV_Flag)


```


#Check R2 & Remove Standards if Necessary
```{r}
split_data <- H2S_STD_CV_fin %>% dplyr::select(Plate, IDs, Conc, Mean_Abs, CV_Flag) %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_H2S_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Conc ~ Mean_Abs, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ Mean_Abs, data = .))[2],
                                        Intercept = coef(lm(Conc ~ Mean_Abs, data = .))[1]))

 Slope_H2S_raw <- bind_rows(Slope_H2S_raw_list)
  
 Slope_H2S_Std_Crvs <- Slope_H2S_raw %>% dplyr::select(Plate, Curve, R2, Slope, Intercept) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))

 
 R2_Check_Raw <- left_join(Slope_H2S_Std_Crvs, H2S_STD_CV_fin, by = "Plate")
 
 R2_Check <- R2_Check_Raw %>% 
   dplyr::select(Plate,
                 IDs,
                 Mean_Abs,
                 Conc,
                 R2,
                 Slope,
                 Intercept,
                 CV_Flag,
                 R2_Flag) %>%
   distinct()
 
 R2_Check_diff <- R2_Check %>% mutate(Calc_Conc = ((Slope*Mean_Abs) + Intercept)) %>%
   mutate(Conc_Diff = abs(Conc - Calc_Conc)) %>% mutate(RMVBL = ifelse(IDs %in% c("Std 0", "Std 1","Std 5"), "No", "Yes"))

 R2_Check_RMVBL <- R2_Check_diff %>% filter(RMVBL == "Yes")
 
 R2_Check_NRMVBL <- R2_Check_diff %>% filter(RMVBL != "Yes") %>% mutate(STD_ABS_rm = Mean_Abs) %>% mutate(N_HCV = 0, STD_ABS_rmv = STD_ABS_rm, N_stds = 3)
 
 
 R2_Check_na <- R2_Check_RMVBL %>% group_by(Plate) %>% mutate(STD_ABS_rm = ifelse(R2_Flag == "R2_Bad", ifelse(max(Conc_Diff) == Conc_Diff, NA, Mean_Abs), Mean_Abs)) %>% ungroup()
 
 R2_Check_CVrm <- R2_Check_na %>% na.omit() %>% group_by(Plate) %>% mutate(N_stds = length(STD_ABS_rm)) %>% group_by(Plate, CV_Flag) %>%
   mutate(N_HCV = n()) %>% ungroup() %>% group_by(Plate) %>%
   mutate(STD_ABS_rmv = ifelse(
     N_stds == 3,
     ifelse(
       CV_Flag == "CV_High" ,
       ifelse(
         N_HCV == 1 ,
         NA,
         ifelse(max(Conc_Diff) == Conc_Diff, NA, STD_ABS_rm)
       ),
       STD_ABS_rm
     ),
     STD_ABS_rm
   ))
 
  R2_Check_bind <- rbind(R2_Check_NRMVBL,R2_Check_CVrm) %>% na.omit()
 
 R2_Check_fin <- R2_Check_bind[order(R2_Check_bind$Plate,R2_Check_bind$IDs), ]
 
 
 
 split_data_rmv <- R2_Check_fin %>% dplyr::select(Plate, IDs, Conc, STD_ABS_rmv, CV_Flag, R2_Flag) %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
 
 Slope_H2S_raw_list_rm <- map(split_data_rmv, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Conc ~ STD_ABS_rmv, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ STD_ABS_rmv, data = .))[2],
                                        Intercept = coef(lm(Conc ~ STD_ABS_rmv, data = .))[1]))

 Slope_H2S_raw_rm <- bind_rows(Slope_H2S_raw_list_rm)
  
 Slope_H2S_Std_Crvs_rm <- Slope_H2S_raw_rm %>% dplyr::select(Plate, Curve, R2, Slope, Intercept, CV_Flag) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))
 
 
 
```



```{r Plot all Standard Curves together, echo=FALSE,message=FALSE,fig.keep='none'}
##this is to see if there are any stds or plates that look off
#Remove check standards which have NAs in the Conc column


H2S_Curve <- ggplot(Slope_H2S_raw_rm, aes(STD_ABS_rmv, Conc, color = Plate)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "H2S Standard Curves", x = "Absorbance", y = "Concentration (uM)")
H2S_Curve


STD_run_date <- left_join(Run_dates, Slope_H2S_Std_Crvs_rm)
new_log_stds <- STD_run_date %>% dplyr::select(Plate,
   Curve,
   R2,
   Slope,
   Intercept,
   Run_Date) %>% mutate(R2 = round(R2, digits = 4), Slope = round(Slope, digits = 4), Intercept = round(Intercept, digits = 4), )


log <- read.csv(log_path)
log_2 <- log %>% mutate(Run_Date = as.Date(Run_Date))
head(log_2)
log_3 <- log_2[, -c(1)]
log_4 <- rbind(log_3, new_log_stds)

unique_log <- log_4 %>% distinct(.keep_all = TRUE)

Slopes_chk <- ggplot(unique_log, aes(x = Run_Date, y = Slope)) +
    ylim(50,175) +
  #need to find range for slopes
  #geom_abline(intercept = 60, slope = 0 )+
  #geom_abline(intercept = 150, slope = 0 )+
  geom_point(size = 4) +
  theme_bw()

Slopes_chk

write.csv(unique_log, log_path)


```

#Return Flags For Standard Curve
```{r}

STD_CRV_flags <- full_join(Slope_H2S_Std_Crvs_rm, Run_dates, by = join_by(Plate))

ifelse(STD_CRV_flags$CV_Flag == "CV_High", paste0("CV High", "-", STD_CRV_flags$Plate), paste0("CV Good", "-", STD_CRV_flags$Plate))

ifelse(STD_CRV_flags$R2_Flag == "R2_Bad", paste0("R2 Bad", "-", STD_CRV_flags$Plate), paste0("R2 Good", "-", STD_CRV_flags$Plate))

```

## Remove High CV Check Standards
```{r}
H2S_Chk_Stds <- dat %>% filter(IDs %like% "ChkStd")

#Add in STD values
H2S_chk_stds <- full_join(H2S_Chk_Stds, chk_standards, by = join_by("IDs"))

#Calculate CVs
H2S_CHK_STD_cv <- H2S_chk_stds  %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_CHK_STD_cv_cor <- H2S_CHK_STD_cv %>% mutate(Mean_diff = abs(Abs - Mean_Abs)) %>% group_by(Plate, IDs) %>% mutate(Abs_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Abs), Abs))

H2S_CHK_STD_cv_new <- H2S_CHK_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>% mutate(Abs_SD2 = sd(Abs_rm, na.rm = TRUE)) %>% mutate(CV2 = (Abs_SD2 /Mean_Abs2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_CHK_STD_CV_fin <- H2S_CHK_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate,                                                                                                      Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Abs,
                                                CV_Flag) 

```

##Compare Check Standards to Standards

```{r, echo = FALSE}
H2S_CHKs_a <- H2S_CHK_STD_CV_fin %>%
  mutate(Std_ID = ifelse(
    IDs == "ChkStd 0",
    paste0("Std 0", Plate),
    ifelse(
      IDs == "ChkStd 3",
      paste0("Std 3", Plate),
      ifelse(IDs == "ChkStd 4", paste0("Std 4", Plate), "ID Not Found")
    )
  )) %>%
  mutate(
    Chk_Abs = Abs,
    Chk_Removed = Removed,
    Chk_IDs = IDs,
    Chk_Mean_Abs = Mean_Abs,
    Chk_CV_Flag = CV_Flag
  ) %>% dplyr::select(Std_ID,
                      Chk_Abs,
                      Chk_Removed,
                      Chk_IDs,
                      Chk_Mean_Abs,
                      Chk_CV_Flag,
                      Plate) %>% filter(Chk_Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Chk_CI_low = t.test(Chk_Abs)$conf.int[1],
    Chk_CI_high = t.test(Chk_Abs)$conf.int[2]
  ) %>%
  ungroup() %>%
  dplyr::select(Std_ID, Chk_IDs, Chk_CI_low, Chk_CI_high, Chk_CV_Flag, Chk_Mean_Abs, Plate) %>% distinct(.keep_all = TRUE)

H2S_CHKs_1 <- H2S_CHKs_a %>% dplyr::select(-Plate, -Chk_Mean_Abs)

H2S_STD_forCHK_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 0", "Std 3", "Std 4")) %>% mutate(Std_ID = ifelse(
  IDs == "Std 0",
  paste0("Std 0", Plate),
  ifelse(
    IDs == "Std 3",
    paste0("Std 3", Plate),
    ifelse(IDs == "Std 4", paste0("Std 4", Plate), "ID Not Found")
  )
)) %>% filter(Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Std_CI_low = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[1]),
    Std_CI_high = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Plate, Std_ID, IDs, Std_CI_low, Std_CI_high, CV_Flag, Mean_Abs) %>% distinct(.keep_all = TRUE)

H2S_STD_forCHK <- H2S_STD_forCHK_a %>% dplyr::select(-Mean_Abs)

Chk_Std_join <- full_join(H2S_CHKs_1, H2S_STD_forCHK, by = join_by(Std_ID))

Chk_Std_flags <- Chk_Std_join %>% mutate(
  Chk_Flag_Std = ifelse(
    Chk_CI_low <= Std_CI_high |
      Std_CI_low <= Chk_CI_high,
    "Chk_Good",
    "Chk_Bad"
  )
) %>% mutate(CHK_Flag = paste0(gsub(" ", "", IDs), Chk_Flag_Std)) %>%
  mutate(CHK_CV_Flag = paste0("Chk",gsub(" ", "", IDs), Chk_CV_Flag))

Chk_Flags <- Chk_Std_flags %>% group_by(Plate) %>% mutate(Chk_Flag = ifelse(any(CHK_Flag %like% "Bad"), "Chk_Bad", "Chk_Good"), Chk_CV_Flag = ifelse(any(CHK_CV_Flag %like% "High"), "Chk_CV_High", "Chk_CV_Good")) %>% ungroup() %>% mutate(CHK_CV_Flags = ifelse(CHK_CV_Flag %like% "High", paste0(gsub(" ", "", IDs), Chk_CV_Flag), "Good")) %>% mutate(CHKs_Flag = ifelse(CHK_Flag %like% "Bad", CHK_Flag, "Good")) 


H2S_CHKs_plot <- H2S_CHKs_a %>% rename(IDs = Chk_IDs, Std_CI_low = Chk_CI_low, Std_CI_high = Chk_CI_high, Mean_Abs = Chk_Mean_Abs) %>% mutate(Type = "Check") %>% dplyr::select(-Chk_CV_Flag)

H2S_Std_chkplot <- H2S_STD_forCHK_a %>% dplyr::select(IDs, Std_CI_low, Std_CI_high, Mean_Abs, Std_ID, Plate) %>% mutate(Type = "Standard")

Chk_flags_plot <- Chk_Flags %>% dplyr::select(Std_ID, CHKs_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(Check_Flag = ifelse(CHKs_Flag == "Good", "Good", "Bad"))

H2S_Chks_bind <- rbind(H2S_CHKs_plot, H2S_Std_chkplot)

H2S_Chks_bind_1 <- left_join(H2S_Chks_bind, Chk_flags_plot, by = join_by(Std_ID)) 

H2S_Chks_plot_bind <- H2S_Chks_bind_1 %>% mutate(Plot_ID = ifelse(IDs %like% "Chk", substr(IDs, 4, 8), IDs))
 

H2S_chk_plot <- ggplot(data = H2S_Chks_plot_bind, aes(x = Plot_ID, y = Mean_Abs, col = Type, fill = Check_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  geom_errorbar(aes(ymin = Std_CI_low, ymax = Std_CI_high), stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  facet_wrap(.~Plate) +
  scale_color_manual(values =c("Standard" = "lightgreen", "Check" = "cornflowerblue" )) +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Check Stds: H2S") +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1))


H2S_chk_plot

```

##Display Any Check Flags
```{r}
Chk_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHKs_Flag) %>% filter(CHKs_Flag != "Good")

CHK_CV_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags) %>% filter(CHK_CV_Flags != "Good")

ifelse(length(Chk_Flag_1$CHKs_Flag) >= 1, paste0(Chk_Flag_1$CHKs_Flag, "-", Chk_Flag_1$Plate), paste0("All Checks Good"))

ifelse(length(CHK_CV_Flag_1$CHK_CV_Flags) >= 1, paste0(CHK_CV_Flag_1$CHK_CV_Flags, "-", CHK_CV_Flag_1$Plate), paste0("All Checks CV Good"))

```


## Remove High CV Matrix
```{r}
H2S_Mtx_Stds <- dat %>% filter(IDs %like% "ppt")

#Add in STD values
H2S_mtx_stds <- full_join(H2S_Mtx_Stds, mtx_standards, by = join_by("IDs"))

#Calculate CVs
H2S_MTX_STD_cv <- H2S_mtx_stds  %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_MTX_STD_cv_cor <- H2S_MTX_STD_cv %>% mutate(Mean_diff = abs(Abs - Mean_Abs)) %>% group_by(Plate, IDs) %>% mutate(Abs_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Abs), Abs))

H2S_MTX_STD_cv_new <- H2S_MTX_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>% mutate(Abs_SD2 = sd(Abs_rm, na.rm = TRUE)) %>% mutate(CV2 = (Abs_SD2 /Mean_Abs2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_MTX_STD_CV_fin <- H2S_MTX_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate,                                                                                                      Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Abs,
                                                CV_Flag) 

```

##Compare Matrix to Standards

```{r, echo = FALSE}
H2S_MTXs_a <- H2S_MTX_STD_CV_fin %>%
  mutate(Std_ID = paste0(gsub(" ", "",Plate), "MTX_STD_5")) %>%
  mutate(
    Mtx_Abs = Abs,
    Mtx_Removed = Removed,
    Mtx_IDs = IDs,
    Mtx_Mean_Abs = Mean_Abs,
    Mtx_CV_Flag = CV_Flag
  ) %>% dplyr::select(Std_ID,
                      Mtx_Abs,
                      Mtx_Removed,
                      Mtx_IDs,
                      Mtx_Mean_Abs,
                      Mtx_CV_Flag,
                      Plate) %>% filter(Mtx_Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Mtx_CI_low = t.test(Mtx_Abs)$conf.int[1],
    Mtx_CI_high = t.test(Mtx_Abs)$conf.int[2]
  ) %>%
  ungroup() %>%
  dplyr::select(Std_ID, Mtx_IDs, Mtx_CI_low, Mtx_CI_high, Mtx_CV_Flag, Mtx_Mean_Abs, Plate) %>% distinct(.keep_all = TRUE)

H2S_MTXs_1 <- H2S_MTXs_a %>% dplyr::select(-Plate, -Mtx_Mean_Abs)

H2S_STD_forMTX_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 5")) %>% mutate(Std_ID = paste0(gsub(" ", "",Plate),"MTX_STD_5")) %>% filter(Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Std_CI_low = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[1]),
    Std_CI_high = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Plate, Std_ID, IDs, Std_CI_low, Std_CI_high, CV_Flag, Mean_Abs) %>% distinct(.keep_all = TRUE)

H2S_STD_forMTX <- H2S_STD_forMTX_a %>% dplyr::select(-Mean_Abs)

Mtx_Std_join <- full_join(H2S_MTXs_1, H2S_STD_forMTX, by = join_by(Std_ID))

Mtx_Std_flags <- Mtx_Std_join %>% mutate(
  Mtx_Flag_Std = ifelse(
    Mtx_CI_low <= Std_CI_high |
      Std_CI_low <= Mtx_CI_high,
    "Mtx_Good",
    "Mtx_Bad"
  )
) %>% mutate(MTX_Flag = paste0(gsub(" ", "", IDs), Mtx_Flag_Std)) %>%
  mutate(MTX_CV_Flag = paste0("Mtx",gsub(" ", "", IDs), Mtx_CV_Flag))

Mtx_Flags <- Mtx_Std_flags %>% group_by(Plate) %>% mutate(Mtx_Flag = ifelse(any(MTX_Flag %like% "Bad"), "Mtx_Bad", "Mtx_Good"), Mtx_CV_Flag = ifelse(any(MTX_CV_Flag %like% "High"), "Mtx_CV_High", "Mtx_CV_Good")) %>% ungroup() %>% mutate(MTX_CV_Flags = ifelse(MTX_CV_Flag %like% "High", paste0(gsub(" ", "", IDs), Mtx_CV_Flag), "Good")) %>% mutate(MTXs_Flag = ifelse(MTX_Flag %like% "Bad", MTX_Flag, "Good")) 


H2S_MTXs_plot <- H2S_MTXs_a %>% rename(IDs = Mtx_IDs, Std_CI_low = Mtx_CI_low, Std_CI_high = Mtx_CI_high, Mean_Abs = Mtx_Mean_Abs) %>% mutate(Type = "Matrix Check") %>% dplyr::select(-Mtx_CV_Flag)

H2S_Std_mtxplot <- H2S_STD_forMTX_a %>% dplyr::select(IDs, Std_CI_low, Std_CI_high, Mean_Abs, Std_ID, Plate) %>% mutate(Type = "Standard")

Mtx_flags_plot <- Mtx_Flags %>% dplyr::select(Std_ID, MTXs_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(Matrix_Flag = ifelse(MTXs_Flag == "Good", "Good", "Bad"))

H2S_Mtxs_bind <- rbind(H2S_MTXs_plot, H2S_Std_mtxplot)

H2S_Mtxs_bind_1 <- left_join(H2S_Mtxs_bind, Mtx_flags_plot, by = join_by(Std_ID)) 

H2S_Mtxs_plot_bind <- H2S_Mtxs_bind_1 %>% mutate(Plot_ID = ifelse(IDs %like% "Mtx", substr(IDs, 4, 8), IDs), Plot_ID = ifelse(Plot_ID == "Std 5", Plot_ID, "10ppt Std 5"))
 

H2S_mtx_plot <- ggplot(data = H2S_Mtxs_plot_bind, aes(x = Plot_ID, y = Mean_Abs, col = Type, fill = Matrix_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  geom_errorbar(aes(ymin = Std_CI_low, ymax = Std_CI_high), stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  facet_wrap(.~Plate) +
  scale_color_manual(values =c("Standard" = "lightgreen", "Matrix Check" = "cornflowerblue" )) +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Matrix Effects Checks with 95% CI: H2S") +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1))


H2S_mtx_plot

```


##Display Any Matrix Flags
```{r}
Mtx_Flag_1 <- Mtx_Flags %>% dplyr::select(Plate, MTXs_Flag) %>% filter(MTXs_Flag != "Good")

MTX_CV_Flag_1 <- Mtx_Flags %>% dplyr::select(Plate, MTX_CV_Flags) %>% filter(MTX_CV_Flags != "Good")

ifelse(length(Mtx_Flag_1$MTXs_Flag) >= 1, paste0(Mtx_Flag_1$MTXs_Flag, "-", Mtx_Flag_1$Plate), paste0("All Matrix Checks Good"))

ifelse(length(MTX_CV_Flag_1$MTX_CV_Flags) >= 1, paste0(MTX_CV_Flag_1$MTX_CV_Flags, "-", MTX_CV_Flag_1$Plate), paste0("All Matrix CV Good"))

```


## Calculate Sulfide Concentrations & Add Concentration Flags
```{r Calculate Sulfide Concentrations,echo=FALSE,include=FALSE}

High_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 5") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(High_DL = Mean_Abs)

Low_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 1") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(Low_DL = Mean_Abs)

#filter out stds and other experiments
raw_data <- dat %>%  
      filter(!str_detect(IDs, "Std")) %>%  
      filter(!str_detect(IDs, "MC")) %>%  
      filter(str_detect(IDs, "L") | str_detect(IDs, "l"))

Slopes_calc <- Slope_H2S_Std_Crvs_rm %>% dplyr::select(Plate, Slope, Intercept)

joined_data <- right_join(Slopes_calc, raw_data, by = join_by(Plate))

calc_data <- joined_data %>% mutate(Conc = ((Slope*Abs) + Intercept)*Dilution) %>% mutate(Conc = ifelse(Conc <= 0, 0, Conc))

calc_data_1 <- right_join(High_Std, calc_data, by = join_by(Plate))

dl_data <- right_join(Low_Std, calc_data_1, by = join_by(Plate))

data_dl_flags <- dl_data %>% mutate(Conc_Flag = ifelse(Abs < Low_DL, "bdl", ifelse(Abs > High_DL, "adl", "Within_Range")))
```

## Remove High CV Samples and Average
```{r}

#Calculate CVs
H2S_SPL_cv <- data_dl_flags  %>% group_by(Plate, IDs) %>% mutate(Mean_Conc = mean(Conc)) %>% mutate(Conc_SD = sd(Conc)) %>% mutate(CV = (Conc_SD / Mean_Conc) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_SPL_cv_cor <- H2S_SPL_cv %>% mutate(Mean_diff = abs(Conc - Mean_Conc)) %>% group_by(Plate, IDs) %>% mutate(Conc_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Conc), Conc))

H2S_SPL_cv_new <- H2S_SPL_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Conc2 = mean(Conc_rm, na.rm = TRUE)) %>% mutate(Conc_SD2 = sd(Conc_rm, na.rm = TRUE)) %>% mutate(CV2 = (Conc_SD2 /Mean_Conc2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Conc_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_SPL_CV_fin <- H2S_SPL_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Conc,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Conc2,
                                                   Conc_SD2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Conc = Mean_Conc2, CV_Flag = CV_Flag2, SD_Conc = Conc_SD2) %>% 
                                  dplyr::select(Plate,                                                                                                      Conc,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Conc,
                                                SD_Conc,
                                                CV_Flag) 

```

## Plot Samples

```{r}


H2S_Spl_pre_viz <- H2S_SPL_cv_cor %>%
  filter(!str_detect(IDs, "Dup")) %>%
  filter(!str_detect(IDs, "Spike")) %>%
  dplyr::select(Plate, IDs, Mean_Conc, Conc_SD, CV_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(IDs = ifelse(Plate %like% "rr", paste0(IDs, "_RR"), IDs))

H2S_Spl_Plot_Pre <- ggplot(H2S_Spl_pre_viz, aes(y = Mean_Conc, x = IDs, fill = CV_Flag)) +
  geom_bar(stat = 'identity',
           position = position_dodge2(),
           linewidth = 0.75) +
  scale_fill_manual(values = c(
    "CV_Good" = "darkgreen",
    "CV_High" = "darkred"
  )) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  geom_errorbar(
    aes(ymin = Mean_Conc - Conc_SD, ymax = Mean_Conc + Conc_SD),
    width = 0.3,
    position = position_dodge(.1)
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ggtitle("Sample triplicate means and sd dev before bad reps removed")


H2S_Spl_Plot_Pre

H2S_Spl_rmv_viz <- H2S_SPL_CV_fin %>%
  filter(!str_detect(IDs, "Dup")) %>%
  filter(!str_detect(IDs, "Spike")) %>%
  dplyr::select(Plate, IDs, Mean_Conc, SD_Conc, CV_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(IDs = ifelse(Plate %like% "rr", paste0(IDs, "_RR"), IDs))

H2S_Spl_Plot_CV <- ggplot(H2S_Spl_rmv_viz, aes(y = Mean_Conc, x = IDs, fill = CV_Flag)) +
  geom_bar(stat = 'identity',
           position = position_dodge(),
           linewidth = 0.75) +
  scale_fill_manual(values = c(
    "CV_Good" = "darkgreen",
    "CV_High" = "darkred"
  )) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  geom_errorbar(
    aes(ymin = Mean_Conc - SD_Conc, ymax = Mean_Conc + SD_Conc),
    width = 0.3,
    position = position_dodge(.1)
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ggtitle("Sample triplicate means and sd dev after bad reps removed")


H2S_Spl_Plot_CV



```


## Check the dups for QAQC

```{r,echo=FALSE}

#pull out any rows that have "Dup" in the ID column
DUP_start <- H2S_SPL_CV_fin %>%  mutate(Dup_ID = gsub(" ", "", substr(IDs, 1, 6)),
                                        Dup = ifelse(IDs %like% "Dup", "Dup", "Org")) %>% mutate(Dup_ID = ifelse(Plate %like% "rr", paste0(Dup_ID, "_RR"), Dup_ID))

dups_only <- DUP_start %>% filter(Dup == "Dup") %>% dplyr::select(Dup_ID, Mean_Conc, SD_Conc) %>% rename(Dup_Conc = Mean_Conc, Dup_SD = SD_Conc) %>% distinct(.keep_all = TRUE)

no_dups <- DUP_start %>% filter(!Dup == "Dup")

dup_spls_l <- dups_only %>% dplyr::select(Dup_ID) %>% distinct(.keep_all = TRUE)

dup_spls <- no_dups %>% filter(Dup_ID %in% as.vector(dup_spls_l$Dup_ID)) %>% dplyr::select(Dup_ID, Mean_Conc, SD_Conc, Plate, IDs) %>% distinct(.keep_all = TRUE)

dup_join <- full_join(dups_only, dup_spls, by = join_by(Dup_ID))

dup_flags <- dup_join %>% mutate(Diff_P = (abs(100 * (Mean_Conc - Dup_Conc) /
                                                 Mean_Conc))) %>% mutate(Dup_Flag = ifelse(Diff_P >= dups_perc_diff, "Dup_Bad", "Dup_Good"))

dup_flags_1 <- dup_flags %>% dplyr::select(Dup_ID, Dup_Flag)

Dup_for_plot_1 <- DUP_start %>% dplyr::select(Plate, Mean_Conc, Dup_ID, IDs) %>% distinct(.keep_all = TRUE) %>% filter(Dup_ID %in% as.vector(dup_spls_l$Dup_ID)) %>% left_join(dup_flags_1) %>% mutate(Dup = ifelse(IDs %like% "Dup", "Duplicate", "Original"))

Dup_Plot <- ggplot(Dup_for_plot_1,
                   aes(
                     x = Dup_ID,
                     y = Mean_Conc,
                     fill = Dup_Flag,
                     col = Dup
                   )) +
  geom_bar(
    stat = 'identity',
    position = position_dodge2(preserve = "single"),
    linewidth = 0.9
  ) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  scale_fill_manual(values = c(
    "Dup_Good" = "darkgreen",
    "Dup_Bad" = "darkred"
  )) +
  scale_color_manual(values = c(
    "Duplicate" = "goldenrod1",
    "Original" = "cornflowerblue"
  )) +
  theme_classic()+
  ggtitle("H2S Duplicates") 

Dup_Plot


```

## Return Dup Flags

```{r}
Dup_Flag_1 <- dup_flags %>% group_by(Plate)  %>% mutate(Dup_Flags = ifelse("Dup_Bad" %in% Dup_Flag, "Dup_Bad", "Dup_Good")) %>% dplyr::select(Plate, Dup_Flags) %>% rename(Dup_Flag = Dup_Flags) %>% distinct(.keep_all = TRUE)

Dup_Flags_1 <-Dup_Flag_1 %>% mutate(Dup_Flags = paste0(Dup_Flag, "-", Plate))

paste0(Dup_Flags_1$Dup_Flags)


```

## Check the spks for QAQC (THIS SECTION NEEDS WORK)
```{r}
#pull out any rows that have "Spk" in the ID column
SPK_start <- H2S_SPL_CV_fin %>%  mutate(Spk_ID = gsub(" ", "", substr(IDs, 1,6)), Spk = ifelse(IDs %like% "Spike", "Spk", "Org")) %>% mutate(Spk_ID = ifelse(Plate %like% "rr", paste0(Spk_ID, "_RR"), Spk_ID))

spks_only <- SPK_start %>% filter(Spk == "Spk") %>% dplyr::select(Spk_ID, Mean_Conc, SD_Conc) %>% rename(Spk_Conc = Mean_Conc, Spk_SD = SD_Conc) %>% distinct(.keep_all = TRUE)

no_spks <- SPK_start %>% filter(!Spk == "Spk")

spls_l <- no_spks %>% dplyr::select(Spk_ID)  %>% distinct(.keep_all = TRUE)

spk_spls_l <- spks_only %>% dplyr::select(Spk_ID) %>% distinct(.keep_all = TRUE)

spk_spls <- no_spks %>% filter(Spk_ID %in% as.vector(spk_spls_l$Spk_ID)) %>% dplyr::select(Spk_ID, Mean_Conc, SD_Conc, Plate, IDs) %>% distinct(.keep_all = TRUE)

SPK_only <- spks_only %>% filter(Spk_ID %in% as.vector(spls_l$Spk_ID))

spk_join <- full_join(SPK_only, spk_spls, by = join_by(Spk_ID))

spk_flags <- spk_join %>% mutate(Diff_P = (abs(100*(Spk_Conc/Mean_Conc)))) %>% mutate(Spk_Flag = ifelse(Diff_P >= high_recovery_cutoff | Diff_P <= low_recovery_cutoff, "Spk_Bad", "Spk_Good"))

spk_flags_1 <- spk_flags %>% dplyr::select(Spk_ID, Spk_Flag)

Spk_for_plot_1 <- SPK_start %>% dplyr::select(Plate, Mean_Conc, Spk_ID, IDs) %>% distinct(.keep_all = TRUE) %>% filter(Spk_ID %in% as.vector(spk_spls_l$Spk_ID)) %>% left_join(spk_flags_1, by = join_by(Spk_ID)) %>% mutate(Spk = ifelse(IDs %like% "Spike", "Spike", "Original")) %>% filter(Spk_ID %in% as.vector(spls_l$Spk_ID))

Spk_Plot <- ggplot(Spk_for_plot_1,
                   aes(
                     x = Spk_ID,
                     y = Mean_Conc,
                     fill = Spk_Flag,
                     col = Spk
                   )) +
  geom_bar(
    stat = 'identity',
    position = position_dodge2(preserve = "single"),
    linewidth = 0.9
  ) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  scale_fill_manual(values = c(
    "Spk_Good" = "darkgreen",
    "Spk_Bad" = "darkred"
  )) +
  scale_color_manual(values = c(
    "Spike" = "goldenrod1",
    "Original" = "cornflowerblue"
  )) +
  theme_classic()+
  ggtitle("H2S Spikes") 

Spk_Plot


```

##Display Spike Flags
```{r}
ifelse(all(as.vector(spk_spls_l$Spk_ID) %in% as.vector(spls_l$Spk_ID)), paste0("Spikes Paired"), paste0("Spike Missing Pair"))

Spk_Flag_1 <- spk_flags %>% group_by(Plate)  %>% mutate(Spk_Flags = ifelse("Spk_Bad" %in% Spk_Flag, "Spk_Bad", "Spk_Good")) %>% dplyr::select(Plate, Spk_Flags) %>% rename(Spk_Flag = Spk_Flags) %>% distinct(.keep_all = TRUE)

Spk_Flags_1 <-Spk_Flag_1 %>% mutate(Spk_Flags = paste0(Spk_Flag, "-", Plate))

paste0(Spk_Flags_1$Spk_Flags)

bad <- Spk_Flags_1 %>% filter(Spk_Flags %like% "Bad")

length_bad <- length(as.vector(bad$Spk_Flags))

length_all <-  length(as.vector(Spk_Flags_1$Spk_Flags))

ifelse(length_bad/length_all > 0.4, paste0("<60% of Spikes Pass, Rerun"), paste0("Spikes Pass") )

```


# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")



H2S_spls_md <- H2S_SPL_CV_fin %>%
  filter(!str_detect(IDs, "Dup")) %>%
  filter(!str_detect(IDs, "Spike"))

H2S_spl_rrn <- H2S_spls_md %>%
  filter(Plate %like% "rr") %>% 
  dplyr::select(IDs, Mean_Conc) %>%
  rename(RRN_Conc = Mean_Conc) %>% 
  distinct(.keep_all = TRUE) 

RRN_IDs <- as.vector(H2S_spl_rrn$IDs)

H2S_spl_og <- H2S_spls_md %>%
  filter(!Plate %like% "rr") %>% 
  dplyr::select(IDs, Mean_Conc, Plate) %>% 
  distinct(.keep_all = TRUE) %>%
  mutate(Rerun = ifelse(IDs %in% RRN_IDs, "RRN", "Org"))

H2S_spl_rr_join <- full_join(H2S_spl_rrn,H2S_spl_og, by = join_by(IDs) ) 

Samples <-  H2S_spl_rr_join %>% mutate(Conc = ifelse(Rerun == "RRN", RRN_Conc, Mean_Conc)) %>% dplyr::select(IDs, Conc, Plate) %>% rename(Sample_ID = IDs)



#check to see if all samples are present in the metadata
all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}


#merge metadata with sample run data
merged_data <- metadata %>%
  mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
  fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included?

```


#Add Flags
```{r}
Chk_Flags_All <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags) %>% group_by(Plate) %>% mutate(CHKSTD_Flags = paste0(ifelse(CHK_CV_Flags %like% "Good", "", CHK_CV_Flags))) %>% dplyr::select(Plate, CHKSTD_Flags) %>% mutate(CHKSTD_Flags = ifelse(CHKSTD_Flags == "", NA, CHKSTD_Flags)) %>% na.omit()

Mtx_Flags_All <- Mtx_Flags %>% dplyr::select(Plate, Mtx_Flag, CV_Flag) %>% rename(MTXCV_Flag = CV_Flag) %>% mutate(Mtx_Flag = ifelse(Mtx_Flag %like% "Good", NA ,Mtx_Flag)) %>% mutate(Mtx_Flag = ifelse(MTXCV_Flag %like% "Good", NA ,MTXCV_Flag)) %>% na.omit()

STD_Flags_All <- STD_CRV_flags %>% dplyr::select(Plate, R2_Flag, CV_Flag) %>% mutate(STDCV_Flags = paste0("STD", CV_Flag)) %>% dplyr::select(-CV_Flag) %>% mutate(R2_Flag = ifelse(R2_Flag %like% "Good", NA, R2_Flag), STDCV_Flags = ifelse(STDCV_Flags %like% "Good", NA, STDCV_Flags)) %>% na.omit()

Spk_Flags_All <- Spk_Flag_1  %>% group_by(Plate) %>% mutate(SPK_Flags = paste0(ifelse(Spk_Flag %like% "Good", NA, Spk_Flag))) %>% dplyr::select(Plate, SPK_Flags) %>% na.omit()

Dup_Flags_All <-  Dup_Flag_1  %>% group_by(Plate) %>% mutate(DUP_Flags = paste0(ifelse(Dup_Flag %like% "Good", NA, Dup_Flag))) %>% dplyr::select(Plate, DUP_Flags) %>% na.omit()

Flags_Joined_1 <- left_join(Spk_Flags_All, Dup_Flags_All, by = join_by(Plate))

Flags_Joined_2 <- left_join(Flags_Joined_1, Chk_Flags_All, by = join_by(Plate))
  
Flags_Joined_3 <- left_join(Flags_Joined_2, Mtx_Flags_All, by = join_by(Plate))
                              
Flags_Joined <- left_join(Flags_Joined_3, STD_Flags_All, by = join_by(Plate))

Flags_Joined[Flags_Joined =="NA"] <- NA

Flags <- Flags_Joined %>% replace(is.na(.), "") %>% mutate(H2S_Flags = paste0(SPK_Flags, " ", DUP_Flags, " ",CHKSTD_Flags, " ", Mtx_Flag, " ", MTXCV_Flag," ", R2_Flag," ", STDCV_Flags )) %>% dplyr::select(Plate, H2S_Flags)

flagged_data <- left_join(merged_data, Flags, by = join_by(Plate))


df_all_clean <- flagged_data %>% dplyr::select(-Plate)

```

```{r,echo=FALSE}

#pull out any rows that have "d" in the SampleID column
spks <- dat1 %>%  filter(str_detect(IDs, "Spike")) 

#remove the Spike from these IDs so we will have duplicates 
spks$IDs<-gsub(" Spike","",spks$IDs)
#head(spks)
colnames(spks) <- c('IDs', 'mean_spk')


#put it back together with the old data set and look for duplicates 
QAspks <- merge(dat2, spks)
#head(QAspks)

#now we need to calculate the spike concentration and calculate the spike recovery 
#spike for these samples was 10 uL of the 100uM standard
QAspks$unspkd <- (QAspks$H2S_mean/QAspks$Dilution)*((sample_vol )/Con1) #gives us the total S2- in the sample in umoles
QAspks$spkd <-    (QAspks$mean_spk/QAspks$Dilution)*((sample_vol + spkvol)/Con1)        ##total S2- in spiked sample in umoles 
QAspks$expctd_spkd <-  (QAspks$unspkd + spk_Conc)
QAspks$spk_recovery <-    (QAspks$spkd/QAspks$expctd_spkd)*100
QAspks$spks_flag <-  ifelse(QAspks$spk_recovery > low_recovery_cutoff & QAspks$spk_recovery < high_recovery_cutoff, 'OK', 'NO, rerun')  
 

#head(QAspks)

#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 
spksbar <- ggplot(data = QAspks, aes(x = IDs, y = spk_recovery, fill=spks_flag)) +facet_wrap(~Plate)+
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK"="darkgreen","NO, rerun"="darkred")) + 
        theme_classic() + labs(x= "Sample ID", y="Spike Recovery (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=low_recovery_cutoff, linetype="dashed", color = "black", linewidth=1) + 
        geom_hline(yintercept=high_recovery_cutoff, linetype="dashed", color = "black", linewidth=1)

spksbar

#check for any no's that would warrant reruns! 

Badspks <- subset(QAspks, spks_flag == "NO, rerun")

#write out a flag to the sample dataframe if any spks have recovery out of range 
for(i in 1:length(plates)){
  Badspks1<-  subset(Badspks,Plate %in% plates[i])
  dat$QAQC_flag<-if(nrow(Badspks1) > 0){
    ifelse(dat$Plate %in% qplate$Plate[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Spk recovery out of range"),"Spk recovery out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
}}

for(i in 1:length(plates)){
  Badspks1<-  subset(Badspks,Plate %in% plates[i])
  dat1$QAQC_flag<-if(nrow(Badspks1) > 0){
    ifelse(dat1$Plate %in% qplate$Plate[i], ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, "; Spk recovery out of range"),"Spk recovery out of range"),paste0( dat1$QAQC_flag,"") )                   
  }else{
    ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, ""),"")
}}

spks_percent <- (sum(QAspks$spks_flag == "OK")/nrow(QAspks))*100
#report out if flags indicate need for rerun
ifelse(spks_percent >= 60  , ">60% of Spikes are within range",
       "<60% of Spikes are out of range - REASSESS")
```

## Export full data then just final data 
```{r,echo=FALSE,include=FALSE}
dat_IDs <- dat1

#Read out the summarized data
head(dat_IDs)
write.csv(dat_IDs, file= summarizeddata_path)

#Read out all the data in dat 
head(dat)
write.csv(dat, file= fulldata_path)

#Now take out the absorbance and stuff 
head(dat)
dat3 <- dat[ ,c(1,4,6,8)]
head(dat3)
write.csv(dat3, file= shortdata_path)

write.csv(H2S_DiluteRerun, Need_dilute)
write.csv(H2S_HighCVRerun, high_CV)

#save to QAQC Data to QAQC File
#write.csv(qlog, file= log_path)


```

### END