---
title: "CO2xN: Porewater Sulfide"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
    number_sections: false
geometry: left=2cm,right=2cm,top=1cm,bottom=2cm
---
## Code Set up 
```{r setup, include=FALSE}

library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
library(purrr)
library(tidyverse)
library(here)
library(data.table)
library(matrixStats)
library(gridExtra)
library(grid)
library(plater)
library(raster)
library(knitr)  
library(readxl)
library(ggrepel)

```

## run information
```{r Information to be changed}

##Before or After Reruns (YES or NO)
After_Reruns = "YES"
#If NO will not write final data file
  #And Will Write a Rerun File

###things that need to be changed

sample_month = "May"
sample_year = "2025"
Run_Date = as.Date("05/27/2025", format = "%m/%d/%Y")
Run_by = "Melanie Giessner"#Instrument user
Script_run_by ="Brenden Blakley"#Code user

Run_notes=""#any notes from run

#Import Plate Data

plates<- c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7", "Plate8", "Plate9", "Plate10", "Plate_rr")

Raw_plates <- c("Raw Data/20250714_LTREB_H2S_Plate1.csv",
                "Raw Data/20250714_LTREB_H2S_Plate2.csv",
                "Raw Data/20250714_LTREB_H2S_Plate3.csv",
                "Raw Data/20250714_LTREB_H2S_Plate4.csv",
                "Raw Data/20250714_LTREB_H2S_Plate5.csv",
                "Raw Data/20250714_LTREB_H2S_Plate6.csv",
                "Raw Data/20250715_LTREB_H2S_Plate7.csv",
                "Raw Data/20250715_LTREB_H2S_Plate8.csv",
                "Raw Data/20250715_LTREB_H2S_Plate9.csv",
                "Raw Data/20250715_LTREB_H2S_Plate10.csv", 
                "Raw Data/20250716_LTREB_H2S_reruns.csv")

#Import Run Date for Each Plate for Std curve qaqc
Run_dates <- tibble(
    Plate = c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate8","Plate9","Plate10", "Plate_rr"),
    Run_Date = c("07/14/2025","07/14/2025","07/14/2025","07/14/2025","07/14/2025","07/14/2025","07/15/2025","07/15/2025", "07/15/2025", "07/15/2025", "07/16/2025")
  )

  Run_dates <- Run_dates %>% mutate(Run_Date = as.Date(Run_Date, format = "%m/%d/%Y"))


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Processed Data/LTREB_2M_H2S_QAQC_2025.csv"

# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_CO2xN_Porewater_H2S_202507.csv"
rerun_path <- "Processed Data/GCReW_CO2xN_Porewater_H2S_202507_reruns.csv"


```

```{r Set Up Code - constants and QAQC cutoffs, include=FALSE}

#Flag cutoffs
  r2_cutoff = 0.985            #this is the level below which we want to rerun or consider a curve 
  cv_flag_stds = 10           #this is the maximum cv allowed for standards
  p_value_chkstds = 0.05      #the p-value for the t-test between check standards must be greater than this
  p_value_MC = 0.05           #the p-value for the t-test between matrix checks and top std must be greater than this
  cv_flag_sample = 10         #this is the maximum cv allowed for samples
  dups_perc_diff = 15.5       #this is the maximum percent difference allowed between duplicates
  high_recovery_cutoff = 120  #this is the maximum percent recovery of SO4 allowed in spiked samples
  low_recovery_cutoff = 80    #this is the minimum percent recovery of SO4 allowed in spiked samples

#Standard concentrations - Update if running different standard curve:
# standard units are in uM
  standards <- tibble(
    IDs = c("Std 0", "Std 1", "Std 2", "Std 3", "Std 4", "Std 5"),
    Conc = c(0, 5.0, 12.5, 25.0, 50.0, 100.0)
  )  #uM
  
    chk_standards <- tibble(
    IDs = c("ChkStd 0", "ChkStd 3", "ChkStd 4"),
    Conc = c(0, 25.0, 50.0)
  )  #uM
    
    mtx_standards <- tibble(
    IDs = c("MC: 10ppt S5"),
    Conc = c(100)
  )  #uM

#Spike concentration calc
  #spike for these samples was 50 uL of the 100uM standard
  Con1 <- 1000000               #conversion factor value for spike volumes (uL -> L)
  spk_std <- 100                # uM S2- standard used
  spkvol <- 50                  # uL volume of spike added
  spk_Conc <- spk_std*(spkvol/Con1)  # umoles of S2- added to each spiked sample
  sample_vol <- 250             # the sample volume without the spike is 250 uL
  samp_vol_a <- 300             # the sample volume 

#Top standard Concentration- Update if running different standard curve: 
   Top_STD = 100
   
   
#Metadata file path
Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

#Significant Digits
sigdigs <- 3

```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata

raw_metadata <- read.csv(Raw_Metadata)
#clean up columns in metadata file:
metadata <- raw_metadata %>%
  filter(grepl("2-Meter", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  ) %>%
  dplyr::select(-Temp_Treatment) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2_N",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2",
      ifelse(CO2_Treatment == "Ambient" &
               N_Treatment == "Amended"  , "N" , "Ambient")
    )
  )) 

```


##Read in data
```{r read in plates,echo=FALSE,include=FALSE}

#read in the csv files
dat<-read_plates(
  files = Raw_plates,#list of all file paths
  plate_names = plates ,  #list of plate names          
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
) %>%
  rename("Abs"=values,
         "IDs"=values.2,
         "Dilution"=values.3)
head(dat)
  
```

##Remove High CV Stds
```{r, echo=FALSE}

#filter for STDs
H2S_stds_raw <- dat %>% subset(IDs %like% "Std" | IDs %like% "Chk" | IDs %like% "MC")

#Add in STD values
H2S_stds <- left_join(H2S_stds_raw, standards, by = "IDs")

#Calculate CVs and add flags
H2S_STD_cv <- H2S_stds %>% group_by(Plate, IDs) %>% mutate(CV = cv(Abs)) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

#calculate new cv getting best cv from points if cv was high
H2S_STD_cv_new <- H2S_STD_cv %>% group_by(Plate, IDs) %>%
  mutate(CV_New = map_dbl(row_number(), ~cv(Abs[-.x]))) %>%    #Calculate CVs without 1 point
  mutate(Removed = ifelse(             #Choose Lowest CV if CV is high
  CV_Flag == "CV_High",
  ifelse(
    CV_New == min(CV_New),
   "RMVD",
    "KEPT"), "CV_Good"),
CV2 = ifelse(CV_Flag == "CV_High", min(CV_New), CV),
Abs_rm = ifelse(Removed == "RMVD", NA, Abs)) %>%
  mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>%
  mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

#organize data
H2S_STD_CV_fin <- H2S_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2, CV = CV2) %>% 
                                  dplyr::select(Plate, 
                                                Wells,                                                                                    Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Conc,
                                                Mean_Abs,
                                                CV,
                                                CV_Flag)


```


##Check R2 & Remove Standards if Necessary
```{r, echo=FALSE}
split_data <- H2S_STD_CV_fin %>% filter(IDs %like% "Std") %>% filter(!IDs %like% "Chk") %>% dplyr::select(Plate, IDs, Conc, Mean_Abs, CV_Flag) %>% #Split Data by Plate
      distinct() %>%
      group_by(Plate) %>%
      group_split()
  
#run regression and get coefficients by Run
Slope_H2S_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Mean_Abs ~ Conc, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Mean_Abs ~ Conc, data = .))[2],
                                        Intercept = coef(lm(Mean_Abs ~ Conc, data = .))[1]))


 Slope_H2S_raw <- bind_rows(Slope_H2S_raw_list)
  
 #check R2 and flag
 Slope_H2S_Std_Crvs <- Slope_H2S_raw %>% dplyr::select(Plate, Curve, R2, Slope, Intercept) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))

 #join standard curve CVs and regression values
 R2_Check_Raw <- left_join(Slope_H2S_Std_Crvs, H2S_STD_CV_fin, by = "Plate")
 
 R2_Check <- R2_Check_Raw %>% 
   dplyr::select(Plate,
                 IDs,
                 Mean_Abs,
                 Conc,
                 R2,
                 Slope,
                 Intercept,
                 CV_Flag,
                 CV,
                 R2_Flag) %>% filter(IDs %like% "Std") %>% filter(!IDs %like% "Chk") %>%
   distinct()
 
 #calculate concentrations from standard curve and separate removable points
 
 R2_Split <- R2_Check %>% mutate(RMVBL = ifelse(IDs %in% c("Std 0", "Std 1", "Std 5"), "No", "Yes")) %>% group_by(Plate) %>% group_split()
 
 R2_New <- map(
   R2_Split,
   ~ .x %>% mutate(
     Curve = "H2S",
     R2_New = map_dbl(row_number(), ~
                        summary(lm(
                          Mean_Abs[-.x] ~ Conc[-.x], data = pick(Mean_Abs, Conc)
                        ))$adj.r.squared) ,
     Slope_New = map_dbl(row_number(), ~
                           coef(lm(
                             Mean_Abs[-.x] ~ Conc[-.x], data = pick(Mean_Abs, Conc)
                           ))[2]) ,
     Intercept_New = map_dbl(row_number(), ~
                               coef(lm(
                                 Mean_Abs[-.x] ~ Conc[-.x], data = pick(Mean_Abs, Conc)
                               ))[1])
   )
 ) %>% bind_rows()
 
 R2_Check_RMVBL <- R2_New %>% mutate(R2_New = round(R2_New, digits = 8),
                                     CV = round(CV, digits = 8)) %>% group_by(Plate) %>% mutate(R2_New = ifelse(IDs %in% c("Std 0", "Std 1", "Std 5"), NA, R2_New)) %>% mutate(R2_RMV = ifelse(
                                       R2_Flag == "R2_Bad",
                                       ifelse(R2_New == max(R2_New, na.rm = TRUE), "RMVD", "KEPT"),
                                       "R2_Good"
                                     )) %>% mutate(R2_RMV_CV = ifelse(
                                       R2_RMV == "R2_Good",
                                       ifelse(CV_Flag == "CV_High", ifelse(CV == max(CV), "RMVD", "KEPT"), "KEPT"),
                                       R2_RMV
                                     )) %>% mutate(R2_RMVD = ifelse(is.na(R2_RMV_CV), "KEPT", R2_RMV_CV)) %>% dplyr::select(-R2_RMV, -R2_RMV_CV)
 
 
   
  R2_Check_fin <- R2_Check_RMVBL %>% mutate(STD_ABS_rmv = ifelse(R2_RMVD == "RMVD", NA, Mean_Abs)) %>% group_by(Plate) %>% mutate(Points = ifelse("RMVD" %in% R2_RMVD, "Point_Removed", "ALL_KEPT")) %>% dplyr::select(Plate, IDs, Conc, STD_ABS_rmv, CV_Flag, R2_Flag, Points) %>% ungroup() %>% na.omit()
 
 #separate plates for new regression
 split_data_rmv <- R2_Check_fin %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
 
 #get new standard curve values
 Slope_H2S_raw_list_rm <- map(split_data_rmv, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(STD_ABS_rmv ~ Conc, data = .))$adj.r.squared , 
                                        Slope = coef(lm(STD_ABS_rmv ~ Conc, data = .))[2],
                                        Intercept = coef(lm(STD_ABS_rmv ~ Conc, data = .))[1]))

 Slope_H2S_raw_rm <- bind_rows(Slope_H2S_raw_list_rm)
  
 Slope_H2S_Std_Crvs_rm <- Slope_H2S_raw_rm %>% dplyr::select(Plate, Curve, R2, Slope, Intercept, CV_Flag, Points) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))
 
 
 
```


## Plot Standard Curves
```{r Plot all Standard Curves together, echo=FALSE}
##this is to see if there are any stds or plates that look off

#Plot Std Crvs by plate
H2S_Curve <- ggplot(Slope_H2S_raw_rm, aes(STD_ABS_rmv, Conc, color = Plate, linetype = Points)) + geom_point(size = 3) +
  geom_smooth(method = "lm", formula = "y ~ x", se = FALSE) + labs(title = "H2S Standard Curves", x = "Absorbance", y = "Concentration (uM)") +
  theme_classic()
H2S_Curve

#Add dates to plate and format data
STD_run_date <- left_join(Run_dates, Slope_H2S_Std_Crvs_rm, by = join_by(Plate))
new_log_stds <- STD_run_date %>% dplyr::select(Plate,
   Curve,
   R2,
   Slope,
   Intercept,
   Run_Date) %>% mutate(R2 = round(R2, digits = 4), Slope = round(Slope, digits = 4), Intercept = round(Intercept, digits = 4), ) %>% mutate(Run = "Current Runs")


#read in log file and format
log <- read.csv(log_path)
log_2 <- log %>% mutate(Run_Date = as.Date(as.character(Run_Date), format = "%Y-%m-%d"), Run = "Past Runs")
log_3 <- log_2[, -c(1)]
log_4 <- rbind(log_3, new_log_stds)

unique_log <- log_4 %>% distinct(.keep_all = TRUE)

#Calculate Mean and SD and calculate 2SD range for slope and intercept
mean_slope <- mean(unique_log$Slope, na.rm = TRUE)
sd_slope <- sd(unique_log$Slope, na.rm = TRUE)

hi_2sd_slope <- mean_slope + 2*sd_slope

lo_2sd_slope <- mean_slope - 2*sd_slope
  
mean_int <- mean(unique_log$Intercept, na.rm = TRUE)
sd_int <- sd(unique_log$Intercept, na.rm = TRUE)

hi_2sd_int <- mean_int + 2*sd_int

lo_2sd_int <- mean_int - 2*sd_int

#Plot Slope log
Slopes_chk <- ggplot(unique_log, aes(x = Run_Date, y = Slope, col = Run)) +
    ylim(0.0045,0.018) +
  scale_color_manual(values = c("Past Runs" = "grey27", "Current Runs" = "cyan3")) +
  geom_abline(intercept = mean_slope - 2*sd_slope, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_abline(intercept = mean_slope + 2*sd_slope, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_point(size = 2) +
  theme_classic()+
  ggtitle("Slope Log") 

#Plot Intercept log
ints_chk <- ggplot(unique_log, aes(x = Run_Date, y = Intercept, col = Run)) +
  ylim(0.03, 0.15) +
  scale_color_manual(values = c("Past Runs" = "grey27", "Current Runs" = "cyan3")) +
  geom_abline(intercept = mean_int - 2*sd_int, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_abline(intercept = mean_int + 2*sd_int, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_point(size = 2) +
  theme_classic()+
  ggtitle("Intercept Log") 

ints_chk

Slopes_chk

write.csv(unique_log, log_path)

#Get done of original runs
Date_1 <-as.Date(new_log_stds$Run_Date[which(new_log_stds$Plate == "Plate1")]) 

#Check if Std Crvs are Within range and see if dates are close enough for combination
stdcrv_tagged <- new_log_stds %>% mutate(
  In_Range_Slope = ifelse(
    Slope >= hi_2sd_slope |
      Slope <= lo_2sd_slope,
    "Out of Range",
    "In Range"
  ),
  In_Range_Int = ifelse(
    Intercept >= hi_2sd_int |
      Intercept <= lo_2sd_int,
    "Out of Range",
    "In Range"
  )
) %>% mutate(
  Slope = ifelse(In_Range_Slope == "Out of Range", NA, Slope),
  Intercept = ifelse(In_Range_Int == "Out of Range", NA, Intercept)
) %>% mutate(Date_Diff = abs(difftime(Date_1, Run_Date, units = "days"))) %>% mutate(Date_Close = ifelse(Date_Diff <= 2, "Yes", "No")) %>% mutate(Slope_Flag = ifelse(is.na(Slope), "Slope_Bad" , "Slope_Good"), Int_Flag = ifelse(is.na(Intercept), "Intercept_Bad" , "Intercept_Good")) %>% mutate(
  Slope = ifelse(is.na(Intercept), NA, Slope),
  Intercept = ifelse(is.na(Slope), NA, Intercept),
  Curve_RMVD = ifelse(is.na(Slope) |
                        is.na(Intercept), "RMVD", "KEPT")
)

#Calculate Combo Std Crv
stdcrv_filtered <- stdcrv_tagged %>% filter(Date_Close == "Yes")

mean_slope_fin <- mean(stdcrv_filtered$Slope, na.rm = TRUE)
  
mean_int_fin <- mean(stdcrv_filtered$Intercept, na.rm = TRUE)

#Attach Slope and Intercept to Plates
Std_crv_calc <- stdcrv_tagged %>% mutate(Slope_Calc = ifelse(Date_Close == "Yes", mean_slope_fin, Slope), Int_Calc = ifelse(Date_Close == "Yes", mean_int_fin, Intercept)) %>% dplyr::select(Plate, Slope_Calc, Int_Calc, Curve_RMVD, Slope_Flag, Int_Flag) %>% as.data.frame()

Curve_Flags <- Std_crv_calc %>% dplyr::select(Plate, Slope_Flag, Int_Flag) %>% distinct()

Curve_Flags_forJoin <- Curve_Flags %>% mutate(Slope_Flag = ifelse(Slope_Flag %like% "Good", "", Slope_Flag), Int_Flag = ifelse(Int_Flag %like% "Good", "", Int_Flag)) 
```

## Return Flags For Standard Curve
```{r, echo=FALSE}

STD_CRV_flags <- full_join(Slope_H2S_Std_Crvs_rm, Run_dates, by = join_by(Plate))

ifelse(STD_CRV_flags$CV_Flag == "CV_High", paste0("CV High", "-", STD_CRV_flags$Plate), paste0("CV Good", "-", STD_CRV_flags$Plate))

ifelse(STD_CRV_flags$R2_Flag == "R2_Bad", paste0("R2 Bad", "-", STD_CRV_flags$Plate), paste0("R2 Good", "-", STD_CRV_flags$Plate))

ifelse(Std_crv_calc$Curve_RMVD == "RMVD", paste0("Curve_Removed-", Std_crv_calc$Plate), paste0("Curve_Kept-", Std_crv_calc$Plate))


```

## Method Detection Limit
```{r, echo=FALSE}

#get only lowest standard
 STD_MDL_df <- H2S_STD_CV_fin %>% filter(IDs == "Std 1"
                                         #, Removed != "RMVD"
                                         ) %>% dplyr::select(Plate, Mean_Abs, IDs) %>% distinct()

 STD0_df <- H2S_STD_CV_fin %>% filter(IDs == "Std 0"
                                         #, Removed != "RMVD"
                                         ) %>% dplyr::select(Plate, Mean_Abs, IDs) %>% distinct()
#calculate the standard deviation of abs
stddev_mdl <- sd(STD_MDL_df$Mean_Abs)

#Calculate Mean Std 0
std0_mean <- mean(STD0_df$Mean_Abs)

#calculate n
n <- length(STD_MDL_df$Mean_Abs)

#Identify the proper Student's T for n-1 
StudentsT <-qt(.99,(n-1))

#Multiply Standard deviation by Student's T
STD_var <- (stddev_mdl*StudentsT)
MDL <- STD_var + std0_mean 

MDL_df <- as.data.frame(MDL)

MDL_Plot <- ggplot() +
  geom_bar(stat = 'identity', data = STD_MDL_df, aes(x = Plate, y = Mean_Abs, fill = IDs)) +
  geom_bar(data = STD0_df, aes(x = Plate, y = Mean_Abs, fill = IDs), stat = 'identity') +
  scale_fill_manual(name = "Standard:", values = c("Std 0" = "lightgreen", "Std 1" = "cornflowerblue")) +
  geom_hline(yintercept  = MDL, linetype = "dashed", linewidth = 1.5, color = "blue3") +
  geom_label_repel(data = MDL_df, aes(x = 1.25, y = MDL + 0.015), label = paste0("MDL: ",     signif(MDL, digits = sigdigs)),  alpha = 0.7, color = "blue3") +
  ggtitle(label = "H2S Minimum Detection Limit") +
  labs(y = "Mean Absorbance") +
  theme_classic()

MDL_Plot
```

## Compare Check Standards to Standards
```{r, echo = FALSE}

H2S_CHK_STD_CV_fin <- H2S_STD_CV_fin %>% filter(IDs %like% "ChkStd" | IDs %like% "ppt")
#Tag Check Standards and perform T Test
H2S_CHKs_a <- H2S_CHK_STD_CV_fin %>%
  mutate(Std_ID = ifelse(
    IDs == "ChkStd 0",
    paste0("Std 0", Plate),
    ifelse(
      IDs == "ChkStd 3",
      paste0("Std 3", Plate),
      ifelse(IDs == "ChkStd 4", paste0("Std 4", Plate), ifelse(IDs %like% "S5", paste0("Std 5", Plate), "ID Not Found"))
    )
  )) %>%
  mutate(
    Chk_Abs = Abs,
    Chk_Removed = Removed,
    Chk_IDs = IDs,
    Chk_Mean_Abs = Mean_Abs,
    Chk_CV_Flag = CV_Flag
  ) %>% dplyr::select(Std_ID,
                      Chk_Abs,
                      Chk_Removed,
                      Chk_IDs,
                      Chk_Mean_Abs,
                      Chk_CV_Flag,
                      Plate) %>% filter(Chk_Removed != "RMVD") %>%
  group_by(Std_ID) %>%
  mutate(
    Chk_CI_low = ifelse(length(unique(Chk_Abs)) == 1, Chk_Abs, t.test(Chk_Abs)$conf.int[1]),
    Chk_CI_high = ifelse(length(unique(Chk_Abs)) == 1, Chk_Abs, t.test(Chk_Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Std_ID, Chk_IDs, Chk_CI_low, Chk_CI_high, Chk_CV_Flag, Chk_Mean_Abs, Plate) %>% distinct(.keep_all = TRUE)

H2S_CHKs_1 <- H2S_CHKs_a %>% dplyr::select(-Plate, -Chk_Mean_Abs)

#Get Standard Values and set values for standards with no variance
H2S_STD_forCHK_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 0", "Std 3", "Std 4", "Std 5")) %>% mutate(Std_ID = ifelse(
  IDs == "Std 0",
  paste0("Std 0", Plate),
  ifelse(
    IDs == "Std 3",
    paste0("Std 3", Plate),
    ifelse(IDs == "Std 4", paste0("Std 4", Plate), ifelse(IDs == "Std 5", paste0("Std 5", Plate), "ID Not Found"))
  )
)) %>% filter(Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Std_CI_low = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[1]),
    Std_CI_high = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Plate, Std_ID, IDs, Std_CI_low, Std_CI_high, CV_Flag, Mean_Abs) %>% distinct(.keep_all = TRUE)

H2S_STD_forCHK <- H2S_STD_forCHK_a %>% dplyr::select(-Mean_Abs)

#join check standards with standards
Chk_Std_join <- full_join(H2S_CHKs_1, H2S_STD_forCHK, by = join_by(Std_ID))

#flag data
Chk_Std_flags <- Chk_Std_join %>% mutate(
  Chk_Flag_Std = ifelse(
    Chk_CI_low <= Std_CI_high |
      Std_CI_low <= Chk_CI_high,
    "Chk_Good",
    "Chk_Bad"
  )
) %>% mutate(CHK_Flag = paste0(gsub(" ", "", IDs), Chk_Flag_Std)) %>%
  mutate(CHK_CV_Flag = paste0("Chk",gsub(" ", "", IDs), Chk_CV_Flag))

#Organize Data for Plotting and Flagging
Chk_Flags <- Chk_Std_flags %>% group_by(Plate) %>% mutate(Chk_Flag = ifelse(any(CHK_Flag %like% "Bad"), "Chk_Bad", "Chk_Good"), Chk_CV_Flag = ifelse(any(CHK_CV_Flag %like% "High"), "Chk_CV_High", "Chk_CV_Good")) %>% ungroup() %>% mutate(CHK_CV_Flags = ifelse(CHK_CV_Flag %like% "High", paste0(gsub(" ", "", IDs), Chk_CV_Flag), "Good")) %>% mutate(CHKs_Flag = ifelse(CHK_Flag %like% "Bad", CHK_Flag, "Good")) 

H2S_CHKs_plot <- H2S_CHKs_a %>% rename(IDs = Chk_IDs, Std_CI_low = Chk_CI_low, Std_CI_high = Chk_CI_high, Mean_Abs = Chk_Mean_Abs) %>% mutate(Type = "Check") %>% dplyr::select(-Chk_CV_Flag)

H2S_Std_chkplot <- H2S_STD_forCHK_a %>% dplyr::select(IDs, Std_CI_low, Std_CI_high, Mean_Abs, Std_ID, Plate) %>% mutate(Type = "Standard")

Chk_flags_plot <- Chk_Flags %>% dplyr::select(Std_ID, CHKs_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(Check_Flag = ifelse(CHKs_Flag == "Good", "Good", "Bad"))

H2S_Chks_bind <- rbind(H2S_CHKs_plot, H2S_Std_chkplot)

H2S_Chks_bind_1 <- left_join(H2S_Chks_bind, Chk_flags_plot, by = join_by(Std_ID)) 

H2S_Chks_plot_bind_a <- H2S_Chks_bind_1 %>% mutate(Plot_ID = ifelse(IDs %like% "Chk" | IDs %like% "MC", substr(IDs, 4, 8), IDs)) %>% mutate(Plot_ID = ifelse(Plot_ID %like% "10pp", "Matrix: 10ppt", Plot_ID ))
H2S_Chks_plot_bind <-H2S_Chks_plot_bind_a %>% filter(str_detect(Type, "Check")) %>% mutate(Plot_ID =as.factor(Plot_ID))

Check_Val <- H2S_Chks_plot_bind_a %>% filter(!str_detect(Plot_ID, "Check")) %>% group_by(Plot_ID) %>% mutate(Z = mean(Mean_Abs)) %>% dplyr::select(Plot_ID, Z) %>% distinct(.keep_all = TRUE) %>%  filter(!Plot_ID %like% "10pp") %>% mutate(Plot_ID = ifelse(Plot_ID %like% "Std 5", "Matrix: 10ppt", Plot_ID )) %>% mutate(Plot_ID = as.factor(Plot_ID))
 
#Plot Check Standards
H2S_chk_plot <- ggplot(data = H2S_Chks_plot_bind, aes(x = Plate, y = Mean_Abs)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75, aes(fill = Check_Flag)) +
      geom_hline(data = Check_Val, aes(yintercept = Z), linetype = "dashed", linewidth = 1, color = "darkblue") +
  facet_wrap(.~Plot_ID) +
  ylim(0, max(H2S_Chks_plot_bind$Mean_Abs) + 0.2) +
  geom_label_repel(data = Check_Val, aes( x = "Plate1", y = Z + 0.25), label = "Mean Target Arbsorbance",  alpha = 0.7, color = "darkblue") +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Check Stds: H2S with Confidence Interval") +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1))


H2S_chk_plot

```

## Display Any Check Flags
```{r, echo=FALSE}
Chk_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHKs_Flag) %>% filter(CHKs_Flag != "Good")

CHK_CV_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags) %>% filter(CHK_CV_Flags != "Good")

ifelse(length(Chk_Flag_1$CHKs_Flag) >= 1, paste0(Chk_Flag_1$CHKs_Flag, "-", Chk_Flag_1$Plate), paste0("All Checks Good"))

ifelse(length(CHK_CV_Flag_1$CHK_CV_Flags) >= 1, paste0(CHK_CV_Flag_1$CHK_CV_Flags, "-", CHK_CV_Flag_1$Plate), paste0("All Checks CV Good"))

```


##Calculate Sulfide Concentrations & Add Concentration Flags
```{r Calculate Sulfide Concentrations,echo=FALSE,include=FALSE}

High_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 5") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% ungroup() %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(High_DL = Mean_Abs)

#filter out stds and other experiments
raw_data <- dat %>%  
      filter(!str_detect(IDs, "Std")) %>%  
      filter(!str_detect(IDs, "MC"))

#Calculate Concentrations and fix zeroes/negative values and flag conc

joined_data <- right_join(Std_crv_calc, raw_data, by = join_by(Plate))

calc_data <- joined_data %>% mutate(Conc = ((Abs - Int_Calc) / Slope_Calc)*Dilution) %>% mutate(Conc = ifelse(Conc <= 0, 0, Conc))

dl_data <- right_join(High_Std, calc_data, by = join_by(Plate))

```

##Remove High CV Samples and Average
```{r, echo=FALSE}

#Calculate CVs
H2S_SPL_cv <- dl_data  %>% group_by(Plate, IDs) %>% mutate(Mean_Conc = mean(Conc)) %>% mutate(Conc_SD = sd(Conc)) %>% mutate(CV = cv(Conc)) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Conc_Flag = ifelse(Mean_Abs < MDL, "bdl", ifelse(Mean_Abs > High_DL, "adl", "Within_Range"))) %>% ungroup()

#calculate new cv getting best cv from points if cv was high
H2S_SPL_cv_new <- H2S_SPL_cv %>% group_by(Plate, IDs) %>%
  mutate(CV_New = map_dbl(row_number(), ~cv(Conc[-.x]))) %>%    #Calculate CVs without 1 point
  mutate(Removed = ifelse(             #Choose Lowest CV if CV is high
  CV_Flag == "CV_High",
  ifelse(
    CV_New == min(CV_New),
   "RMVD",
    "KEPT"), "CV_Good"),
CV2 = ifelse(CV_Flag == "CV_High", min(CV_New), CV),
Conc_rm = ifelse(Removed == "RMVD", NA, Conc)) %>%
  mutate(Mean_Conc2 = mean(Conc_rm, na.rm = TRUE)) %>%
  mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_sample, "CV_High", "CV_Good")) %>%
  mutate(Conc_SD2 = sd(Conc_rm, na.rm = TRUE)) %>% ungroup()

H2S_SPL_CV_fin <- H2S_SPL_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Conc,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Conc2,
                                                   Conc_SD2,
                                                   CV2,
                                                   CV_Flag2,
                                                   Conc_Flag) %>% 
  mutate(Mean_Conc = Mean_Conc2, CV_Flag = CV_Flag2, SD_Conc = Conc_SD2, CV = CV2) %>% 
                                  dplyr::select(Plate,                                                                                                      Conc,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Conc,
                                                SD_Conc,
                                                CV,
                                                CV_Flag,
                                                Conc_Flag) 

```

## Plot Samples
```{r, echo=FALSE}

#Filter samples and mark reruns
H2S_Spl_pre_viz <- H2S_SPL_cv_new %>%  mutate(IDs = toupper(IDs)) %>%
      filter(str_detect(IDs, "L")) %>%
  filter(!str_detect(IDs, "DUP")) %>%
  filter(!str_detect(IDs, "SPIKE")) %>%
  dplyr::select(Plate, IDs, Mean_Conc, Conc_SD, CV_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(IDs = ifelse(Plate %like% "rr", paste0(IDs, "_RR"), IDs))

#Plot Samples before corrections
H2S_Spl_Plot_Pre <- ggplot(H2S_Spl_pre_viz, aes(y = Mean_Conc, x = IDs, fill = CV_Flag)) +
  geom_bar(stat = 'identity',
           position = position_dodge2(),
           linewidth = 0.75,
           col = "white") +
  scale_fill_manual(values = c(
    "CV_Good" = "green2",
    "CV_High" = "pink2"
  )) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  geom_errorbar(
    aes(ymin = Mean_Conc - Conc_SD, ymax = Mean_Conc + Conc_SD),
    width = 0.3,
    position = position_dodge(.1)
  ) +
  theme_classic() +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ggtitle("Sample triplicate means and sd dev before bad reps removed")


H2S_Spl_Plot_Pre

H2S_Spl_rmv_viz <- H2S_SPL_CV_fin %>%  mutate(IDs = toupper(IDs)) %>%
      filter(str_detect(IDs, "L")) %>%
  filter(!str_detect(IDs, "DUP")) %>%
  filter(!str_detect(IDs, "SPIKE")) %>%
  dplyr::select(Plate, IDs, Mean_Conc, SD_Conc, CV_Flag, Conc_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(IDs = ifelse(Plate %like% "rr", paste0(IDs, "_RR"), IDs))

#Plot samples after Corrections
H2S_Spl_Plot_CV <- ggplot(H2S_Spl_rmv_viz, aes(y = Mean_Conc, x = IDs, fill = CV_Flag)) +
  geom_bar(aes(col = Conc_Flag), stat = 'identity',
           position = position_dodge(),
           linewidth = 0.75) +
  scale_fill_manual(values = c(
    "CV_Good" = "green2",
    "CV_High" = "pink2"
  )) +
  scale_color_manual(values = c("Within_Range" = "darkgreen", "adl" = "red3", "bdl" = "blue3" )) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  geom_errorbar(
    aes(ymin = Mean_Conc - SD_Conc, ymax = Mean_Conc + SD_Conc),
    width = 0.3,
    position = position_dodge(.1)
  ) +
  theme_classic() +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ggtitle("Sample triplicate means and sd dev after bad reps removed")


H2S_Spl_Plot_CV



```


## Duplicate QAQC

```{r,echo=FALSE}

#pull out any rows that have "Dup" in the ID column and tag reruns
DUP_start <- H2S_SPL_CV_fin %>%  mutate(Dup_ID = gsub(" ", "", substr(IDs, 1, 6)),
                                        Dup = ifelse(IDs %like% "Dup", "Dup", "Org")) %>% mutate(Dup_ID = ifelse(Plate %like% "rr", paste0(Dup_ID, "_RR"), Dup_ID))

#separate dups from originals and tag for pairing 

dups_only <- DUP_start %>% filter(Dup == "Dup") %>% dplyr::select(Dup_ID, Mean_Conc, SD_Conc, Plate) %>% rename(Dup_Conc = Mean_Conc, Dup_SD = SD_Conc) %>% distinct(.keep_all = TRUE)

no_dups <- DUP_start %>% filter(!Dup == "Dup")

dup_spls_l <- dups_only %>% dplyr::select(Dup_ID) %>% distinct(.keep_all = TRUE)

dup_spls <- no_dups %>% filter(Dup_ID %in% as.vector(dup_spls_l$Dup_ID)) %>% dplyr::select(Dup_ID, Mean_Conc, SD_Conc, Plate, IDs) %>% distinct(.keep_all = TRUE)

dup_join <- full_join(dups_only, dup_spls, by = join_by(Dup_ID, Plate)) %>% na.omit()

#Calculate Percent difference and Flag
dup_flags <- dup_join %>% 
  mutate(Diff_P = (abs(100 * (Mean_Conc - Dup_Conc) / mean(c(Mean_Conc, Dup_Conc))))) %>% 
  mutate(Dup_Flag = ifelse(Diff_P >= dups_perc_diff, "Dup_Bad", "Dup_Good"))

#oraganize data for flagging and plotting
dup_flags_1 <- dup_flags %>% dplyr::select(Dup_ID, Dup_Flag, Diff_P, Plate)

Dup_for_plot_1 <- DUP_start %>% dplyr::select(Plate, Mean_Conc, Dup_ID, IDs) %>% distinct(.keep_all = TRUE) %>% filter(Dup_ID %in% as.vector(dup_spls_l$Dup_ID)) %>%
  left_join(dup_flags_1, by = join_by(Dup_ID, Plate)) %>%
  mutate(Dup = ifelse(IDs %like% "Dup", "Duplicate", "Original")) %>%
  dplyr::select(-IDs, -Mean_Conc, -Dup) %>% distinct(.keep_all = TRUE) %>% na.omit

#plot Data
Dup_Plot <- ggplot(Dup_for_plot_1,
                   aes(
                     x = Plate,
                     y = Diff_P
                   )) +
  geom_bar(
    aes(fill = Dup_Flag),
    stat = 'identity',
    position = position_dodge2(preserve = "single"),
    linewidth = 0.9
  ) +
  geom_label_repel(data = as.data.frame(1), aes(x = "Plate1", y=17.5), label= "Percent Difference Cutoff", color = "red3",  alpha = 0.7) + 
  geom_hline(aes(yintercept = dups_perc_diff), linetype = "dashed", linewidth = 1.5, color = "red3") +
  labs(y = "Percent Difference", x = "Sample ID") +
  scale_fill_manual(values = c(
    "Dup_Good" = "darkgreen",
    "Dup_Bad" = "darkred"
  )) +
  theme_classic()+
  theme(legend.position = "bottom") +
  ggtitle("H2S Duplicates") 

Dup_Plot


```

## Display Dup Flags

```{r, echo=FALSE}
Dup_Flag_1 <- dup_flags %>% group_by(Plate)  %>% mutate(Dup_Flags = ifelse("Dup_Bad" %in% Dup_Flag, "Dup_Bad", "Dup_Good")) %>% dplyr::select(Plate, Dup_Flags) %>% rename(Dup_Flag = Dup_Flags) %>% distinct(.keep_all = TRUE)

Dup_Flags_1 <-Dup_Flag_1 %>% mutate(Dup_Flags = paste0(Dup_Flag, "-", Plate))

Dup_Flags_2 <- Dup_Flags_1 %>% dplyr::select(Plate, Dup_Flag) %>% group_by(Plate) %>% mutate(Duplicate_Flag = ifelse("Dup_Bad" %in% Dup_Flag, "Dup_Bad", "")) %>% dplyr::select(Plate, Duplicate_Flag) %>% rename(Dup_Flag = Duplicate_Flag) %>% distinct(.keep_all = TRUE)

Bad_Dups <- dup_flags %>% filter(str_detect(Dup_Flag, "Dup_Bad"))
                                 
Bad_D_N <- length(Bad_Dups$Dup_Flag)

Total_D_N <- length(dup_flags$Dup_Flag)

ifelse(Bad_D_N/Total_D_N > 0.4, paste0("<60% of Dups Pass, Rerun"), paste0(">60% Dups Pass") )

Duplicate_Flag <-  ifelse(Bad_D_N/Total_D_N > 0.4, paste0(" >40% Dups Fail"), "" )

paste0(Dup_Flags_1$Dup_Flags)


```

## Spike QAQC
```{r, echo=FALSE}
#pull out any rows that have "Spk" in the ID column
SPK_start <- H2S_SPL_CV_fin %>%  mutate(Spk_ID = gsub(" ", "", substr(IDs, 1, 6)),
                                        Spk = ifelse(IDs %like% "Spike", "Spk", "Org")) %>% mutate(Spk_ID = ifelse(Plate %like% "rr", paste0(Spk_ID, "_RR"), Spk_ID))
#separate spikes from originals
spks_only <- SPK_start %>% filter(Spk == "Spk") %>% dplyr::select(Spk_ID, Mean_Conc, SD_Conc, Dilution, Plate) %>% rename(Spk_Conc = Mean_Conc,
                                                                                                                   Spk_SD = SD_Conc,
                                                                                                                   Spk_Dilution = Dilution) %>% distinct(.keep_all = TRUE)

no_spks <- SPK_start %>% filter(!Spk == "Spk")

spls_l <- no_spks %>% dplyr::select(Spk_ID)  %>% distinct(.keep_all = TRUE)

spk_spls_l <- spks_only %>% dplyr::select(Spk_ID) %>% distinct(.keep_all = TRUE)

spk_spls <- no_spks %>% filter(Spk_ID %in% as.vector(spk_spls_l$Spk_ID)) %>% dplyr::select(Spk_ID, Mean_Conc, SD_Conc, Plate, IDs, Dilution) %>% distinct(.keep_all = TRUE)

SPK_only <- spks_only %>% filter(Spk_ID %in% as.vector(spls_l$Spk_ID))

spk_join <- full_join(SPK_only, spk_spls, by = join_by(Spk_ID, Plate))

#Calculate recovery and flag data
spk_flags <- spk_join %>% mutate(
  Spl_mol = (sample_vol / Con1) * (Mean_Conc / Dilution),
  Spk_mol = ((sample_vol + spkvol) / Con1) * (Spk_Conc / Spk_Dilution), 
  Spk_exp = Spl_mol + spk_Conc, 
  Spk_recovery = (Spk_mol/Spk_exp) * 100) %>% 
  mutate(
    Spk_Flag = ifelse(
      Spk_recovery >= high_recovery_cutoff |
      Spk_recovery <= low_recovery_cutoff,
      "Spk_Bad",
      "Spk_Good"
    )
  )

#organize data for flagging and plotting
spk_flags_1 <- spk_flags %>% dplyr::select(Spk_ID, Spk_Flag, Spk_recovery, Plate) %>% na.omit()

Spk_for_plot_a <- SPK_start %>% dplyr::select(Plate, Mean_Conc, Spk_ID, IDs) %>%
  distinct(.keep_all = TRUE) %>%
  filter(Spk_ID %in% as.vector(spk_spls_l$Spk_ID)) %>%
  left_join(spk_flags_1, by = join_by(Spk_ID, Plate)) %>%
  mutate(Spk = ifelse(IDs %like% "Spike", "Spike", "Original")) %>%
  filter(Spk_ID %in% as.vector(spls_l$Spk_ID))

Spk_for_plot_1 <- Spk_for_plot_a %>% dplyr::select(Plate, Spk_ID, Spk_recovery, Spk_Flag) %>% na.omit() %>% distinct(.keep_all = TRUE)

#plot data
Spk_Plot <- ggplot(Spk_for_plot_1,
                   aes(
                     x = Plate,
                     y = Spk_recovery
                   )) +
  geom_rect(aes(xmin=0, xmax=Inf, ymin=low_recovery_cutoff, ymax=high_recovery_cutoff), alpha = 0.25, fill ="palegreen") +
  geom_bar( aes(fill = Spk_Flag),
    stat = 'identity',
    position = position_dodge2(preserve = "single"),
    linewidth = 0.9
  ) +
  geom_hline(aes(yintercept = high_recovery_cutoff), linetype = "dashed", color = "green3", linewidth = 1.5) +
    geom_hline(aes(yintercept = low_recovery_cutoff), linetype = "dashed", color = "green3", linewidth = 1.5) +
  geom_label_repel(data = as.data.frame(1),aes(x = "Plate1", y=high_recovery_cutoff - 7.5), label= "Acceptable Recovery Range", color = "darkgreen",  alpha = 0.7) + 
  labs(y = "Spike Recovery (%)", x = "Plate") +
  scale_fill_manual(values = c(
    "Spk_Good" = "darkgreen",
    "Spk_Bad" = "darkred"
  )) +
  theme_classic() +
    theme(legend.position = "bottom") +
  ggtitle("H2S Spikes")

Spk_Plot


```

## Display Spike Flags
```{r, echo=FALSE}

#Check if all Spikes are paired
ifelse(all(as.vector(spk_spls_l$Spk_ID) %in% as.vector(spls_l$Spk_ID)), paste0("Spikes Paired"), paste0("Spike Missing Pair"))

#Format Data frame
Spk_Flag_1 <- spk_flags  %>% dplyr::select(Plate, IDs, Spk_Flag) %>% na.omit()

#Attach Plate to Flag
Spk_Flags_1 <-Spk_Flag_1 %>% group_by(Plate) %>% mutate(Spk_Flags = ifelse( "Spk_Bad" %in% Spk_Flag, paste0("Spk_Bad", "-", Plate), paste0("Spk_Good", "-", Plate))) %>% ungroup()
#If any Spikes in Plate are bad, then flag plate
Spk_Flags_2 <- Spk_Flags_1 %>% dplyr::select(Plate, Spk_Flag) %>% group_by(Plate) %>% mutate(Spike_Flag = ifelse("Spk_Bad" %in% Spk_Flag, "Spk_Bad", "")) %>% dplyr::select(Plate, Spike_Flag) %>% rename(Spk_Flag = Spike_Flag) %>% distinct(.keep_all = TRUE)

Spk_disp <- Spk_Flags_1 %>% dplyr::select(Plate, Spk_Flags) %>% distinct(.keep_all = TRUE) %>% na.omit()

paste0(Spk_disp$Spk_Flags)

bad <- Spk_Flag_1 %>% filter(Spk_Flag %like% "Bad")

length_bad <- length(as.vector(bad$Spk_Flag))

length_all <-  length(as.vector(Spk_Flag_1$Spk_Flag))

ifelse(length_bad/length_all > 0.4, paste0("<60% of Spikes Pass, Rerun"), paste0("Spikes Pass") )

Spike_Flag <- ifelse(length_bad/length_all > 0.4, paste0(">40% Spikes Fail"), "" )

```


## Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}
#Get Samples Only
H2S_spls_md <- H2S_SPL_CV_fin %>%  mutate(IDs = toupper(IDs)) %>%
      filter(str_detect(IDs, "L")) %>%
  filter(!str_detect(IDs, "DUP")) %>%
  filter(!str_detect(IDs, "SPIKE"))

#Check for reruns and replace
H2S_spl_rrn <- H2S_spls_md %>%
  filter(Plate %like% "rr") %>% 
  dplyr::select(IDs, Mean_Conc, Conc_Flag) %>%
  rename(RRN_Conc = Mean_Conc, Conc_Flag_rr = Conc_Flag) %>% 
  distinct(.keep_all = TRUE) 

RRN_IDs <- as.vector(H2S_spl_rrn$IDs)

H2S_spl_og <- H2S_spls_md %>%
  filter(!Plate %like% "rr") %>% 
  dplyr::select(IDs, Mean_Conc, Conc_Flag, Plate, CV_Flag, Dilution, CV) %>% 
  distinct(.keep_all = TRUE) %>%
  mutate(Rerun = ifelse(IDs %in% RRN_IDs, "RRN", "Org"))

H2S_spl_rr_join <- full_join(H2S_spl_rrn,H2S_spl_og, by = join_by(IDs) ) 

Samples_Tagged <-  H2S_spl_rr_join %>% mutate(
  Conc = ifelse(Rerun == "RRN", RRN_Conc, Mean_Conc),
  Conc_Flag = ifelse(Rerun == "RRN", Conc_Flag_rr, Conc_Flag)
) %>% dplyr::select(IDs, Conc, Conc_Flag, CV_Flag, Plate, Dilution, CV) %>% rename(Sample_ID = IDs) %>% ungroup() %>% group_by(Sample_ID) %>% mutate(
  Min_CV = ifelse(CV == min(CV), "KEEP", "RMV"), # get if CV is lowest of reruns
  Within_Range = "Within_Range" %in% Conc_Flag, #Get if any are within range
  BDL = "bdl" %in% Conc_Flag, #get if any are below detection limit
  Max_Dil = max(Dilution), #get maximum dilution of reruns
  Min_Dil = min(Dilution),#get minimum dilution of reruns
  All_Good = all(Conc_Flag == "Within_Range") & all(CV_Flag == "CV_Good"), #get if all are within range and good cv
  CV_Check = CV == min(CV[which(Dilution == Max_Dil)]), #Get if CV is lowest of highest dilution for reruns
  Number = length(Conc) #get number of times ran
) %>% ungroup() %>%
  mutate(Remove = ifelse(Number == 1, #check if sample was reran
                         "KEEP", 
                         ifelse(All_Good, #check if multiple samples meet all criteria
    ifelse(Min_CV == "KEEP", #Keep only lowest CV
           "KEEP1",
           "RMV1"),
    ifelse(Conc_Flag == "Within_Range" & CV_Flag == "CV_Good", #If only one sample meets all criteria keep that one
      "KEEP2",
      ifelse(!(Within_Range) & BDL, #Check if none are within range but are under detection limit
        ifelse(Dilution == Min_Dil, #keep lowest dilution
               "KEEP3",
               "RMV3"),
        ifelse(!(Within_Range) & CV_Check, #if all are above detection limit keep lowest CV of highest dilution
               "KEEP4",
               "RMV4")
      )
    )
  )))

Samples <- Samples_Tagged %>% filter(Remove %like% "KEEP")

#check to see if all samples are present in the metadata
all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}


#merge metadata with sample run data
merged_data <- metadata %>%
  mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
  fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included?

```


##Add Flags
```{r, echo=FALSE}

#Get only bad flags for each plate and check
Chk_Flags_All <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags, Chk_Flag, Chk_IDs) %>% group_by(Plate) %>% mutate(CHKSTD_Flags = ifelse(Chk_Flag %like% "Good", NA,paste0(Chk_IDs, "_",Chk_Flag))) %>% dplyr::select(Plate, CHKSTD_Flags) %>% mutate(CHKSTD_Flags = ifelse(CHKSTD_Flags == "", NA, CHKSTD_Flags)) %>% na.omit() %>% distinct(.keep_all = TRUE)

STD_Flags_All <- STD_CRV_flags %>% dplyr::select(Plate, R2_Flag, CV_Flag) %>% mutate(STDCV_Flags = paste0("STD", CV_Flag)) %>% dplyr::select(-CV_Flag) %>% mutate(
  R2_Flag = ifelse(R2_Flag %like% "Good", NA, R2_Flag),
  STDCV_Flags = ifelse(STDCV_Flags %like% "Good", NA, STDCV_Flags)
) %>% na.omit()

Flags_Int <- tibble(Plate = c(plates))

#Join Flags

Flags_Joined_1 <-left_join(Flags_Int, Dup_Flags_2, by = join_by(Plate))

Flags_Joined_2 <- left_join(Flags_Joined_1, Spk_Flags_2, by = join_by(Plate))

Flags_Joined_3 <- left_join(Flags_Joined_2, Chk_Flags_All, by = join_by(Plate))

Flags_Joined_4 <- left_join(Flags_Joined_3, STD_Flags_All, by = join_by(Plate))

Flags_Joined <- left_join(Flags_Joined_4, Curve_Flags_forJoin, by = join_by(Plate))


Flags_Joined[Flags_Joined == "NA"] <- NA

#Merge Flags into one column
Flags <- Flags_Joined %>% replace(is.na(.), "") %>% mutate(
  H2S_Flags = paste0(
    Spk_Flag,
    " ",
    Dup_Flag,
    " ",
    CHKSTD_Flags,
    " ",
    R2_Flag,
    " ",
    STDCV_Flags,
    " ",
    Slope_Flag,
    " ",
    Int_Flag
  )
) %>% dplyr::select(Plate, H2S_Flags)

#Add Flags to data
flagged_data <- left_join(merged_data, Flags, by = join_by(Plate))


df_all_clean <- flagged_data %>% dplyr::select(-Plate)

```

##Organize Data  
```{r, echo=FALSE}
#check results
#head(Samples)

#pull out what we need

df_all_clean_cols <- df_all_clean %>%
  dplyr::select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot_Name,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    Conc,
    Conc_Flag,
    CV_Flag,
    H2S_Flags
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = Run_notes,# new column with notes about the run
    Analysis_rundate = Run_Date) 

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
    H2S_Conc_uM = Conc,
    H2S_Conc_Flag = Conc_Flag,
    Plot = Plot_Name,
    H2S_CV_Flag = CV_Flag
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  dplyr::select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    H2S_Conc_uM,
    H2S_Conc_Flag,
    H2S_CV_Flag,
    H2S_Flags,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)

```


## Visualize Data
```{r Visualize Data, echo=FALSE}

H2S_forplot <- All_Clean_Data

#ensure points are in correct order

H2S_forplot <- H2S_forplot[order(H2S_forplot$Depth_cm), ]

color1 <- c("L1", "L2", "L3", "L4")
color2 <- c("L5", "L6", "L8", "L10")
color3 <- c("L7", "L9", "L11", "L12")
color4 <- c("L13", "L14", "L16", "L15")
color5 <- c("L17", "L18", "L19", "L20")

H2S_forplot <- H2S_forplot %>%
  mutate(
    Plot_Color = case_when(
      Plot %in% color1 ~ "Color1",
      Plot %in% color2 ~ "Color2",
      Plot %in% color3 ~ "Color3",
      Plot %in% color4 ~ "Color4",
      Plot %in% color5 ~ "Color5"
    )
  ) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2 & eN",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2 & aN",
      ifelse(
        CO2_Treatment == "Ambient" &
          N_Treatment == "Amended"  ,
        "aCO2 & eN" ,
        "aCO2 & aN"
      )
    )
  ))  %>%
  mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2"))  %>%
  mutate(N_Treatment = ifelse(N_Treatment == "Ambient", "aN", "eN")) 

viz_H2S_plot2 <- ggplot(data = H2S_forplot, aes(
  x = factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")),
  y = H2S_Conc_uM,
  fill = as.factor(Treatment),
  color = as.factor(Treatment)
)) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(padding = 0.25,preserve = "single")
  ) +
  facet_grid(Depth_cm ~ N_Treatment, scales = "fixed") +
  scale_fill_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "white",
    "aCO2 & aN" = "white"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
    
  ) +
  scale_color_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "#DB5461",
    "aCO2 & aN" = "#454ADE"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
  ) +
  theme_classic() +
  labs(x = " ",
       y = "H2S (uM)",
       title = "CO2xN: Porewater Sulfide"
       ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  ) +
coord_flip()
print(viz_H2S_plot2)

```



```{r, echo=FALSE}
#ensure points are in correct order

#make plot
viz_H2S_plot_2 <- ggplot(data = H2S_forplot, aes(x = H2S_Conc_uM, y = as.numeric(Depth_cm), color = Plot_Color)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(
    values = c(
      "Color1" = "#FFD500",
      "Color2" = "#00D0FF",
      "Color3" = "#FF0075",
      "Color4" = "#80d280",
      "Color5" = "#ff6040"
    ),
    labels = c("1", "2", "3", "4", "5")
  ) +
  facet_grid(Treatment ~ ., scales = "fixed") +
  labs(
    x = "H2S (uM)",
    y = "Depth (cm)",
    title = "CO2xN: Porewater H2S by Replicate",
    color = "Replicate"
  ) +
  scale_y_reverse(breaks = c(20, 40, 80), limits = c(90, 10)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_H2S_plot_2
```


##Export Data
```{r, echo=FALSE}

  Rerun_Data <- All_Clean_Data %>% filter(H2S_Conc_Flag != "Within Range" | H2S_CV_Flag %like% "High")

if(After_Reruns == "YES"){
  write.csv(All_Clean_Data, final_path)} else {
  write.csv(Rerun_Data, rerun_path) }
#end
```


### END