---
title: "CO2xN: Porewater Sulfide"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
    number_sections: false
geometry: left=2cm,right=2cm,top=1cm,bottom=2cm
---
## Code Set up 
```{r setup, include=FALSE}

library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
library(purrr)
library(tidyverse)
library(here)
library(data.table)
library(matrixStats)
library(gridExtra)
library(grid)
library(plater)
library(raster)
library(knitr)  
library(readxl)


```

## run information
```{r Information to be changed}

###things that need to be changed
Date_Run = "20250527"
Run_by = "Melanie Giessner"#Instrument user
Script_run_by ="Brenden Blakley"#Code user

Run_notes=""#any notes from run

#Import Plate Data

plates<- c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate_rr")

Raw_plates <- c("Raw Data/20250527_LTREB_H2S_Plate1.csv","Raw Data/20250527_LTREB_H2S_Plate2.csv","Raw Data/20250527_LTREB_H2S_Plate3.csv","Raw Data/20250527_LTREB_H2S_Plate4.csv","Raw Data/20250527_LTREB_H2S_Plate5.csv","Raw Data/20250527_LTREB_H2S_Plate6.csv","Raw Data/20250527_LTREB_H2S_Plate7.csv","Raw Data/20250602_H2S_LTREBrr.csv")

#Import Run Date for Each Plate
Run_dates <- tibble(
    Plate = c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate_rr"),
    Run_Date = c("05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","06/02/2025")
  )

  Run_dates <- Run_dates %>% mutate(Run_Date = as.Date(Run_Date, format = "%m/%d/%Y"))


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Processed Data/LTREB_2M_H2S_QAQC_2025.csv"

# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_CO2xN_Porewater_H2S_202505.csv"

#Stds that should be excluded
#stds_to_remove<-data.frame(Plate=c("Plate4","STD2"),IDs=c("Std 4","Std 3"))


```


```{r, echo=FALSE}


#file path name for processed data
summarizeddata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_Summary_Data.csv")

fulldata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_Data.csv")

shortdata_path=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_short_Data.csv")

Need_dilute=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_Need_dilute_Data.csv")

high_CV=paste0("2025/",Experiment,"/QAQC'd Data/GENX2_H2S_",Date_Run,"_",file_name,"_high_CV_Data.csv")


```

```{r Set Up Code - constants and QAQC cutoffs, include=FALSE}

#Flag cutoffs
  r2_cutoff = 0.985            #this is the level below which we want to rerun or consider a curve 
  cv_flag_stds = 10           #this is the maximum cv allowed for standards
  p_value_chkstds = 0.05      #the p-value for the t-test between check standards must be greater than this
  p_value_MC = 0.05           #the p-value for the t-test between matrix checks and top std must be greater than this
  cv_flag_sample = 10         #this is the maximum cv allowed for samples
  dups_perc_diff = 15.5       #this is the maximum percent difference allowed between duplicates
  high_recovery_cutoff = 120  #this is the maximum percent recovery of SO4 allowed in spiked samples
  low_recovery_cutoff = 80    #this is the minimum percent recovery of SO4 allowed in spiked samples

#Standard concentrations - Update if running different standard curve:
# standard units are in uM
  standards <- tibble(
    IDs = c("Std 0", "Std 1", "Std 2", "Std 3", "Std 4", "Std 5"),
    Conc = c(0, 5.0, 12.5, 25.0, 50.0, 100.0)
  )  #uM
  
    chk_standards <- tibble(
    IDs = c("ChkStd 0", "ChkStd 3", "ChkStd 4"),
    Conc = c(0, 25.0, 50.0)
  )  #uM
    
    mtx_standards <- tibble(
    IDs = c("MC: 10ppt S5"),
    Conc = c(100)
  )  #uM

#Spike concentration calc
  #spike for these samples was 50 uL of the 100uM standard
  Con1 <- 1000000               #conversion factor value for spike volumes (uL -> L)
  spk_std <- 100                # uM S2- standard used
  spkvol <- 50                  # uL volume of spike added
  spk_Conc <- spk_std*(spkvol/Con1)  # umoles of S2- added to each spiked sample
  sample_vol <- 250             # the sample volume without the spike is 250 uL

#Top standard Concentration- Update if running different standard curve: 
   Top_STD = 100

```


## read in data
```{r read in plates,echo=FALSE,include=FALSE}

#read in the csv files
dat<-read_plates(
  files = Raw_plates,#list of all file paths
  plate_names = plates ,  #list of plate names          
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
) %>%
  rename("Abs"=values,
         "IDs"=values.2,
         "Dilution"=values.3)
head(dat)
  
```

#Remove High CV Stds
```{r}

#filter for STDs
H2S_stds_raw <- dat %>% subset(IDs %like% "Std") %>% subset(!IDs %like% "Chk")

#Add in STD values
H2S_stds <- left_join(H2S_stds_raw, standards, by = "IDs")

#Calculate CVs
H2S_STD_cv <- H2S_stds %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_STD_cv_cor <- H2S_STD_cv %>% mutate(Mean_diff = abs(Abs - Mean_Abs)) %>% group_by(Plate, IDs) %>% mutate(Abs_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Abs), Abs))

H2S_STD_cv_new <- H2S_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>% mutate(Abs_SD2 = sd(Abs_rm, na.rm = TRUE)) %>% mutate(CV2 = (Abs_SD2 /
                                                                                                                                                                                Mean_Abs2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_STD_CV_fin <- H2S_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate, 
                                                Wells,                                                                                                                Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Conc,
                                                Mean_Abs,
                                                CV_Flag)


```


#Check R2 & Remove Standards if Necessary
```{r}
split_data <- H2S_STD_CV_fin %>% dplyr::select(Plate, IDs, Conc, Mean_Abs, CV_Flag) %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_H2S_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Conc ~ Mean_Abs, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ Mean_Abs, data = .))[2],
                                        Intercept = coef(lm(Conc ~ Mean_Abs, data = .))[1]))

 Slope_H2S_raw <- bind_rows(Slope_H2S_raw_list)
  
 Slope_H2S_Std_Crvs <- Slope_H2S_raw %>% dplyr::select(Plate, Curve, R2, Slope, Intercept) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))

 
 R2_Check_Raw <- left_join(Slope_H2S_Std_Crvs, H2S_STD_CV_fin, by = "Plate")
 
 R2_Check <- R2_Check_Raw %>% 
   dplyr::select(Plate,
                 IDs,
                 Mean_Abs,
                 Conc,
                 R2,
                 Slope,
                 Intercept,
                 CV_Flag,
                 R2_Flag) %>%
   distinct()
 
 R2_Check_diff <- R2_Check %>% mutate(Calc_Conc = ((Slope*Mean_Abs) + Intercept)) %>%
   mutate(Conc_Diff = abs(Conc - Calc_Conc)) %>% mutate(RMVBL = ifelse(IDs %in% c("Std 0", "Std 1","Std 5"), "No", "Yes"))

 R2_Check_RMVBL <- R2_Check_diff %>% filter(RMVBL == "Yes")
 
 R2_Check_NRMVBL <- R2_Check_diff %>% filter(RMVBL != "Yes") %>% mutate(STD_ABS_rm = Mean_Abs) %>% mutate(N_HCV = 0, STD_ABS_rmv = STD_ABS_rm, N_stds = 3)
 
 
 R2_Check_na <- R2_Check_RMVBL %>% group_by(Plate) %>% mutate(STD_ABS_rm = ifelse(R2_Flag == "R2_Bad", ifelse(max(Conc_Diff) == Conc_Diff, NA, Mean_Abs), Mean_Abs)) %>% ungroup()
 
 R2_Check_CVrm <- R2_Check_na %>% na.omit() %>% group_by(Plate) %>% mutate(N_stds = length(STD_ABS_rm)) %>% group_by(Plate, CV_Flag) %>%
   mutate(N_HCV = n()) %>% ungroup() %>% group_by(Plate) %>%
   mutate(STD_ABS_rmv = ifelse(
     N_stds == 3,
     ifelse(
       CV_Flag == "CV_High" ,
       ifelse(
         N_HCV == 1 ,
         NA,
         ifelse(max(Conc_Diff) == Conc_Diff, NA, STD_ABS_rm)
       ),
       STD_ABS_rm
     ),
     STD_ABS_rm
   ))
 
  R2_Check_bind <- rbind(R2_Check_NRMVBL,R2_Check_CVrm) %>% na.omit()
 
 R2_Check_fin <- R2_Check_bind[order(R2_Check_bind$Plate,R2_Check_bind$IDs), ]
 
 
 
 split_data_rmv <- R2_Check_fin %>% dplyr::select(Plate, IDs, Conc, STD_ABS_rmv, CV_Flag, R2_Flag) %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
 
 Slope_H2S_raw_list_rm <- map(split_data_rmv, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Conc ~ STD_ABS_rmv, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ STD_ABS_rmv, data = .))[2],
                                        Intercept = coef(lm(Conc ~ STD_ABS_rmv, data = .))[1]))

 Slope_H2S_raw_rm <- bind_rows(Slope_H2S_raw_list_rm)
  
 Slope_H2S_Std_Crvs_rm <- Slope_H2S_raw_rm %>% dplyr::select(Plate, Curve, R2, Slope, Intercept, CV_Flag) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))
 
 
 
```



```{r Plot all Standard Curves together, echo=FALSE,message=FALSE,fig.keep='none'}
##this is to see if there are any stds or plates that look off
#Remove check standards which have NAs in the Conc column


H2S_Curve <- ggplot(Slope_H2S_raw_rm, aes(STD_ABS_rmv, Conc, color = Plate)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "H2S Standard Curves", x = "Absorbance", y = "Concentration (uM)")
H2S_Curve


STD_run_date <- left_join(Run_dates, Slope_H2S_Std_Crvs_rm )
new_log_stds <- STD_run_date %>% dplyr::select(Plate,
   Curve,
   R2,
   Slope,
   Intercept,
   Run_Date) %>% mutate(R2 = round(R2, digits = 4), Slope = round(Slope, digits = 4), Intercept = round(Intercept, digits = 4), )


log <- read.csv(log_path)
log_2 <- log %>% mutate(Run_Date = as.Date(Run_Date))
head(log_2)
log_3 <- log_2[, -c(1)]
log_4 <- rbind(log_3, new_log_stds)

unique_log <- log_4 %>% distinct(.keep_all = TRUE)

Slopes_chk <- ggplot(unique_log, aes(x = Run_Date, y = Slope)) +
    ylim(50,175) +
  geom_abline(intercept = 60, slope = 0 )+
    geom_abline(intercept = 150, slope = 0 )+
  geom_point(size = 4) +
  theme_bw()

Slopes_chk

write.csv(unique_log, log_path)


```

#Return Flags For Standard Curve
```{r}

STD_CRV_flags <- full_join(Slope_H2S_Std_Crvs_rm, Run_dates, by = join_by(Plate))

ifelse(STD_CRV_flags$CV_Flag == "CV_High", paste0("CV High", "-", STD_CRV_flags$Plate), paste0("CV Good", "-", STD_CRV_flags$Plate))

ifelse(STD_CRV_flags$R2_Flag == "R2_Bad", paste0("R2 Bad", "-", STD_CRV_flags$Plate), paste0("R2 Good", "-", STD_CRV_flags$Plate))

```



```{r Plot standard graphs, echo = FALSE, warning = FALSE, message = FALSE, fig.width=8, fig.height=12}

##Plot std graphs together
ggarrange(slope1,int1,Rsq1, nrow=2, ncol=2)

#pick the best std curve
std_curve<-qplate %>% filter(Intercept<(mean(qlog$Intercept)+ (2*sd(qlog$Intercept)))&Intercept>(mean(qlog$Intercept)- (2*sd(qlog$Intercept))))%>% filter(Slope<(mean(qlog$Slope)+ (2*sd(qlog$Slope)))&Slope>(mean(qlog$Slope)- (2*sd(qlog$Slope)))) %>% filter(R2 == max(R2))

knit_print(std_curve)
#stop knit if no Std curve satisfies the conditions
if (nrow(std_curve)==0) {
  stop("Error: No Std Curve is good -Reassess. Exiting early.")
}
```

## Remove High CV Check Standards
```{r}
H2S_Chk_Stds <- dat %>% filter(IDs %like% "ChkStd")

#Add in STD values
H2S_chk_stds <- full_join(H2S_Chk_Stds, chk_standards, by = join_by("IDs"))

#Calculate CVs
H2S_CHK_STD_cv <- H2S_chk_stds  %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_CHK_STD_cv_cor <- H2S_CHK_STD_cv %>% mutate(Mean_diff = abs(Abs - Mean_Abs)) %>% group_by(Plate, IDs) %>% mutate(Abs_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Abs), Abs))

H2S_CHK_STD_cv_new <- H2S_CHK_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>% mutate(Abs_SD2 = sd(Abs_rm, na.rm = TRUE)) %>% mutate(CV2 = (Abs_SD2 /Mean_Abs2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_CHK_STD_CV_fin <- H2S_CHK_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate,                                                                                                      Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Abs,
                                                CV_Flag) 

```

##Compare Check Standards to Standards

```{r, echo = FALSE}
H2S_CHKs_a <- H2S_CHK_STD_CV_fin %>%
  mutate(Std_ID = ifelse(
    IDs == "ChkStd 0",
    paste0("Std 0", Plate),
    ifelse(
      IDs == "ChkStd 3",
      paste0("Std 3", Plate),
      ifelse(IDs == "ChkStd 4", paste0("Std 4", Plate), "ID Not Found")
    )
  )) %>%
  mutate(
    Chk_Abs = Abs,
    Chk_Removed = Removed,
    Chk_IDs = IDs,
    Chk_Mean_Abs = Mean_Abs,
    Chk_CV_Flag = CV_Flag
  ) %>% dplyr::select(Std_ID,
                      Chk_Abs,
                      Chk_Removed,
                      Chk_IDs,
                      Chk_Mean_Abs,
                      Chk_CV_Flag,
                      Plate) %>% filter(Chk_Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Chk_CI_low = t.test(Chk_Abs)$conf.int[1],
    Chk_CI_high = t.test(Chk_Abs)$conf.int[2]
  ) %>%
  ungroup() %>%
  dplyr::select(Std_ID, Chk_IDs, Chk_CI_low, Chk_CI_high, Chk_CV_Flag, Chk_Mean_Abs, Plate) %>% distinct(.keep_all = TRUE)

H2S_CHKs_1 <- H2S_CHKs_a %>% dplyr::select(-Plate, -Chk_Mean_Abs)

H2S_STD_forCHK_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 0", "Std 3", "Std 4")) %>% mutate(Std_ID = ifelse(
  IDs == "Std 0",
  paste0("Std 0", Plate),
  ifelse(
    IDs == "Std 3",
    paste0("Std 3", Plate),
    ifelse(IDs == "Std 4", paste0("Std 4", Plate), "ID Not Found")
  )
)) %>% filter(Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Std_CI_low = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[1]),
    Std_CI_high = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Plate, Std_ID, IDs, Std_CI_low, Std_CI_high, CV_Flag, Mean_Abs) %>% distinct(.keep_all = TRUE)

H2S_STD_forCHK <- H2S_STD_forCHK_a %>% dplyr::select(-Mean_Abs)

Chk_Std_join <- full_join(H2S_CHKs_1, H2S_STD_forCHK, by = join_by(Std_ID))

Chk_Std_flags <- Chk_Std_join %>% mutate(
  Chk_Flag_Std = ifelse(
    Chk_CI_low <= Std_CI_high |
      Std_CI_low <= Chk_CI_high,
    "Chk_Good",
    "Chk_Bad"
  )
) %>% mutate(CHK_Flag = paste0(gsub(" ", "", IDs), Chk_Flag_Std)) %>%
  mutate(CHK_CV_Flag = paste0("Chk",gsub(" ", "", IDs), Chk_CV_Flag))

Chk_Flags <- Chk_Std_flags %>% group_by(Plate) %>% mutate(Chk_Flag = ifelse(any(CHK_Flag %like% "Bad"), "Chk_Bad", "Chk_Good"), Chk_CV_Flag = ifelse(any(CHK_CV_Flag %like% "High"), "Chk_CV_High", "Chk_CV_Good")) %>% ungroup() %>% mutate(CHK_CV_Flags = ifelse(CHK_CV_Flag %like% "High", paste0(gsub(" ", "", IDs), Chk_CV_Flag), "Good")) %>% mutate(CHKs_Flag = ifelse(CHK_Flag %like% "Bad", CHK_Flag, "Good")) 


H2S_CHKs_plot <- H2S_CHKs_a %>% rename(IDs = Chk_IDs, Std_CI_low = Chk_CI_low, Std_CI_high = Chk_CI_high, Mean_Abs = Chk_Mean_Abs) %>% mutate(Type = "Check") %>% dplyr::select(-Chk_CV_Flag)

H2S_Std_chkplot <- H2S_STD_forCHK_a %>% dplyr::select(IDs, Std_CI_low, Std_CI_high, Mean_Abs, Std_ID, Plate) %>% mutate(Type = "Standard")

Chk_flags_plot <- Chk_Flags %>% dplyr::select(Std_ID, CHKs_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(Check_Flag = ifelse(CHKs_Flag == "Good", "Good", "Bad"))

H2S_Chks_bind <- rbind(H2S_CHKs_plot, H2S_Std_chkplot)

H2S_Chks_bind_1 <- left_join(H2S_Chks_bind, Chk_flags_plot, by = join_by(Std_ID)) 

H2S_Chks_plot_bind <- H2S_Chks_bind_1 %>% mutate(Plot_ID = ifelse(IDs %like% "Chk", substr(IDs, 4, 8), IDs))
 

H2S_chk_plot <- ggplot(data = H2S_Chks_plot_bind, aes(x = Plot_ID, y = Mean_Abs, col = Type, fill = Check_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  geom_errorbar(aes(ymin = Std_CI_low, ymax = Std_CI_high), stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  facet_wrap(.~Plate) +
  scale_color_manual(values =c("Standard" = "lightgreen", "Check" = "cornflowerblue" )) +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Check Stds: H2S") +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1))


H2S_chk_plot

```

##Display Any Check Flags
```{r}
Chk_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHKs_Flag) %>% filter(CHKs_Flag != "Good")

CHK_CV_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags) %>% filter(CHK_CV_Flags != "Good")

ifelse(length(Chk_Flag_1$CHKs_Flag) >= 1, paste0(Chk_Flag_1$CHKs_Flag, "-", Chk_Flag_1$Plate), paste0("All Checks Good"))

ifelse(length(CHK_CV_Flag_1$CHK_CV_Flags) >= 1, paste0(CHK_CV_Flag_1$CHK_CV_Flags, "-", CHK_CV_Flag_1$Plate), paste0("All Checks CV Good"))

```


## Remove High CV Matrix
```{r}
H2S_Mtx_Stds <- dat %>% filter(IDs %like% "ppt")

#Add in STD values
H2S_mtx_stds <- full_join(H2S_Mtx_Stds, mtx_standards, by = join_by("IDs"))

#Calculate CVs
H2S_MTX_STD_cv <- H2S_mtx_stds  %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

H2S_MTX_STD_cv_cor <- H2S_MTX_STD_cv %>% mutate(Mean_diff = abs(Abs - Mean_Abs)) %>% group_by(Plate, IDs) %>% mutate(Abs_rm = ifelse(CV_Flag == "CV_High", ifelse(max(Mean_diff) == Mean_diff, NA, Abs), Abs))

H2S_MTX_STD_cv_new <- H2S_MTX_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE)) %>% mutate(Abs_SD2 = sd(Abs_rm, na.rm = TRUE)) %>% mutate(CV2 = (Abs_SD2 /Mean_Abs2) * 100) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_MTX_STD_CV_fin <- H2S_MTX_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate,                                                                                                      Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Abs,
                                                CV_Flag) 

```

##Compare Matrix to Standards

```{r, echo = FALSE}
H2S_MTXs_a <- H2S_MTX_STD_CV_fin %>%
  mutate(Std_ID = paste0(gsub(" ", "",Plate), "MTX_STD_5")) %>%
  mutate(
    Mtx_Abs = Abs,
    Mtx_Removed = Removed,
    Mtx_IDs = IDs,
    Mtx_Mean_Abs = Mean_Abs,
    Mtx_CV_Flag = CV_Flag
  ) %>% dplyr::select(Std_ID,
                      Mtx_Abs,
                      Mtx_Removed,
                      Mtx_IDs,
                      Mtx_Mean_Abs,
                      Mtx_CV_Flag,
                      Plate) %>% filter(Mtx_Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Mtx_CI_low = t.test(Mtx_Abs)$conf.int[1],
    Mtx_CI_high = t.test(Mtx_Abs)$conf.int[2]
  ) %>%
  ungroup() %>%
  dplyr::select(Std_ID, Mtx_IDs, Mtx_CI_low, Mtx_CI_high, Mtx_CV_Flag, Mtx_Mean_Abs, Plate) %>% distinct(.keep_all = TRUE)

H2S_MTXs_1 <- H2S_MTXs_a %>% dplyr::select(-Plate, -Mtx_Mean_Abs)

H2S_STD_forMTX_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 5")) %>% mutate(Std_ID = paste0(gsub(" ", "",Plate),"MTX_STD_5")) %>% filter(Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Std_CI_low = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[1]),
    Std_CI_high = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Plate, Std_ID, IDs, Std_CI_low, Std_CI_high, CV_Flag, Mean_Abs) %>% distinct(.keep_all = TRUE)

H2S_STD_forMTX <- H2S_STD_forMTX_a %>% dplyr::select(-Mean_Abs)

Mtx_Std_join <- full_join(H2S_MTXs_1, H2S_STD_forMTX, by = join_by(Std_ID))

Mtx_Std_flags <- Mtx_Std_join %>% mutate(
  Mtx_Flag_Std = ifelse(
    Mtx_CI_low <= Std_CI_high |
      Std_CI_low <= Mtx_CI_high,
    "Mtx_Good",
    "Mtx_Bad"
  )
) %>% mutate(MTX_Flag = paste0(gsub(" ", "", IDs), Mtx_Flag_Std)) %>%
  mutate(MTX_CV_Flag = paste0("Mtx",gsub(" ", "", IDs), Mtx_CV_Flag))

Mtx_Flags <- Mtx_Std_flags %>% group_by(Plate) %>% mutate(Mtx_Flag = ifelse(any(MTX_Flag %like% "Bad"), "Mtx_Bad", "Mtx_Good"), Mtx_CV_Flag = ifelse(any(MTX_CV_Flag %like% "High"), "Mtx_CV_High", "Mtx_CV_Good")) %>% ungroup() %>% mutate(MTX_CV_Flags = ifelse(MTX_CV_Flag %like% "High", paste0(gsub(" ", "", IDs), Mtx_CV_Flag), "Good")) %>% mutate(MTXs_Flag = ifelse(MTX_Flag %like% "Bad", MTX_Flag, "Good")) 


H2S_MTXs_plot <- H2S_MTXs_a %>% rename(IDs = Mtx_IDs, Std_CI_low = Mtx_CI_low, Std_CI_high = Mtx_CI_high, Mean_Abs = Mtx_Mean_Abs) %>% mutate(Type = "Matrix Check") %>% dplyr::select(-Mtx_CV_Flag)

H2S_Std_mtxplot <- H2S_STD_forMTX_a %>% dplyr::select(IDs, Std_CI_low, Std_CI_high, Mean_Abs, Std_ID, Plate) %>% mutate(Type = "Standard")

Mtx_flags_plot <- Mtx_Flags %>% dplyr::select(Std_ID, MTXs_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(Matrix_Flag = ifelse(MTXs_Flag == "Good", "Good", "Bad"))

H2S_Mtxs_bind <- rbind(H2S_MTXs_plot, H2S_Std_mtxplot)

H2S_Mtxs_bind_1 <- left_join(H2S_Mtxs_bind, Mtx_flags_plot, by = join_by(Std_ID)) 

H2S_Mtxs_plot_bind <- H2S_Mtxs_bind_1 %>% mutate(Plot_ID = ifelse(IDs %like% "Mtx", substr(IDs, 4, 8), IDs), Plot_ID = ifelse(Plot_ID == "Std 5", Plot_ID, "10ppt Std 5"))
 

H2S_mtx_plot <- ggplot(data = H2S_Mtxs_plot_bind, aes(x = Plot_ID, y = Mean_Abs, col = Type, fill = Matrix_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  geom_errorbar(aes(ymin = Std_CI_low, ymax = Std_CI_high), stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  facet_wrap(.~Plate) +
  scale_color_manual(values =c("Standard" = "lightgreen", "Matrix Check" = "cornflowerblue" )) +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Matrix Effects Checks: H2S") +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1))


H2S_mtx_plot

```


##Display Any Matrix Flags
```{r}
Mtx_Flag_1 <- Mtx_Flags %>% dplyr::select(Plate, MTXs_Flag) %>% filter(MTXs_Flag != "Good")

MTX_CV_Flag_1 <- Mtx_Flags %>% dplyr::select(Plate, MTX_CV_Flags) %>% filter(MTX_CV_Flags != "Good")

ifelse(length(Mtx_Flag_1$MTXs_Flag) >= 1, paste0(Mtx_Flag_1$MTXs_Flag, "-", Mtx_Flag_1$Plate), paste0("All Matrix Checks Good"))

ifelse(length(MTX_CV_Flag_1$MTX_CV_Flags) >= 1, paste0(MTX_CV_Flag_1$MTX_CV_Flags, "-", MTX_CV_Flag_1$Plate), paste0("All Matrix CV Good"))

```


## Calculate Sulfide Concentrations & Add Concentration Flags
```{r Calculate Sulfide Concentrations,echo=FALSE,include=FALSE}

High_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 5") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(High_DL = Mean_Abs)

Low_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 1") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(Low_DL = Mean_Abs)

#filter out stds and other experiments
raw_data <- dat %>%  
      filter(!str_detect(IDs, "Std")) %>%  
      filter(!str_detect(IDs, "MC")) %>%  
      filter(str_detect(IDs, "L") | str_detect(IDs, "l"))

Slopes_calc <- Slope_H2S_Std_Crvs_rm %>% dplyr::select(Plate, Slope, Intercept)

joined_data <- right_join(Slopes_calc, raw_data, by = join_by(Plate))

calc_data <- joined_data %>% mutate(Conc = (Slope*Abs) + Intercept) %>% mutate(Conc = ifelse(Conc <= 0, 0, Conc))

calc_data_1 <- right_join(High_Std, calc_data, by = join_by(Plate))

dl_data <- right_join(Low_Std, calc_data_1, by = join_by(Plate))

data_dl_flags <- dl_data %>% mutate(Conc_Flag = ifelse(Abs < Low_DL, "bdl", ifelse(Abs > High_DL, "adl", "Within_Range")))
```

  
```{r Calculate Averages across wells and sd,echo=FALSE}

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1 <- dat %>%
  group_by(IDs) %>%
  summarise(H2S_mean = mean(H2S_Conc_Final), H2S_sd = sd(H2S_Conc_Final), H2S_cv = cv(H2S_Conc_Final),
            Plate=first(Plate), 
            QAQC_flag = first(QAQC_flag),
            Dilution = first(Dilution))

#Flag high cvs 
dat1$H2S_cv_flag <- ifelse(dat1$H2S_cv > cv_flag_sample, 'High CV', 'within range')

# to flag high dilutions 
dat1$H2S_flag <- ifelse(dat1$H2S_mean <=0, 'bdl', ifelse(dat1$H2S_mean > ((mean(std5$Abs)-std_curve$Intercept)/std_curve$Slope)*(dat1$Dilution), "adl", "Within_Range"))

head(dat1)

#plot data and sd's just to check and see what they look like - just a quick first look 
H2S <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color=H2S_flag))+ 
  geom_point(size=4)+ 
       scale_color_manual(values = c("abl" = "red", "bdl" = "orange", "Within_Range" = "black")) +  theme_classic() + 
  labs(y="Sulfide (uM)", x="Sample ID",title=paste(Project, Experiment," Sulfide Data")) + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
H2S
#quick look at high cvs
H2S_original <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color = H2S_cv_flag))+ 
  geom_point(size=4) +  theme_classic() + 
  scale_color_manual(values = c("Within range" = "darkgreen", "High CV" = "darkred"))+
  labs(y="Sulfide (uM)", x="Sample ID") + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70)) + 
  ggtitle("Sample triplicate means and sd dev before bad reps removed") 
H2S_original

```


## Remove bad reps

```{r remove bad reps,echo=FALSE}

##auto remove bad reps
# filter High CV Samples
dat1_HCV <- dat1 %>%  
  filter(str_detect(H2S_cv_flag, "High CV")) 
dat_HCV <- subset(dat, (IDs %in% dat1_HCV$IDs ))
#make dataframe for fixed data
dat_HCV_fixed<-data.frame(matrix(ncol = 8, nrow =0 ))
colnames(dat_HCV_fixed)<-c("Wells", "Abs", "IDs", "Dilution", "QAQC_flag", "Conc", "H2S_Conc_Final", "H2S_info")

#make a loop to filter each high cv sample
for(x in 1:nrow(dat1_HCV)){ 
  sing_ID<- dat_HCV %>% subset(IDs == dat1_HCV$IDs[x])
  #make a dataframe for cvs
  cv_trial<-data.frame(matrix(ncol = 3, nrow =1 ))
        
  #make a loop to find the cv for each deleted point
  for(i in 1:nrow(sing_ID)){
    sing_ID1<-sing_ID[-i,] 
    cv_trial[i]<-cv(sing_ID1$Abs)
  }
        
  # delete the point that gives the lowest cv
  sing_ID<-sing_ID[-which.min(cv_trial),]
  dat_HCV_fixed<-rbind(dat_HCV_fixed,sing_ID)
}

#summarize by sampleID so that we can calculate the mean and std. dev. of the three wells 
dat1_HCV_fixed <- dat_HCV_fixed %>%
  group_by(IDs) %>%
  summarise(H2S_mean = mean(H2S_Conc_Final), H2S_sd = sd(H2S_Conc_Final), H2S_cv = cv(H2S_Conc_Final),
            Plate=first(Plate), 
            QAQC_flag = first(QAQC_flag),
            Dilution = first(Dilution))

#Flag high cvs 
dat1_HCV_fixed$H2S_cv_flag <- ifelse(dat1_HCV_fixed$H2S_cv > cv_flag_sample, 'High CV', 'within range')

# to flag high dilutions 
dat1_HCV_fixed$H2S_flag <- ifelse(dat1_HCV_fixed$H2S_mean <=0, 'bdl', ifelse(dat1_HCV_fixed$H2S_mean > ((mean(std5$Abs)-std_curve$Intercept)/std_curve$Slope)*(dat1_HCV_fixed$Dilution), "adl", "Within_Range"))

#recombine data
dat1 <- subset(dat1, (!IDs %in% dat1_HCV_fixed$IDs ))
dat1 <- rbind(dat1_HCV_fixed, dat1)
head(dat1)


#plot data and sd's just to check and see what they look like - just a quick first look 
H2S <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color=H2S_flag))+ 
  geom_point(size=4)+ 
       scale_color_manual(values = c("abl" = "red", "bdl" = "orange", "Within_Range" = "black")) +  theme_classic() + 
  labs(y="Sulfide (uM)", x="Sample ID",title=paste(Project, Experiment," Sulfide Data")) + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70))
H2S

H2S_fixed <- ggplot(dat1, aes(x=IDs, y=H2S_mean, color = H2S_cv_flag))+ 
  geom_point(size=4) +  theme_classic() + 
  scale_color_manual(values = c("Within range" = "darkgreen", "High CV" = "darkred"))+
  labs(y="Sulfide (uM)", x="Sample ID") + 
  geom_errorbar(aes(ymin=H2S_mean-H2S_sd,
                    ymax=H2S_mean+H2S_sd),width=0.3,position=position_dodge(.1))+
  guides(x = guide_axis(angle = 70)) + 
  ggtitle("Sample triplicate means and sd dev after bad reps removed") 
H2S_fixed

#If samples with cv >10 rerun those samples 
H2S_HighCVRerun <- subset(dat1, H2S_cv_flag == "High CV")

#filter the samples that need to be run at a higher dilution 
H2S_DiluteRerun <- subset(dat1, H2S_flag == "adl")
ADL<-subset(dat1 ,(IDs %in% H2S_DiluteRerun$IDs))
ADL<-filter(ADL,H2S_flag=="Within_Range") 
H2S_DiluteRerun<-H2S_DiluteRerun %>% 
  subset((!IDs %in% ADL$IDs)) %>% 
  subset((!IDs %like% " Dup")) %>% 
  subset((!IDs %like% " Spike"))
#filter the samples that need to be run at a lower dilution 
H2S_bdl <- subset(dat1,H2S_flag == "bdl"& Dilution!=1)
BDL<-subset(dat1 ,(IDs %in% H2S_bdl$IDs))
BDL<-filter(BDL,H2S_flag=="Within_Range") 
H2S_bdl<-H2S_bdl %>% 
  subset((!IDs %in% BDL$IDs)) %>% 
  subset((!IDs %like% " Dup")) %>% 
  subset((!IDs %like% " Spike"))

knitr::kable(H2S_HighCVRerun, caption = "High CV Samples")
knitr::kable(H2S_DiluteRerun, caption = "Samples Above the Detection Limit")
knitr::kable(H2S_bdl, caption = "Samples Below the Detection Limit")


```

## Check the dups for QAQC

```{r,echo=FALSE}

#Show me the data that we have from the calculations 
#head(dat1)

#pull out any rows that have "Dup" in the ID column
dups <- dat1 %>%  
  filter(str_detect(IDs, "Dup")) 
#head(dups)

#remove these from dat1
dat2 <- dat1 %>%  
  filter(!str_detect(IDs, "Dup")) %>%  
  filter(!str_detect(IDs, "Spike")) 
#head(dat2)

#remove the dup from these IDs so we will have duplicates 
dups$IDs<-gsub(" Dup",'',dups$IDs)
#head(dups)
colnames(dups) <- c('IDs', 'mean_dup')


#put it back together with the old data set and look for duplicates 
QAdups <- merge(dat2, dups)
#head(QAdups)

QAdups$dups_chk <- ((abs(QAdups$H2S_mean-QAdups$mean_dup))/((QAdups$H2S_mean+QAdups$mean_dup)/2))*100
QAdups$dups_chk <- ifelse(!is.na(QAdups$dups_chk),QAdups$dups_chk,0 )
QAdups$dups_flag <-  ifelse(QAdups$dups_chk <dups_perc_diff, 'OK', 'Rerun')

#head(QAdups)

#plot dups output as a bar graph to easily check - want any over 10% to be red need to work on this 
dupsbar<-ggplot(data = QAdups, aes(x = IDs, y = dups_chk, fill=dups_flag)) +
          geom_bar(stat = 'identity') + facet_wrap(~Plate)+
          scale_fill_manual(values=c("OK" ='darkgreen' , "Rerun" = 'darkred')) + 
          theme_classic() + labs(x= "Sample ID", y="Difference Between Duplicates (%)") + 
          theme(legend.position="none") +  geom_hline(yintercept=10, linetype="dashed", color = "black", linewidth=1)

dupsbar

#check for any one's that would warrant reruns 
Baddups <- subset(QAdups, dups_flag == "Rerun")

#write out a flag to the sample dataframe if any dups have percent differences out of range
for(i in 1:length(plates)){
  Baddups1<-  subset(Baddups,Plate %in% plates[i])
  dat$QAQC_flag<-if(nrow(Baddups1) > 0){
    ifelse(dat$Plate %in% qplate$Plate[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Dup perc diff out of range"),"Dup perc diff out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
}}

for(i in 1:length(plates)){
  Baddups1<-  subset(Baddups,Plate %in% plates[i])
  dat1$QAQC_flag<-if(nrow(Baddups1) > 0){
    ifelse(dat1$Plate %in% qplate$Plate[i], ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, "; Dup perc diff out of range"),"Dup perc diff out of range"),paste0( dat1$QAQC_flag,"") )                   
  }else{
    ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, ""),"")
}}

dups_percent <- (sum(QAdups$dups_flag == "OK")/nrow(QAdups))*100
#report out if flags indicate need for rerun
ifelse(dups_percent >= 60  , ">60% of Duplicates are within <10%",
       "<60% of Duplicates are within <10% - REASSESS")

```

## Check the spks for QAQC

```{r,echo=FALSE}

#pull out any rows that have "d" in the SampleID column
spks <- dat1 %>%  filter(str_detect(IDs, "Spike")) 

#remove the Spike from these IDs so we will have duplicates 
spks$IDs<-gsub(" Spike","",spks$IDs)
#head(spks)
colnames(spks) <- c('IDs', 'mean_spk')


#put it back together with the old data set and look for duplicates 
QAspks <- merge(dat2, spks)
#head(QAspks)

#now we need to calculate the spike concentration and calculate the spike recovery 
#spike for these samples was 10 uL of the 100uM standard
QAspks$unspkd <- (QAspks$H2S_mean/QAspks$Dilution)*((sample_vol )/Con1) #gives us the total S2- in the sample in umoles
QAspks$spkd <-    (QAspks$mean_spk/QAspks$Dilution)*((sample_vol + spkvol)/Con1)        ##total S2- in spiked sample in umoles 
QAspks$expctd_spkd <-  (QAspks$unspkd + spk_Conc)
QAspks$spk_recovery <-    (QAspks$spkd/QAspks$expctd_spkd)*100
QAspks$spks_flag <-  ifelse(QAspks$spk_recovery > low_recovery_cutoff & QAspks$spk_recovery < high_recovery_cutoff, 'OK', 'NO, rerun')  
 

#head(QAspks)

#plot spk recoveries output as a bar graph to easily check - want any over 10% to be red need to work on this 
spksbar <- ggplot(data = QAspks, aes(x = IDs, y = spk_recovery, fill=spks_flag)) +facet_wrap(~Plate)+
        geom_bar(stat = 'identity') + 
        scale_fill_manual(values=c("OK"="darkgreen","NO, rerun"="darkred")) + 
        theme_classic() + labs(x= "Sample ID", y="Spike Recovery (%)") + 
        theme(legend.position="none") +  
        geom_hline(yintercept=low_recovery_cutoff, linetype="dashed", color = "black", linewidth=1) + 
        geom_hline(yintercept=high_recovery_cutoff, linetype="dashed", color = "black", linewidth=1)

spksbar

#check for any no's that would warrant reruns! 

Badspks <- subset(QAspks, spks_flag == "NO, rerun")

#write out a flag to the sample dataframe if any spks have recovery out of range 
for(i in 1:length(plates)){
  Badspks1<-  subset(Badspks,Plate %in% plates[i])
  dat$QAQC_flag<-if(nrow(Badspks1) > 0){
    ifelse(dat$Plate %in% qplate$Plate[i], ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, "; Spk recovery out of range"),"Spk recovery out of range"),paste0( dat$QAQC_flag,"") )                   
  }else{
    ifelse(dat$QAQC_flag != "", paste0(dat$QAQC_flag, ""),"")
}}

for(i in 1:length(plates)){
  Badspks1<-  subset(Badspks,Plate %in% plates[i])
  dat1$QAQC_flag<-if(nrow(Badspks1) > 0){
    ifelse(dat1$Plate %in% qplate$Plate[i], ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, "; Spk recovery out of range"),"Spk recovery out of range"),paste0( dat1$QAQC_flag,"") )                   
  }else{
    ifelse(dat1$QAQC_flag != "", paste0(dat1$QAQC_flag, ""),"")
}}

spks_percent <- (sum(QAspks$spks_flag == "OK")/nrow(QAspks))*100
#report out if flags indicate need for rerun
ifelse(spks_percent >= 60  , ">60% of Spikes are within range",
       "<60% of Spikes are out of range - REASSESS")
```

## Export full data then just final data 
```{r,echo=FALSE,include=FALSE}
dat_IDs <- dat1

#Read out the summarized data
head(dat_IDs)
write.csv(dat_IDs, file= summarizeddata_path)

#Read out all the data in dat 
head(dat)
write.csv(dat, file= fulldata_path)

#Now take out the absorbance and stuff 
head(dat)
dat3 <- dat[ ,c(1,4,6,8)]
head(dat3)
write.csv(dat3, file= shortdata_path)

write.csv(H2S_DiluteRerun, Need_dilute)
write.csv(H2S_HighCVRerun, high_CV)

#save to QAQC Data to QAQC File
#write.csv(qlog, file= log_path)


```

### END
