---
title: "CO2xN: Porewater Sulfide"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
    number_sections: false
geometry: left=2cm,right=2cm,top=1cm,bottom=2cm
---
## Code Set up 
```{r setup, include=FALSE}

library(dplyr)
library(broom)
library(ggplot2)
library(ggpubr)
library(stringr)
library(purrr)
library(tidyverse)
library(here)
library(data.table)
library(matrixStats)
library(gridExtra)
library(grid)
library(plater)
library(raster)
library(knitr)  
library(readxl)


```

## run information
```{r Information to be changed}

##Before or After Reruns (YES or NO)
After_Reruns = "YES"
#If NO will not write final data file
  #And Will Write a Rerun File

###things that need to be changed

sample_month = "May"
sample_year = "2025"
Run_Date = as.Date("05/27/2025", format = "%m/%d/%Y")
Run_by = "Melanie Giessner"#Instrument user
Script_run_by ="Brenden Blakley"#Code user

Run_notes=""#any notes from run

#Import Plate Data

plates<- c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate_rr")

Raw_plates <- c("Raw Data/20250527_LTREB_H2S_Plate1.csv","Raw Data/20250527_LTREB_H2S_Plate2.csv","Raw Data/20250527_LTREB_H2S_Plate3.csv","Raw Data/20250527_LTREB_H2S_Plate4.csv","Raw Data/20250527_LTREB_H2S_Plate5.csv","Raw Data/20250527_LTREB_H2S_Plate6.csv","Raw Data/20250527_LTREB_H2S_Plate7.csv","Raw Data/20250602_H2S_LTREBrr.csv")

#Import Run Date for Each Plate for Std curve qaqc
Run_dates <- tibble(
    Plate = c("Plate1","Plate2","Plate3","Plate4","Plate5","Plate6","Plate7","Plate_rr"),
    Run_Date = c("05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","05/27/2025","06/02/2025")
  )

  Run_dates <- Run_dates %>% mutate(Run_Date = as.Date(Run_Date, format = "%m/%d/%Y"))


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Processed Data/LTREB_2M_H2S_QAQC_2025.csv"

# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_CO2xN_Porewater_H2S_202505.csv"
rerun_path <- "Processed Data/GCReW_CO2xN_Porewater_H2S_202505_reruns.csv"

#Stds that should be excluded
#stds_to_remove<-data.frame(Plate=c("Plate4","STD2"),IDs=c("Std 4","Std 3"))


```

```{r Set Up Code - constants and QAQC cutoffs, include=FALSE}

#Flag cutoffs
  r2_cutoff = 0.985            #this is the level below which we want to rerun or consider a curve 
  cv_flag_stds = 10           #this is the maximum cv allowed for standards
  p_value_chkstds = 0.05      #the p-value for the t-test between check standards must be greater than this
  p_value_MC = 0.05           #the p-value for the t-test between matrix checks and top std must be greater than this
  cv_flag_sample = 10         #this is the maximum cv allowed for samples
  dups_perc_diff = 15.5       #this is the maximum percent difference allowed between duplicates
  high_recovery_cutoff = 120  #this is the maximum percent recovery of SO4 allowed in spiked samples
  low_recovery_cutoff = 80    #this is the minimum percent recovery of SO4 allowed in spiked samples

#Standard concentrations - Update if running different standard curve:
# standard units are in uM
  standards <- tibble(
    IDs = c("Std 0", "Std 1", "Std 2", "Std 3", "Std 4", "Std 5"),
    Conc = c(0, 5.0, 12.5, 25.0, 50.0, 100.0)
  )  #uM
  
    chk_standards <- tibble(
    IDs = c("ChkStd 0", "ChkStd 3", "ChkStd 4"),
    Conc = c(0, 25.0, 50.0)
  )  #uM
    
    mtx_standards <- tibble(
    IDs = c("MC: 10ppt S5"),
    Conc = c(100)
  )  #uM

#Spike concentration calc
  #spike for these samples was 50 uL of the 100uM standard
  Con1 <- 1000000               #conversion factor value for spike volumes (uL -> L)
  spk_std <- 100                # uM S2- standard used
  spkvol <- 50                  # uL volume of spike added
  spk_Conc <- spk_std*(spkvol/Con1)  # umoles of S2- added to each spiked sample
  sample_vol <- 250             # the sample volume without the spike is 250 uL
  samp_vol_a <- 300             # the sample volume 

#Top standard Concentration- Update if running different standard curve: 
   Top_STD = 100
   
   
#Metadata file path
Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata

raw_metadata <- read.csv(Raw_Metadata)
#clean up columns in metadata file:
#Pull out only SMARTX Rows
#Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("2-Meter", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  ) %>%
  dplyr::select(-Temp_Treatment) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2_N",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2",
      ifelse(CO2_Treatment == "Ambient" &
               N_Treatment == "Amended"  , "N" , "Ambient")
    )
  )) 
#Bring in the collection date and any field notes if relevant
#collection_metadata <- read.csv(Collection_Dates)
#then we will merge the collection date for SMARTX based on the project and the chambers

```



## read in data
```{r read in plates,echo=FALSE,include=FALSE}

#read in the csv files
dat<-read_plates(
  files = Raw_plates,#list of all file paths
  plate_names = plates ,  #list of plate names          
  well_ids_column = "Wells",    # name to give column of well IDs (optional)
  sep = ","                     # separator used in the csv file (optional)
) %>%
  rename("Abs"=values,
         "IDs"=values.2,
         "Dilution"=values.3)
head(dat)
  
```

#Remove High CV Stds
```{r, echo=FALSE}

#filter for STDs
H2S_stds_raw <- dat %>% subset(IDs %like% "Std") %>% subset(!IDs %like% "Chk")

#Add in STD values
H2S_stds <- left_join(H2S_stds_raw, standards, by = "IDs")

#Calculate CVs and add flags
H2S_STD_cv <- H2S_stds %>% group_by(Plate, IDs) %>% mutate(CV = cv(Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

#calculate new cv getting best cv from points if cv was high
H2S_STD_cv_cor <- H2S_STD_cv %>% group_by(Plate, IDs) %>% mutate(CV_WO1 = cv(Abs[-1]), CV_WO2 = cv(Abs[-2]), CV_WO3 = cv(Abs[-3])) %>% mutate(RMV = ifelse(CV_Flag == "CV_High", ifelse(CV_WO1 == min(CV_WO1, CV_WO2, CV_WO3), "RMV1", ifelse(CV_WO2 == min(CV_WO1, CV_WO2, CV_WO3), "RMV2", ifelse(CV_WO3 == min(CV_WO1, CV_WO2, CV_WO3), "RMV3", "Point Not Found"))), "CV_Good")) %>% mutate(Abs_rm = case_when(
  RMV == "RMV1" ~ c(NA, Abs[2], Abs[3]),
  RMV == "RMV2" ~ c(Abs[1], NA, Abs[3]),
  RMV == "RMV3" ~ c(Abs[1], Abs[2], NA),
  RMV == "CV_Good" ~ c(Abs[1], Abs[2], Abs[3]),
  RMV == "Point Not Found" ~ NA)
  
)


#Calculate new CVs and note if removed
H2S_STD_cv_new <- H2S_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE), CV2 = cv(Abs_rm, na.rm = TRUE)) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()


#organize data
H2S_STD_CV_fin <- H2S_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate, 
                                                Wells,                                                                                                                Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Conc,
                                                Mean_Abs,
                                                CV_Flag)


```


#Check R2 & Remove Standards if Necessary
```{r, echo=FALSE}
split_data <- H2S_STD_CV_fin %>% dplyr::select(Plate, IDs, Conc, Mean_Abs, CV_Flag) %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_H2S_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Conc ~ Mean_Abs, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ Mean_Abs, data = .))[2],
                                        Intercept = coef(lm(Conc ~ Mean_Abs, data = .))[1]))


 Slope_H2S_raw <- bind_rows(Slope_H2S_raw_list)
  
 #check R2 and flag
 Slope_H2S_Std_Crvs <- Slope_H2S_raw %>% dplyr::select(Plate, Curve, R2, Slope, Intercept) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))

 #join standard curve CVs and regression values
 R2_Check_Raw <- left_join(Slope_H2S_Std_Crvs, H2S_STD_CV_fin, by = "Plate")
 
 R2_Check <- R2_Check_Raw %>% 
   dplyr::select(Plate,
                 IDs,
                 Mean_Abs,
                 Conc,
                 R2,
                 Slope,
                 Intercept,
                 CV_Flag,
                 R2_Flag) %>%
   distinct()
 
 #calculate concentrations from standard curve and separate removable points
 R2_Check_diff <- R2_Check %>% mutate(Calc_Conc = ((Slope*Mean_Abs) + Intercept)) %>%
   mutate(Conc_Diff = abs(Conc - Calc_Conc)) %>% mutate(RMVBL = ifelse(IDs %in% c("Std 0", "Std 1","Std 5"), "No", "Yes"))

 R2_Check_RMVBL <- R2_Check_diff %>% filter(RMVBL == "Yes")
 
 R2_Check_NRMVBL <- R2_Check_diff %>% filter(RMVBL != "Yes") %>% mutate(STD_ABS_rm = Mean_Abs) %>% mutate(N_HCV = 0, STD_ABS_rmv = STD_ABS_rm, N_stds = 3)
 
 #remove standard curve point if r2 is bad
 R2_Check_na <- R2_Check_RMVBL %>% group_by(Plate) %>% mutate(STD_ABS_rm = ifelse(R2_Flag == "R2_Bad", ifelse(max(Conc_Diff) == Conc_Diff, NA, Mean_Abs), Mean_Abs)) %>% ungroup()
 
 #remove high cv standard curve point if no other points were removed
 R2_Check_CVrm <- R2_Check_na %>% na.omit() %>% group_by(Plate) %>% mutate(N_stds = length(STD_ABS_rm)) %>% group_by(Plate, CV_Flag) %>%
   mutate(N_HCV = n()) %>% ungroup() %>% group_by(Plate) %>%
   mutate(STD_ABS_rmv = ifelse(
     N_stds == 3,
     ifelse(
       CV_Flag == "CV_High" ,
       ifelse(
         N_HCV == 1 ,
         NA,
         ifelse(max(Conc_Diff) == Conc_Diff, NA, STD_ABS_rm)
       ),
       STD_ABS_rm
     ),
     STD_ABS_rm
   ))
 
  R2_Check_bind <- rbind(R2_Check_NRMVBL,R2_Check_CVrm) %>% na.omit()
 
 R2_Check_fin <- R2_Check_bind[order(R2_Check_bind$Plate,R2_Check_bind$IDs), ]
 
 
 #separate plates for regression
 split_data_rmv <- R2_Check_fin %>% dplyr::select(Plate, IDs, Conc, STD_ABS_rmv, CV_Flag, R2_Flag) %>%
      distinct() %>%
      group_by(Plate) %>%
      group_split()
 
 #get new standard curve values
 Slope_H2S_raw_list_rm <- map(split_data_rmv, ~ .x %>% mutate(Curve = "H2S",
                                        R2 = summary(lm(Conc ~ STD_ABS_rmv, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ STD_ABS_rmv, data = .))[2],
                                        Intercept = coef(lm(Conc ~ STD_ABS_rmv, data = .))[1]))

 Slope_H2S_raw_rm <- bind_rows(Slope_H2S_raw_list_rm)
  
 Slope_H2S_Std_Crvs_rm <- Slope_H2S_raw_rm %>% dplyr::select(Plate, Curve, R2, Slope, Intercept, CV_Flag) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Good", "R2_Bad"))
 
 
 
```



```{r Plot all Standard Curves together, echo=FALSE,message=FALSE,fig.keep='none'}
##this is to see if there are any stds or plates that look off

H2S_Curve <- ggplot(Slope_H2S_raw_rm, aes(STD_ABS_rmv, Conc, color = Plate)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "H2S Standard Curves", x = "Absorbance", y = "Concentration (uM)")
H2S_Curve


STD_run_date <- left_join(Run_dates, Slope_H2S_Std_Crvs_rm)
new_log_stds <- STD_run_date %>% dplyr::select(Plate,
   Curve,
   R2,
   Slope,
   Intercept,
   Run_Date) %>% mutate(R2 = round(R2, digits = 4), Slope = round(Slope, digits = 4), Intercept = round(Intercept, digits = 4), ) %>% mutate(Run = "Current Runs")


log <- read.csv(log_path)
log_2 <- log %>% mutate(Run_Date = as.Date(as.character(Run_Date), format = "%Y-%m-%d"), Run = "Past Runs")
head(log_2)
log_3 <- log_2[, -c(1)]
log_4 <- rbind(log_3, new_log_stds)

unique_log <- log_4 %>% distinct(.keep_all = TRUE)

mean_slope <- mean(unique_log$Slope, na.rm = TRUE)
sd_slope <- sd(unique_log$Slope, na.rm = TRUE)

hi_2sd_slope <- mean_slope + 2*sd_slope

lo_2sd_slope <- mean_slope - 2*sd_slope
  
mean_int <- mean(unique_log$Intercept, na.rm = TRUE)
sd_int <- sd(unique_log$Intercept, na.rm = TRUE)

hi_2sd_int <- mean_int + 2*sd_int

lo_2sd_int <- mean_int - 2*sd_int

Slopes_chk <- ggplot(unique_log, aes(x = Run_Date, y = Slope, col = Run)) +
    ylim(50,125) +
  scale_color_manual(values = c("Past Runs" = "grey27", "Current Runs" = "cyan3")) +
  geom_abline(intercept = mean_slope - 2*sd_slope, slope = 0 )+
  geom_abline(intercept = mean_slope + 2*sd_slope, slope = 0 )+
  geom_point(size = 2) +
  theme_bw()

ints_chk <- ggplot(unique_log, aes(x = Run_Date, y = Intercept, col = Run)) +
    ylim(-15, 0) +
  scale_color_manual(values = c("Past Runs" = "grey27", "Current Runs" = "cyan3")) +
  geom_abline(intercept = mean_int - 2*sd_int, slope = 0 )+
  geom_abline(intercept = mean_int + 2*sd_int, slope = 0 )+
  geom_point(size = 2) +
  theme_bw()

ints_chk

Slopes_chk

write.csv(unique_log, log_path)

stdcrv_filtered <- new_log_stds %>% mutate(In_Range_Slope = ifelse(Slope >= hi_2sd_slope |  Slope <= lo_2sd_slope, "Out of Range", "In Range"), In_Range_Int = ifelse(Intercept >= hi_2sd_int |  Intercept <= lo_2sd_int, "Out of Range", "In Range")) %>% filter(In_Range_Slope == "In Range" & In_Range_Int == "In Range")

mean_slope_fin <- mean(stdcrv_filtered$Slope, na.rm = TRUE)
  
mean_int_fin <- mean(stdcrv_filtered$Intercept, na.rm = TRUE)


```

#Return Flags For Standard Curve
```{r, echo=FALSE}

STD_CRV_flags <- full_join(Slope_H2S_Std_Crvs_rm, Run_dates, by = join_by(Plate))

ifelse(STD_CRV_flags$CV_Flag == "CV_High", paste0("CV High", "-", STD_CRV_flags$Plate), paste0("CV Good", "-", STD_CRV_flags$Plate))

ifelse(STD_CRV_flags$R2_Flag == "R2_Bad", paste0("R2 Bad", "-", STD_CRV_flags$Plate), paste0("R2 Good", "-", STD_CRV_flags$Plate))

```

## Remove High CV Check Standards
```{r, echo=FALSE}
H2S_Chk_Stds <- dat %>% filter(IDs %like% "ChkStd")

#Add in STD values
H2S_chk_stds <- full_join(H2S_Chk_Stds, chk_standards, by = join_by("IDs"))

#Calculate CVs
H2S_CHK_STD_cv <- H2S_chk_stds  %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

#remove replicates to fix CV 
H2S_CHK_STD_cv_cor <- H2S_CHK_STD_cv %>% mutate(Column = substr(Wells, 1,1)) %>% group_by(Plate, IDs, Column) %>% mutate(CV_WO1 = cv(Abs[-1]), CV_WO2 = cv(Abs[-2]), CV_WO3 = cv(Abs[-3])) %>% mutate(RMV = ifelse(CV_Flag == "CV_High", ifelse(CV_WO1 == min(CV_WO1, CV_WO2, CV_WO3), "RMV1", ifelse(CV_WO2 == min(CV_WO1, CV_WO2, CV_WO3), "RMV2", ifelse(CV_WO3 == min(CV_WO1, CV_WO2, CV_WO3), "RMV3", "Point Not Found"))), "CV_Good")) %>% mutate(Abs_rm = case_when(
RMV == "RMV1" ~ c(NA, Abs[2], Abs[3]),
RMV == "RMV2" ~ c(Abs[1], NA, Abs[3]),
RMV == "RMV3" ~ c(Abs[1], Abs[2], NA),
RMV == "CV_Good" ~ c(Abs[1], Abs[2], Abs[3]),
RMV == "Point Not Found" ~ NA)

)

H2S_CHK_STD_cv_new <- H2S_CHK_STD_cv_cor %>% group_by(Plate, IDs, Column) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE), CV2 = cv(Abs_rm, na.rm = TRUE)) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_CHK_STD_CV_fin <- H2S_CHK_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate,                                                                                                      Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Abs,
                                                CV_Flag) 

```

##Compare Check Standards to Standards

```{r, echo = FALSE}

#Tag Check Standards and perform T Test
H2S_CHKs_a <- H2S_CHK_STD_CV_fin %>%
  mutate(Std_ID = ifelse(
    IDs == "ChkStd 0",
    paste0("Std 0", Plate),
    ifelse(
      IDs == "ChkStd 3",
      paste0("Std 3", Plate),
      ifelse(IDs == "ChkStd 4", paste0("Std 4", Plate), "ID Not Found")
    )
  )) %>%
  mutate(
    Chk_Abs = Abs,
    Chk_Removed = Removed,
    Chk_IDs = IDs,
    Chk_Mean_Abs = Mean_Abs,
    Chk_CV_Flag = CV_Flag
  ) %>% dplyr::select(Std_ID,
                      Chk_Abs,
                      Chk_Removed,
                      Chk_IDs,
                      Chk_Mean_Abs,
                      Chk_CV_Flag,
                      Plate) %>% filter(Chk_Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Chk_CI_low = t.test(Chk_Abs)$conf.int[1],
    Chk_CI_high = t.test(Chk_Abs)$conf.int[2]
  ) %>%
  ungroup() %>%
  dplyr::select(Std_ID, Chk_IDs, Chk_CI_low, Chk_CI_high, Chk_CV_Flag, Chk_Mean_Abs, Plate) %>% distinct(.keep_all = TRUE)

H2S_CHKs_1 <- H2S_CHKs_a %>% dplyr::select(-Plate, -Chk_Mean_Abs)

#Get Standard Values and set values for standards with no variance
H2S_STD_forCHK_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 0", "Std 3", "Std 4")) %>% mutate(Std_ID = ifelse(
  IDs == "Std 0",
  paste0("Std 0", Plate),
  ifelse(
    IDs == "Std 3",
    paste0("Std 3", Plate),
    ifelse(IDs == "Std 4", paste0("Std 4", Plate), "ID Not Found")
  )
)) %>% filter(Removed == "KEPT") %>%
  group_by(Std_ID) %>%
  mutate(
    Std_CI_low = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[1]),
    Std_CI_high = ifelse(length(unique(Abs)) == 1, Abs, t.test(Abs)$conf.int[2])
  ) %>%
  ungroup() %>%
  dplyr::select(Plate, Std_ID, IDs, Std_CI_low, Std_CI_high, CV_Flag, Mean_Abs) %>% distinct(.keep_all = TRUE)

H2S_STD_forCHK <- H2S_STD_forCHK_a %>% dplyr::select(-Mean_Abs)

#join check standards with standards
Chk_Std_join <- full_join(H2S_CHKs_1, H2S_STD_forCHK, by = join_by(Std_ID))

#flag data
Chk_Std_flags <- Chk_Std_join %>% mutate(
  Chk_Flag_Std = ifelse(
    Chk_CI_low <= Std_CI_high |
      Std_CI_low <= Chk_CI_high,
    "Chk_Good",
    "Chk_Bad"
  )
) %>% mutate(CHK_Flag = paste0(gsub(" ", "", IDs), Chk_Flag_Std)) %>%
  mutate(CHK_CV_Flag = paste0("Chk",gsub(" ", "", IDs), Chk_CV_Flag))

#Organize Data for Plotting and Flagging
Chk_Flags <- Chk_Std_flags %>% group_by(Plate) %>% mutate(Chk_Flag = ifelse(any(CHK_Flag %like% "Bad"), "Chk_Bad", "Chk_Good"), Chk_CV_Flag = ifelse(any(CHK_CV_Flag %like% "High"), "Chk_CV_High", "Chk_CV_Good")) %>% ungroup() %>% mutate(CHK_CV_Flags = ifelse(CHK_CV_Flag %like% "High", paste0(gsub(" ", "", IDs), Chk_CV_Flag), "Good")) %>% mutate(CHKs_Flag = ifelse(CHK_Flag %like% "Bad", CHK_Flag, "Good")) 


H2S_CHKs_plot <- H2S_CHKs_a %>% rename(IDs = Chk_IDs, Std_CI_low = Chk_CI_low, Std_CI_high = Chk_CI_high, Mean_Abs = Chk_Mean_Abs) %>% mutate(Type = "Check") %>% dplyr::select(-Chk_CV_Flag)

H2S_Std_chkplot <- H2S_STD_forCHK_a %>% dplyr::select(IDs, Std_CI_low, Std_CI_high, Mean_Abs, Std_ID, Plate) %>% mutate(Type = "Standard")

Chk_flags_plot <- Chk_Flags %>% dplyr::select(Std_ID, CHKs_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(Check_Flag = ifelse(CHKs_Flag == "Good", "Good", "Bad"))

H2S_Chks_bind <- rbind(H2S_CHKs_plot, H2S_Std_chkplot)

H2S_Chks_bind_1 <- left_join(H2S_Chks_bind, Chk_flags_plot, by = join_by(Std_ID)) 

H2S_Chks_plot_bind_a <- H2S_Chks_bind_1 %>% mutate(Plot_ID = as.factor(ifelse(IDs %like% "Chk", substr(IDs, 4, 8), IDs))) 

H2S_Chks_plot_bind <-H2S_Chks_plot_bind_a %>% filter(str_detect(Type, "Check"))

Check_Val <- H2S_Chks_plot_bind_a %>% filter(!str_detect(Plot_ID, "Check")) %>% group_by(Plot_ID) %>% mutate(Z = mean(Mean_Abs)) %>% dplyr::select(Plot_ID, Z) %>% distinct(.keep_all = TRUE)
 
#Plot Check Standards
H2S_chk_plot <- ggplot(data = H2S_Chks_plot_bind, aes(x = Plate, y = Mean_Abs, fill = Check_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
      geom_hline(data = Check_Val, aes(yintercept = Z), linetype = "dashed") +
 # annotate(data = Check_Val, "text", aes( x = 1.75, y = (Z + Z*0.10)), label = "Target Arbsorbance") +
  geom_errorbar(aes(ymin = Std_CI_low, ymax = Std_CI_high), stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  facet_wrap(.~Plot_ID) +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Check Stds: H2S with Confidence Interval") +
  theme(legend.position = "bottom", axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1))


H2S_chk_plot

```

##Display Any Check Flags
```{r, echo=FALSE}
Chk_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHKs_Flag) %>% filter(CHKs_Flag != "Good")

CHK_CV_Flag_1 <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags) %>% filter(CHK_CV_Flags != "Good")

ifelse(length(Chk_Flag_1$CHKs_Flag) >= 1, paste0(Chk_Flag_1$CHKs_Flag, "-", Chk_Flag_1$Plate), paste0("All Checks Good"))

ifelse(length(CHK_CV_Flag_1$CHK_CV_Flags) >= 1, paste0(CHK_CV_Flag_1$CHK_CV_Flags, "-", CHK_CV_Flag_1$Plate), paste0("All Checks CV Good"))

```


## Remove High CV Matrix
```{r, echo=FALSE}
H2S_Mtx_Stds <- dat %>% filter(IDs %like% "ppt")

#Add in STD values
H2S_mtx_stds <- full_join(H2S_Mtx_Stds, mtx_standards, by = join_by("IDs"))

#Calculate CVs
H2S_MTX_STD_cv <- H2S_mtx_stds  %>% group_by(Plate, IDs) %>% mutate(Mean_Abs = mean(Abs)) %>% mutate(Abs_SD = sd(Abs)) %>% mutate(CV = (Abs_SD / Mean_Abs) * 100) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

#Remove Replicates to Fix CVs
H2S_MTX_STD_cv_cor <- H2S_MTX_STD_cv %>% group_by(Plate, IDs) %>% mutate(CV_WO1 = cv(Abs[-1]), CV_WO2 = cv(Abs[-2]), CV_WO3 = cv(Abs[-3])) %>% mutate(RMV = ifelse(CV_Flag == "CV_High", ifelse(CV_WO1 == min(CV_WO1, CV_WO2, CV_WO3), "RMV1", ifelse(CV_WO2 == min(CV_WO1, CV_WO2, CV_WO3), "RMV2", ifelse(CV_WO3 == min(CV_WO1, CV_WO2, CV_WO3), "RMV3", "Point Not Found"))), "CV_Good")) %>% mutate(Abs_rm = case_when(
RMV == "RMV1" ~ c(NA, Abs[2], Abs[3]),
RMV == "RMV2" ~ c(Abs[1], NA, Abs[3]),
RMV == "RMV3" ~ c(Abs[1], Abs[2], NA),
RMV == "CV_Good" ~ c(Abs[1], Abs[2], Abs[3]),
RMV == "Point Not Found" ~ NA)

)

#Tag Data with replicate removal
H2S_MTX_STD_cv_new <- H2S_MTX_STD_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Abs2 = mean(Abs_rm, na.rm = TRUE), CV2 = cv(Abs_rm, na.rm = TRUE)) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Abs_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_MTX_STD_CV_fin <- H2S_MTX_STD_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Abs,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Abs2,
                                                   CV_Flag2) %>% 
  mutate(Mean_Abs = Mean_Abs2, CV_Flag = CV_Flag2) %>% 
                                  dplyr::select(Plate,                                                                                                      Abs,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Abs,
                                                CV_Flag) 

```

##Compare Matrix to Standards

```{r, echo = FALSE}

#Organize Matrix Check and T Test
H2S_MTXs_1 <- H2S_MTX_STD_CV_fin %>%
  mutate(Std_ID = paste0(gsub(" ", "",Plate), "MTX_STD_5")) %>%
  dplyr::select(
    Plate,
    Std_ID,
    Abs,
    Removed,
    IDs,
    CV_Flag
  ) %>% mutate(MTX = "YES")


#Get Standatds for Matrix Check
H2S_STD_forMTX_a <- H2S_STD_CV_fin %>% filter(IDs %in% c("Std 5")) %>% mutate(Std_ID = paste0(gsub(" ", "",Plate),"MTX_STD_5")) %>% dplyr::select(Plate, Std_ID,
    Abs,
    Removed,
    IDs,
    CV_Flag) %>% mutate(MTX = "NO") 

H2S_mtxstd_join <- rbind(H2S_MTXs_1, H2S_STD_forMTX_a)

H2S_ttest <- H2S_mtxstd_join %>% group_by(Std_ID) %>% mutate(T_Test = t.test(na.omit(ifelse(MTX == "YES" & Removed == "KEPT", Abs, NA)), na.omit(ifelse(MTX == "NO" & Removed == "KEPT", Abs, NA)), )$p.value) %>% ungroup()

Mtx_Std_flags <- H2S_ttest %>% mutate(MTX_Flag = ifelse(T_Test >= p_value_MC, "Mtx_Good", "Mtx_Bad")) %>% mutate(MTX_CV_Flag = ifelse(MTX == "YES" & CV_Flag == "CV_High", paste0("Mtx_CV_High"), paste0("Mtx_CV_Good")))
  
MTX_plot1 <- Mtx_Std_flags %>% mutate(ID_stat = paste0(Std_ID, IDs)) 

MTX_plota <- MTX_plot1 %>% group_by(ID_stat) %>% filter(Removed != "RMVD") %>% mutate(Mean_Abs = mean(Abs), SD_2 = 2* sd(Abs)) %>% ungroup() %>% dplyr::select(Plate, Mean_Abs, SD_2, IDs, MTX_Flag, Std_ID) %>% distinct() %>% mutate(IDs = ifelse(IDs == "Std 5", "Standard", "Matrix Check"), Matrix_Flag = ifelse(MTX_Flag %like% "Bad", "Bad", "Good"), Plot_ID = "Matrix Check") 

MTX_plot2 <- MTX_plota %>% filter(str_detect(IDs, "Matrix Check"))

MTX_Val <- MTX_plota %>% filter(!str_detect(IDs, "Matrix Check")) %>% group_by(IDs) %>% mutate(Z = mean(Mean_Abs))

Z_mtx = mean(MTX_plota$Mean_Abs)




H2S_mtx_plot <- ggplot(data = MTX_plot2, aes(x = Plate, y = Mean_Abs, fill = Matrix_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  geom_errorbar(aes(ymin = Mean_Abs- SD_2, ymax = Mean_Abs + SD_2), stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  geom_hline(data = MTX_Val, aes(yintercept = Z), linetype = "dashed" ) +
    annotate("text", x = 1.75 , y= (Z_mtx + 0.1 * Z_mtx), label= "Target Absorbance", colour = "Black") + 
  facet_wrap(.~Plot_ID) +
  scale_fill_manual(values = c("Good" = "darkgreen", "Bad" = "darkred")) +
  theme_classic() + labs(x = " ", y = "Absorbance", title =
                           "Matrix Effects Checks with 2 SD: H2S") +
  theme(legend.position = "bottom")


H2S_mtx_plot

```


##Display Any Matrix Flags
```{r, echo=FALSE}

Mtx_Flags_org <- Mtx_Std_flags %>% dplyr::select(Plate, MTX_Flag, MTX_CV_Flag) %>% distinct()

Mtx_Flag_1 <- Mtx_Flags_org %>% dplyr::select(Plate, MTX_Flag) %>% filter(!MTX_Flag %like% "Good")

MTX_CV_Flag_1 <- Mtx_Flags_org %>% dplyr::select(Plate, MTX_CV_Flag) %>% filter(!MTX_CV_Flag %like% "Good")

ifelse(length(Mtx_Flag_1$MTX_Flag) >= 1, paste0(Mtx_Flag_1$MTX_Flag, "-", Mtx_Flag_1$Plate), paste0("All Matrix Checks Good"))

ifelse(length(MTX_CV_Flag_1$MTX_CV_Flag) >= 1, paste0(MTX_CV_Flag_1$MTX_CV_Flag, "-", MTX_CV_Flag_1$Plate), paste0("All Matrix CV Good"))

```


## Calculate Sulfide Concentrations & Add Concentration Flags
```{r Calculate Sulfide Concentrations,echo=FALSE,include=FALSE}

High_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 5") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(High_DL = Mean_Abs)

Low_Std <- H2S_STD_CV_fin %>% filter(IDs == "Std 1") %>% mutate(Mean_Abs = round(Mean_Abs, digits = 4)) %>% dplyr::select(Plate, Mean_Abs) %>% distinct(.keep_all = TRUE) %>% rename(Low_DL = Mean_Abs)

#filter out stds and other experiments
raw_data <- dat %>%  
      filter(!str_detect(IDs, "Std")) %>%  
      filter(!str_detect(IDs, "MC")) %>%  
      filter(str_detect(IDs, "L") | str_detect(IDs, "l"))

#Calculate Concentrations and fix zeroes/negative values and flag conc
Slopes_calc <- Slope_H2S_Std_Crvs_rm %>% dplyr::select(Plate, Slope, Intercept)

joined_data <- right_join(Slopes_calc, raw_data, by = join_by(Plate))

calc_data <- joined_data %>% mutate(Conc = ((mean_slope_fin*Abs) + mean_int_fin)*Dilution) %>% mutate(Conc = ifelse(Conc <= 0, 0, Conc))

calc_data_1 <- right_join(High_Std, calc_data, by = join_by(Plate))

dl_data <- right_join(Low_Std, calc_data_1, by = join_by(Plate))

data_dl_flags <- dl_data %>% mutate(Conc_Flag = ifelse(Abs < Low_DL, "bdl", ifelse(Abs > High_DL, "adl", "Within_Range")))
```

## Remove High CV Samples and Average
```{r, echo=FALSE}

#Calculate CVs
H2S_SPL_cv <- data_dl_flags  %>% group_by(Plate, IDs) %>% mutate(Mean_Conc = mean(Conc)) %>% mutate(Conc_SD = sd(Conc)) %>% mutate(CV = cv(Conc)) %>% mutate(CV_Flag = ifelse(CV >= cv_flag_stds, "CV_High", "CV_Good")) %>% ungroup()

#Remove Replicates to fix CV
H2S_SPL_cv_cor <- H2S_SPL_cv %>% group_by(Plate, IDs) %>% mutate(CV_WO1 = cv(Conc[-1]), CV_WO2 = cv(Conc[-2]), CV_WO3 = cv(Conc[-3])) %>% mutate(RMV = ifelse(CV_Flag == "CV_High", ifelse(CV_WO1 == min(CV_WO1, CV_WO2, CV_WO3), "RMV1", ifelse(CV_WO2 == min(CV_WO1, CV_WO2, CV_WO3), "RMV2", ifelse(CV_WO3 == min(CV_WO1, CV_WO2, CV_WO3), "RMV3", "Point Not Found"))), "CV_Good")) %>% mutate(Conc_rm = case_when(
RMV == "RMV1" ~ c(NA, Conc[2], Conc[3]),
RMV == "RMV2" ~ c(Conc[1], NA, Conc[3]),
RMV == "RMV3" ~ c(Conc[1], Conc[2], NA),
RMV == "CV_Good" ~ c(Conc[1], Conc[2], Conc[3]),
RMV == "Point Not Found" ~ NA)

)

#Tag Data with CV reomved
H2S_SPL_cv_new <- H2S_SPL_cv_cor %>% group_by(Plate, IDs) %>% mutate(Mean_Conc2 = mean(Conc_rm, na.rm = TRUE), Conc_SD2 = sd(Conc_rm, na.rm = TRUE),  CV2 = cv(Conc_rm, na.rm = TRUE)) %>% mutate(CV_Flag2 = ifelse(CV2 >= cv_flag_stds, "CV_High", "CV_Good")) %>% mutate(Removed = ifelse(Conc_rm == "NA", "RMVD", "KEPT")) %>% mutate(Removed = replace_na(Removed, "RMVD")) %>% ungroup()

H2S_SPL_CV_fin <- H2S_SPL_cv_new %>% dplyr::select(Plate,
                                                   Wells,
                                                   Conc,
                                                   Removed,
                                                   IDs,
                                                   Dilution,
                                                   Conc,
                                                   Mean_Conc2,
                                                   Conc_SD2,
                                                   CV_Flag2,
                                                   Conc_Flag) %>% 
  mutate(Mean_Conc = Mean_Conc2, CV_Flag = CV_Flag2, SD_Conc = Conc_SD2) %>% 
                                  dplyr::select(Plate,                                                                                                      Conc,
                                                Removed,
                                                IDs,
                                                Dilution,
                                                Mean_Conc,
                                                SD_Conc,
                                                CV_Flag,
                                                Conc_Flag) 

```

## Plot Samples
```{r, echo=FALSE}

#Filter samples and mark reruns
H2S_Spl_pre_viz <- H2S_SPL_cv_cor %>%
  filter(!str_detect(IDs, "Dup")) %>%
  filter(!str_detect(IDs, "Spike")) %>%
  dplyr::select(Plate, IDs, Mean_Conc, Conc_SD, CV_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(IDs = ifelse(Plate %like% "rr", paste0(IDs, "_RR"), IDs))

#Plot Samples before corrections
H2S_Spl_Plot_Pre <- ggplot(H2S_Spl_pre_viz, aes(y = Mean_Conc, x = IDs, fill = CV_Flag)) +
  geom_bar(stat = 'identity',
           position = position_dodge2(),
           linewidth = 0.75) +
  scale_fill_manual(values = c(
    "CV_Good" = "darkgreen",
    "CV_High" = "darkred"
  )) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  geom_errorbar(
    aes(ymin = Mean_Conc - Conc_SD, ymax = Mean_Conc + Conc_SD),
    width = 0.3,
    position = position_dodge(.1)
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ggtitle("Sample triplicate means and sd dev before bad reps removed")


H2S_Spl_Plot_Pre

H2S_Spl_rmv_viz <- H2S_SPL_CV_fin %>%
  filter(!str_detect(IDs, "Dup")) %>%
  filter(!str_detect(IDs, "Spike")) %>%
  dplyr::select(Plate, IDs, Mean_Conc, SD_Conc, CV_Flag, Conc_Flag) %>% distinct(.keep_all = TRUE) %>% mutate(IDs = ifelse(Plate %like% "rr", paste0(IDs, "_RR"), IDs))

#Plot samples after Corrections
H2S_Spl_Plot_CV <- ggplot(H2S_Spl_rmv_viz, aes(y = Mean_Conc, x = IDs, fill = CV_Flag, col = Conc_Flag)) +
  geom_bar(stat = 'identity',
           position = position_dodge(),
           linewidth = 0.75) +
  scale_fill_manual(values = c(
    "CV_Good" = "darkgreen",
    "CV_High" = "darkred"
  )) +
  scale_color_manual(values = c("Within_Range" = "darkgreen", "adl" = "red", "bdl" = "blue" )) +
  labs(y = "Sulfide (uM)", x = "Sample ID") +
  geom_errorbar(
    aes(ymin = Mean_Conc - SD_Conc, ymax = Mean_Conc + SD_Conc),
    width = 0.3,
    position = position_dodge(.1)
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  ggtitle("Sample triplicate means and sd dev after bad reps removed")


H2S_Spl_Plot_CV



```


## Check the dups for QAQC

```{r,echo=FALSE}

#pull out any rows that have "Dup" in the ID column and tag reruns
DUP_start <- H2S_SPL_CV_fin %>%  mutate(Dup_ID = gsub(" ", "", substr(IDs, 1, 6)),
                                        Dup = ifelse(IDs %like% "Dup", "Dup", "Org")) %>% mutate(Dup_ID = ifelse(Plate %like% "rr", paste0(Dup_ID, "_RR"), Dup_ID))

#separate dups from originals and tag for pairing 

dups_only <- DUP_start %>% filter(Dup == "Dup") %>% dplyr::select(Dup_ID, Mean_Conc, SD_Conc) %>% rename(Dup_Conc = Mean_Conc, Dup_SD = SD_Conc) %>% distinct(.keep_all = TRUE)

no_dups <- DUP_start %>% filter(!Dup == "Dup")

dup_spls_l <- dups_only %>% dplyr::select(Dup_ID) %>% distinct(.keep_all = TRUE)

dup_spls <- no_dups %>% filter(Dup_ID %in% as.vector(dup_spls_l$Dup_ID)) %>% dplyr::select(Dup_ID, Mean_Conc, SD_Conc, Plate, IDs) %>% distinct(.keep_all = TRUE)

dup_join <- full_join(dups_only, dup_spls, by = join_by(Dup_ID))

#Calculate Percent difference and Flag
dup_flags <- dup_join %>% 
  mutate(Diff_P = (abs(100 * (Mean_Conc - Dup_Conc) / mean(c(Mean_Conc, Dup_Conc))))) %>% 
  mutate(Dup_Flag = ifelse(Diff_P >= dups_perc_diff, "Dup_Bad", "Dup_Good"))

#oraganize data for flagging and plotting
dup_flags_1 <- dup_flags %>% dplyr::select(Dup_ID, Dup_Flag, Diff_P)

Dup_for_plot_1 <- DUP_start %>% dplyr::select(Plate, Mean_Conc, Dup_ID, IDs) %>% distinct(.keep_all = TRUE) %>% filter(Dup_ID %in% as.vector(dup_spls_l$Dup_ID)) %>% left_join(dup_flags_1) %>% mutate(Dup = ifelse(IDs %like% "Dup", "Duplicate", "Original")) %>% dplyr::select(-IDs, -Mean_Conc, -Dup) %>% distinct(.keep_all = TRUE)

#plot Data
Dup_Plot <- ggplot(Dup_for_plot_1,
                   aes(
                     x = Dup_ID,
                     y = Diff_P,
                     fill = Dup_Flag
                   )) +
  geom_bar(
    stat = 'identity',
    position = position_dodge2(preserve = "single"),
    linewidth = 0.9
  ) +
  geom_hline(aes(yintercept = dups_perc_diff), linetype = "dashed") +
    annotate("text",x = 1.75, y=17.5, label= "Percent Difference Cutoff", colour = "Black") + 
  labs(y = "Percent Difference", x = "Sample ID") +
  scale_fill_manual(values = c(
    "Dup_Good" = "darkgreen",
    "Dup_Bad" = "darkred"
  )) +
  theme_classic()+
  ggtitle("H2S Duplicates") 

Dup_Plot


```

## Return Dup Flags

```{r, echo=FALSE}
Dup_Flag_1 <- dup_flags %>% group_by(Plate)  %>% mutate(Dup_Flags = ifelse("Dup_Bad" %in% Dup_Flag, "Dup_Bad", "Dup_Good")) %>% dplyr::select(Plate, Dup_Flags) %>% rename(Dup_Flag = Dup_Flags) %>% distinct(.keep_all = TRUE)

Dup_Flags_1 <-Dup_Flag_1 %>% mutate(Dup_Flags = paste0(Dup_Flag, "-", Plate))

Bad_Dups <- dup_flags %>% filter(str_detect(Dup_Flag, "Dup_Bad"))
                                 
Bad_D_N <- length(Bad_Dups$Dup_Flag)

Total_D_N <- length(dup_flags$Dup_Flag)

ifelse(Bad_D_N/Total_D_N > 0.4, paste0("<60% of Dups Pass, Rerun"), paste0("Dups Pass") )

Duplicate_Flag <-  ifelse(Bad_D_N/Total_D_N > 0.4, paste0(" >40% Dups Fail"), "" )

paste0(Dup_Flags_1$Dup_Flags)


```

## Check the spks for QAQC
```{r, echo=FALSE}
#pull out any rows that have "Spk" in the ID column
SPK_start <- H2S_SPL_CV_fin %>%  mutate(Spk_ID = gsub(" ", "", substr(IDs, 1, 6)),
                                        Spk = ifelse(IDs %like% "Spike", "Spk", "Org")) %>% mutate(Spk_ID = ifelse(Plate %like% "rr", paste0(Spk_ID, "_RR"), Spk_ID))
#separate spikes from originals
spks_only <- SPK_start %>% filter(Spk == "Spk") %>% dplyr::select(Spk_ID, Mean_Conc, SD_Conc, Dilution) %>% rename(Spk_Conc = Mean_Conc,
                                                                                                                   Spk_SD = SD_Conc,
                                                                                                                   Spk_Dilution = Dilution) %>% distinct(.keep_all = TRUE)

no_spks <- SPK_start %>% filter(!Spk == "Spk")

spls_l <- no_spks %>% dplyr::select(Spk_ID)  %>% distinct(.keep_all = TRUE)

spk_spls_l <- spks_only %>% dplyr::select(Spk_ID) %>% distinct(.keep_all = TRUE)

spk_spls <- no_spks %>% filter(Spk_ID %in% as.vector(spk_spls_l$Spk_ID)) %>% dplyr::select(Spk_ID, Mean_Conc, SD_Conc, Plate, IDs, Dilution) %>% distinct(.keep_all = TRUE)

SPK_only <- spks_only %>% filter(Spk_ID %in% as.vector(spls_l$Spk_ID))

spk_join <- full_join(SPK_only, spk_spls, by = join_by(Spk_ID))

#Calculate recovery and flag data
spk_flags <- spk_join %>% mutate(
  Spl_mol = (sample_vol / Con1) * (Mean_Conc / Dilution),
  Spk_mol = ((sample_vol + spkvol) / Con1) * (Spk_Conc / Spk_Dilution), 
  Spk_exp = Spl_mol + spk_Conc, 
  Spk_recovery = (Spk_mol/Spk_exp) * 100) %>% 
  mutate(
    Spk_Flag = ifelse(
      Spk_recovery >= high_recovery_cutoff |
      Spk_recovery <= low_recovery_cutoff,
      "Spk_Bad",
      "Spk_Good"
    )
  )

#organize data for flagging and plotting
spk_flags_1 <- spk_flags %>% dplyr::select(Spk_ID, Spk_Flag, Spk_recovery)

Spk_for_plot_a <- SPK_start %>% dplyr::select(Plate, Mean_Conc, Spk_ID, IDs) %>% distinct(.keep_all = TRUE) %>% filter(Spk_ID %in% as.vector(spk_spls_l$Spk_ID)) %>% left_join(spk_flags_1, by = join_by(Spk_ID)) %>% mutate(Spk = ifelse(IDs %like% "Spike", "Spike", "Original")) %>% filter(Spk_ID %in% as.vector(spls_l$Spk_ID))

Spk_for_plot_1 <- Spk_for_plot_a %>% dplyr::select(Plate, Spk_ID, Spk_recovery, Spk_Flag) %>% distinct(.keep_all = TRUE)

#plot data
Spk_Plot <- ggplot(Spk_for_plot_1,
                   aes(
                     x = Plate,
                     y = Spk_recovery,
                     fill = Spk_Flag
                   )) +
  geom_bar(
    stat = 'identity',
    position = position_dodge2(preserve = "single"),
    linewidth = 0.9
  ) +
  geom_hline(aes(yintercept = high_recovery_cutoff), linetype = "dashed") +
    geom_hline(aes(yintercept = low_recovery_cutoff), linetype = "dashed") +
  annotate("text",x = 1.25, y=115, label= "Acceptable Recovery Range", colour = "Black") + 
  labs(y = "Spike Recovery (%)", x = "Plate") +
  scale_fill_manual(values = c(
    "Spk_Good" = "darkgreen",
    "Spk_Bad" = "darkred"
  )) +
  theme_classic() +
  ggtitle("H2S Spikes")

Spk_Plot


```

##Display Spike Flags
```{r, echo=FALSE}
ifelse(all(as.vector(spk_spls_l$Spk_ID) %in% as.vector(spls_l$Spk_ID)), paste0("Spikes Paired"), paste0("Spike Missing Pair"))

Spk_Flag_1 <- spk_flags %>% mutate(Spk_Flags = Spk_Flag) %>% dplyr::select(Plate, IDs, Spk_Flags) %>% rename(Spk_Flag = Spk_Flags)

Spk_Flags_1 <-Spk_Flag_1 %>% mutate(Spk_Flags = paste0(Spk_Flag, "-", Plate))

paste0(Spk_Flags_1$Spk_Flags)

bad <- Spk_Flags_1 %>% filter(Spk_Flags %like% "Bad")

length_bad <- length(as.vector(bad$Spk_Flags))

length_all <-  length(as.vector(Spk_Flags_1$Spk_Flags))

ifelse(length_bad/length_all > 0.4, paste0("<60% of Spikes Pass, Rerun"), paste0("Spikes Pass") )

Spike_Flag <- ifelse(length_bad/length_all > 0.4, paste0(">40% Spikes Fail"), "" )

```


# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")


#Get Samples Only
H2S_spls_md <- H2S_SPL_CV_fin %>%
  filter(!str_detect(IDs, "Dup")) %>%
  filter(!str_detect(IDs, "Spike"))

#Check for reruns and replace
H2S_spl_rrn <- H2S_spls_md %>%
  filter(Plate %like% "rr") %>% 
  dplyr::select(IDs, Mean_Conc, Conc_Flag) %>%
  rename(RRN_Conc = Mean_Conc, Conc_Flag_rr = Conc_Flag) %>% 
  distinct(.keep_all = TRUE) 

RRN_IDs <- as.vector(H2S_spl_rrn$IDs)

H2S_spl_og <- H2S_spls_md %>%
  filter(!Plate %like% "rr") %>% 
  dplyr::select(IDs, Mean_Conc, Conc_Flag, Plate) %>% 
  distinct(.keep_all = TRUE) %>%
  mutate(Rerun = ifelse(IDs %in% RRN_IDs, "RRN", "Org"))

H2S_spl_rr_join <- full_join(H2S_spl_rrn,H2S_spl_og, by = join_by(IDs) ) 

Samples <-  H2S_spl_rr_join %>% mutate(Conc = ifelse(Rerun == "RRN", RRN_Conc, Mean_Conc), Conc_Flag = ifelse(Rerun == "RRN", Conc_Flag_rr, Conc_Flag)) %>% dplyr::select(IDs, Conc, Conc_Flag, Plate) %>% rename(Sample_ID = IDs)



#check to see if all samples are present in the metadata
all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}


#merge metadata with sample run data
merged_data <- metadata %>%
  mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
  fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included?

```


#Add Flags
```{r, echo=FALSE}

#Get only bad flags for each plate and check
Chk_Flags_All <- Chk_Flags %>% dplyr::select(Plate, CHK_CV_Flags) %>% group_by(Plate) %>% mutate(CHKSTD_Flags = paste0(ifelse(
  CHK_CV_Flags %like% "Good", "", CHK_CV_Flags
))) %>% dplyr::select(Plate, CHKSTD_Flags) %>% mutate(CHKSTD_Flags = ifelse(CHKSTD_Flags == "", NA, CHKSTD_Flags)) %>% na.omit()

Mtx_Flags_All <- Mtx_Flags_org %>% dplyr::select(Plate, MTX_Flag, MTX_CV_Flag) %>% rename(MTXCV_Flag = MTX_CV_Flag, Mtx_Flag = MTX_Flag) %>% mutate(Mtx_Flag = ifelse(Mtx_Flag %like% "Good", NA , Mtx_Flag)) %>% mutate(Mtx_Flag = ifelse(MTXCV_Flag %like% "Good", NA , MTXCV_Flag)) %>% na.omit()

STD_Flags_All <- STD_CRV_flags %>% dplyr::select(Plate, R2_Flag, CV_Flag) %>% mutate(STDCV_Flags = paste0("STD", CV_Flag)) %>% dplyr::select(-CV_Flag) %>% mutate(
  R2_Flag = ifelse(R2_Flag %like% "Good", NA, R2_Flag),
  STDCV_Flags = ifelse(STDCV_Flags %like% "Good", NA, STDCV_Flags)
) %>% na.omit()

Flags_Int <- tibble(Plate = c(plates))

#Join Flags

Flags_Joined_1 <- left_join(Flags_Int, Mtx_Flags_All, by = join_by(Plate))

Flags_Joined_2 <- left_join(Flags_Joined_1, Chk_Flags_All, by = join_by(Plate))

Flags_Joined <- left_join(Flags_Joined_2, STD_Flags_All, by = join_by(Plate))


Flags_Joined[Flags_Joined == "NA"] <- NA

#Merge Flags into one column
Flags <- Flags_Joined %>% replace(is.na(.), "") %>% mutate(
  H2S_Flags = paste0(
    Spike_Flag,
    " ",
    Duplicate_Flag,
    " ",
    CHKSTD_Flags,
    " ",
    Mtx_Flag,
    " ",
    MTXCV_Flag,
    " ",
    R2_Flag,
    " ",
    STDCV_Flags
  )
) %>% dplyr::select(Plate, H2S_Flags)

#Add Flags to data
flagged_data <- left_join(merged_data, Flags, by = join_by(Plate))


df_all_clean <- flagged_data %>% dplyr::select(-Plate)

```
#Organize Data  
```{r, echo=FALSE}
#check results
#head(Samples)

#pull out what we need

df_all_clean_cols <- df_all_clean %>%
  dplyr::select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot_Name,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    Conc,
    Conc_Flag,
    H2S_Flags
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = Run_notes,# new column with notes about the run
    Analysis_rundate = Run_Date) 

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
    H2S_Conc_uM = Conc,
    H2S_Conc_Flag = Conc_Flag,
    Plot = Plot_Name
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  dplyr::select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    H2S_Conc_uM,
    H2S_Conc_Flag,
    H2S_Flags,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)

```


## Visualize Data
```{r Visualize Data, echo=FALSE}

H2S_forplot <- All_Clean_Data

#ensure points are in correct order

H2S_forplot <- H2S_forplot[order(H2S_forplot$Depth_cm), ]

color1 <- c("L1", "L2", "L3", "L4")
color2 <- c("L5", "L6", "L8", "L10")
color3 <- c("L7", "L9", "L11", "L12")
color4 <- c("L13", "L14", "L16", "L15")
color5 <- c("L17", "L18", "L19", "L20")

H2S_forplot <- H2S_forplot %>%
  mutate(
    Plot_Color = case_when(
      Plot %in% color1 ~ "Color1",
      Plot %in% color2 ~ "Color2",
      Plot %in% color3 ~ "Color3",
      Plot %in% color4 ~ "Color4",
      Plot %in% color5 ~ "Color5"
    )
  ) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2 & eN",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2 & aN",
      ifelse(
        CO2_Treatment == "Ambient" &
          N_Treatment == "Amended"  ,
        "aCO2 & eN" ,
        "aCO2 & aN"
      )
    )
  ))  %>%
  mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2"))  %>%
  mutate(N_Treatment = ifelse(N_Treatment == "Ambient", "aN", "eN")) 

viz_H2S_plot2 <- ggplot(data = H2S_forplot, aes(
  x = factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")),
  y = H2S_Conc_uM,
  fill = as.factor(Treatment),
  color = as.factor(Treatment)
)) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(padding = 0.25,preserve = "single")
  ) +
  facet_grid(Depth_cm ~ N_Treatment, scales = "fixed") +
  scale_fill_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "white",
    "aCO2 & aN" = "white"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
    
  ) +
  scale_color_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "#DB5461",
    "aCO2 & aN" = "#454ADE"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
  ) +
  theme_classic() +
  labs(x = " ",
       y = "H2S (uM)",
       title = "CO2xN: Porewater Sulfide"
       ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  ) +
coord_flip()
print(viz_H2S_plot2)

```



```{r, echo=FALSE}
#ensure points are in correct order

#make plot
viz_H2S_plot_2 <- ggplot(data = H2S_forplot, aes(x = H2S_Conc_uM, y = as.numeric(Depth_cm), color = Plot_Color)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(
    values = c(
      "Color1" = "#FFD500",
      "Color2" = "#00D0FF",
      "Color3" = "#FF0075",
      "Color4" = "#80d280",
      "Color5" = "#ff6040"
    ),
    labels = c("1", "2", "3", "4", "5")
  ) +
  facet_grid(Treatment ~ ., scales = "fixed") +
  labs(
    x = "H2S (uM)",
    y = "Depth (cm)",
    title = "CO2xN: Porewater H2S by Replicate",
    color = "Replicate"
  ) +
  scale_y_reverse(breaks = c(20, 40, 80), limits = c(90, 10)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_H2S_plot_2
```


##Export Data
```{r, echo=FALSE}

if(After_Reruns == "YES"){
write.csv(All_Clean_Data, final_path)} else { write.csv(All_Clean_Data, rerun_path) }
#end
```


### END