---
title: "CO2xN: Porewater Methane"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:  
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information
```{r run information, include=TRUE}
cat("Run Information: Melanie Giessner ") #lets you know what section you're in
#set the run date & user name
sample_year <- "2026"
sample_month <- "March & April"
user <- "Melanie Giessner"
#identify the files you want to read in
#read in as a list to accommodate multiple runs in a month

GHG_files <- c("Raw Data/20260406_GENX_GHGdata.csv" )


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Processed Data/ShimadzuGC_Slope_QAQC.csv"


# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_GENX_Porewater_CH4_20260406.csv"

#Gas in Water Calculations
# "TempC" column value in Celsius (equilibration temperature) and volume of gas (in liters) in syringe (not changing currently)
TempC = 25 #Equilibration Temperature (Celsius)

volume_g = 0.015 #volume of gas (in liters) in syringe when equilibrized
volume_w = 0.015 #volume of water (in liters) in syringe when equilibrized (should be equal to volume_g)

#record any notes about the run or anything other info here:
run_notes <- ""

#Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

cat(run_notes)
```


##Set Up

```{r setup, include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  fuzzyjoin,
  ggplot2,
  ggpubr,
  stringr,
  here)

##Low & High Standard Curve Cutoffs
CH4_low_std_crv_cutoff <- 10
CH4_low_std_crv_area_cutoff <- 7000
CH4_mid_std_crv_cutoff <- 100
CH4_mid_std_crv_area_cutoff <- 70000
CH4_high_std_crv_cutoff <- 1006

CO2_low_std_crv_cutoff <- 2000
CO2_low_std_crv_area_cutoff <- 900000
CO2_mid_std_crv_cutoff <- 20000
CO2_mid_std_crv_area_cutoff <- 7000000
CO2_high_std_crv_cutoff <- 50100
  
N2O_low_std_crv_cutoff <- 5
N2O_low_std_crv_area_cutoff <- 13000
N2O_mid_std_crv_cutoff <- 50.2
N2O_mid_std_crv_area_cutoff <- 100000
N2O_high_std_crv_cutoff <- 500
  
r2_cutoff <- 0.990

##Gas In Water Calculation
R <- 0.08206
CTK <- 273.15

#Metadata file path
#Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

##Detection Limits
adl_cutoff <- 20000
bdl_cutoff <- 50

##Check Standard Concentration
level_CH4 <- 10      #ppm
level_CO2 <- 2000      #ppm
level_N2O <- 5      #ppm
chks_flag = 60
blks_flag = 60

#Volumes
exetainer_volume =12


```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata

# raw_metadata <- read.csv(Raw_Metadata)
# #clean up columns in metadata file:
# #Pull out only SMARTX Rows
# #Create a Sample_Name column to match the SEAL
# metadata <- raw_metadata %>%
#   filter(grepl("2-Meter", Nickname, ignore.case = TRUE)) %>%
#   mutate(Sample_Name = paste0(Sample.ID)) %>%
#   mutate(Month = sample_month) %>%
#   mutate(Year = sample_year) %>%
#   rename(
#     Sample_ID = Sample.ID,
#     Plot_Name = Plot,
#   ) %>%
#   select(-Temp_Treatment) %>%
#   mutate(Treatment = ifelse(
#     CO2_Treatment == "Elevated" &
#       N_Treatment == "Amended",
#     "eCO2_N",
#     ifelse(
#       CO2_Treatment == "Elevated" &
#         N_Treatment == "Ambient",
#       "eCO2",
#       ifelse(CO2_Treatment == "Ambient" &
#                N_Treatment == "Amended"  , "N" , "Ambient")
#     )
#   )) 
#Bring in the collection date and any field notes if relevant
#collection_metadata <- read.csv(Collection_Dates)
#then we will merge the collection date for SMARTX based on the project and the chambers

```

# Read in data files
```{r , include=FALSE}

#Read in Nov Data
raw_list <- map(GHG_files, read.csv)


# Combining Files
raw <- bind_rows(raw_list)



#rename column
colnames(raw)[1] <- "Sample_ID"

#head(stds_ch4) 

```

# Assess standard curves
```{r , echo=FALSE}
stds_ch4 <-  raw %>%
  filter(str_detect(Type, "Standard")) %>% filter(!Conc_CH4 == "unknown")

#Correct data class
stds_ch4[, "Conc_CH4"] <- as.numeric(stds_ch4[, "Conc_CH4"])

#get the area of the top of the low curve
CH4_low_std_crv_area_cutoff <-stds_ch4$CH4_Area[stds_ch4$Conc_CH4 == CH4_low_std_crv_cutoff]
#Separate Into Low and High

mnmx_stds <- stds_ch4 %>% filter(Conc_CH4 > 0) %>% group_by(Run_Date) %>% mutate(CH4_Min = min(Conc_CH4), CH4_Max = max(Conc_CH4)) %>% select(Run_Date, CH4_Min, CH4_Max) %>% unique()

low_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 > CH4_low_std_crv_cutoff)

low_stds_ch4$Conc_CH4 <- as.numeric(low_stds_ch4$Conc_CH4)


high_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 < CH4_low_std_crv_cutoff)

high_stds_ch4$Conc_CH4 <- as.numeric(high_stds_ch4$Conc_CH4)

#High Curve for CH4
CH4_Curve <- ggplot(high_stds_ch4, aes(CH4_Area , Conc_CH4,   color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High CH4 Std Curve")
CH4_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_ch4_grouped <- high_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list <- map(split_data, ~ .x %>% mutate(CH4_Curve = "High",
                                        CH4_R2 = summary(lm(Conc_CH4 ~ CH4_Area  , data = .))$adj.r.squared , 
                                        CH4_Slope = coef(lm(Conc_CH4 ~ CH4_Area  , data = .))[2],
                                        CH4_Intercept = coef(lm(Conc_CH4 ~ CH4_Area  , data = .))[1]))

 Slope_CH4_raw <- bind_rows(Slope_CH4_raw_list)
  
 Slope_CH4_High <- Slope_CH4_raw %>% select(Run_Date, CH4_Curve, CH4_R2, CH4_Slope, CH4_Intercept) %>% distinct()


                              
#low Curve for CH4
CH4_Curve <- ggplot(low_stds_ch4, aes(CH4_Area , Conc_CH4,   color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low CH4 Std Curve")
CH4_Curve

low_CH4_lm <- lm(low_stds_ch4$CH4_Area ~ low_stds_ch4$Conc_CH4)
low_cf <- coef(low_CH4_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_ch4_grouped <- low_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list_l <- map(split_data_l, ~ .x %>% mutate(CH4_Curve = "Low",
                                        CH4_R2 = summary(lm(Conc_CH4 ~ CH4_Area  , data = .))$adj.r.squared , 
                                        CH4_Slope = coef(lm(Conc_CH4 ~ CH4_Area  , data = .))[2],
                                        CH4_Intercept = coef(lm(Conc_CH4 ~ CH4_Area  , data = .))[1]))

 Slope_CH4_raw_l <- bind_rows(Slope_CH4_raw_list_l)
  
 Slope_CH4_Low <- Slope_CH4_raw_l %>% select(Run_Date, CH4_Curve, CH4_R2, CH4_Slope, CH4_Intercept) %>% distinct()
 
 Slope_CH4 <- rbind(Slope_CH4_High, Slope_CH4_Low)
 
 Slope_CH4_m <- left_join(Slope_CH4, mnmx_stds)



#check r2

ifelse(Slope_CH4_High$CH4_R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_CH4_High$Run_Date), paste0("High_R2 is Good","-", Slope_CH4_High$Run_Date))

ifelse(Slope_CH4_Low$CH4_R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_CH4_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_CH4_Low$Run_Date))

High_R2_Flags <- Slope_CH4_High %>% mutate(Flag = ifelse(CH4_R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, CH4_Curve)


Low_R2_Flags <- Slope_CH4_Low %>% mutate(Flag = ifelse(CH4_R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, CH4_Curve)

Flags_df <- rbind(High_R2_Flags, Low_R2_Flags)
```


```{r CO2, echo=FALSE}
stds_CO2 <-  raw %>%
  filter(str_detect(Type, "Standard")) %>% filter(!Conc_CO2 == "unknown")

#Correct data class
stds_CO2[, "Conc_CO2"] <- as.numeric(stds_CO2[, "Conc_CO2"])

#get the area of the top of the low curve
CO2_low_std_crv_area_cutoff <-stds_CO2$CO2_Area[stds_CO2$Conc_CO2 == CO2_low_std_crv_cutoff]
#Separate Into Low and High

mnmx_stds <- stds_CO2 %>% filter(Conc_CO2 > 0) %>% group_by(Run_Date) %>% mutate(CO2_Min = min(Conc_CO2), CO2_Max = max(Conc_CO2)) %>% select(Run_Date, CO2_Min, CO2_Max) %>% unique()

low_stds_CO2 <- stds_CO2  %>%
  filter(!Conc_CO2 > CO2_low_std_crv_cutoff)

low_stds_CO2$Conc_CO2 <- as.numeric(low_stds_CO2$Conc_CO2)


high_stds_CO2 <- stds_CO2  %>%
  filter(!Conc_CO2 < CO2_low_std_crv_cutoff)

high_stds_CO2$Conc_CO2 <- as.numeric(high_stds_CO2$Conc_CO2)

#High Curve for CO2
CO2_Curve <- ggplot(high_stds_CO2, aes(CO2_Area , Conc_CO2,   color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High CO2 Std Curve")
CO2_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_CO2_grouped <- high_stds_CO2 %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_CO2_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_CO2_raw_list <- map(split_data, ~ .x %>% mutate(CO2_Curve = "High",
                                        CO2_R2 = summary(lm(Conc_CO2 ~ CO2_Area  , data = .))$adj.r.squared , 
                                        CO2_Slope = coef(lm(Conc_CO2 ~ CO2_Area  , data = .))[2],
                                        CO2_Intercept = coef(lm(Conc_CO2 ~ CO2_Area  , data = .))[1]))

 Slope_CO2_raw <- bind_rows(Slope_CO2_raw_list)
  
 Slope_CO2_High <- Slope_CO2_raw %>% select(Run_Date, CO2_Curve, CO2_R2, CO2_Slope, CO2_Intercept) %>% distinct()


                              
#low Curve for CO2
CO2_Curve <- ggplot(low_stds_CO2, aes(CO2_Area , Conc_CO2,   color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low CO2 Std Curve")
CO2_Curve

low_CO2_lm <- lm(low_stds_CO2$CO2_Area ~ low_stds_CO2$Conc_CO2)
low_cf <- coef(low_CO2_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_CO2_grouped <- low_stds_CO2 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_CO2_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_CO2_raw_list_l <- map(split_data_l, ~ .x %>% mutate(CO2_Curve = "Low",
                                        CO2_R2 = summary(lm(Conc_CO2 ~ CO2_Area  , data = .))$adj.r.squared , 
                                        CO2_Slope = coef(lm(Conc_CO2 ~ CO2_Area  , data = .))[2],
                                        CO2_Intercept = coef(lm(Conc_CO2 ~ CO2_Area  , data = .))[1]))

 Slope_CO2_raw_l <- bind_rows(Slope_CO2_raw_list_l)
  
 Slope_CO2_Low <- Slope_CO2_raw_l %>% select(Run_Date, CO2_Curve, CO2_R2, CO2_Slope, CO2_Intercept) %>% distinct()
 
 Slope_CO2 <- rbind(Slope_CO2_High, Slope_CO2_Low)
 
 Slope_CO2_m <- left_join(Slope_CO2, mnmx_stds)



#check r2

ifelse(Slope_CO2_High$CO2_R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_CO2_High$Run_Date), paste0("High_R2 is Good","-", Slope_CO2_High$Run_Date))

ifelse(Slope_CO2_Low$CO2_R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_CO2_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_CO2_Low$Run_Date))

High_R2_Flags <- Slope_CO2_High %>% mutate(Flag = ifelse(CO2_R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, CO2_Curve)


Low_R2_Flags <- Slope_CO2_Low %>% mutate(Flag = ifelse(CO2_R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, CO2_Curve)

Flags_df <- rbind(High_R2_Flags, Low_R2_Flags)
```


```{r N20, echo=FALSE}
stds_N2O <-  raw %>%
  filter(str_detect(Type, "Standard")) %>% filter(!Conc_N2O == "unknown")

#Correct data class
stds_N2O[, "Conc_N2O"] <- as.numeric(stds_N2O[, "Conc_N2O"])

#get the area of the top of the low curve
N2O_low_std_crv_area_cutoff <-stds_N2O$N2O_Area[stds_N2O$Conc_N2O == N2O_low_std_crv_cutoff]
#Separate Into Low and High

mnmx_stds <- stds_N2O %>% filter(Conc_N2O > 0) %>% group_by(Run_Date) %>% mutate(N2O_Min = min(Conc_N2O), N2O_Max = max(Conc_N2O)) %>% select(Run_Date, N2O_Min, N2O_Max) %>% unique()

low_stds_N2O <- stds_N2O  %>%
  filter(!Conc_N2O > N2O_low_std_crv_cutoff)

low_stds_N2O$Conc_N2O <- as.numeric(low_stds_N2O$Conc_N2O)


high_stds_N2O <- stds_N2O  %>%
  filter(!Conc_N2O < N2O_low_std_crv_cutoff)

high_stds_N2O$Conc_N2O <- as.numeric(high_stds_N2O$Conc_N2O)

#High Curve for N2O
N2O_Curve <- ggplot(high_stds_N2O, aes(N2O_Area , Conc_N2O,   color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High N2O Std Curve")
N2O_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_N2O_grouped <- high_stds_N2O %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_N2O_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_N2O_raw_list <- map(split_data, ~ .x %>% mutate(N2O_Curve = "High",
                                        N2O_R2 = summary(lm(Conc_N2O ~ N2O_Area  , data = .))$adj.r.squared , 
                                        N2O_Slope = coef(lm(Conc_N2O ~ N2O_Area  , data = .))[2],
                                        N2O_Intercept = coef(lm(Conc_N2O ~ N2O_Area  , data = .))[1]))

 Slope_N2O_raw <- bind_rows(Slope_N2O_raw_list)
  
 Slope_N2O_High <- Slope_N2O_raw %>% select(Run_Date, N2O_Curve, N2O_R2, N2O_Slope, N2O_Intercept) %>% distinct()


                              
#low Curve for N2O
N2O_Curve <- ggplot(low_stds_N2O, aes(N2O_Area , Conc_N2O,   color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low N2O Std Curve")
N2O_Curve

low_N2O_lm <- lm(low_stds_N2O$N2O_Area ~ low_stds_N2O$Conc_N2O)
low_cf <- coef(low_N2O_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_N2O_grouped <- low_stds_N2O %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_N2O_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_N2O_raw_list_l <- map(split_data_l, ~ .x %>% mutate(N2O_Curve = "Low",
                                        N2O_R2 = summary(lm(Conc_N2O ~ N2O_Area  , data = .))$adj.r.squared , 
                                        N2O_Slope = coef(lm(Conc_N2O ~ N2O_Area  , data = .))[2],
                                        N2O_Intercept = coef(lm(Conc_N2O ~ N2O_Area  , data = .))[1]))

 Slope_N2O_raw_l <- bind_rows(Slope_N2O_raw_list_l)
  
 Slope_N2O_Low <- Slope_N2O_raw_l %>% select(Run_Date, N2O_Curve, N2O_R2, N2O_Slope, N2O_Intercept) %>% distinct()
 
 Slope_N2O <- rbind(Slope_N2O_High, Slope_N2O_Low)
 
 Slope_N2O_m <- left_join(Slope_N2O, mnmx_stds)



#check r2

ifelse(Slope_N2O_High$N2O_R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_N2O_High$Run_Date), paste0("High_R2 is Good","-", Slope_N2O_High$Run_Date))

ifelse(Slope_N2O_Low$N2O_R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_N2O_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_N2O_Low$Run_Date))

High_R2_Flags <- Slope_N2O_High %>% mutate(Flag = ifelse(N2O_R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, N2O_Curve)


Low_R2_Flags <- Slope_N2O_Low %>% mutate(Flag = ifelse(N2O_R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, N2O_Curve)

Flags_df <- rbind(High_R2_Flags, Low_R2_Flags)
```

```{r log}
#add new standard curve data to log
Slope_CH4_1 <- Slope_CH4 %>% mutate(curve = paste0("Slope_CH4_", CH4_Curve), R2 = CH4_R2, slope = CH4_Slope, intercept = CH4_Intercept, Analysis = "CH4", Run_by = user, Curve_top = case_when(
  CH4_Curve == "Low" ~ paste(CH4_low_std_crv_cutoff,"ppm"),
  CH4_Curve == "High" ~ paste(CH4_high_std_crv_cutoff,"ppm"),
  .default = as.character(CH4_mid_std_crv_cutoff)
)) %>% select(Analysis, Run_Date, Run_by, Curve_top, curve, R2, slope, intercept)
Slope_CO2_1 <- Slope_CO2 %>% mutate(curve = paste0("Slope_CO2_", CO2_Curve), R2 = CO2_R2, slope = CO2_Slope, intercept = CO2_Intercept, Analysis = "CO2", Run_by = user, Curve_top = case_when(
  CO2_Curve == "Low" ~ paste(CO2_low_std_crv_cutoff,"ppm"),
  CO2_Curve == "High" ~ paste(CO2_high_std_crv_cutoff,"ppm"),
  .default = as.character(CO2_mid_std_crv_cutoff)
)) %>% select(Analysis, Run_Date, Run_by, Curve_top, curve, R2, slope, intercept)
Slope_N2O_1 <- Slope_N2O %>% mutate(curve = paste0("Slope_N2O_", N2O_Curve), R2 = N2O_R2, slope = N2O_Slope, intercept = N2O_Intercept, Analysis = "N2O", Run_by = user, Curve_top = case_when(
  N2O_Curve == "Low" ~ paste(N2O_low_std_crv_cutoff,"ppm"),
  N2O_Curve == "High" ~ paste(N2O_high_std_crv_cutoff,"ppm"),
  .default = as.character(N2O_mid_std_crv_cutoff)
)) %>% select(Analysis, Run_Date, Run_by, Curve_top, curve, R2, slope, intercept)

Slope <- rbind(Slope_CH4_1, Slope_CO2_1, Slope_N2O_1)

 log <- read.csv(log_path)
 log <- log[, -c(1)]
 head(log)
 
 log <- rbind(log, Slope)

#remove possible duplicates from multiple QAQC attempts

 unique_log <- log %>% distinct(.keep_all = TRUE)%>%
      mutate(Run_Date = as.Date(Run_Date,"%m/%d/%Y"), curve= gsub("L","l",curve), curve= gsub("H","h",curve))

 Slopes_chk_CH4 <- ggplot(unique_log %>% filter( grepl("CH4",curve,ignore.case = T)), aes(x = Run_Date, y = slope, col = curve)) +
   geom_point(size = 4) +
   #ylim(400, 800) +
   theme_bw()

 Slopes_chk_CH4
 
 Slopes_chk_CO2 <- ggplot(unique_log %>% filter( grepl("CO2",curve,ignore.case = T)), aes(x = Run_Date, y = slope, col = curve)) +
   geom_point(size = 4) +
   #ylim(400, 800) +
   theme_bw()

 Slopes_chk_CO2
 
  Slopes_chk_N2O <- ggplot(unique_log %>% filter( grepl("N2O",curve,ignore.case = T)), aes(x = Run_Date, y = slope, col = curve)) +
   geom_point(size = 4) +
   #ylim(400, 800) +
   theme_bw()

 Slopes_chk_N2O
 
 
  int_chk_CH4 <- ggplot(unique_log %>% filter( grepl("CH4",curve,ignore.case = T)), aes(x = Run_Date, y = intercept, col = curve)) +
   geom_point(size = 4) +
   #ylim(400, 800) +
   theme_bw()

 int_chk_CH4
 
   int_chk_CO2 <- ggplot(unique_log %>% filter( grepl("CO2",curve,ignore.case = T)), aes(x = Run_Date, y = intercept, col = curve)) +
   geom_point(size = 4) +
   #ylim(400, 800) +
   theme_bw()

 int_chk_CO2
 
  int_chk_N2O <- ggplot(unique_log %>% filter( grepl("N2O",curve,ignore.case = T)), aes(x = Run_Date, y = intercept, col = curve)) +
   geom_point(size = 4) +
   #ylim(400, 800) +
   theme_bw()

 int_chk_N2O
 #write.csv(unique_log, log_path)
```


## Now calculate the CH4 in ppm  
```{r Calculate Concetration, echo=FALSE}
#head(raw)

#pull out Samples
Samples_1 <-  raw %>%  
  filter(!str_detect(Type, "Lab air")) %>%  
  filter(!str_detect(Type, "Standard")) %>%  
  filter(!str_detect(Type, "Blank")) %>%  
  filter(!str_detect(Type, "Chk_STD")) %>%  
   filter(!str_detect(Type, "ChkStd")) %>%
  filter(!str_detect(Type, "NA")) %>%
  
  mutate(CH4_Curve = ifelse(CH4_Area >= CH4_low_std_crv_area_cutoff, "High", "Low"), CO2_Curve = ifelse(CO2_Area >= CO2_low_std_crv_area_cutoff, "High", "Low"), N2O_Curve = ifelse(N2O_Area >= N2O_low_std_crv_area_cutoff, "High", "Low"))
#head(Samples)

Samples_2 <- merge(Samples_1, Slope_CH4_m)
Samples_2 <- merge(Samples_2, Slope_CO2_m)
Samples_2 <- merge(Samples_2, Slope_N2O_m)
#Calculate CH4 concentrations in ppm
Samples <- Samples_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area*CH4_Slope) + CH4_Intercept ,CO2_Conc_ppm = (CO2_Area*CO2_Slope)+ CO2_Intercept, N2O_Conc_ppm = (N2O_Area*N2O_Slope) + N2O_Intercept )

##########make flags for any dilutions needed 


Samples$CH4_Flag <- ifelse(Samples$CH4_Conc_ppm > Samples$CH4_Max, "ADL", ifelse(Samples$CH4_Conc_ppm < Samples$CH4_Min, "BDL", "Within Range"))
Samples$CO2_Flag <- ifelse(Samples$CO2_Conc_ppm > Samples$CO2_Max, "ADL", ifelse(Samples$CO2_Conc_ppm < Samples$CO2_Min, "BDL", "Within Range"))
Samples$N2O_Flag <- ifelse(Samples$N2O_Conc_ppm > Samples$N2O_Max, "ADL", ifelse(Samples$N2O_Conc_ppm < Samples$N2O_Min, "BDL", "Within Range"))
#head(Samples)



```

# Aanlayze the Check Standards 
```{r Asses Check Standards, echo=FALSE}

chks_1 <-  raw %>%
  filter(!str_detect(Type, "Lab air")) %>%
  filter(!str_detect(Type, "Standard")) %>%
  filter(!str_detect(Type, "Blank")) %>%
  filter(!str_detect(Type, "Sample")) %>%
  filter(!str_detect(Type, "NA")) %>%
  mutate(CH4_Curve = "Low",CO2_Curve = "Low",N2O_Curve = "Low")
#head(chks)


chks_2 <- merge(chks_1, Slope_CH4_m)
chks_2 <- merge(chks_2, Slope_CO2_m)
chks_2 <- merge(chks_2, Slope_N2O_m)

chks <- chks_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area*CH4_Slope) + CH4_Intercept ,CO2_Conc_ppm = (CO2_Area*CO2_Slope)+ CO2_Intercept, N2O_Conc_ppm = (N2O_Area*N2O_Slope) + N2O_Intercept )

#calculate percent difference of check standards
chks<- chks %>% mutate(
  Ch4_diff = (((chks$CH4_Conc_ppm - level_CH4) / ((chks$CH4_Conc_ppm + level_CH4) /2)) * 100),
  CH4_diff_flag =  ifelse(abs(Ch4_diff) < 10, 'YES', 'NO, rerun'),
  CO2_diff = (((chks$CO2_Conc_ppm - level_CO2) / ((chks$CO2_Conc_ppm + level_CO2) /2)) * 100),
  CO2_diff_flag =  ifelse(abs(CO2_diff) < 10, 'YES', 'NO, rerun'),
  N2O_diff = (((chks$N2O_Conc_ppm - level_N2O) / ((chks$N2O_Conc_ppm + level_N2O) /2)) * 100),
  N2O_diff_flag =  ifelse(abs(N2O_diff) < 10, 'YES', 'NO, rerun'),
  )

#calculate the percent of check standards that are within the range based on the flag
CH4_chks_percent <- (sum(chks$CH4_diff_flag == "YES") / nrow(chks)) * 100
CO2_chks_percent <- (sum(chks$CO2_diff_flag == "YES") / nrow(chks)) * 100
N2O_chks_percent <- (sum(chks$N2O_diff_flag == "YES") / nrow(chks)) * 100
#report out if flags indicate need for rerun

if (CH4_chks_percent >= chks_flag) {
  message(">60% of CH4 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CH4 Check Standards are within range of expected concentration - REASSESS")
}

if (CO2_chks_percent >= chks_flag) {
  message(">60% of CO2 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CO2 Check Standards are within range of expected concentration - REASSESS")
}

if (N2O_chks_percent >= chks_flag) {
  message(">60% of N2O Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of N2O Check Standards are within range of expected concentration - REASSESS")
}

#now plot the Ch4 concentrations and the CO2 concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

ch4_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CH4_Conc_ppm, fill =
                                      CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: CH4") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_CH4,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


ch4_chk

CO2_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CO2_Conc_ppm, fill =
                                      CO2_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: CO2") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_CO2,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


CO2_chk

N2O_chk <-  ggplot(data = chks, aes(x = Run_Date, y = N2O_Conc_ppm, fill =
                                      N2O_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: N2O") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_N2O,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


N2O_chk

#Add Check Standard Flags

flag_chk_CH4 <- ifelse(CH4_chks_percent >= chks_flag,
                "",
               "; CH4 Check Standards out Of Range")
flag_chk_CO2 <- ifelse(CO2_chks_percent >= chks_flag,
                "",
               "; CO2 Check Standards out Of Range")
flag_chk_N2O <- ifelse(N2O_chks_percent >= chks_flag,
                "",
               "; N2O Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk_CH4,flag_chk_CO2,flag_chk_N2O))

```

# Analyze the Blanks 
```{r Analyze Blanks, echo=FALSE}
cat("Assess Blanks")

blanks <-  raw %>%
  filter(!str_detect(Type, "Lab air")) %>%
  filter(!str_detect(Type, "Standard")) %>%
  filter(!str_detect(Type, "Sample")) %>%
  filter(!str_detect(Type, "Chk_STD")) %>%
  filter(!str_detect(Type, "ChkStd")) %>%
  filter(!str_detect(Type, "NA"))


#Pull out samples from df_all to calc quantile
samples_blanks <- raw %>%
  filter(str_detect(Type, c("Sample")))

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

blk_flag_CH4 <- quantile(samples_blanks$CH4_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data

blk_flag_CO2 <- quantile(samples_blanks$CO2_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data

blk_flag_N2O <- quantile(samples_blanks$N2O_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data

blanks<-blanks %>% mutate(CH4_diff_flag =  ifelse(blanks$CH4_Area <= blk_flag_CH4, 'YES', 'NO, rerun'),
                          CO2_diff_flag =  ifelse(blanks$CO2_Area <= blk_flag_CO2, 'YES', 'NO, rerun'),
                          N2O_diff_flag =  ifelse(blanks$N2O_Area <= blk_flag_N2O, 'YES', 'NO, rerun'))

#calculate the percent of check standards that are within the range based on the flag
CH4_blks_percent <- (sum(blanks$CH4_diff_flag == "YES") / nrow(blanks)) *
  100
CO2_blks_percent <- (sum(blanks$CO2_diff_flag == "YES") / nrow(blanks)) *
  100
N2O_blks_percent <- (sum(blanks$N2O_diff_flag == "YES") / nrow(blanks)) *
  100

#report out if flags indicate need for rerun
if (CH4_blks_percent >= blks_flag) {
  message(">60% of CH4 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CH4 blaks are lower 25% quartile of samples - REASSESS")
}

#report out if flags indicate need for rerun
if (CO2_blks_percent >= blks_flag) {
  message(">60% of CO2 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CO2 blaks are lower 25% quartile of samples - REASSESS")
}

#report out if flags indicate need for rerun
if (N2O_blks_percent >= blks_flag) {
  message(">60% of N2O Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of N2O blaks are lower 25% quartile of samples - REASSESS")
}


#plotting the blanks compared to the lower 25% of conc (what the flag is)
CH4_blks_plot <-  ggplot(data = blanks, aes(x = Run_Date, y = CH4_Area, fill =
                                                CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CH4 Area", title = "CH4 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CH4),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(CH4_blks_plot)

#CO2
CO2_blks_plot <-  ggplot(data = blanks, aes(x = Run_Date, y = CO2_Area, fill =
                                                CO2_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CO2 Area", title = "CO2 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CO2),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(CO2_blks_plot)

#N2O
N2O_blks_plot <-  ggplot(data = blanks, aes(x = Run_Date, y = N2O_Area, fill =
                                                N2O_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "N2O Area", title = "N2O Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_N2O),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(N2O_blks_plot)
#find average of run blanks for flagging samples later
# Compute averages
blk_avg_CH4 <- mean(blanks$CH4_Area, na.rm = TRUE)
blk_avg_CO2 <- mean(blanks$CO2_Area, na.rm = TRUE)
blk_avg_N2O <- mean(blanks$N2O_Area, na.rm = TRUE)
# Create a data frame
blank_avgs <- data.frame(Test = c("CH4","CO2","N2O"),
                         Blank_Mean_Area = c(blk_avg_CH4,blk_avg_CO2,blk_avg_N2O))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks_CH4 <-  ifelse(CH4_blks_percent >= blks_flag,
                 "",
                 '; CH4 Blanks out of Range')

flag_blks_CO2 <-  ifelse(CO2_blks_percent >= blks_flag,
                 "",
                 '; CO2 Blanks out of Range')

flag_blks_N2O <-  ifelse(N2O_blks_percent >= blks_flag,
                 "",
                 '; N2O Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks_CH4,flag_blks_CO2,flag_blks_N2O))




```

# Dilution correct samples and Check Range
```{r, echo=FALSE}

#Correct if volume other than 12mL of Sample was added
Samples$CH4_Conc_ppm_dilcorr <- ((0.5 * Samples$CH4_Conc_ppm) * (exetainer_volume + Samples$Sample_Volume)) /
  (Samples$Sample_Volume)
Samples$CO2_Conc_ppm_dilcorr <- ((0.5 * Samples$CO2_Conc_ppm) * (exetainer_volume + Samples$Sample_Volume)) /
  (Samples$Sample_Volume)
Samples$N2O_Conc_ppm_dilcorr <- ((0.5 * Samples$N2O_Conc_ppm) * (exetainer_volume + Samples$Sample_Volume)) /
  (Samples$Sample_Volume)
#check results
#head(Samples)

#quick first look at the samples
ch4_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm, fill =
                                             CH4_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range (Not Dil Cor)") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

 ch4_samples
 
 CO2_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CO2_Conc_ppm, fill =
                                             CO2_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "CO2 (ppm)", title = "CO2: Green = Within Range (Not Dil Cor)") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

 CO2_samples
 
 N2O_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = N2O_Conc_ppm, fill =
                                             N2O_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "N2O (ppm)", title = "N2O: Green = Within Range (Not Dil Cor)") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

 N2O_samples
# #corrected for different dilutions
# ch4_samples_cor <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm_dilcorr, fill =
#                                              CH4_Flag)) +
#   geom_bar(stat = 'identity') +
#   scale_fill_manual(values = c("Within Range" = "green4",
#                     "ADL" = "red3",
#                     "BDL" = "blue3")) +
#   theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range") +
#   theme(axis.text.x = element_text(
#     angle = 90,
#     vjust = 0.5,
#     hjust = 1
#   )) +
#   theme(legend.position = "none")
# 

# ch4_samples_cor


```


# Calculate gas in water  
```{r, echo=FALSE}

# Convert ppm CH4 to total uM CH4 in the sample headspace
# use Ideal Gas Law:
#   hdsp umol CH4 = ppm CH4*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

#Calculate CH4 uM from ppm

Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

Samples$CO2_Conc_uM <- (Samples$CO2_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

Samples$N2O_Conc_uM <- (Samples$N2O_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w
#old equation
#Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * ((1 * volume / 1000) / (R * (TempC + CTK)))) / (volume / 1000)


#quick first look at the samples
ch4_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "CH4 (uM)", title = "CH4 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


ch4_samples1

#quick first look at the samples
CO2_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = CO2_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "CO2 (uM)", title = "CO2 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


CO2_samples1

#quick first look at the samples
N2O_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = N2O_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "N2O (uM)", title = "N2O Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


N2O_samples1

```

# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

# cat("Merge Samples with Metadata")
# 
# #check to see if all samples are present in the metadata
# all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)
# 
# if (all_present) {
#   message("All sample IDs are present in data")
# } else {
#   message("Some sample IDs are missing from data.")
#   
#   # Optional: Which ones are missing?
#   missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
#   print(missing_ids)
# }
# 
# 
# #merge metadata with sample run data
# merged_data <- metadata %>%
#   mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
#   fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))
# 
# #might want to include a flag or something to see that all the samples are included?
# 
# #add flags
# 
# merged_2 <- merge(merged_data, Flags_df)
# merged_data <-merged_2 %>% mutate(CH4_Flags = Flag)
# 
# 
# 
# df_all_clean <- merged_data

```

## Organize Data  
```{r, echo=FALSE}
# #check results
# #head(Samples)
# 
# #pull out what we need
# 
# df_all_clean_cols <- df_all_clean %>%
#   select(
#     Site,
#     Project,
#     Year,
#     Month,
#     Community,
#     Nickname,
#     Plot_Name,
#     Depth_cm,
#     CO2_Treatment,
#     N_Treatment,
#     CH4_Conc_ppm_dilcorr,
#     CH4_Conc_uM,
#     CH4_Flag,
#     CH4_Flags,
#     Run_Date
#   )
# 
# final_data_labeled <- df_all_clean_cols %>%
#   mutate(
#     Run_notes = run_notes,# new column with notes about the run
#     Analysis_rundate = Run_Date) 
# 
# #rename columns
# org_labeled_fin_data <- final_data_labeled %>%
#   rename(
#     CH4_Conc_ppm = CH4_Conc_ppm_dilcorr,
#     CH4_Conc_Flag = CH4_Flag,
#     CH4_QAQC_Flags = CH4_Flags,
#     Plot = Plot_Name
#   )
# 
# #Order Columns for Final Data Frame
# All_Clean_Data <- org_labeled_fin_data %>%
#   select(
#     Site,
#     Project,
#     Year,
#     Month,
#     Community,
#     Nickname,
#     Plot,
#     Depth_cm,
#     CO2_Treatment,
#     N_Treatment,
#     CH4_Conc_ppm,
#     CH4_Conc_uM,
#     CH4_Conc_Flag,
#     CH4_QAQC_Flags,
#     Analysis_rundate,
#     Run_notes
#   )
# #head(Samples1)

```


## Visualize Data
```{r Visualize Data, echo=FALSE}

# CH4_forplot <- All_Clean_Data
# 
# #ensure points are in correct order
# 
# CH4_forplot <- CH4_forplot[order(CH4_forplot$Depth_cm), ]
# 
# color1 <- c("L1", "L2", "L3", "L4")
# color2 <- c("L5", "L6", "L8", "L10")
# color3 <- c("L7", "L9", "L11", "L12")
# color4 <- c("L13", "L14", "L16", "L15")
# color5 <- c("L17", "L18", "L19", "L20")
# 
# CH4_forplot <- CH4_forplot %>%
#   mutate(
#     Plot_Color = case_when(
#       Plot %in% color1 ~ "Color1",
#       Plot %in% color2 ~ "Color2",
#       Plot %in% color3 ~ "Color3",
#       Plot %in% color4 ~ "Color4",
#       Plot %in% color5 ~ "Color5"
#     )
#   ) %>%
#   mutate(Treatment = ifelse(
#     CO2_Treatment == "Elevated" &
#       N_Treatment == "Amended",
#     "eCO2 & eN",
#     ifelse(
#       CO2_Treatment == "Elevated" &
#         N_Treatment == "Ambient",
#       "eCO2 & aN",
#       ifelse(
#         CO2_Treatment == "Ambient" &
#           N_Treatment == "Amended"  ,
#         "aCO2 & eN" ,
#         "aCO2 & aN"
#       )
#     )
#   ))  %>%
#   mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2"))  %>%
#   mutate(N_Treatment = ifelse(N_Treatment == "Ambient", "aN", "eN")) 
# #need to work out how best to plot these data
# 
# viz_CH4_plot2 <- ggplot(data = CH4_forplot, aes(
#   x = factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")),
#   y = CH4_Conc_uM,
#   fill = as.factor(Treatment),
#   color = as.factor(Treatment)
# )) +
#   geom_bar(
#     stat = "identity",
#     position = position_dodge2(padding = 0.25,preserve = "single")
#   ) +
#   facet_grid(Depth_cm ~ N_Treatment, scales = "fixed") +
#   scale_fill_manual(values = c(
#     "eCO2 & eN" = "#DB5461",
#     "eCO2 & aN" = "#454ADE", 
#     "aCO2 & eN" = "white",
#     "aCO2 & aN" = "white"
#   ),
#     name = "Treatment:",
#     breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
#     labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
#     
#   ) +
#   scale_color_manual(values = c(
#     "eCO2 & eN" = "#DB5461",
#     "eCO2 & aN" = "#454ADE", 
#     "aCO2 & eN" = "#DB5461",
#     "aCO2 & aN" = "#454ADE"
#   ),
#     name = "Treatment:",
#     breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
#     labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
#   ) +
#   theme_classic() +
#   labs(x = " ",
#        y = "CH4 (uM)",
#        title = "CO2xN: Porewater Methane"
#        ) +
#   theme(legend.position = "right") +
#   scale_x_discrete(drop = TRUE) +
#   theme(
#     axis.text.x = element_text(angle = 0, hjust = 0.5),
#     panel.border = element_rect(
#       color = "black",
#       fill = NA,
#       linewidth = 0.8
#     )
#   ) +
# coord_flip()
# print(viz_CH4_plot2)

```



```{r, echo=FALSE}
# #ensure points are in correct order
# 
# #make plot
# viz_CH4_plot_2 <- ggplot(data = CH4_forplot, aes(x = CH4_Conc_uM, y = as.numeric(Depth_cm), color = Plot_Color)) +
#   geom_point(size = 3) +
#   geom_path(linewidth = 1) +
#   scale_color_manual(
#     values = c(
#       "Color1" = "#FFD500",
#       "Color2" = "#00D0FF",
#       "Color3" = "#FF0075",
#       "Color4" = "#80d280",
#       "Color5" = "#ff6040"
#     ),
#     labels = c("1", "2", "3", "4", "5")
#   ) +
#   facet_grid(Treatment ~ ., scales = "fixed") +
#   labs(
#     x = "CH4 (uM)",
#     y = "Depth (cm)",
#     title = "CO2xN: Porewater CH4 by Replicate",
#     color = "Replicate"
#   ) +
#   scale_y_reverse(breaks = c(20, 40, 80), limits = c(90, 10)) +
#   theme_classic() +
#   theme(
#     axis.text.x = element_text(angle = 0, hjust = 0.5),
#     panel.border = element_rect(
#       color = "black",
#       fill = NA,
#       linewidth = 0.8
#     )
#   )
# 
# viz_CH4_plot_2
```
##Export Data
```{r, echo=FALSE}
final_data<-Samples %>% dplyr::select(Run_Date,N2O_Curve,CO2_Curve,CH4_Curve,Sample_ID,N2O_Area,CO2_Area,CH4_Area,N2O_Conc_uM,CO2_Conc_uM,CH4_Conc_uM,N2O_Flag,CO2_Flag,CH4_Flag,Notes)
write.csv(final_data, final_path)
#end
```

#end