---
title: "CO2xN: Porewater Ammonium"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information 
```{r run information, include=TRUE}
cat("Run Information: ") #lets you know what section you're in 
#set the run date & user name 
  run_date <- "01/30/2025"
  sample_year <- "2025"
  sample_month <- "Sept"
  user <- "Victoria Ellis"

#identify the files you want to read in 
  #read in as a list to accommodate multiple runs in a month
  NH3_files <- c("Raw Data/LTREB_2M_Sept_2025_Run1.csv",
                 "Raw Data/LTREB_2M_Sept_2025_Run2.csv")

# Define the file path for QAQC log file - NO Need to change just check year 
  file_path <- "Raw Data/SEAL_CO2xN_QAQC_Log_2025.csv"
  final_path <- "Processed Data/CO2xN_Porewater_NH4_202509.csv"

#record any notes about the run or anything other info here: 
  run_notes <- ""
  
  #Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

  cat(run_notes)
```

##Setup
```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#load/install packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  DescTools,
  purrr,
  ggplot2,
  ggpubr,
  tidyr,
  tidyverse,
  lubridate,
  tinytex,
  stringr,
  writexl,
  readr,
  readxl,
  purrr,
  tinytex,
  broom,
  here, 
  data.table,
  raster)

Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
P_mw <- 30.973762  # molecular weight of P 
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 

#Detection limit and top standards for flagging 
NOx_dl <- 0.025  #mgL
NH3_dl <- 0.02  #mgL
PO4_dl <- 0.003 #mgL

NOx_top <- 1  #mgL
NH3_top <- 2  #mgL
PO4_top <- 0.3 #mgL

#Check Standard concnetrations 
NOx_CCV <- 0.5 
NH3_CCV <- 1.0 
PO4_CCV <- 0.15 

#Spike Concentrations 
NOx_Spk <- 0.5 #ppm or mg/L
NH3_Spk <- 1
PO4_Spk <- 0.152 

#expected ranges for sample concentrations used for flags 
r2_cutoff = 0.990
chk_flag = 0.25
chk_conc_flag = 15 #cutoff level for percent difference of check stds vs. expected concentration 
chks_flag = 60
rep_flag = 25  #this is a 25% error between samples
#blank_flag is calculated based on samples later in this code

#pe check Concentrations
NH3_pe <- 1.034
NOx_pe <- 1.51
PO4_pe <- 0.824

```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata)

#clean up columns in metadata file:
  #Pull out only SMARTX Rows 
  #Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("2-Meter", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  ) %>%
  dplyr::select(-Temp_Treatment) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2_N",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2",
      ifelse(CO2_Treatment == "Ambient" &
               N_Treatment == "Amended"  , "N" , "Ambient")
    )
  )) 
#Bring in the collection date and any field notes if relevant 
#collection_metadata <- read.csv(Collection_Dates)
  #then we will merge the collection date for SMARTX based on the project and the chambers 

```

\newpage

## Import Data & Clean
```{r Import Data, include=FALSE}
cat("Import Data")

#set file path for data - in run info chunk 
path <- ("file path")

#Read in Raw Data 
data_NH3 <- map(NH3_files, read.csv)

# Combining Files
df_combo <- bind_rows(data_NH3)

# Select needed columns & rename 
df <- df_combo %>%
  dplyr::select(Sample_Name = 4, 
         Run_Number = 5,
         Conc = 6, 
         Absorbance = 7, 
         Dilution = 9, 
         Unit = 12,
         Test = 13, 
         Run_Time = 15)

#Checking column headers
head(df)

df_all <- df %>% mutate(Run_Number = as.numeric(Run_Number),
         Run_Date = format(as.POSIXct(Run_Time,format='%m/%d/%Y %H:%M'),format='%m/%d/%Y'))

#Need to add some extra code to deal with July and Sept being run in the same file 

#Initialize Flags
DF_Flags <- df_all %>% dplyr::select(Run_Date) %>% distinct() %>% na.omit()


```



## Assessing standard Curves
# Assess Standard Curves
```{r, echo=FALSE}
split_data <- df_all %>% dplyr::select(Sample_Name, Conc, Absorbance, Test, Run_Date) %>% filter(Sample_Name %ilike% "Standard") %>%
  mutate(Conc = ifelse(Test == "Ammonia 2", case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard .0389" ~ 0.0389,
    Sample_Name == "Standard .1000" ~ 0.1,
    Sample_Name == "Standard .2000" ~ 0.2,
    Sample_Name == "Standard .5000" ~ 0.5,
    Sample_Name == "Standard 1.0000" ~ 1.0,
    Sample_Name == "Standard 1.5000" ~ 1.5,
    Sample_Name == "Standard 2.0000" ~ 2),
    ifelse(Test == "o-PHOS 0.3", 
           case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard .006" ~ 0.0060,
    Sample_Name == "Standard .0150" ~ 0.0150,
    Sample_Name == "Standard .03" ~ 0.0300,
    Sample_Name == "Standard .075" ~ 0.0750,
    Sample_Name == "Standard .15" ~ 0.1500,
    Sample_Name == "Standard .225" ~ 0.2250,
    Sample_Name == "Standard .3" ~ 0.3000),
    "Test Not Found"))) %>%
  group_by(Run_Date, Test) %>% 
  na.omit() %>%
  group_split()
  
#run regression and get coefficients by Run
 
Slope_raw_list <- map(split_data, ~ .x %>% mutate(
                                        R2 = summary(lm(Conc ~ Absorbance, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ Absorbance, data = .))[2],
                                        Intercept = coef(lm(Conc ~ Absorbance, data = .))[1]))


 Slope_raw <- bind_rows(Slope_raw_list)
  
 #check R2 and flag
 Slope_Std_Crvs <- Slope_raw %>% dplyr::select(Run_Date, R2, Slope, Intercept, Test) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Pass", "R2_Fail"))

 #join standard curve CVs and regression values
 
 R2_Check <- Slope_Std_Crvs %>% 
   dplyr::select(Run_Date,
                 R2,
                 Slope,
                 Intercept,
                 Test,
                 R2_Flag) %>%
   distinct()
 
 For_crv_plot <- Slope_raw %>% mutate(Conc =  as.numeric(Conc))

 Curves_Plot <- ggplot(For_crv_plot, aes(Conc, Absorbance, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + facet_grid(.~Test, scales = "free") + labs(title = "Standard Curves") + theme_classic()
 
Curves_Plot
 

R2_Flags <- R2_Check %>% dplyr::select(Test, Run_Date, R2_Flag)
 
```
  #Plot standards data 
```{r Assess Standard Curves, echo=FALSE}
cat("Assess Standard Curves")

#Plot standard Curve or Curves 
  #NH3 
std_curve_NH3 <- ggplot(stds_NH3, aes(x = Conc, y = Absorbance)) +
  geom_point(size=3) +
  geom_smooth(method = "lm", se = FALSE, color = "darkgoldenrod") +
  theme_bw() + 
  facet_wrap(~Run_Date) +
  labs(x="Concentration (ppm)",
       y="Absorbance",
       title = "NH3 Standard Curve")+ 
  annotate("text", x = 1.7, y = max(stds_NH3$Absorbance), label = paste("r^2 =", round(NH3_r2, 4)),
          color = "black", size = 4)
print(std_curve_NH3)

############## Report on Cutoffs 

#Report out a flag if the run has an R2 lower than appropriate 
#Write out to the user whether or not the r2 is above the cutoff of 0.98
  ifelse(std_curve_NH3_1$R_squared <= r2_cutoff, 
         "NH3 Curve r2 is below cutoff! - REASSESS", 
         "NH3 Curve r2 GOOD - PROCEED")

#check for the a slope QAQC file, if there is not one, make one 
if (file.exists(file_path)) {
  # If it exists, read it back into R
  stds_log <- read.csv(file_path)
  print("QAQC log file exists and has been read into the code.")
  } else {
  # If it does not exist, create the CSV file
  stds_log <- as.data.frame(matrix(ncol = 6, nrow = 0))
  colnames(stds_log) <- c("Test", "Intercept", "Slope", "R_squared", "run_date", "user")
  
  #maybe add here the last slopes from the previous year
  
  # Write add_log to CSV
  write.csv(stds_log, file = file_path, row.names = FALSE)
  print("QAQC log file does not exist. It has been created.")
  }
  
  #getting rid of first "X" column
  #stds_log <- stds_log %>%
  #  select(-1)
  
#create a dataframe that can be appended to the QAQC log 
add_log <- as.data.frame(bind_rows(std_curve_NH3_1))
add_log$run_date <- run_date
add_log$user <- user

#compare slopes to previous runs (from log) in order to assess drift 
stds_log$run_date <- as.character(stds_log$run_date)
add_log$run_date  <- as.character(add_log$run_date)     #getting the run_date column in the same type to merge

#Filter to only rows in Slopes that are NOT already in log (by run_date + analyte) but need to add if none
if (nrow(stds_log) == 0) {
  # If log is empty, everything in add_log is new
  new_rows <- add_log
} else {
  # Otherwise, compare to existing log
  new_rows <- anti_join(add_log, stds_log, by = c("run_date", "Test"))
}


# Append the new, non-duplicate rows to log
log <- bind_rows(stds_log, new_rows)


#Plot the slopes through time 
Slopes_chk <- ggplot(log, aes(run_date, Slope, col=Test)) + 
  geom_point(size=4) + 
  geom_line(aes(group = Test)) + 
  theme_bw() + 
  labs(title="Slope Drift Assessment", x="Run Date", y="Slope") +
  scale_color_manual(values=c( "darkgoldenrod"))+
  ylim(1.5,2.3)
Slopes_chk

#averaging the slopes from the log and printing
avg_slopes <- log %>%
  group_by(Test) %>%
  summarise(avg_slope = mean(Slope, na.rm = TRUE))
knitr::kable(avg_slopes, caption = "Average Slope by Analyte", digits = 3)


#write out the log file with the added lines for this run date 
write.csv(log, file_path)

#write out a flag to the sample dataframe if the r2 is above the cutoff of 0.98
df_all <- df_all %>%
  mutate(
    NH3_flag = if (NH3_r2 <= r2_cutoff) {
      "NH3 r2 low"
    } else {
      ""
    }
  )



```
 
 \newpage


## Dilution Corrections - ensure the latest dilution is kept

#Dilutions
```{r Dilution Corrections, echo=FALSE}
df_all_cor <- df_all %>% na.omit() %>% group_by(Run_Number, Run_Date, Sample_Name, Test) %>% mutate(Keep = ifelse(Dilution == max(Dilution), "Keep", "Rmv")) %>% filter(Keep == "Keep") %>% ungroup() %>% mutate(
     Pair_ID = ifelse(
       grepl("Duplicate", Sample_Name, ignore.case = TRUE),
       paste0(lag(Sample_Name), "_", lag(Run_Number)),  # pair only if same run
       paste0(Sample_Name, "_", Run_Number)
     ))
```

```{r Dilution Corrections, echo=FALSE}

# cat("Dilution Corrections")
# 
# #Calculate the concentration when accounting for the dilution factor if applicable
# 
# #checking to see if duplicate samples (mislabeled/wrongly inputted/reran)
# df_duplicates <- df_all %>% 
#   filter(!if_all(everything(), is.na)) %>%                    #removing NA rows
#   filter(!(Sample_Name %in% c("Duplicate",                    #removing duplicates, autospikes, ccv/bs 
#                               "CCV", "CCB", "1mg/L ammonia", "DI",
#                               "Blank", "Auto Spike", 
#                               "0.15 mg/L")) &
#   !startsWith(Sample_Name, "Standard ") &               #removing standards
#   !startsWith(Sample_Name, "Nitrite ") &                #removing standards
#   !startsWith(Sample_Name, "Nitrate ") &                #removing standards
#   !startsWith(Sample_Name, "Abs_Chk_") &                #removing abs_chks
#   !startsWith(Sample_Name, "peChk_") &                 #removing peChks
#   !grepl("^\\d{1,2}/\\d{1,2}/\\d{4}\\s+\\d{1,2}:\\d{2}$", Sample_Name) & #removing naything that is a date
#   !grepl("^\\d+$", Sample_Name)) %>%                    #removing any rows with just numbers in the sample name column
#   group_by(Test, Sample_Name)%>%                              #grouping by test & sample name to avoid fake dups
#   filter(n() > 1) %>%                                         #count how many times each Sample_Name appears
#   ungroup() 
# 
# #pulling them out into a string and printing them out
# dil_dups_string <- df_duplicates %>%
#   distinct(Sample_Name) %>%
#   pull(Sample_Name) %>%
#   paste(collapse = ", ")
# 
# #Report duplicated samples
# if (dil_dups_string != "") {
#   message("Duplicated samples: ", dil_dups_string)
#   
#   bad_dups <- df_duplicates %>%
#     group_by(Sample_Name) %>%
#     summarise(
#       n_dups = n(),
#       has_dilution = any(Dilution > 0, na.rm = TRUE)
#     ) %>%
#     filter(n_dups > 1 & !has_dilution) %>%
#     pull(Sample_Name)
#   
#   if (length(bad_dups) > 0) {
#     message("Duplicated samples with NO dilution present (possible input error): ",
#             paste(bad_dups, collapse = ", "))
#   } else {
#     message("All duplicated samples have valid dilutions. No naming issues detected.")
#   }
#   
# } else {
#   message("No duplicated samples.")
# }
# 
# #This is to pick out the dilutions and keep the last run
# df_all_cor <- df_all %>%
#   arrange(row_number()) %>%
#   mutate(
#     Dilution = case_when(
#       Dilution == "0" ~ 1.0,
#       TRUE ~ as.numeric(Dilution)
#     ),
#     Pair_ID = ifelse(
#       grepl("Duplicate", Sample_Name, ignore.case = TRUE),
#       paste0(lag(Sample_Name), "_", lag(Run_Number)),  # pair only if same run
#       paste0(Sample_Name, "_", Run_Number)
#     ),
#     Original_Name = Sample_Name
#   ) %>%
#   group_by(Pair_ID, Test, Run_Number) %>%  # group by Pair_ID, Test, and Run_Number
#   filter(
#     Dilution == ifelse(any(Dilution > 1, na.rm = TRUE), max(Dilution, na.rm = TRUE), 1)
#   ) %>%
#   ungroup()
# 
# 
# 

```


## Performance Check
```{r Performance Check, echo = FALSE}

NH3_peChk <- df_all_cor %>%
  filter(Test == "Ammonia 2", 
         Pair_ID %ilike% "pe") %>% group_by(Run_Date) %>%
  mutate(
    Mean_Chk_Conc = mean(Conc, na.rm = TRUE),
    SD_Chk_Conc = sd(Conc, na.rm = TRUE),
    CV = cv(Conc),
    pct_diff_Chk = 100*(abs(Mean_Chk_Conc - NH3_pe) / NH3_pe),
    NH3_pe_flag = ifelse(abs(pct_diff_Chk) < 25, "YES, Pass", "NO, rerun"),
    Run_Date = Run_Date) %>% dplyr::select(Pair_ID, Mean_Chk_Conc, Run_Date, pct_diff_Chk, NH3_pe_flag) %>% distinct()

#flag write out 

Print <- NH3_peChk %>% mutate(Print = ifelse(pct_diff_Chk >= chk_flag, 
       paste0(Run_Date, "--NH3 pe Check has a % Difference >25% - REASSESS"),  
       paste0(Run_Date, "--NH3 pe Check has a % Difference <25% - PROCEED")))

Print$Print

PE_Plot <- ggplot(data = NH3_peChk, aes(x = Run_Date, y = pct_diff_Chk, fill = NH3_pe_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES, Pass" = "darkgreen", "NO, rerun" = "darkred"), name = "PE Check Pass?") +
  geom_hline(yintercept = 25 , linetype = "dashed" ) +
  theme_classic() + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8))+
   labs(title = "Performance Check") +
  ylab("Percent Difference") +
  xlab("Run Date")

PE_Plot



```

\newpage

## Analyze the Check Standards
```{r Check Standards, echo=FALSE}

cat("Analyze Check Standards")

#Pull out check standards from raw file 
NH3_chks <- df_all %>%
  filter(str_detect(Sample_Name, "1mg/L ammonia"))%>% 
  mutate(rep = row_number())

#RSV of standards 
chks_NH3_rsd <- NH3_chks %>% group_by(Run_Date) %>% mutate(Mean_Conc = mean(Conc), SD_Conc = sd(Conc), CHK_rsd = (SD_Conc)/Mean_Conc, PCT_Diff = 100*abs(Conc - NH3_CCV)/NH3_CCV, NH3_diff_flag = ifelse(PCT_Diff <= chk_conc_flag, 'YES', 'NO, rerun'), RSD_flag = ifelse(CHK_rsd >= chk_flag, 
       paste0(Run_Date,"--","NH3 CHECK STANDARD RSD TOO HIGH - REASSESS"),
       paste0(Run_Date,"--","NH3 Check Standard RSD within Range - PROCEED")), CHK_60_Flag = ifelse(length(NH3_diff_flag[which(NH3_diff_flag == 'No, rerun')])/length(NH3_diff_flag) >= 40,paste0(Run_Date,"-- ", "<60% of NH3 Check Standards are within range of expected concentration - REASSESS") , paste0(Run_Date,"-- ",">60% of NH3 Check Standards are within range of expected concentration - PROCEED"))) %>% dplyr::select(Run_Date, CHK_rsd, Conc, Mean_Conc, SD_Conc, PCT_Diff,NH3_diff_flag, RSD_flag, CHK_60_Flag) %>% ungroup() %>% distinct()
#write out to user about whether or not to continue 


DISP_RSD_Flag <- chks_NH3_rsd %>% dplyr::select(RSD_flag) %>% distinct()
DISP_RSD_Flag

#calculate percent difference between check standards & expected concentration 
#flag if the percent difference is over X% (defined in setup)
#calculate percent difference of check standards 


#plot the check standards

Plot_data <- chks_NH3_rsd %>% dplyr::select(Conc, Run_Date, NH3_diff_flag) %>% distinct()

NH3_Check_Plot <- ggplot(data = Plot_data, 
                         aes(x = Run_Date,
                             y = Conc,
                             fill=NH3_diff_flag)
                         ) +
       geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred"), name ="% Difference <10%") +
       theme_classic() + 
       labs(x= "Run_Date", y="NH3  (mg/L)", title="Check Stds: NH3") + 
       theme(legend.position="bottom") +  
       geom_hline(yintercept=NH3_CCV, 
                  linetype="dashed", 
                  color = "black",
                  linewidth=1) 

NH3_Check_Plot


#calculate the percent of check standards that are within the range based on the flag 
NH3_chks_percent <- (sum(NH3_chks$NH3_diff_flag == "YES")/nrow(NH3_chks))*100

#report out if flags indicate need for rerun

CHK_60_flag <- chks_NH3_rsd %>% dplyr::select(CHK_60_Flag) %>% distinct()
CHK_60_flag
#write out a flag to the sample dataframe if less than 60% of the checks are within the expected CV
if (NH3_chks_percent <= chks_flag) {  
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 checks out of range"),
    "NH3 checks out of range"
  )
}

```

\newpage

## Analyze Blanks 
```{r Analyze Blanks, echo=FALSE}

cat("Assess Blanks")

#Pull out check standards from raw file 
NH3_blks <- df_all %>%
  filter(Test == "Ammonia 2")%>% 
  filter(str_detect(Sample_Name, "Blank"))%>% 
  mutate(rep = row_number())


#Pull out samples from df_all to calc quantile
samples <- df_all %>%  
  filter(str_detect(Sample_Name, c("L"))) %>%  
  filter(!str_detect(Sample_Name, c("1mg/L ammonia"))) #have to do this because Phrag ids start with L 

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

NH3_only <- samples %>%
  filter(Test == "Ammonia 2")
blk_flag_NH3 <- NH3_only %>% group_by(Run_Date) %>% mutate(Quant = quantile(Conc, prob=c(.25))   ) %>% dplyr::select(Run_Date, Quant) %>% ungroup() %>% distinct()
                                                           
NH3_blks_qt <- left_join(NH3_blks, blk_flag_NH3, by = join_by(Run_Date))
                                           
#this gives you the lower 25% quartile of the data 
NH3_flagged <- NH3_blks_qt %>% group_by(Run_Date) %>% mutate(NH3_diff_flag = ifelse(Conc <= Quant, 'YES', 'NO, rerun'), BLK_60_Flag = ifelse(length(NH3_diff_flag[which(NH3_diff_flag == 'No, rerun')])/length(NH3_diff_flag) >= 40,paste0(Run_Date,"-- ", "<60% of NH3 Blanks are below the lower 25% quartile of samples - REASSESS") , paste0(Run_Date,"-- ",">60% of NH3 Blanks are below the lower 25% quartile of samples - PROCEED"))) %>% ungroup()

#calculate the percent of check standards that are within the range based on the flag 

#report out if flags indicate need for rerun
BLK_60_disp <- NH3_flagged %>% dplyr::select(BLK_60_Flag)  %>% distinct()

BLK_60_disp

Blks_Plot_df <- NH3_flagged %>% dplyr::select(Run_Date, Conc, NH3_diff_flag, Quant)

#plotting the blanks compared to the lower 25% of conc (what the flag is)
NH3_blks_plot <-  ggplot(data = Blks_Plot_df , aes(x = Run_Date, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
  facet_grid(.~Run_Date, scales = "free_x") +
       theme_classic() + labs(x= " ", y="NH3  (mg/L)", title="NH3 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept=as.numeric(Quant)), linetype="dashed",
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

print(NH3_blks_plot)


#find average of run blanks for flagging samples later 
# Compute averages
blk_avg_NH3 <- mean(NH3_blks$Conc, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(
  Test = c("NH3"),
  Blank_Mean_Conc = c(blk_avg_NH3)
)

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Concentration of Blanks", digits = 4)


#write out a flag to the sample dataframe if more than 60% of the blanks are above the lower 25% quantile of samples
if (NH3_blks_percent <= chks_flag) {  # assuming you have tn_chks_percent similarly
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 blanks out of range"),
    "NH3 blanks out of range"
  )
}


```

\newpage

## Analyze Duplicates
```{r Analyze Duplicates, echo=FALSE}

cat("Analyze Duplicates")

Duplicates <- df_all_cor %>% filter(Sample_Name == "Duplicate") %>% mutate(Sample_Num = Run_Number - 1) %>% mutate(Date_Num = paste0(Sample_Num,"_", Run_Date, "_", Test)) %>% dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% rename(Dup_Conc = Conc, Dup_Absorbance = Absorbance, Dup_Dilution = Dilution)

Dup_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test))

Dup_Joined <- left_join(Duplicates, Dup_ready, by = join_by(Date_Num))

Dup_plot_df <- Dup_Joined %>% filter(!Sample_Name %ilike% "pe") %>% group_by(Date_Num) %>% mutate(Duplicate_CV = cv(c(Dup_Conc, Conc))) %>% mutate(Dup_Flag = ifelse(Duplicate_CV >= 10, "NO, rerun", "YES")) %>% ungroup() %>% group_by(Sample_Name) %>% ungroup() %>% group_by(Sample_Name) %>% mutate(Num_Runs = length(Sample_Name), Max_Dil = max(Dilution), Rerun = ifelse(Num_Runs != 1 , ifelse(Dilution == Max_Dil, "KEEP", "REMOVE"), "KEEP")) %>% filter(Rerun == "KEEP")

Dup_Plot <-  ggplot(Dup_plot_df, aes(x = Run_Date, y = Duplicate_CV, fill =  Dup_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "NO, rerun" = "darkred"), name = "CV below 10% ?") +
  geom_hline(yintercept = 10, linetype = "dashed" ) +
  facet_grid(Test~Run_Date, scales = "free")  + 
  theme_classic() + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8)) +
   labs(title = "Duplicate Check")

Dup_Plot


Dup_Flags <- Dup_plot_df %>% group_by(Run_Date, Test) %>% mutate(Dup_Flags = ifelse(length(which(Dup_Flag == "DUP PASS"))/(length(Dup_Flag)) >= 0.6,  "Dups Pass", "Dups Fail")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Dup_Flags) %>% distinct()



```

\newpage

## Spikes
```{r Analyze Spikes, echo=FALSE}

#getting spikes and samples pulled 
df_NH3_spk <- df_all_cor %>%
  filter(Test == "Ammonia 2") %>%
  filter(
     !grepl("CCV|CCB|Blank|1mg/L ammonia|Standard|Duplicate", Sample_Name, ignore.case = TRUE),
     #^removes dup, ccv, ccb, standards
     !is.na(Sample_Name),                                         # Remove NA
     !grepl("^\\d+$", Sample_Name),                               # removes all numbers
     !grepl("^\\d{1,2}/\\d{1,2}/\\d{2,4}", Sample_Name)) %>%      # removes all date/time
  mutate(`Spike?` = if_else(                                      # adds duplicate column
     grepl("Auto Spike", Sample_Name, ignore.case = TRUE),        # keeps sample ID if there
     lag(Sample_Name),                                            # get previous row's SampleID
     Sample_Name)) %>%                                            # keeps sample ID if not duplicate
  group_by(`Spike?`) %>%                                          # groups by duplicate
  filter(n() > 1) %>%                                             # filters anything that appears more than once
  ungroup()  %>%
  transmute(
    Pair_ID = `Spike?`,
    Sample_Name,
    Conc,
    Type = if_else(str_detect(Sample_Name, "Auto Spike"), "Auto_Spike", "Original"))
 

#checking and flagging against the expected spike conc
df_spk_NH3_chk <- df_NH3_spk %>%
  dplyr::select(Pair_ID, Conc, Type) %>%
  pivot_wider(
    names_from = Type,
    values_from = Conc,
    values_fn = mean,  # in case there are duplicates
    names_prefix = "" ) %>%
  rename(
    Spike_Conc = Auto_Spike,
    OG_Conc = Original) %>%
  filter(!is.na(Spike_Conc), !is.na(OG_Conc)) %>%
  rowwise() %>%# keep only rows with both
   mutate(
    Exp_Conc = OG_Conc + NH3_Spk,
    Mean_Conc = mean(c(Spike_Conc, Exp_Conc), na.rm = TRUE),
    Pct_Diff = abs(Spike_Conc - Exp_Conc) / Mean_Conc * 100,
    SD_Conc = sd(c(Spike_Conc, Exp_Conc)),
    CV = (SD_Conc / Mean_Conc) * 100,
    Spike_diff_flag = ifelse(CV < 50, "YES", "NO, check"),
    Test = "Ammonia 2") %>%
    ungroup()

# Combining Auto Spike data frames
df_spk_combo <- bind_rows(df_spk_NH3_chk)

#calculate the percent of check standards that are within the range based on the flag
NH3_spk_percent <- (sum(df_spk_NH3_chk$Spike_diff_flag == "YES")/nrow(df_spk_NH3_chk))*100
 
#spk flag report out
ifelse(NH3_spk_percent >= chks_flag, 
       ">60% of Spikes have a CV <50% - PROCEED",
       "<60% of Carbon Spikes have a CV <50% - REASSESS")

#plot spikes output as a bar graph to easily check - want any over 10% to be red need to work on this 
NH3_spk_plot <- ggplot(data = df_spk_NH3_chk, aes(x =Pair_ID, y =CV, fill=Spike_diff_flag)) +
        geom_bar(stat = 'identity') + 
        theme_classic() + 
        labs(x= "Sample ID", y="NH3 N(mg/L)", title = "NH3 Auto Spike CV") + 
        scale_fill_manual(values = c("YES" = "darkcyan", "NO, check" = "darkred")) +
        theme(legend.position="none")  + 
        geom_hline(yintercept=50, linetype="dashed", 
            color = "black", size=1)  +
        guides(fill=guide_legend(title="CV Between Spks <50%"))+
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
 
print(NH3_spk_plot)


#write out a flag to the sample dataframe if more than 60% of the dups have CVs out of range
if (NH3_spk_percent <= chks_flag) {  
    df_all_cor$NH3_flag <- ifelse(
    df_all_cor$NH3_flag != "",
    paste0(df_all_cor$NH3_flag, "; NH3 spikes out of range"),
    "NH3 spikes out of range"
  )
}

```

\newpage

## Matrix Effects  
```{r Analyze Matrix Effects, echo=FALSE}

df_matrix <- df_spk_combo%>%
  filter(str_detect(Pair_ID, "10ppt")) 
  
df_matrix_NH3 <- df_matrix %>%
  filter(str_detect(Test, "Ammonia")) 
  ifelse(
  all(df_matrix_NH3$Spike_diff_flag == "YES"),
  "NO NH3 Matrix Effect, PROCEED",
  ">20% CV in ASW NH3 matrix effect checks - REASSESS"
)

#write out a flag to the sample dataframe if more than 60% of the dups have CVs out of range
if (any(df_matrix$Spike_diff_flag[df_matrix$Test == "Ammonia 2"] != "YES")) {
  
    df_all_cor$NOx_flag <- ifelse(
    df_all_cor$NOx_flag != "",
    paste0(df_all_cor$NOx_flag, "; NH3 matrix check out of range"),
    "NH3 matrix check out of range"
  )

}


```

## Unit Converted Data Column Added (mg/L to uM )
```{r Add unit conversion, include=FALSE}

#Pull out samples 
samples <- df_all_cor %>%  
   filter(str_detect(Sample_Name, c("L"))) %>%  
  filter(!str_detect(Sample_Name, c("1mg/L ammonia"))) #have to do this because Phrag ids start with L   

#Convert values based on the Test ID 
samples <- samples %>%
  mutate(
    Conc_uM = case_when(
      Test == "Vanadium NOx" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Ammonia 2" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "o-PHOS 0.3" ~ (((as.numeric(samples$Conc))/Con1)/P_mw)*Con2,
      TRUE ~ as.numeric(NA)  # fallback in case of unexpected value
    ),
    # Replace negatives with 0 in 'value' and 'value_converted'
    Conc = pmax(Conc, 0),
    Conc_uM = pmax(Conc_uM, 0)
  )

head(samples)

```

## Sample Flagging - Within range of standard curve
```{r Sample Flagging, echo=FALSE}

cat("Sample Flagging")

#Flagging data if the concentration is outside the standards range 
samples_flagged <- samples %>%
  mutate(
     Conc_flag = case_when(
        Test == "Vanadium NOx" & Conc/Dilution < NOx_dl ~ "bdl",
        Test == "Vanadium NOx" & Conc/Dilution > NOx_top    ~ "adl",
        Test == "Vanadium NOx"               ~ "Within_Range",
      
      Test == "Ammonia 2" & Conc/Dilution < NH3_dl ~ "bdl",
      Test == "Ammonia 2" & Conc/Dilution > NH3_top    ~ "adl",
      Test == "Ammonia 2"                ~ "Within_Range",
      
      Test == "o-PHOS 0.3" & Conc/Dilution < PO4_dl  ~ "bdl",
      Test == "o-PHOS 0.3" & Conc/Dilution > PO4_top   ~ "adl",
      Test == "o-PHOS 0.3"                ~ "Within_Range",
      
      TRUE ~ NA_character_  # fallback for unexpected values
    )
  )

```

## Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata 
all_present <- all(metadata$Sample_Name %in% samples_flagged$Sample_Name)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, samples_flagged$Sample_Name)
  print(missing_ids)
}


#merge metadata with sample run data 
merged_data <- samples_flagged %>%
  left_join(metadata, by = c("Sample_Name" = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included? 

df_all_clean <- merged_data 

```

\newpage



## Organize Data
```{r Organize Data, include=FALSE}

cat("Export Processed Data")

df_all_clean_cols <- merged_data %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Sample_Name,
    Plot_Name,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    Conc,
    Conc_uM,
    Conc_flag,
    NH3_flag
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = run_notes,# new column with notes about the run
    Analysis_rundate = run_date) 

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
   NH3_Conc = Conc,
   NH3_Conc_uM = Conc_uM,
    Plot = Plot_Name,
   Sample_ID = Sample_Name
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Sample_ID,
    Plot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    NH3_Conc,
    NH3_Conc_uM,
    Conc_flag,
    NH3_flag,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)


```

## Visualize Data
```{r Visualize Data, echo=FALSE}

NH3_forplot <- All_Clean_Data

#ensure points are in correct order

NH3_forplot <- NH3_forplot[order(NH3_forplot$Depth_cm), ]

color1 <- c("L1", "L2", "L3", "L4")
color2 <- c("L5", "L6", "L8", "L10")
color3 <- c("L7", "L9", "L11", "L12")
color4 <- c("L13", "L14", "L16", "L15")
color5 <- c("L17", "L18", "L19", "L20")

NH3_forplot <- NH3_forplot %>%
  mutate(
    Plot_Color = case_when(
      Plot %in% color1 ~ "Color1",
      Plot %in% color2 ~ "Color2",
      Plot %in% color3 ~ "Color3",
      Plot %in% color4 ~ "Color4",
      Plot %in% color5 ~ "Color5"
    )
  ) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2 & eN",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2 & aN",
      ifelse(
        CO2_Treatment == "Ambient" &
          N_Treatment == "Amended"  ,
        "aCO2 & eN" ,
        "aCO2 & aN"
      )
    )
  ))  %>%
  mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2"))  %>%
  mutate(N_Treatment = ifelse(N_Treatment == "Ambient", "aN", "eN")) 
#need to work out how best to plot these data

viz_NH3_plot2 <- ggplot(data = NH3_forplot, aes(
  x = factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")),
  y = NH3_Conc,
  fill = as.factor(Treatment),
  color = as.factor(Treatment)
)) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(padding = 0.25,preserve = "single")
  ) +
  facet_grid(Depth_cm ~ N_Treatment, scales = "fixed") +
  scale_fill_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "white",
    "aCO2 & aN" = "white"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
    
  ) +
  scale_color_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "#DB5461",
    "aCO2 & aN" = "#454ADE"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
  ) +
  theme_classic() +
  labs(x = " ",
       y = "NH3 (mg/L)",
       title = "CO2xN: Porewater Ammonium"
       ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  ) +
coord_flip()
print(viz_NH3_plot2)

```



```{r, echo=FALSE}
#ensure points are in correct order

#make plot
viz_NH3_plot_2 <- ggplot(data = NH3_forplot, aes(x = NH3_Conc, y = as.numeric(Depth_cm), color = Plot_Color)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(
    values = c(
      "Color1" = "#FFD500",
      "Color2" = "#00D0FF",
      "Color3" = "#FF0075",
      "Color4" = "#80d280",
      "Color5" = "#ff6040"
    ),
    labels = c("1", "2", "3", "4", "5")
  ) +
  facet_grid(Treatment ~ ., scales = "fixed") +
  labs(
    x = "NH3 (mg/L)",
    y = "Depth (cm)",
    title = "CO2xN: Porewater NH3 by Replicate",
    color = "Replicate"
  ) +
  scale_y_reverse(breaks = c(20, 40, 80), limits = c(90, 10)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_NH3_plot_2
```


## Export Processed Data
```{r Export Processed Data, include=FALSE}

#Write out data frame 
  write.csv(All_Clean_Data, final_path)
  

```


#end