---
title: "CO2xN: Porewater Ammonium"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information 
```{r run information, include=TRUE}
cat("Run Information: ") #lets you know what section you're in 
#set the run date & user name 
  run_date <- "01/30/2025"
  sample_year <- "2025"
  sample_month <- "Sept"
  user <- "Victoria Ellis"

#identify the files you want to read in 
  #read in as a list to accommodate multiple runs in a month
  NH3_files <- c("Raw Data/LTREB_2M_Sept_2025_Run1.csv",
                 "Raw Data/LTREB_2M_Sept_2025_Run2.csv")

# Define the file path for QAQC log file - NO Need to change just check year 
  final_path <- "Processed Data/CO2xN_Porewater_NH4_202509.csv"

#record any notes about the run or anything other info here: 
  run_notes <- ""
  
  #Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

  cat(run_notes)
```

##Setup
```{r setup, include=FALSE}

#let you know which section you are in 
cat("Setup")

#load/install packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  DescTools,
  purrr,
  ggplot2,
  ggpubr,
  tidyr,
  tidyverse,
  lubridate,
  tinytex,
  stringr,
  writexl,
  readr,
  readxl,
  purrr,
  tinytex,
  broom,
  here, 
  data.table,
  raster,
  curl,
  googledrive,
  httr,
  gargle)

Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

#file_path_QAQC = here("..", "GCReW", "SEAL_Check_QAQC_Log.csv")

#Coefficients / constants that are needed for calculations 
N_mw <- 14.0067    # molecular weight of N 
P_mw <- 30.973762  # molecular weight of P 
Con1 <- 1000       # conversion factor value
Con2 <- 1000000    # conversion factor value 

#Detection limit and top standards for flagging 
NOx_dl <- 0.025  #mgL
NH3_dl <- 0.02  #mgL
PO4_dl <- 0.003 #mgL

NOx_top <- 1  #mgL
NH3_top <- 2  #mgL
PO4_top <- 0.3 #mgL

#Check Standard concnetrations 
NOx_CCV <- 0.5 
NH3_CCV <- 1.0 
PO4_CCV <- 0.15 

#Spike Concentrations 
NOx_Spk <- 0.5 #ppm or mg/L
NH3_Spk <- 1
PO4_Spk <- 0.152 

#expected ranges for sample concentrations used for flags 
r2_cutoff = 0.990
chk_flag = 0.25
chk_conc_flag = 15 #cutoff level for percent difference of check stds vs. expected concentration 
chks_flag = 60
rep_flag = 25  #this is a 25% error between samples
#blank_flag is calculated based on samples later in this code

#pe check Concentrations
NH3_pe <- 1.034
NOx_pe <- 1.51
PO4_pe <- 0.824

```


## Enable Google Drive Access and Retrieve QAQC Files
```{r Enable Google Drive, echo = FALSE}


my_token <- readRDS("Token/httr-oauth") 

drive_auth(token = my_token, scopes = c("https://www.googleapis.com/auth/drive.file", "https://www.googleapis.com/auth/drive"), email = "blakleybd@si.edu")


drive_download("GCReW_SEAL_StdCrv_QAQC_Log.csv", overwrite = TRUE)
file_path <-"GCReW_SEAL_StdCrv_QAQC_Log.csv"

drive_download("SEAL_Check_QAQC_Log.csv", overwrite = TRUE)
file_path_QAQC <- "SEAL_Check_QAQC_Log.csv"


# Get Authorization/Refresh Token for Offline Use
 # google_client <- gargle::gargle_oauth_client_from_json("S:/Biogeochemistry/People/Brenden Blakley/client_secret_949417754999-vi67ohqh65kq1ugiui68sfrl9e7a4quq.apps.googleusercontent.com.json")
 # app <- httr::oauth_app(appname = "myapp", key = google_client$id, secret = google_client$secret)
 # endpoint <- httr::oauth_endpoints("google")
 # token <- httr::oauth2.0_token(
 #     endpoint = endpoint,
 #     app = app,
 #     scope = "https://www.googleapis.com/auth/drive",
 #     cache = "Token/httr-oauth" # Saves the token for future use
 # )



```

##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata)

#clean up columns in metadata file:
  #Pull out only SMARTX Rows 
  #Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("2-Meter", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  ) %>%
  dplyr::select(-Temp_Treatment) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2_N",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2",
      ifelse(CO2_Treatment == "Ambient" &
               N_Treatment == "Amended"  , "N" , "Ambient")
    )
  )) 
#Bring in the collection date and any field notes if relevant 
#collection_metadata <- read.csv(Collection_Dates)
  #then we will merge the collection date for SMARTX based on the project and the chambers 

```

\newpage

## Import Data & Clean
```{r Import Data, include=FALSE}
cat("Import Data")

#set file path for data - in run info chunk 
path <- ("file path")

#Read in Raw Data 
data_NH3 <- map(NH3_files, read.csv)

# Combining Files
df_combo <- bind_rows(data_NH3)

# Select needed columns & rename 
df <- df_combo %>%
  dplyr::select(Sample_Name = 4, 
         Run_Number = 5,
         Conc = 6, 
         Absorbance = 7, 
         Dilution = 9, 
         Unit = 12,
         Test = 13, 
         Run_Time = 15)

#Checking column headers
head(df)

df_all <- df %>% mutate(Run_Number = as.numeric(Run_Number),
         Run_Date = format(as.POSIXct(Run_Time,format='%m/%d/%Y %H:%M'),format='%m/%d/%Y'))

#Need to add some extra code to deal with July and Sept being run in the same file 

#Initialize Flags
DF_Flags <- df_all %>% dplyr::select(Run_Date) %>% distinct() %>% na.omit()


```



## Assessing standard Curves
# Assess Standard Curves
```{r, echo=FALSE}
split_data <- df_all %>% dplyr::select(Sample_Name, Conc, Absorbance, Test, Run_Date) %>% filter(Sample_Name %ilike% "Standard") %>%
  mutate(Conc = ifelse(Test == "Ammonia 2", case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard .0389" ~ 0.0389,
    Sample_Name == "Standard .1000" ~ 0.1,
    Sample_Name == "Standard .2000" ~ 0.2,
    Sample_Name == "Standard .5000" ~ 0.5,
    Sample_Name == "Standard 1.0000" ~ 1.0,
    Sample_Name == "Standard 1.5000" ~ 1.5,
    Sample_Name == "Standard 2.0000" ~ 2),
    ifelse(Test == "o-PHOS 0.3", 
           case_when(
    Sample_Name == "Standard 0" ~  0.0,
    Sample_Name == "Standard 1" ~  0.0,
    Sample_Name == "Standard .006" ~ 0.0060,
    Sample_Name == "Standard .0150" ~ 0.0150,
    Sample_Name == "Standard .03" ~ 0.0300,
    Sample_Name == "Standard .075" ~ 0.0750,
    Sample_Name == "Standard .15" ~ 0.1500,
    Sample_Name == "Standard .225" ~ 0.2250,
    Sample_Name == "Standard .3" ~ 0.3000),
    "Test Not Found"))) %>%
  group_by(Run_Date, Test) %>% 
  na.omit() %>%
  group_split()
  
#run regression and get coefficients by Run
 
Slope_raw_list <- map(split_data, ~ .x %>% mutate(
                                        R2 = summary(lm(Conc ~ Absorbance, data = .))$adj.r.squared , 
                                        Slope = coef(lm(Conc ~ Absorbance, data = .))[2],
                                        Intercept = coef(lm(Conc ~ Absorbance, data = .))[1]))


 Slope_raw <- bind_rows(Slope_raw_list)
  
 #check R2 and flag
 Slope_Std_Crvs <- Slope_raw %>% dplyr::select(Run_Date, R2, Slope, Intercept, Test) %>% distinct() %>% mutate(R2_Flag = ifelse(R2 >= r2_cutoff, "R2_Pass", "R2_Fail"))

 #join standard curve CVs and regression values
 
 R2_Check <- Slope_Std_Crvs %>% 
   dplyr::select(Run_Date,
                 R2,
                 Slope,
                 Intercept,
                 Test,
                 R2_Flag) %>%
   distinct()
 
 For_crv_plot <- Slope_raw %>% mutate(Conc =  as.numeric(Conc))

 Curves_Plot <- ggplot(For_crv_plot, aes(Conc, Absorbance, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", formula = "y~x",se = FALSE) + facet_grid(.~Test, scales = "free") + labs(title = "Standard Curves") + theme_classic()
 
Curves_Plot
 

R2_Flags <- R2_Check %>% dplyr::select(Test, Run_Date, R2_Flag)
 
```

 
 \newpage
## Update Standard Log
```{r, echo = FALSE}

#log <- tibble(Test = c(), Intercept= c(), Slope= c(), Quadratic= c(), Run_Date= c(), User= c()) #create standard log in proper format

old_log <- read.csv(file_path) %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% mutate(Run = "Past Runs")#read in log and format date

add_log <- R2_Check %>% dplyr::select(Test, Run_Date, R2, Slope, Intercept) %>%
  mutate(Test = case_when(
    Test == "Ammonia 2" ~ "NH3",
    Test == "Vanadium NOx" ~ "NOx",
    Test == "o-PHOS 0.3" ~ "PO4"
  )) %>% mutate(Run_Date = as.Date(as.character(Run_Date), format = "%m/%d/%Y")) %>% mutate(User = user) %>% mutate(Run = "Current Runs")#Format new run data for log


new_log <- rbind(old_log, add_log) %>% mutate(across(where(is.numeric), \(x) round(x, digits = 4))) %>% distinct() #bind rows, round data, and remove possible duplicate data from rerunning code
mean_slope = mean(new_log$Slope)

sd_slope = sd(new_log$Slope)

mean_int = mean(new_log$Intercept)

sd_int = sd(new_log$Intercept)
#Plot Slope log
Slopes_chk <- ggplot(new_log, aes(x = Run_Date, y = Slope, col = Run)) +
    ylim(1.25,2.25) +
  scale_color_manual(values = c("Past Runs" = "grey27", "Current Runs" = "cyan3")) +
  geom_abline(intercept = mean_slope - 2*sd_slope, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_abline(intercept = mean_slope + 2*sd_slope, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_point(size = 2) +
  theme_classic()+
  ggtitle("Slope Log") 

#Plot Intercept log
ints_chk <- ggplot(new_log, aes(x = Run_Date, y = Intercept, col = Run)) +
  ylim(-0.09, 0.01) +
  scale_color_manual(values = c("Past Runs" = "grey27", "Current Runs" = "cyan3")) +
  geom_abline(intercept = mean_int - 2*sd_int, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_abline(intercept = mean_int + 2*sd_int, slope = 0, linetype = "dashed", linewidth = 1.5, col = "red" )+
  geom_point(size = 2) +
  theme_classic()+
  ggtitle("Intercept Log") 

Slopes_chk

ints_chk

```

## Write Standard Log to Google Drive
```{r Write Standard Log, echo = FALSE}

write.csv(new_log, file_path, row.names = FALSE) #write out log

drive_upload(file_path, name = "GCReW_SEAL_StdCrv_QAQC_Log.csv", overwrite = TRUE)

```

## Dilution Corrections - ensure the latest dilution is kept

#Dilutions
```{r Dilution Corrections, echo=FALSE}
df_all_cor <- df_all %>% na.omit() %>% group_by(Run_Number, Run_Date, Sample_Name, Test) %>% mutate(Keep = ifelse(Dilution == max(Dilution), "Keep", "Rmv")) %>% filter(Keep == "Keep") %>% ungroup() %>% mutate(
     Pair_ID = ifelse(
       grepl("Duplicate", Sample_Name, ignore.case = TRUE),
       paste0(lag(Sample_Name), "_", lag(Run_Number)),  # pair only if same run
       paste0(Sample_Name, "_", Run_Number)
     ))
```




## Performance Check
```{r Performance Check, echo = FALSE}

NH3_peChk <- df_all_cor %>%
  filter(Test == "Ammonia 2", 
         Pair_ID %ilike% "pe") %>% group_by(Run_Date) %>%
  mutate(
    Mean_Chk_Conc = mean(Conc, na.rm = TRUE),
    SD_Chk_Conc = sd(Conc, na.rm = TRUE),
    CV = cv(Conc),
    pct_diff_Chk = 100*(abs(Mean_Chk_Conc - NH3_pe) / NH3_pe),
    NH3_pe_flag = ifelse(abs(pct_diff_Chk) < chk_flag * 100, "YES, Pass", "NO, rerun"),
    Run_Date = Run_Date) %>% dplyr::select(Pair_ID, Mean_Chk_Conc, Run_Date, pct_diff_Chk, Test, NH3_pe_flag) %>% distinct()

#flag write out 

Print <- NH3_peChk %>% mutate(Print = ifelse(pct_diff_Chk >= chk_flag * 100, 
       paste0(Run_Date, "--NH3 pe Check has a % Difference >25% - REASSESS"),  
       paste0(Run_Date, "--NH3 pe Check has a % Difference <25% - PROCEED")))

Print$Print

PE_Plot <- ggplot(data = NH3_peChk, aes(x = Run_Date, y = pct_diff_Chk, fill = NH3_pe_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES, Pass" = "darkgreen", "NO, rerun" = "darkred"), name = "PE Check Pass?") +
  geom_hline(yintercept = 25 , linetype = "dashed" ) +
  theme_classic() + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8))+
   labs(title = "Performance Check") +
  ylab("Percent Difference") +
  xlab("Run Date")

PE_Plot

Flags_PeCheck <- NH3_peChk %>% mutate(PeCHK_Flag = ifelse(pct_diff_Chk >= chk_flag * 100, 
      "NH3 pe Check has a % Difference >25% - REASSESS",  
    "NH3 pe Check has a % Difference <25% - PROCEED")) %>% dplyr::select(Test, Run_Date, PeCHK_Flag)



```

\newpage

## Analyze the Check Standards
```{r Check Standards, echo=FALSE}

cat("Analyze Check Standards")

#Pull out check standards from raw file 
NH3_chks <- df_all %>%
  filter(str_detect(Sample_Name, "1mg/L ammonia"))%>% 
  mutate(rep = row_number())

#RSV of standards 
chks_NH3_rsd <- NH3_chks %>% group_by(Run_Date) %>% mutate(
  Mean_Conc = mean(Conc),
  SD_Conc = sd(Conc),
  CHK_rsd = (SD_Conc) / Mean_Conc,
  PCT_Diff = 100 * abs(Conc - NH3_CCV) / NH3_CCV,
  NH3_diff_flag = ifelse(PCT_Diff <= chk_conc_flag, 'YES', 'NO, rerun'),
  RSD_flag = ifelse(
    CHK_rsd >= chk_flag,
    paste0(Run_Date, "--", "NH3 CHECK STANDARD RSD TOO HIGH - REASSESS"),
    paste0(Run_Date, "--", "NH3 Check Standard RSD within Range - PROCEED")
  ),
  CHK_60_Flag = ifelse(
    length(NH3_diff_flag[which(NH3_diff_flag == 'No, rerun')]) / length(NH3_diff_flag) >= 40,
    paste0(
      Run_Date,
      "-- ",
      "<60% of NH3 Check Standards are within range of expected concentration - REASSESS"
    ) ,
    paste0(
      Run_Date,
      "-- ",
      ">60% of NH3 Check Standards are within range of expected concentration - PROCEED"
    )
  ),
  CHKRSD_Flag = ifelse(
    CHK_rsd >= chk_flag,
    "NH3 CHECK STANDARD RSD TOO HIGH - REASSESS",
    "NH3 Check Standard RSD within Range - PROCEED"
  ),
  CHK_Flag = ifelse(
    length(NH3_diff_flag[which(NH3_diff_flag == 'No, rerun')]) / length(NH3_diff_flag) >= 40,
      "<60% of NH3 Check Standards Pass - REASSESS" ,
      ">60% of NH3 Check Standards Pass - PROCEED"
    
  )
) %>% dplyr::select(
  Run_Date,
  Test,
  CHK_rsd,
  Conc,
  Mean_Conc,
  SD_Conc,
  PCT_Diff,
  NH3_diff_flag,
  RSD_flag,
  CHK_60_Flag,
  CHKRSD_Flag,
  CHK_Flag
) %>% ungroup() %>% distinct()
#write out to user about whether or not to continue 


DISP_RSD_Flag <- chks_NH3_rsd %>% dplyr::select(RSD_flag) %>% distinct()
DISP_RSD_Flag

#calculate percent difference between check standards & expected concentration 
#flag if the percent difference is over X% (defined in setup)
#calculate percent difference of check standards 


#plot the check standards

Plot_data <- chks_NH3_rsd %>% dplyr::select(Conc, Run_Date, NH3_diff_flag) %>% distinct()

NH3_Check_Plot <- ggplot(data = Plot_data, 
                         aes(x = Run_Date,
                             y = Conc,
                             fill=NH3_diff_flag)
                         ) +
       geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) + 
       scale_fill_manual(values = c("YES" = "chartreuse4", "NO, rerun" = "darkred"), name ="% Difference <10%") +
       theme_classic() + 
       labs(x= "Run_Date", y="NH3  (mg/L)", title="Check Stds: NH3") +
  facet_grid(.~Run_Date, scales = "free_x") +
       theme(legend.position="bottom") +  
       geom_hline(yintercept=NH3_CCV, 
                  linetype="dashed", 
                  color = "black",
                  linewidth=1) 

NH3_Check_Plot


#calculate the percent of check standards that are within the range based on the flag 
NH3_chks_percent <- (sum(NH3_chks$NH3_diff_flag == "YES")/nrow(NH3_chks))*100

#report out if flags indicate need for rerun

CHK_60_flag <- chks_NH3_rsd %>% dplyr::select(CHK_60_Flag) %>% distinct()
CHK_60_flag

Flags_Check <- chks_NH3_rsd %>% dplyr::select(Test, Run_Date, CHKRSD_Flag, CHK_Flag) %>% distinct()


```

\newpage

## Analyze Blanks 
```{r Analyze Blanks, echo=FALSE}

cat("Assess Blanks")

#Pull out check standards from raw file 
NH3_blks <- df_all %>%
  filter(Test == "Ammonia 2")%>% 
  filter(str_detect(Sample_Name, "Blank"))%>% 
  mutate(rep = row_number())


#Pull out samples from df_all to calc quantile
samples <- df_all %>%  
  filter(str_detect(Sample_Name, c("L"))) %>%  
  filter(!str_detect(Sample_Name, c("1mg/L ammonia"))) #have to do this because Phrag ids start with L 

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

NH3_only <- samples %>%
  filter(Test == "Ammonia 2")
blk_flag_NH3 <- NH3_only %>% group_by(Run_Date) %>% mutate(Quant = quantile(Conc, prob=c(.25))   ) %>% dplyr::select(Run_Date, Quant) %>% ungroup() %>% distinct()
                                                           
NH3_blks_qt <- left_join(NH3_blks, blk_flag_NH3, by = join_by(Run_Date))
                                           
#this gives you the lower 25% quartile of the data 
NH3_flagged <- NH3_blks_qt %>% group_by(Run_Date) %>% mutate(NH3_diff_flag = ifelse(Conc <= Quant, 'YES', 'NO, rerun'), BLK_60_Flag = ifelse(length(NH3_diff_flag[which(NH3_diff_flag == 'No, rerun')])/length(NH3_diff_flag) >= 40,paste0(Run_Date,"-- ", "<60% of NH3 Blanks are below the lower 25% quartile of samples - REASSESS") , paste0(Run_Date,"-- ",">60% of NH3 Blanks are below the lower 25% quartile of samples - PROCEED")), Blank_Flag = ifelse(length(NH3_diff_flag[which(NH3_diff_flag == 'No, rerun')])/length(NH3_diff_flag) >= 40,"<60% of NH3 Blanks Pass - REASSESS" , ">60% of NH3 Blanks Pass - PROCEED")) %>% ungroup()

#calculate the percent of check standards that are within the range based on the flag 

#report out if flags indicate need for rerun
BLK_60_disp <- NH3_flagged %>% dplyr::select(BLK_60_Flag)  %>% distinct()

BLK_60_disp

Blks_Plot_df <- NH3_flagged %>% dplyr::select(Run_Date, Conc, NH3_diff_flag, Quant)

#plotting the blanks compared to the lower 25% of conc (what the flag is)
NH3_blks_plot <-  ggplot(data = Blks_Plot_df , aes(x = Run_Date, y = Conc, fill=NH3_diff_flag)) +
       geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) + 
       scale_fill_manual(values = c("YES" = "deepskyblue3", "NO, rerun" = "darkred")) +
  facet_grid(.~Run_Date, scales = "free_x") +
       theme_classic() + labs(x= " ", y="NH3  (mg/L)", title="NH3 Blanks") + 
       theme(legend.position="bottom") +  
       geom_hline(aes(yintercept=as.numeric(Quant)), linetype="dashed",
                color = "black", linewidth=1)  + 
       guides(fill=guide_legend(title="Blank Conc <25% Quartile Samples"))

print(NH3_blks_plot)


#find average of run blanks for flagging samples later 
# Compute averages
blk_avg_NH3 <- mean(NH3_blks$Conc, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(
  Test = c("NH3"),
  Blank_Mean_Conc = c(blk_avg_NH3)
)

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Concentration of Blanks", digits = 4)

Flags_Blank <- NH3_flagged %>% dplyr::select(Test, Run_Date, Blank_Flag) %>% distinct()



```

\newpage

## Analyze Duplicates
```{r Analyze Duplicates, echo=FALSE}

cat("Analyze Duplicates")

Duplicates <- df_all_cor %>% filter(Sample_Name == "Duplicate") %>% mutate(Sample_Num = Run_Number - 1) %>% mutate(Date_Num = paste0(Sample_Num,"_", Run_Date, "_", Test)) %>% dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% rename(Dup_Conc = Conc, Dup_Absorbance = Absorbance, Dup_Dilution = Dilution)

Dup_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test))

Dup_Joined <- left_join(Duplicates, Dup_ready, by = join_by(Date_Num))

Dup_plot_df <- Dup_Joined %>% filter(!Sample_Name %ilike% "pe") %>% group_by(Date_Num) %>% mutate(Duplicate_CV = cv(c(Dup_Conc, Conc))) %>% mutate(Dup_Flag = ifelse(Duplicate_CV >= 10, "NO, rerun", "YES")) %>% ungroup() %>% group_by(Sample_Name) %>% ungroup() %>% group_by(Sample_Name) %>% mutate(Num_Runs = length(Sample_Name), Max_Dil = max(Dilution), Rerun = ifelse(Num_Runs != 1 , ifelse(Dilution == Max_Dil, "KEEP", "REMOVE"), "KEEP")) %>% filter(Rerun == "KEEP") %>% mutate(Dilution = ifelse(Dilution == 0, 1, Dilution)) %>% 
  mutate(
     Conc_flag = case_when(
        Test == "Vanadium NOx" & Conc/Dilution < NOx_dl ~ "bdl",
        Test == "Vanadium NOx" & Conc/Dilution > NOx_top    ~ "adl",
        Test == "Vanadium NOx"               ~ "Within_Range",
      
      Test == "Ammonia 2" & Conc/Dilution < NH3_dl ~ "bdl",
      Test == "Ammonia 2" & Conc/Dilution > NH3_top    ~ "adl",
      Test == "Ammonia 2"                ~ "Within_Range",
      
      Test == "o-PHOS 0.3" & Conc/Dilution < PO4_dl  ~ "bdl",
      Test == "o-PHOS 0.3" & Conc/Dilution > PO4_top   ~ "adl",
      Test == "o-PHOS 0.3"                ~ "Within_Range",
      
      TRUE ~ NA_character_  # fallback for unexpected values
    )
  )

Dup_Plot <-  ggplot(Dup_plot_df, aes(x = Run_Date, y = Duplicate_CV, fill =  Dup_Flag, col = Conc_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "NO, rerun" = "darkred"), name = "CV below 10% ?") +
  geom_hline(yintercept = 10, linetype = "dashed" ) +
  facet_grid(.~Run_Date, scales = "free")  + 
  theme_classic() + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8)) +
   labs(title = "Duplicate Check")

Dup_Plot


Flags_Duplicates <- Dup_plot_df %>% group_by(Run_Date, Test) %>% mutate(Dup_Flags = ifelse(length(which(Dup_Flag == "YES"))/(length(Dup_Flag)) >= 0.6,  ">60% of Dups Pass - PROCEED", "<60% of Dups Pass - REASSESS")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Dup_Flags) %>% distinct()



```

\newpage

## Spikes

```{r}
Spikes <- df_all_cor %>% filter(Sample_Name == "Auto Spike") %>% mutate(Sample_Num = Run_Number - 1) %>% mutate(Date_Num = paste0(Sample_Num,"_", Run_Date, "_", Test)) %>% dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% rename(Spike_Conc = Conc, Spike_Absorbance = Absorbance, Spike_Dilution = Dilution)

Spike_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test))

Spike_Joined <- left_join(Spikes, Spike_ready, by = join_by(Date_Num))

Spike_Val <- tibble(
    Test = c("Ammonia 2", "o-PHOS 0.3"),
    Spike = c(NH3_Spk, PO4_Spk)
  )

Spike_Joined2 <- left_join(Spike_Joined, Spike_Val, by = join_by(Test))

Spike_plot_df <- Spike_Joined2 %>% filter(!Sample_Name %ilike% "ppt") %>% group_by(Date_Num) %>% mutate(Spike_Exp_Recovery = Conc + Spike) %>% 
  mutate(Spike_CV = cv(c(Spike_Exp_Recovery, Spike_Conc))) %>% 
  mutate(
    Spike_Flag = ifelse(
        Spike_CV >= 50,    #SEAL Uses 50% CV cutoff to check Spikes
      "No, rerun",
      "YES"
      )) %>% ungroup()

Spike_Plot <-  ggplot(Spike_plot_df, aes(x = Run_Date, y = Spike_CV, fill =  Spike_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "No, rerun" = "darkred"), name = "Spike CV <50% ?") +
  geom_hline(yintercept = c(50), linetype = "dashed" ) +
  facet_grid(.~Run_Date, scales = "free_x") + 
  theme_classic()  + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8)) +
   labs(title = "Spike Check")


Spike_Plot


Flags_Spike <- Spike_plot_df %>% group_by(Run_Date, Test) %>% mutate(Spike_Flags = ifelse(length(which(Spike_Flag == "YES"))/(length(Spike_Flag)) >= 0.6,  ">60% of Spikes have a CV <50% - PROCEED", ">60% of Spikes have a CV <50% - REASSESS")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Spike_Flags) %>% distinct()
```



\newpage

## Matrix Effects  

```{r}
Matrix <- df_all_cor %>% filter(Sample_Name == "Auto Spike") %>% mutate(Sample_Num = Run_Number - 1) %>% mutate(Date_Num = paste0(Sample_Num,"_", Run_Date, "_", Test)) %>% dplyr::select(Date_Num, Conc, Absorbance, Dilution) %>% rename(Spike_Conc = Conc, Spike_Absorbance = Absorbance, Spike_Dilution = Dilution)

Spike_ready <- df_all_cor %>% mutate(Date_Num = paste0(Run_Number, "_", Run_Date, "_", Test))

Spike_Joined <- left_join(Spikes, Spike_ready, by = join_by(Date_Num))

Spike_Val <- tibble(
    Test = c("Ammonia 2", "o-PHOS 0.3"),
    Spike = c(NH3_Spk, PO4_Spk)
  )

Spike_Joined2 <- left_join(Spike_Joined, Spike_Val, by = join_by(Test))

Spike_plot_df <- Spike_Joined2 %>% filter(Sample_Name %ilike% "ppt") %>% group_by(Date_Num) %>% mutate(Spike_Exp_Recovery = Conc + Spike) %>% 
  mutate(Spike_CV = cv(c(Spike_Exp_Recovery, Spike_Conc))) %>% 
  mutate(
    Spike_Flag = ifelse(
        Spike_CV >= 50,    #SEAL Uses 50% CV cutoff to check Spikes
      "No, rerun",
      "YES"
      )) %>% ungroup()

Spike_Plot <-  ggplot(Spike_plot_df, aes(x = Run_Date, y = Spike_CV, fill =  Spike_Flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single"), linewidth = 0.75) +
  scale_fill_manual(values = c("YES" = "darkgreen", "No, rerun" = "darkred"), name = "Matrix Spike CV <50% ?") +
  geom_hline(yintercept = c(50), linetype = "dashed" ) +
  facet_grid(.~Run_Date, scales = "free_x") + 
  theme_classic()  + 
    theme(legend.position = "bottom", strip.text = element_text(size = 8)) +
   labs(title = "Matrix Check") +
  ylab("Matrix Spike CV")


Spike_Plot


Flags_Matrix <- Spike_plot_df %>% group_by(Run_Date, Test) %>% mutate(Matrix_Flags = ifelse(length(which(Spike_Flag == "YES"))/(length(Spike_Flag)) >= 0.6,  "Matrix Has CV <50% - PROCEED", "Matrix Has CV >50% - REASSESS")) %>% ungroup() %>% dplyr::select(Test, Run_Date, Matrix_Flags) %>% distinct()
```



## Unit Converted Data Column Added (mg/L to uM )
```{r Add unit conversion, include=FALSE}

#Pull out samples 
samples <- df_all_cor %>%  
   filter(str_detect(Sample_Name, c("L"))) %>%  
  filter(!str_detect(Sample_Name, c("1mg/L ammonia"))) #have to do this because Phrag ids start with L   

#Convert values based on the Test ID 
samples <- samples %>%
  mutate(
    Conc_uM = case_when(
      Test == "Vanadium NOx" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "Ammonia 2" ~ (((as.numeric(samples$Conc))/Con1)/N_mw)*Con2,
      Test == "o-PHOS 0.3" ~ (((as.numeric(samples$Conc))/Con1)/P_mw)*Con2,
      TRUE ~ as.numeric(NA)  # fallback in case of unexpected value
    ),
    # Replace negatives with 0 in 'value' and 'value_converted'
    Conc = pmax(Conc, 0),
    Conc_uM = pmax(Conc_uM, 0)
  )

head(samples)

```

## Sample Flagging - Within range of standard curve
```{r Sample Flagging, echo=FALSE}

cat("Sample Flagging")

#Flagging data if the concentration is outside the standards range 
samples_limit<- samples %>% mutate(Dilution = ifelse(Dilution == 0, 1, Dilution)) %>% 
  mutate(
     Conc_flag = case_when(
        Test == "Vanadium NOx" & Conc/Dilution < NOx_dl ~ "bdl",
        Test == "Vanadium NOx" & Conc/Dilution > NOx_top    ~ "adl",
        Test == "Vanadium NOx"               ~ "Within_Range",
      
      Test == "Ammonia 2" & Conc/Dilution < NH3_dl ~ "bdl",
      Test == "Ammonia 2" & Conc/Dilution > NH3_top    ~ "adl",
      Test == "Ammonia 2"                ~ "Within_Range",
      
      Test == "o-PHOS 0.3" & Conc/Dilution < PO4_dl  ~ "bdl",
      Test == "o-PHOS 0.3" & Conc/Dilution > PO4_top   ~ "adl",
      Test == "o-PHOS 0.3"                ~ "Within_Range",
      
      TRUE ~ NA_character_  # fallback for unexpected values
    )
  )

```

## Add QAQC Flags 
```{r}


Flags_1 <- full_join(Flags_Blank, Flags_Check, by = join_by(Test, Run_Date))

Flags_2 <- full_join(Flags_1, Flags_Duplicates, by = join_by(Test, Run_Date))

Flags_3 <- full_join(Flags_2, Flags_Matrix, by = join_by(Test, Run_Date))

Flags_4 <- full_join(Flags_3, Flags_Spike, by = join_by(Test, Run_Date))

Flags_All <- full_join(Flags_4, Flags_PeCheck, by = join_by(Test, Run_Date))

Flags_Merged <- Flags_All %>% mutate(Blank_Flag = ifelse(Blank_Flag %like% "PROCEED", "", Blank_Flag),
                       CHKRSD_Flag = ifelse(CHKRSD_Flag %like% "PROCEED", "", paste0("-", CHKRSD_Flag)),
                       CHK_Flag = ifelse(CHK_Flag %like% "PROCEED", "",paste0("-", CHK_Flag)),
                       Dup_Flags = ifelse(Dup_Flags %like% "PROCEED", "", paste0("-", Dup_Flags)),
                       Matrix_Flags = ifelse(Matrix_Flags %like% "PROCEED", "", paste0("-", Matrix_Flags)),
                       Spike_Flags = ifelse(Spike_Flags %like% "PROCEED", "", paste0("-", Spike_Flags)),
                       PeCHK_Flag = ifelse(PeCHK_Flag %like% "PROCEED", "", paste0("-", PeCHK_Flag))
                       ) %>%
  mutate(NH3_flag = paste0(Blank_Flag, CHKRSD_Flag, CHK_Flag, Dup_Flags, Matrix_Flags, Spike_Flags, PeCHK_Flag)) %>% dplyr::select(Test, Run_Date, NH3_flag)
  
Flagged_Samples <- samples_limit %>% left_join(Flags_Merged, by = join_by(Test, Run_Date))
```

## Create Log for QAQC
```{r}
#Old_Log_QAQC <- tibble(Test = c(), Mean_PE_Conc= c(), Spike_CV= c(), Top_Std_Absorbance= c(), Run_Date= c()) #create standard log in proper format
Old_Log_QAQC <- read.csv(file_path_QAQC) %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% mutate(Run = "Past Runs")

Top_Standards <- Slope_raw %>% group_by(Test, Run_Date) %>% filter(Sample_Name == "Standard 2.0000") %>% dplyr::select(Test, Run_Date, Absorbance) %>% distinct() %>% ungroup() %>% rename(Top_Std_Absorbance = Absorbance)

PE_Check <- NH3_peChk %>% group_by(Test, Run_Date) %>% dplyr::select(Test, Run_Date, Mean_Chk_Conc) %>% rename(Mean_PE_Conc = Mean_Chk_Conc) %>% ungroup()

Spikes_Log <- Spike_plot_df %>% group_by(Test, Run_Date) %>% dplyr::select(Test, Run_Date, Spike_CV) %>% ungroup()

Joined_QAQC <- left_join(Top_Standards, PE_Check, by = join_by(Test, Run_Date)) %>%
  left_join(Spikes_Log, by = join_by(Test, Run_Date)) %>% mutate(Run = "Current Runs") %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% rbind(Old_Log_QAQC) %>% mutate(Run_Date = as.Date(as.character(Run_Date), tryFormats = c("%m/%d/%Y", "%Y-%m-%d"))) %>% distinct()

Spike_Log_Plot <- ggplot(Joined_QAQC, aes(x = Run_Date, y = Spike_CV)) +
  geom_point(size = 3, aes(col = Run)) +
 # geom_path(linewidth = 1) +
  facet_grid(~Test, scales = "free") +
  labs(title = "Spike") +
  theme_classic()

Spike_Log_Plot

TopStd_Log_Plot <- ggplot(Joined_QAQC, aes(x = Run_Date, y = Top_Std_Absorbance)) +
  geom_point(size = 3, aes(col = Run)) +
 # geom_path(linewidth = 1) +
  facet_grid(~Test, scales = "free") +
  labs(title = "Top Standard Absorbance") +
  theme_classic()

TopStd_Log_Plot

PE_Log_Plot <- ggplot(Joined_QAQC, aes(x = Run_Date, y = Mean_PE_Conc)) +
  geom_point(size = 3, aes(col = Run)) +
  #geom_path(linewidth = 1) +
  facet_grid(~Test, scales = "free") +
  labs(title = "Performance Check") +
  theme_classic()

PE_Log_Plot


```
##Write Log To Google Drive
```{r}

write.csv(Joined_QAQC, file_path_QAQC, row.names = FALSE)

drive_upload(file_path_QAQC, name = "SEAL_Check_QAQC_Log.csv", overwrite = TRUE)


```



## Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata 
all_present <- all(metadata$Sample_Name %in% Flagged_Samples$Sample_Name)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Flagged_Samples$Sample_Name)
  print(missing_ids)
}


#merge metadata with sample run data 
merged_data <- Flagged_Samples %>%
  left_join(metadata, by = c("Sample_Name" = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included? 

df_all_clean <- merged_data 

```


\newpage



## Organize Data
```{r Organize Data, include=FALSE}

cat("Export Processed Data")

df_all_clean_cols <- merged_data %>%
  dplyr::select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Sample_Name,
    Plot_Name,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    Conc,
    Conc_uM,
    Conc_flag,
    NH3_flag,
    Run_Date
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = run_notes,# new column with notes about the run
    Analysis_rundate = Run_Date) 

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
   NH3_Conc = Conc,
   NH3_Conc_uM = Conc_uM,
    Plot = Plot_Name,
   Sample_ID = Sample_Name
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  dplyr::select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Sample_ID,
    Plot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    NH3_Conc,
    NH3_Conc_uM,
    Conc_flag,
    NH3_flag,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)


```

## Visualize Data
```{r Visualize Data, echo=FALSE}

NH3_forplot <- All_Clean_Data

#ensure points are in correct order

NH3_forplot <- NH3_forplot[order(NH3_forplot$Depth_cm), ]

color1 <- c("L1", "L2", "L3", "L4")
color2 <- c("L5", "L6", "L8", "L10")
color3 <- c("L7", "L9", "L11", "L12")
color4 <- c("L13", "L14", "L16", "L15")
color5 <- c("L17", "L18", "L19", "L20")

NH3_forplot <- NH3_forplot %>%
  mutate(
    Plot_Color = case_when(
      Plot %in% color1 ~ "Color1",
      Plot %in% color2 ~ "Color2",
      Plot %in% color3 ~ "Color3",
      Plot %in% color4 ~ "Color4",
      Plot %in% color5 ~ "Color5"
    )
  ) %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      N_Treatment == "Amended",
    "eCO2 & eN",
    ifelse(
      CO2_Treatment == "Elevated" &
        N_Treatment == "Ambient",
      "eCO2 & aN",
      ifelse(
        CO2_Treatment == "Ambient" &
          N_Treatment == "Amended"  ,
        "aCO2 & eN" ,
        "aCO2 & aN"
      )
    )
  ))  %>%
  mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2"))  %>%
  mutate(N_Treatment = ifelse(N_Treatment == "Ambient", "aN", "eN")) 
#need to work out how best to plot these data

viz_NH3_plot2 <- ggplot(data = NH3_forplot, aes(
  x = factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")),
  y = NH3_Conc,
  fill = as.factor(Treatment),
  color = as.factor(Treatment)
)) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(padding = 0.25,preserve = "single")
  ) +
  facet_grid(Depth_cm ~ N_Treatment, scales = "fixed") +
  scale_fill_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "white",
    "aCO2 & aN" = "white"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
    
  ) +
  scale_color_manual(values = c(
    "eCO2 & eN" = "#DB5461",
    "eCO2 & aN" = "#454ADE", 
    "aCO2 & eN" = "#DB5461",
    "aCO2 & aN" = "#454ADE"
  ),
    name = "Treatment:",
    breaks = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN"),
    labels = c("eCO2 & eN", "eCO2 & aN", "aCO2 & aN", "aCO2 & eN")
  ) +
  theme_classic() +
  labs(x = " ",
       y = "NH3 (mg/L)",
       title = "CO2xN: Porewater Ammonium"
       ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  ) +
coord_flip()
print(viz_NH3_plot2)

```



```{r, echo=FALSE}
#ensure points are in correct order

#make plot
viz_NH3_plot_2 <- ggplot(data = NH3_forplot, aes(x = NH3_Conc, y = as.numeric(Depth_cm), color = Plot_Color)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(
    values = c(
      "Color1" = "#FFD500",
      "Color2" = "#00D0FF",
      "Color3" = "#FF0075",
      "Color4" = "#80d280",
      "Color5" = "#ff6040"
    ),
    labels = c("1", "2", "3", "4", "5")
  ) +
  facet_grid(Treatment ~ ., scales = "fixed") +
  labs(
    x = "NH3 (mg/L)",
    y = "Depth (cm)",
    title = "CO2xN: Porewater NH3 by Replicate",
    color = "Replicate"
  ) +
  scale_y_reverse(breaks = c(20, 40, 80), limits = c(90, 10)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_NH3_plot_2
```


## Export Processed Data
```{r Export Processed Data, include=FALSE}

#Write out data frame 
  write.csv(All_Clean_Data, final_path)
  

```


#end