---
title: "GENX: Porewater Methane"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:  
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information
```{r run information, include=TRUE}
cat("Run Information: NAME ") #lets you know what section you're in
#set the run date & user name
sample_year <- "YEAR"
sample_month <- "MONTH"
user <- "NAME"
#identify the files you want to read in
#read in as a list to accommodate multiple runs in a month

GHG_files <- c("Raw Data/20260618_GHGdata_GENX_May_Run1.csv","Raw Data/20260623_GHGdata_GENX_May_Run2.csv" )


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- here("Porewater QAQC Logs", "GCReW_CH4_QAQC_Log.csv")


# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_GENX_Porewater_CH4_202605.csv"

#Samples with Volumes not 12 mL
Sample_Vol_Data <- data.frame(Sample_Name = c(),
                          Sample_Volume = c())

#Gas in Water Calculations
# "TempC" column value in Celsius (equilibration temperature) and volume of gas (in liters) in syringe (not changing currently)
TempC = 25 #Equilibration Temperature (Celsius)

volume_g = 0.015 #volume of gas (in liters) in syringe when equilibrized
volume_w = 0.015 #volume of water (in liters) in syringe when equilibrized (should be equal to volume_g)

#record any notes about the run or anything other info here:
run_notes <- ""

#Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

cat(run_notes)
```


##Set Up

```{r setup, include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  fuzzyjoin,
  ggplot2,
  ggpubr,
  stringr,
  here,
  data.table,
  tidyverse)

##Low & High Standard Curve Cutoffs
low_std_crv_cutoff <- 1000
low_std_crv_area_cutoff <- 600000
r2_cutoff <- 0.990

##Gas In Water Calculation
R <- 0.08206
CTK <- 273.15

#Metadata file path
Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

##Detection Limits
adl_cutoff <- 20000
bdl_cutoff <- 50

##Check Standard Concentration
level_ch4 <- 1000      #ppm
chks_flag = 60
blks_flag = 60



```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in the raw metadata file 
raw_metadata <- read.csv(Raw_Metadata)

#clean up columns in metadata file:
  #Pull out only SMARTX Rows 
  #Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("GENX", Project, ignore.case = TRUE)) %>%
  mutate(Sample_Name = Sample.ID) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
    rename(
    Sample_ID = Sample.ID,
  ) %>%
  select(-N_Treatment)


```

```{r}
#Change the extension .txt to .csv on original data file, open and save it
  #If you can't see file extensions in file explorer go to view > show > file name extensions
#Data File should be more than two columns when imported
DataFile <- read.csv("20260817_GCReW_Phrag_GENX.csv", header = FALSE, row.names = NULL, sep = ",")

Final_File <- "20260817_GCReW_Phrag_GENX_formatted.csv"

RunDate <- as.Date(DataFile$V2[which(DataFile$V1 == "Acquired")[1]], format = "%m/%d/%Y %H:%M")

##CH4 Retention Time Cutoffs
CH4_Low <- 2.00
CH4_High <- 2.07

##CO2 Retention Times Cutoffs
CO2_Low <- 2.63
CO2_High <- 2.70

##N2O Retention Times Cutoffs
N2O_Low <- 2.86
N2O_High <- 2.93

##Standard Type ("Actual" or "Half")
 #LTREB, SMARTX, and GENX PW Samples use "Half", others should use "Actual"
Standard_Type <- "Half"

##Standard Concentrations (ppm)
CH4_LowTri <- 10
CO2_LowTri <- 2000
N2O_LowTri <- 5.05

CH4_MidTri <- 100
CO2_MidTri <- 20000
N2O_MidTri <- 50.2

CH4_HighTri <- 1000
CO2_HighTri <- 50000
N2O_HighTri <- 500

CH4_10K <- 10000
CO2_10K <- 0
N2O_10K <- 0

```

#Get Location of Each Sample in Main Data Frame
```{r}
Numbers <- c(which(DataFile$V1 == "[Header]"), (length(DataFile$V1)+1))
```

#Turn Each Sample into it's own Data Frame
```{r}
n_dfs <- (length(Numbers)-1)

df_list <- replicate(n_dfs, data.frame(), simplify = FALSE)

for(i in 1:(length(Numbers)-1)) {
  df_loop1 <- as.data.frame(DataFile[(Numbers[i]):(Numbers[i+1] - 1),])
  df_list[[i]] <- df_loop1
}
```

#Get Only Area Tables and Retrieve Info
```{r}

Sample_Data <- data.frame(Sample_Name = c(NA), Area1 = c(NA), RTime1 = c(NA), Area2 = c(NA), RTime2 = c(NA), Area3 = c(NA), RTime3 = c(NA), NumPeakFID = c(NA), NumPeakECD = c(NA))

for(j in 1:(length(Numbers)-1)) {
  
#FID
  df_loop2 <- df_list[[j]] #get only 1 sameple
  peak_start <- which(df_loop2$V1 == "[Peak Table(Ch1)]") #find where peak data starts
  peak_end <- which(df_loop2$V1 == "[Peak Table(Ch2)]") #find where peak data ends
  sample <- df_loop2[which(df_loop2$V1 == "Sample Name"),]$V2 #get sample name
  df_peaks <- df_loop2[peak_start:(peak_end-1),] #subset data to get only peak info
  num_peaks <- df_peaks[which(df_peaks$V1 == "# of Peaks"),]$V2 #get number of found peaks (should be 2, possibly 1 if blank/low)
  table_start <- which(df_peaks$V1 == "Peak#") #find where table headers start
  area_col <- which(df_peaks[table_start,] == "Area") #get column of area
  rtime_col <- which(df_peaks[table_start,] == "R.Time") #get column of retention time
  if(num_peaks == 0) {peak1_area <- NA #get peak 1 info if present (should be CH4)
                      peak1_rtime <- NA} else { 
         peak1_area <- df_peaks[which(df_peaks$V1 == 1), area_col]
         peak1_rtime <- df_peaks[which(df_peaks$V1 == 1), rtime_col]}
  if(num_peaks <= 1) {peak2_area <- NA #get peak 2 info if present (should be CO2)
                      peak2_rtime <- NA} else { 
         peak2_area <- df_peaks[which(df_peaks$V1 == 2), area_col]
         peak2_rtime <- df_peaks[which(df_peaks$V1 == 2), rtime_col]}
  
#ECD
  ECD_peak_start <- which(df_loop2$V1 == "[Peak Table(Ch2)]") #Find ECD Peak Data start
  ECD_peak_end <- nrow(df_loop2) #get peak data end
  df_ecd_peaks <- df_loop2[ECD_peak_start:ECD_peak_end,] #get only ecd peak info
  ecd_num_peaks <- df_ecd_peaks[which(df_ecd_peaks$V1 == "# of Peaks"),]$V2  #get number of found peaks (should be 1, possibly 0 if blank)
  ecd_table_start <- which(df_ecd_peaks$V1 == "Peak#") #find ecd table headers
  ecd_area_col <- which(df_ecd_peaks[table_start,] == "Area") #get column of area
  ecd_rtime_col <- which(df_ecd_peaks[table_start,] == "R.Time") #get column of retention time
  if(ecd_num_peaks == 0) {peak3_area <- NA #get peak info if present
                         peak3_rtime <- NA } else {
                           peak3_area <-df_ecd_peaks[which(df_ecd_peaks$V1 == 1), ecd_area_col]
                           peak3_rtime <- df_ecd_peaks[which(df_ecd_peaks$V1 == 1), ecd_rtime_col]}
##Assign Values
  Sample_Data[j,]$Sample_Name <- sample
  Sample_Data[j,]$Area1 <- peak1_area
  Sample_Data[j,]$RTime1 <- peak1_rtime
  Sample_Data[j,]$Area2 <- peak2_area
  Sample_Data[j,]$RTime2 <- peak2_rtime
  Sample_Data[j,]$Area3 <- peak3_area
  Sample_Data[j,]$RTime3 <- peak3_rtime
  Sample_Data[j,]$NumPeakFID <- num_peaks
  Sample_Data[j,]$NumPeakECD <- ecd_num_peaks
  
}

```


##QAQC Peak Count and Retention Time
```{r}
#Flag which samples have the wrong amount of peaks FID should have 2, ECD 1
PeakFlags <- Sample_Data %>% mutate(FID_Peak_Flag = ifelse(NumPeakFID == 0, "Incorrect - No Peaks", ifelse(NumPeakFID == 1, "Incorrect - One Peak", ifelse(NumPeakFID == 2, "Correct", "Incorrect - Too Many Peaks")))) %>% mutate(ECD_Peak_Flag = ifelse(NumPeakECD == 0, "Incorrect - No Peaks", ifelse(NumPeakECD == 1, "Correct", "Incorrect - Too Many Peaks")))

#Flag any retention times that are out of range for compound
RTimeFlags <- PeakFlags %>% mutate(CH4_RTime_Flag = ifelse(RTime1 >= CH4_Low & RTime1 <= CH4_High, "Correct - CH4 RTime", "Incorrect - Wrong RTime")) %>% mutate(CO2_RTime_Flag = ifelse(RTime2 >= CO2_Low & RTime2 <= CO2_High, "Correct - CO2 RTime", "Incorrect - Wrong RTime")) %>% mutate(N2O_RTime_Flag = ifelse(RTime3 >= N2O_Low & RTime3 <= N2O_High, "Correct - N2O RTime", "Incorrect - Wrong RTime" ))

#Display Flags
paste0("If there are too many peaks there is likely a problem with the sample or run, if there are too few they may be low/blanks, verify that concentration is low or get area manually")
Flagged_FID_Peaks <- ifelse(RTimeFlags$FID_Peak_Flag %ilike% "Incorrect",paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$NumPeakFID), NA)
Flagged_FID_Peaks2 <- na.omit(Flagged_FID_Peaks)
paste("FID:",Flagged_FID_Peaks2)

Flagged_ECD_Peaks <- ifelse(RTimeFlags$ECD_Peak_Flag %ilike% "Incorrect",paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$NumPeakECD), NA)
Flagged_ECD_Peaks2 <- na.omit(Flagged_ECD_Peaks)
paste("ECD:", Flagged_ECD_Peaks2)

paste0("If RTime is wrong check that correct number of peaks were found, if not a peak may have been mislabeled, if peak number is correct then there may be a problem with the run, get area manually")
CH4_Flagged_RTimes <- ifelse(RTimeFlags$CH4_RTime_Flag %ilike% "Incorrect", paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$RTime1), NA)
CH4_Flagged_RTimes2 <- na.omit(CH4_Flagged_RTimes)
paste("CH4: ", CH4_Flagged_RTimes2)

CO2_Flagged_RTimes <- ifelse(RTimeFlags$CO2_RTime_Flag %ilike% "Incorrect", paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$RTime2), NA)
CO2_Flagged_RTimes2 <- na.omit(CO2_Flagged_RTimes)
paste("CO2: ", CO2_Flagged_RTimes2)

N2O_Flagged_RTimes <- ifelse(RTimeFlags$N2O_RTime_Flag %ilike% "Incorrect", paste0(RTimeFlags$Sample_Name, "; ", RTimeFlags$RTime3), NA)
N2O_Flagged_RTimes2 <- na.omit(N2O_Flagged_RTimes)
paste("N2O: ", N2O_Flagged_RTimes2)

Flagged_Data <- RTimeFlags




```
##Format Data
```{r}
#Assign 0 to Blanks without peaks
Data_Zero <- Flagged_Data %>% mutate(Area1 =  ifelse(is.na(Area1) & Sample_Name %ilike% "Blank|0ppm", 0, Area1),
                                     Area2 = ifelse(is.na(Area2) & Sample_Name %ilike% "Blank|0ppm|10K", 0, Area2),
                                     Area3 = ifelse(is.na(Area3) & Sample_Name %ilike% "Blank|0ppm|10K", 0, Area3))

#Rename Columns
Data_RNM <- Data_Zero %>% rename(Sample_ID = Sample_Name,
                                    CH4_Area = Area1,
                                    CO2_Area = Area2,
                                    N2O_Area = Area3)

#Get Only Relevant Columns
Data_Sum <- Data_RNM %>% dplyr::select(Sample_ID,
                                       CH4_Area,
                                       CH4_RTime_Flag,
                                       CO2_Area,
                                       CO2_RTime_Flag,
                                       N2O_Area,
                                       N2O_RTime_Flag,
                                       FID_Peak_Flag,
                                       ECD_Peak_Flag)
```

##Add Standard Concentrations
```{r}

#Assign Source Standard Concentrations
Data_Stds <- Data_Sum %>% 
  mutate(CH4_STD = ifelse(Sample_ID %ilike% "HighTri", CH4_HighTri, 
                               ifelse(Sample_ID %ilike% "MidTri", CH4_MidTri, 
                                      ifelse(Sample_ID %ilike% "LowTri", CH4_LowTri, 
                                             ifelse(Sample_ID %ilike% "10K", CH4_10K, ifelse(Sample_ID %ilike% "0ppm", 0, NA)))))) %>% 
  mutate(CO2_STD = ifelse(Sample_ID %ilike% "HighTri", CO2_HighTri, 
                               ifelse(Sample_ID %ilike% "MidTri", CO2_MidTri, 
                                      ifelse(Sample_ID %ilike% "LowTri", CO2_LowTri, 
                                             ifelse(Sample_ID %ilike% "10K", CO2_10K, ifelse(Sample_ID %ilike% "0ppm", 0, NA)))))) %>% 
  mutate(N2O_STD = ifelse(Sample_ID %ilike% "HighTri", N2O_HighTri, 
                               ifelse(Sample_ID %ilike% "MidTri", N2O_MidTri, 
                                      ifelse(Sample_ID %ilike% "LowTri", N2O_LowTri, 
                                             ifelse(Sample_ID %ilike% "10K", N2O_10K, ifelse(Sample_ID %ilike% "0ppm", 0, NA)))))) %>%
  mutate(mL_STD = ifelse(Sample_ID %ilike% "Tri|10K", sub(".*_([0-9]+)mL.*", "\\1", Sample_ID), NA)) %>% #Get mL of source standard
  mutate(mL_STD = as.numeric(mL_STD), #Convert Values
         CH4_STD = as.numeric(CH4_STD),
         CO2_STD = as.numeric(CO2_STD),
         N2O_STD = as.numeric(N2O_STD)) %>% 
  mutate(Conc_CH4 = round(CH4_STD * mL_STD / 24, digits = 4), #Calculate Concentrations of Standards
         Conc_CO2 = round(CO2_STD * mL_STD / 24, digits = 4),
         Conc_N2O = round(N2O_STD * mL_STD / 24, digits = 4),)

```

##Finalize Raw Data
```{r}
raw <-Data_Stds %>% mutate(Standard = Standard_Type) %>%
  mutate(Conc_CH4 = ifelse(Standard == "Half", Conc_CH4 * 2, Conc_CH4), #Adjust standard concentrations for dilutions if necessary
  Conc_CO2 = ifelse(Standard == "Half", Conc_CO2 * 2, Conc_CO2),
   Conc_N2O = ifelse(Standard == "Half", Conc_N2O * 2, Conc_N2O)) %>% 
  mutate(Run_Date = RunDate) %>% #Add Run Date
  mutate(Type = ifelse(Sample_ID %ilike% "Std|0ppm" & !(Sample_ID %ilike% "Chk|Check"), "Standard", ifelse(Sample_ID %ilike% "Lab", "Lab air", ifelse(Sample_ID %ilike% "Blank", "Blank", ifelse(Sample_ID %ilike% "Chk|Check", "ChkStd", "Sample") )))) %>% #Add Type
  dplyr::select(Sample_ID,
                 Type,
                 Run_Date,
                 CH4_Area,
                 Conc_CH4,
                 CH4_RTime_Flag,
                 CO2_Area,
                 Conc_CO2,
                 CO2_RTime_Flag,
                 N2O_Area,
                 Conc_N2O,
                 N2O_RTime_Flag,
                 FID_Peak_Flag,
                 ECD_Peak_Flag) %>%
  mutate(Sample_Volume = ifelse(Sample_ID %in% Sample_Vol_Data$Sample_Name, sample_Vol_Data$Sample_Volume[which(Sample_ID == Sample_Vol_Data$Sample_Name)], 12)) #Select columns for export
```

# Assess standard curves
```{r , echo=FALSE}

stds_ch4 <-  raw %>%
  filter(str_detect(Type, "Standard"))

#Correct data class
stds_ch4[, "Conc_CH4"] <- as.numeric(stds_ch4[, "Conc_CH4"])

#Separate Into Low and High

mnmx_stds <- stds_ch4 %>% filter(Conc_CH4 > 0) %>% group_by(Run_Date) %>% mutate(Min = min(Conc_CH4), Max = max(Conc_CH4)) %>% select(Run_Date, Min, Max) %>% unique()

low_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 > low_std_crv_cutoff)

low_stds_ch4$Conc_CH4 <- as.numeric(low_stds_ch4$Conc_CH4)


high_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 < low_std_crv_cutoff)

high_stds_ch4$Conc_CH4 <- as.numeric(high_stds_ch4$Conc_CH4)

#High Curve for CH4
CH4_Curve <- ggplot(high_stds_ch4, aes(Conc_CH4, CH4_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High CH4 Std Curve")
CH4_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_ch4_grouped <- high_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "High",
                                        R2 = summary(lm(CH4_Area ~ Conc_CH4, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ Conc_CH4, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ Conc_CH4, data = .))[1]))

 Slope_CH4_raw <- bind_rows(Slope_CH4_raw_list)
  
 Slope_CH4_High <- Slope_CH4_raw %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()


                              
#low Curve for CH4
CH4_Curve <- ggplot(low_stds_ch4, aes(Conc_CH4, CH4_Area, color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low CH4 Std Curve")
CH4_Curve

low_CH4_lm <- lm(low_stds_ch4$CH4_Area ~ low_stds_ch4$Conc_CH4)
low_cf <- coef(low_CH4_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_ch4_grouped <- low_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list_l <- map(split_data_l, ~ .x %>% mutate(Curve = "Low",
                                        R2 = summary(lm(CH4_Area ~ Conc_CH4, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ Conc_CH4, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ Conc_CH4, data = .))[1]))

 Slope_CH4_raw_l <- bind_rows(Slope_CH4_raw_list_l)
  
 Slope_CH4_Low <- Slope_CH4_raw_l %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()
 
 Slope_CH4 <- rbind(Slope_CH4_High, Slope_CH4_Low)
 
 Slope_CH4_m <- left_join(Slope_CH4, mnmx_stds)

#add new standard curve data to log
log <- read.csv(log_path)
head(log)
log <- rbind(log, Slope_CH4)

#remove possible duplicates from multiple QAQC attempts 

unique_log <- log %>% distinct(.keep_all = TRUE)

plot_log <- unique_log %>% mutate(Run_Date = mdy(Run_Date)) %>%
  filter(Run_Date >= today() - months(12))

Slopes_chk <- ggplot(plot_log, aes(x = Run_Date, y = Slope, col = Curve)) +
  geom_point(size = 4) +
 # ylim() +
  theme_bw()

Slopes_chk

write.csv(unique_log, log_path, row.names = FALSE)

#check r2

ifelse(Slope_CH4_High$R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_CH4_High$Run_Date), paste0("High_R2 is Good","-", Slope_CH4_High$Run_Date))

ifelse(Slope_CH4_Low$R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_CH4_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_CH4_Low$Run_Date))

High_R2_Flags <- Slope_CH4_High %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)


Low_R2_Flags <- Slope_CH4_Low %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)

Flags_df <- rbind(High_R2_Flags, Low_R2_Flags)

```

## Now calculate the CH4 in ppm  
```{r Calculate Concetration, echo=FALSE}
#head(raw)

#pull out CO2xN Samples
Samples_1 <-  raw %>%  
  filter(!str_detect(Type, "Lab air")) %>%  
  filter(!str_detect(Type, "Standard")) %>%  
  filter(!str_detect(Type, "Blank")) %>%  
  filter(!str_detect(Type, "Chk_STD")) %>%  
   filter(!str_detect(Type, "ChkStd")) %>%
  filter(!str_detect(Type, "NA")) %>%
    filter(str_detect(Sample_ID, c("Ch"))) %>%
  mutate(Curve = ifelse(CH4_Area >= low_std_crv_area_cutoff, "High", "Low"))
#head(Samples)

Samples_2 <- merge(Samples_1, Slope_CH4_m)

#Calculate CH4 concentrations in ppm
Samples <- Samples_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area - Intercept) / Slope)

##########make flags for any dilutions needed 


Samples$CH4_Flag <- ifelse(Samples$CH4_Conc_ppm > Samples$Max, "ADL", ifelse(Samples$CH4_Conc_ppm < Samples$Min, "BDL", "Within Range"))

#head(Samples)

```

# Aanlayze the Check Standards 
```{r Asses Check Standards, echo=FALSE}

chks_1 <-  raw %>%
  filter(!str_detect(Type, "Lab air")) %>%
  filter(!str_detect(Type, "Standard")) %>%
  filter(!str_detect(Type, "Blank")) %>%
  filter(!str_detect(Type, "Sample")) %>%
  filter(!str_detect(Type, "NA")) %>%
  mutate(Curve = "Low")
#head(chks)


chks_2 <- merge(chks_1, Slope_CH4)


chks <- chks_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area - Intercept) / Slope)

#calculate percent difference of check standards
chks$Ch4_diff <- ((chks$CH4_Conc_ppm - level_ch4) / ((chks$CH4_Conc_ppm + level_ch4) /
                                                       2)) * 100
chks$CH4_diff_flag <-  ifelse(chks$Ch4_diff < 10, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CH4_chks_percent <- (sum(chks$CH4_diff_flag == "YES") / nrow(chks)) * 100

#report out if flags indicate need for rerun

if (CH4_chks_percent >= chks_flag) {
  message(">60% of CH4 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CH$ Check Standards are within range of expected concentration - REASSESS")
}

#now plot the Ch4 concentrations and the CO2 concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

ch4_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CH4_Conc_ppm, fill =
                                      CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: CH4") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_ch4,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


ch4_chk

#Add Check Standard Flags

flag_chk <- ifelse(CH4_chks_percent >= chks_flag,
                "",
               "; Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk))

```

# Analyze the Blanks 
```{r Analyze Blanks, echo=FALSE}
cat("Assess Blanks")

CH4_blks <-  raw %>%
  filter(!str_detect(Type, "Lab air")) %>%
  filter(!str_detect(Type, "Standard")) %>%
  filter(!str_detect(Type, "Sample")) %>%
  filter(!str_detect(Type, "Chk_STD")) %>%
  filter(!str_detect(Type, "ChkStd")) %>%
  filter(!str_detect(Type, "NA"))


#Pull out samples from df_all to calc quantile
samples_blanks <- raw %>%
  filter(str_detect(Type, c("Sample"))) %>%
  filter(str_detect(Sample_ID, c("Ch")))#have to do this because CO2xN ids start with L

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

blk_flag_CH4 <- quantile(samples_blanks$CH4_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data
CH4_blks$CH4_diff_flag <-  ifelse(CH4_blks$CH4_Area <= blk_flag_CH4, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CH4_blks_percent <- (sum(CH4_blks$CH4_diff_flag == "YES") / nrow(CH4_blks)) *
  100

#report out if flags indicate need for rerun
if (CH4_blks_percent >= blks_flag) {
  message(">60% of CH4 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CH4 blaks are lower 25% quartile of samples - REASSESS")
}

#plotting the blanks compared to the lower 25% of conc (what the flag is)
CH4_blks_plot <-  ggplot(data = CH4_blks, aes(x = Run_Date, y = CH4_Area, fill =
                                                CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CH4 Area", title = "CH4 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CH4),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(CH4_blks_plot)


#find average of run blanks for flagging samples later
# Compute averages
blk_avg_CH4 <- mean(CH4_blks$CH4_Area, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(Test = c("CH4"),
                         Blank_Mean_Area = c(blk_avg_CH4))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks <-  ifelse(CH4_blks_percent >= blks_flag,
                 "",
                 '; Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks))




```

# Dilution correct samples and Check Range
```{r, echo=FALSE}

#Correct if volume other than 12mL of Sample was added
Samples$CH4_Conc_ppm_dilcorr <- ((0.5 * Samples$CH4_Conc_ppm) * (12 + Samples$Sample_Volume)) /
  (Samples$Sample_Volume)

#check results
#head(Samples)

#quick first look at the samples
ch4_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm, fill =
                                             CH4_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range (Not Dil Cor)") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

#corrected for different dilutions
ch4_samples_cor <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm_dilcorr, fill =
                                             CH4_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "green4",
                    "ADL" = "red3",
                    "BDL" = "blue3")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

ch4_samples

ch4_samples_cor


```


# Calculate gas in water  
```{r, echo=FALSE}

# Convert ppm CH4 to total uM CH4 in the sample headspace
# use Ideal Gas Law:
#   hdsp umol CH4 = ppm CH4*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

#Calculate CH4 uM from ppm

Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

#old equation
#Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * ((1 * volume / 1000) / (R * (TempC + CTK)))) / (volume / 1000)


#quick first look at the samples
ch4_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "CH4 (uM)", title = "CH4 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


ch4_samples1

```

# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata
all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}


#merge metadata with sample run data
merged_data <- metadata %>%
  mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
  fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included?

#add flags

merged_2 <- merge(merged_data, Flags_df)
merged_data <-merged_2 %>% mutate(CH4_Flags = Flag)



df_all_clean <- merged_data

```

## Organize Data  
```{r, echo=FALSE}
#check results
  head(Samples)
 
  #pull out what we need
 
  df_all_clean_cols <- df_all_clean %>%
    select(
     Site,
     Project,
     Year,
     Month,
     Community,
     Nickname,
     Plot,
     Depth_cm,
     CO2_Treatment,
     Temp_Treatment,
      CH4_Conc_ppm_dilcorr,
      CH4_Conc_uM,
      CH4_Flag,
      CH4_Flags,
      Run_Date
    )
 
  final_data_labeled <- df_all_clean_cols %>%
    mutate(
      Run_notes = run_notes, #new column with notes about the run
      Analysis_rundate = Run_Date)
 
  #rename columns
  org_labeled_fin_data <- final_data_labeled %>%
    rename(
      CH4_Conc_ppm = CH4_Conc_ppm_dilcorr,
      CH4_Conc_Flag = CH4_Flag,
      CH4_QAQC_Flags = CH4_Flags
    )
 
  #Order Columns for Final Data Frame
  All_Clean_Data <- org_labeled_fin_data %>%
    select(
     Site,
     Project,
     Year,
     Month,
     Community,
     Nickname,
     Plot,
     Depth_cm,
     CO2_Treatment,
     Temp_Treatment,
      CH4_Conc_ppm,
      CH4_Conc_uM,
      CH4_Conc_Flag,
      CH4_QAQC_Flags,
      Analysis_rundate,
      Run_notes
    )
  head(All_Clean_Data)

```



## Visualize Data CH4
```{r Visualize Data, fig.height=10, echo=FALSE}
cat("Visualize Data")

### Ammonia 
CH4_forplot <- df_all_clean

CH4_forplot$Depth_cm <- 
  as.numeric(CH4_forplot$Depth_cm)

#group the data for plotting
CH4_forplot <- CH4_forplot %>%
  group_by(Depth_cm) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup() %>%
  mutate(Treatment = paste0("+", Temp_Treatment, " " ,"C")) %>% mutate(Chamber = factor(Plot, levels = c(
      "Ch1", "Ch2", "Ch3", "Ch4", "Ch5", "Ch6", "Ch7", "Ch8", "Ch9", "Ch10", "Ch11", "Ch12"))) %>% mutate(Depth1 = paste0(Depth_cm, " cm")) %>% mutate(Depth = factor(Depth1, levels = c("5 cm", "20 cm", "40 cm", "80 cm")))

#create plot
viz_CH4_plot <- ggplot(data = CH4_forplot, aes(
  x = as.factor(Chamber),
  y = CH4_Conc_uM,
  fill = as.factor(Chamber)
)) +
  geom_col(stat = "identity", 
           position = position_dodge2(padding = 0.1, 
                                      width = 0.5, 
                                      preserve =  "single" )) +
  facet_grid(Depth ~ ., 
             scales = "free_x") +
  scale_fill_manual(
    breaks = c(
       "Ch1", "Ch2", "Ch3", "Ch4", "Ch5", "Ch6", "Ch7", "Ch8", "Ch9", "Ch10", "Ch11", "Ch12"),
    values = c(
       "Ch1" = "blue4",
       "Ch2" = "blue3",
       "Ch3" = "turquoise4",
       "Ch4" = "lightseagreen",
       "Ch5" = "mediumseagreen",
       "Ch6" = "limegreen",
       "Ch7" = "yellowgreen",
       "Ch8" = "yellow2",
       "Ch9" = "darkgoldenrod2",
       "Ch10" = "darkorange2",
       "Ch11" = "orangered1",
       "Ch12" = "red2"),
    name = "Chamber:"
  ) +
  theme_classic() +
  labs(
    x = " ",
    y = "CH4 Conc (uM)",
    title = "GENX: Porewater CH4"
  ) +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE, limits = rev) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8),
      legend.key = element_rect("black"),
    legend.position = "none"
    ) + 
  coord_flip()
print(viz_CH4_plot)

```
\newpage

\newpage
```{r Visualize Data by Rep,  fig.height=8.5, echo=FALSE}

#get replicate value and treatment info
CH4_forplot <- CH4_forplot %>%
  mutate(Chamber = case_when(
    Plot == "Ch1"  ~ "Chamber 1",
    Plot == "Ch2"  ~ "Chamber 2",
    Plot == "Ch3"  ~ "Chamber 3",
    Plot == "Ch4"  ~ "Chamber 4",
    Plot == "Ch5"  ~ "Chamber 5",
    Plot == "Ch6"  ~ "Chamber 6",
    Plot == "Ch7"  ~ "Chamber 7",
    Plot == "Ch8"  ~ "Chamber 8",
    Plot == "Ch9"  ~ "Chamber 9",
    Plot == "Ch10"  ~ "Chamber 10",
    Plot == "Ch11"  ~ "Chamber 11",
    Plot == "Ch12"  ~ "Chamber 12"
    
    
  )
           ) %>% mutate(Chamber = factor(levels = c(
             "Chamber 1",
             "Chamber 2",
             "Chamber 3",
             "Chamber 4",
             "Chamber 5",
             "Chamber 6",
             "Chamber 7",
             "Chamber 8",
             "Chamber 9",
             "Chamber 10",
             "Chamber 11",
             "Chamber 12"
           ), Chamber)) %>% arrange(Chamber, Depth_cm)

#make plot
viz_CH4_plot_2 <- ggplot(data = CH4_forplot, 
                            aes(x = CH4_Conc_uM, y = Depth_cm, color = Chamber)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(breaks = c(
       "Ch1", "Ch2", "Ch3", "Ch4", "Ch5", "Ch6", "Ch7", "Ch8", "Ch9", "Ch10", "Ch11", "Ch12"),
    values = c(
       "Chamber 1" = "blue4",
       "Chamber 2" = "blue3",
       "Chamber 3" = "turquoise4",
       "Chamber 4" = "lightseagreen",
       "Chamber 5" = "mediumseagreen",
       "Chamber 6" = "limegreen",
       "Chamber 7" = "yellowgreen",
       "Chamber 8" = "yellow2",
       "Chamber 9" = "darkgoldenrod2",
       "Chamber 10" = "darkorange2",
       "Chamber 11" = "orangered1",
       "Chamber 12" = "red2"),
    name = "Chamber:") +
  facet_wrap(~Chamber, ncol = 3, nrow = 4, scales = "free_x") +
  labs(x = "CH4 Conc (uM)", 
       y = "Depth (cm)", 
       title = "GENX: Porewater CH4 by Chamber") +
  scale_y_reverse(breaks = c(5, 20, 40, 80),
                  limits = c(85, 0)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8),
      legend.position = "none"
    )


viz_CH4_plot_2

```

##Export Data
```{r, echo=FALSE}
write.csv(All_Clean_Data, final_path)
#end
```

#end