---
title: "CO2xCOMM: Porewater Methane"
author: "Month Year Samples"
date: "`r Sys.Date()`"
output:  
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
output_dir: "To Be Reviewed/PDF"
---

\newpage

##Run Information
```{r run information, include=TRUE}
cat("Run Information: Brenden D. Blakley ") #lets you know what section you're in
#set the run date & user name
sample_year <- "2025"
sample_month <- "July"
user <- "Brenden Blakley"
#identify the files you want to read in
#read in as a list to accommodate multiple runs in a month

GHG_files <- c("Raw Data/20250821_GHGdata_LTREB_2M_1M.csv", "Raw Data/20250911_GHGdata_LTREB_1M.csv" )


# Define the file path for QAQC log file - NO Need to change just check year
log_path <- "Processed Data/LTREB_1M_ShimadzuGC_QAQC_2025.csv"

# Define final path for data be sure to change year and month and run number
final_path <- "Processed Data/GCReW_CO2xCOMM_Porewater_CH4_202507.csv"

#Gas in Water Calculations
# "TempC" column value in Celsius (equilibration temperature) and volume of gas (in liters) in syringe
TempC = 25 #Equilibration Temperature (Celsius)

volume_g = 0.015 #volume of gas (in liters) in syringe when equilibrized
volume_w = 0.015 #volume of water (in liters) in syringe when equilibrized (should be equal to volume_g)

#record any notes about the run or anything other info here:
run_notes <- ""

#Collection_Dates = "Raw Data/Sample_Collection_Dates_2024.csv"

cat(run_notes)
```


##Set Up

```{r setup, include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr,
  purrr,
  fuzzyjoin,
  ggplot2,
  ggpubr,
  stringr,
  here)

#Metadata file path
Raw_Metadata = here(".." , "GCReW", "GCReW_Project_Treatment_Metadata.csv")

##Low & High Standard Curve Cutoffs
low_std_crv_cutoff <- 1000
low_std_crv_area_cutoff <- 600000
r2_cutoff <- 0.990

##Gas In Water Calculation
R <- 0.08206
CTK <- 273.15



##Detection Limits (ppm)
adl_cutoff <- 20000
bdl_cutoff <- 50

##Check Standard Concentration
level_ch4 <- 1000      #ppm
chks_flag = 60
blks_flag = 60

#initialize flags
flags <- ""


```


##Read in metadata and create similar sample IDs for matching to samples 
```{r pull in metadata for later, include=FALSE}

#read in metadata

raw_metadata <- read.csv(Raw_Metadata)
#clean up columns in metadata file:
#Pull out only 1 Meter Rows
#Create a Sample_Name column to match the SEAL
metadata <- raw_metadata %>%
  filter(grepl("1-Meter", Nickname, ignore.case = TRUE)) %>%
  mutate(Sample_Name = paste0(Sample.ID)) %>%
  mutate(Month = sample_month) %>%
  mutate(Year = sample_year) %>%
  rename(
    Sample_ID = Sample.ID,
    Plot_Name = Plot,
  )

#Bring in the collection date and any field notes if relevant
#collection_metadata <- read.csv(Collection_Dates)
#then we will merge the collection date for project based on the project and the chambers

```

# Read in data files
```{r , echo=FALSE}
#Read in Nov Data
raw_list <- map(GHG_files, read.csv)


# Combining Files
raw <- bind_rows(raw_list)



#rename column
colnames(raw)[1] <- "Sample_ID"

#head(stds_ch4) 

```

# Assess standard curves
```{r , echo=FALSE}

stds_ch4 <-  raw %>%
  filter(str_detect(Type, "Standard"))

#Correct data class
stds_ch4[, "Conc_CH4"] <- as.numeric(stds_ch4[, "Conc_CH4"])

#Separate Into Low and High

mnmx_stds <- stds_ch4 %>% filter(Conc_CH4 > 0) %>% group_by(Run_Date) %>% mutate(Min = min(Conc_CH4), Max = max(Conc_CH4)) %>% select(Run_Date, Min, Max) %>% unique()

low_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 > low_std_crv_cutoff)

low_stds_ch4$Conc_CH4 <- as.numeric(low_stds_ch4$Conc_CH4)


high_stds_ch4 <- stds_ch4  %>%
  filter(!Conc_CH4 < low_std_crv_cutoff)

high_stds_ch4$Conc_CH4 <- as.numeric(high_stds_ch4$Conc_CH4)

#High Curve for CH4
CH4_Curve <- ggplot(high_stds_ch4, aes(Conc_CH4, CH4_Area, color = Run_Date)) + geom_point(size = 4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "High CH4 Std Curve")
CH4_Curve

 # Convert 'run date' to a factor for grouping
    high_stds_ch4_grouped <- high_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

    
  split_data <- high_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list <- map(split_data, ~ .x %>% mutate(Curve = "High",
                                        R2 = summary(lm(CH4_Area ~ Conc_CH4, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ Conc_CH4, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ Conc_CH4, data = .))[1]))

 Slope_CH4_raw <- bind_rows(Slope_CH4_raw_list)
  
 Slope_CH4_High <- Slope_CH4_raw %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()


                              
#low Curve for CH4
CH4_Curve <- ggplot(low_stds_ch4, aes(Conc_CH4, CH4_Area, color = Run_Date)) + geom_point(size =
                                                                          4) +
  geom_smooth(method = "lm", se = FALSE) + labs(title = "Low CH4 Std Curve")
CH4_Curve

low_CH4_lm <- lm(low_stds_ch4$CH4_Area ~ low_stds_ch4$Conc_CH4)
low_cf <- coef(low_CH4_lm)

 # Convert 'run date' to a factor for grouping
    low_stds_ch4_grouped <- low_stds_ch4 %>%
      mutate(Run_Date = as.factor(Run_Date))

  
  split_data_l <- low_stds_ch4_grouped %>%
      group_by(Run_Date) %>%
      group_split()
  
  
#run regression and get coefficients by Run
 
Slope_CH4_raw_list_l <- map(split_data_l, ~ .x %>% mutate(Curve = "Low",
                                        R2 = summary(lm(CH4_Area ~ Conc_CH4, data = .))$adj.r.squared , 
                                        Slope = coef(lm(CH4_Area ~ Conc_CH4, data = .))[2],
                                        Intercept = coef(lm(CH4_Area ~ Conc_CH4, data = .))[1]))

 Slope_CH4_raw_l <- bind_rows(Slope_CH4_raw_list_l)
  
 Slope_CH4_Low <- Slope_CH4_raw_l %>% select(Run_Date, Curve, R2, Slope, Intercept) %>% distinct()
 
 Slope_CH4 <- rbind(Slope_CH4_High, Slope_CH4_Low)
 
 Slope_CH4_m <- left_join(Slope_CH4, mnmx_stds)

#add new standard curve data to log
log <- read.csv(log_path)
head(log)
log <- log[, -c(1)]
log <- rbind(log, Slope_CH4)

#remove possible duplicates from multiple QAQC attempts 

unique_log <- log %>% distinct(.keep_all = TRUE)

Slopes_chk <- ggplot(log, aes(x = Run_Date, y = Slope, col = Curve)) +
  geom_point(size = 4) +
  ylim(400, 800) +
  theme_bw()

Slopes_chk

write.csv(unique_log, log_path)

#check r2

ifelse(Slope_CH4_High$R2 < r2_cutoff, paste0("High_R2 is Bad","-", Slope_CH4_High$Run_Date), paste0("High_R2 is Good","-", Slope_CH4_High$Run_Date))

ifelse(Slope_CH4_Low$R2 < r2_cutoff, paste0("Low_R2 is Bad","-", Slope_CH4_Low$Run_Date), paste0("Low_R2 is Good","-", Slope_CH4_Low$Run_Date))

High_R2_Flags <- Slope_CH4_High %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)


Low_R2_Flags <- Slope_CH4_Low %>% mutate(Flag = ifelse(R2 < r2_cutoff,"R2 is Bad", "")) %>% select(Flag, Run_Date, Curve)

Flags_df <- rbind(High_R2_Flags, Low_R2_Flags)

```

## Now calculate the CH4 in ppm  
```{r Calculate Concetration, echo=FALSE}
#head(raw)

#pull out CO2xN Samples
Samples_1 <-  raw %>%  
  filter(!str_detect(Type, "Lab air")) %>%  
  filter(!str_detect(Type, "Standard")) %>%  
  filter(!str_detect(Type, "Blank")) %>%  
  filter(!str_detect(Type, "Chk_STD")) %>%  
   filter(!str_detect(Type, "ChkStd")) %>%
  filter(!str_detect(Type, "NA")) %>%
    filter(str_detect(Sample_ID, c("S"))) %>%
  mutate(Curve = ifelse(CH4_Area >= low_std_crv_area_cutoff, "High", "Low"))
#head(Samples)

Samples_2 <- merge(Samples_1, Slope_CH4_m)

#Calculate CH4 concentrations in ppm
Samples <- Samples_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area - Intercept) / Slope)

##########make flags for any dilutions needed 


Samples$CH4_Flag <- ifelse(Samples$CH4_Conc_ppm > Samples$Max, "ADL", ifelse(Samples$CH4_Conc_ppm < Samples$Min, "BDL", "Within Range"))

#head(Samples)

```

# Aanlayze the Check Standards 
```{r Asses Check Standards, echo=FALSE}

chks_1 <-  raw %>%
  filter(!str_detect(Type, "Lab air")) %>%
  filter(!str_detect(Type, "Standard")) %>%
  filter(!str_detect(Type, "Blank")) %>%
  filter(!str_detect(Type, "Sample")) %>%
  filter(!str_detect(Type, "NA")) %>%
  mutate(Curve = "Low")
#head(chks)


chks_2 <- merge(chks_1, Slope_CH4)


chks <- chks_2 %>% 
  mutate(CH4_Conc_ppm = (CH4_Area - Intercept) / Slope)

#calculate percent difference of check standards
chks$Ch4_diff <- ((chks$CH4_Conc_ppm - level_ch4) / ((chks$CH4_Conc_ppm + level_ch4) /
                                                       2)) * 100
chks$CH4_diff_flag <-  ifelse(chks$Ch4_diff < 10, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CH4_chks_percent <- (sum(chks$CH4_diff_flag == "YES") / nrow(chks)) * 100

#report out if flags indicate need for rerun

if (CH4_chks_percent >= chks_flag) {
  message(">60% of CH4 Check Standards are within range of expected concentration - PROCEED")
} else {
  message("<60% of CH$ Check Standards are within range of expected concentration - REASSESS")
}

#now plot the Ch4 concentrations and the CO2 concentrations vs. the expected concentration
#then also make the color the percent difference between the expected and observed concentration

ch4_chk <-  ggplot(data = chks, aes(x = Run_Date, y = CH4_Conc_ppm, fill =
                                      CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c("NO, rerun" = "darkgrey", "YES" = "darkgreen")) +
  theme_classic() + labs(x = " ", y = "Concentration (ppm)", title =
                           "Check Stds: CH4") +
  theme(legend.position = "bottom") +  geom_hline(
    yintercept = level_ch4,
    linetype = "dashed",
    color = "black",
    linewidth = 1
  ) +
  guides(fill = guide_legend(title = "% Difference <10%"))


ch4_chk

#Add Check Standard Flags

flag_chk <- ifelse(CH4_chks_percent >= chks_flag,
                "",
               "; Check Standards out Of Range")

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_chk))

```

# Analyze the Blanks 
```{r Analyze Blanks, echo=FALSE}
cat("Assess Blanks")

CH4_blks <-  raw %>%
  filter(!str_detect(Type, "Lab air")) %>%
  filter(!str_detect(Type, "Standard")) %>%
  filter(!str_detect(Type, "Sample")) %>%
  filter(!str_detect(Type, "Chk_STD")) %>%
  filter(!str_detect(Type, "ChkStd")) %>%
  filter(!str_detect(Type, "NA"))


#Pull out samples from df_all to calc quantile
samples_blanks <- raw %>%
  filter(str_detect(Type, c("Sample"))) %>%
  filter(str_detect(Sample_ID, c("S"))) 

#Calculating the lowest 25% of sample concentrations to compare to the blank concentrations

blk_flag_CH4 <- quantile(samples_blanks$CH4_Area, prob = c(.25))   #this gives you the lower 25% quartile of the data
CH4_blks$CH4_diff_flag <-  ifelse(CH4_blks$CH4_Area <= blk_flag_CH4, 'YES', 'NO, rerun')

#calculate the percent of check standards that are within the range based on the flag
CH4_blks_percent <- (sum(CH4_blks$CH4_diff_flag == "YES") / nrow(CH4_blks)) *
  100

#report out if flags indicate need for rerun
if (CH4_blks_percent >= blks_flag) {
  message(">60% of CH4 Blanks are below the lower 25% quartile of samples - PROCEED")
} else {
  message("<60% of CH4 blaks are lower 25% quartile of samples - REASSESS")
}

#plotting the blanks compared to the lower 25% of conc (what the flag is)
CH4_blks_plot <-  ggplot(data = CH4_blks, aes(x = Run_Date, y = CH4_Area, fill =
                                                CH4_diff_flag)) +
  geom_bar(stat = 'identity',
    position = position_dodge2(preserve = "single")) +
  scale_fill_manual(values = c(
    "YES" = "deepskyblue3",
    "NO, rerun" = "darkred"
  )) +
  theme_classic() + labs(x = " ", y = "CH4 Area", title = "CH4 Blanks") +
  theme(legend.position = "bottom") +
  geom_hline(
    yintercept = as.numeric(blk_flag_CH4),
    linetype = "dashed",
    color = "black",
    linewidth = 1
  )  +
  guides(fill = guide_legend(title = "Blank Conc <25% Quartile Samples"))

print(CH4_blks_plot)


#find average of run blanks for flagging samples later
# Compute averages
blk_avg_CH4 <- mean(CH4_blks$CH4_Area, na.rm = TRUE)


# Create a data frame
blank_avgs <- data.frame(Test = c("CH4"),
                         Blank_Mean_Area = c(blk_avg_CH4))

# Pretty print
knitr::kable(blank_avgs, caption = "Mean Area of Blanks", digits = 4)


#write out a flag if more than 60% of the blanks are above the lower 25% quantile of samples
flag_blks <-  ifelse(CH4_blks_percent >= blks_flag,
                 "",
                 '; Blanks out of Range')

Flags_df <- Flags_df %>% mutate(Flag = paste0(Flag, flag_blks))




```

# Dilution correct samples and Check Range
```{r, echo=FALSE}

#Correct if volume other than 12mL of Sample was added
Samples$CH4_Conc_ppm_dilcorr <- ((0.5 * Samples$CH4_Conc_ppm) * (12 + Samples$Sample_Volume)) /
  (Samples$Sample_Volume)

#check results
#head(Samples)

#quick first look at the samples
ch4_samples <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm, fill =
                                             CH4_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "darkgreen",
                    "ADL" = "darkred",
                    "BDL" = "darkblue")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range (Not Dil Cor)") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

#corrected for different dilutions
ch4_samples_cor <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_ppm_dilcorr, fill =
                                             CH4_Flag)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c("Within Range" = "green4",
                    "ADL" = "red3",
                    "BDL" = "blue3")) +
  theme_classic() + labs(x = " ", y = "CH4 (ppm)", title = "CH4: Green = Within Range") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")

ch4_samples

ch4_samples_cor


```


# Calculate gas in water  
```{r, echo=FALSE}

# Convert ppm CH4 to total uM CH4 in the sample headspace
# use Ideal Gas Law:
#   hdsp umol CH4 = ppm CH4*(PV/RT)
#      P = 1 (pressure in atm)
#      R = 0.082 (gas constant)
#      V = 0.015 (gas syringe volume in L)
#      T = "TempC" column value in Kelvin (equilibration temperature)

#Calculate CH4 uM from ppm

Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * (( 1 * volume_g )/(R * (TempC + CTK) )))/ volume_w

#old equation
#Samples$CH4_Conc_uM <- (Samples$CH4_Conc_ppm_dilcorr * ((1 * volume / 1000) / (R * (TempC + CTK)))) / (volume / 1000)


#quick first look at the samples
ch4_samples1 <-  ggplot(data = Samples, aes(x = Sample_ID, y = CH4_Conc_uM)) +
  geom_bar(stat = 'identity', fill = "darkblue") +
  theme_classic() +
  labs(x = " ", y = "CH4 (umol/L)", title = "CH4 Gas in Water") +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  theme(legend.position = "none")


ch4_samples1

```

# Merge samples with Metadata & check all samples are present 
```{r check sample ids with metadata, echo=FALSE}

cat("Merge Samples with Metadata")

#check to see if all samples are present in the metadata
all_present <- all(metadata$Sample_Name %in% Samples$Sample_ID)

if (all_present) {
  message("All sample IDs are present in data")
} else {
  message("Some sample IDs are missing from data.")
  
  # Optional: Which ones are missing?
  missing_ids <- setdiff(metadata$Sample_Name, Samples$Sample_ID)
  print(missing_ids)
}


#merge metadata with sample run data
merged_data <- metadata %>%
  mutate(Sample_ID = paste0("\\b", Sample_ID, "\\b")) %>%
  fuzzyjoin::regex_left_join(Samples, ., by = c(Sample_ID = "Sample_Name"))

#might want to include a flag or something to see that all the samples are included?

#add flags

merged_data$CH4_Flags <- flags

df_all_clean <- merged_data

```



## Organize Data  
```{r, echo=FALSE}
#check results
#head(Samples)

#pull out what we need

df_all_clean_cols <- df_all_clean %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot_Name,
    Depth_cm,
    Run_Date,
    CO2_Treatment,
    N_Treatment,
    CH4_Conc_ppm_dilcorr,
    CH4_Conc_uM,
    CH4_Flag,
    CH4_Flags
  )

final_data_labeled <- df_all_clean_cols %>%
  mutate(
    Run_notes = run_notes,# new column with notes about the run
    Analysis_rundate = Run_Date 
  )

#rename columns
org_labeled_fin_data <- final_data_labeled %>%
  rename(
    CH4_Conc_ppm = CH4_Conc_ppm_dilcorr,
    CH4_Conc_Flag = CH4_Flag,
    CH4_QAQC_Flags = CH4_Flags,
    Plot = Plot_Name
  )

#Order Columns for Final Data Frame
All_Clean_Data <- org_labeled_fin_data %>%
  select(
    Site,
    Project,
    Year,
    Month,
    Community,
    Nickname,
    Plot,
    Depth_cm,
    CO2_Treatment,
    N_Treatment,
    CH4_Conc_ppm,
    CH4_Conc_uM,
    CH4_Conc_Flag,
    CH4_QAQC_Flags,
    Analysis_rundate,
    Run_notes
  )
#head(Samples1)


```

## Visualize Data
```{r Visualize Data, echo=FALSE}
cat("Visualize Data")

### Ammonia 
CH4_forplot <- All_Clean_Data

#need to work out how best to plot these data 

 CH4_forplot <- CH4_forplot %>%
  mutate(Treatment = ifelse(
    CO2_Treatment == "Elevated" &
      Community == "SPPA",
    "eCO2_SPPA",
    ifelse(
      CO2_Treatment == "Elevated" &
        Community == "SCAM",
      "eCO2_SCAM",
      ifelse(CO2_Treatment == "Ambient" &
               Community == "SPPA"  , "aCO2_SPPA" , "aCO2_SCAM")
    )
  )) %>%
  mutate(CO2_Treatment = ifelse(CO2_Treatment == "Ambient", "aCO2", "eCO2")) %>%
  mutate(Order = ifelse(CO2_Treatment == "Ambient", "1", "2"))
 
 CH4_forplot <- CH4_forplot[order(CH4_forplot$CO2_Treatment), ]
 
viz_CH4_plot2 <- ggplot(data = CH4_forplot, aes(
  x =  factor(as.factor(CO2_Treatment), levels = c("eCO2", "aCO2")), 
  y = CH4_Conc_uM,
  color = as.factor(Treatment),
  fill = as.factor(Treatment)
)) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(padding = 0.5, preserve = "single")
  ) +
  facet_grid(Depth_cm ~ ., scales = "fixed", space = "free_x") +
  scale_fill_manual(
    values = c(
    "eCO2_SPPA" = "#EAC435",
    "eCO2_SCAM" = "#149911", 
    "aCO2_SPPA" = "white",
    "aCO2_SCAM" = "white"
  ),
    name = "Treatment:",
    breaks = c("eCO2_SPPA", "eCO2_SCAM", "aCO2_SPPA", "aCO2_SCAM"),
    labels = c("eCO2_SPPA", "eCO2_SCAM", "aCO2_SPPA", "aCO2_SCAM") 
  ) +
  scale_color_manual(values = c(
    "eCO2_SPPA" = "#EAC435",
    "eCO2_SCAM" = "#149911", 
    "aCO2_SPPA" = "#EAC435",
    "aCO2_SCAM" = "#149911"
  ),
    name = "Treatment:",
    breaks = c("eCO2_SPPA", "eCO2_SCAM", "aCO2_SPPA", "aCO2_SCAM"),
    labels = c("eCO2_SPPA", "eCO2_SCAM", "aCO2_SPPA", "aCO2_SCAM") 
  ) +
  theme_classic() +
  labs(x = " ",
       y = "CH4 (uM)",
       title = "CO2xCOMM: Porewater Methane",
       fill = "CO2 Treatment:",
       color = "Community:") +
  theme(legend.position = "right") +
  scale_x_discrete(drop = TRUE) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  ) +
coord_flip()
print(viz_CH4_plot2)


```
```{r, echo = FALSE}
#ensure points are in correct order
CH4_forplot <- CH4_forplot[order(CH4_forplot$Depth_cm), ]

color1 <- c("SC1", "SC2", "SP1", "SP2")
color2 <- c("SC6", "SC4", "SP4", "SP5")
color3 <- c("SC9", "SC8", "SP8", "SP9")
color4 <- c("SC10", "SC12", "SP10", "SP11")
color5 <- c("SC14", "SC15", "SP14", "SP13")

CH4_forplot <- CH4_forplot %>%
  mutate(
    Plot_Color = case_when(
      Plot %in% color1 ~ "Color1",
      Plot %in% color2 ~ "Color2",
      Plot %in% color3 ~ "Color3",
      Plot %in% color4 ~ "Color4",
      Plot %in% color5 ~ "Color5"
    )
  )

#make plot
viz_CH4_plot_2 <- ggplot(data = CH4_forplot, aes(x = CH4_Conc_uM, y = as.numeric(Depth_cm), color = Plot_Color)) +
  geom_point(size = 3) +
  geom_path(linewidth = 1) +
  scale_color_manual(
    values = c(
      "Color1" = "#FFD500",
      "Color2" = "#00D0FF",
      "Color3" = "#FF0075",
      "Color4" = "#80d280",
      "Color5" = "#ff6040"
    ),
    labels = c("1", "2", "3", "4", "5")
  ) +
  facet_grid(Treatment ~ ., scales = "fixed") +
  labs(
    x = "CH4 (uM)",
    y = "Depth (cm)",
    title = "CO2xCOMM: Porewater CH4 by Replicate",
    color = "Replicate"
  ) +
  scale_y_reverse(breaks = c(20, 40, 80), limits = c(90, 10)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.8
    )
  )

viz_CH4_plot_2
```


```{r, echo = FALSE}
write.csv(All_Clean_Data, final_path)
```


#end
